---
title: "06-08-2025"
author: "Antsa"
format: html
editor: visual
---

# I. DATA ACQUISITION AND PREPARATION

## I.1. Initial Data Loading

Loading essential R packages (dplyr, ggplot2, corrplot, etc.) and importing the leaf dataset with proper French encoding. The dataset uses semicolons as separators and commas for decimals, requiring specific import parameters.

```{r}
# Load necessary packages
library(pacman)
p_load(
  dplyr, readr, lubridate, tidyr, stringr, skimr,
  VIM, summarytools, corrplot, janitor, purrr
)

# Set working directory
setwd("C:/Users/U078-N800/Desktop/Ml_Internship/Copy")

# Load the dataset (semicolon as separator and comma for decimal)
leaf_data <- read.csv("BD_Leaf.csv", sep = ";", dec = ",", header = TRUE, stringsAsFactors = FALSE , fileEncoding = "latin1")


# Display basic information about the dataset
cat("Dataset dimensions:", dim(leaf_data), "\n")
cat("Number of rows:", nrow(leaf_data), "\n")
cat("Number of columns:", ncol(leaf_data), "\n")

# Display column names
cat("\nColumn names:\n")
print(colnames(leaf_data))

```

## I.2. Data Types Overview

```{r}
# 2. DATA TYPES OVERVIEW
# ===============================
cat("2. DATA TYPES OVERVIEW\n")
cat("======================\n")
str(leaf_data)
cat("\n")

# Show first few rows
cat("First 3 rows of the dataset:\n")
print(head(leaf_data, 3))
cat("\n")

str(leaf_data)
```

## I.3. Data Type Identification and Conversion

```{r}
# ===============================
cat("2. DATA TYPE IDENTIFICATION AND CONVERSION\n")
cat("==========================================\n")

# Define data types based on domain knowledge
factor_variables <- c(
  "Campagne",           # Campaign
  "Note",              # Note
  "Parc.",             # Parcelle
  "Ferme",             # Farm
  "Variete",           # Variety
  "Cycle",             # Cycle
  "Periode",           # Period
  "Validation_periode", # Period validation
  "Parcelle_de_campagne", # Campaign parcelle
  "Validation_recolte", # Harvest validation
  "Irr.",              # Irrigation type
  "Mois.recolte",      # Harvest month
  "Mois_prelevement"   # Sampling month
)

# Date variables (with their meanings)
date_variables <- c(
  "Impl",              # Date d'implantation (planting date)
  "Prel.",             # Date de prélèvement des samples (sampling date)
  "Date_analyse",      # Analysis date
  "Date_recolte",     # Harvest date
  "Date.commun.récolte"
)

# Numerical variables
numerical_variables <- c(
  "Surf",              # Surface
  "A_ce_jour",         # To this day
  "Tc.Ha_A.1",         # Tc/Ha A-1
  "RH.C_A.1",          # RH%C A-1
  "Age_prel.",         # Age at sampling
  "H2O_feuilles",      # Leaf water content
  "H20_gaines",        # Sheath water content
  "N.",                # Nitrogen %
  "X.P",               # Phosphorus %
  "X.K",               # Potassium %
  "Tc.Ha_A",           # Tc/Ha A
  "RH.C_A",            # RH%C A
  "T_canne_reelle",    # Real cane tons
  "T_pol_canne",       # Pol tons per cane
  "Age.en.Sem.",       # Age in weeks
  "Age.recolte.A",     # Harvest age A
  "Surface.A",         # Surface A
  "Age.des.surf",      # Surface age
  "Num_Ssemaine",      # Week number
  "A.ce.jour"
)

# Check which variables actually exist in the dataset
existing_factors <- intersect(factor_variables, names(leaf_data))
existing_dates <- intersect(date_variables, names(leaf_data))
existing_numerics <- intersect(numerical_variables, names(leaf_data))

cat("FACTOR VARIABLES (", length(existing_factors), " found):\n")
print(existing_factors)
cat("\nDATE VARIABLES (", length(existing_dates), " found):\n")
print(existing_dates)
cat("\nNUMERICAL VARIABLES (", length(existing_numerics), " found):\n")
print(existing_numerics)

# Variables not classified
all_classified <- c(existing_factors, existing_dates, existing_numerics)
unclassified <- setdiff(names(leaf_data), all_classified)
if(length(unclassified) > 0) {
  cat("\nUNCLASSIFIED VARIABLES:\n")
  print(unclassified)
  
  # Suggest classification for unclassified variables based on their content
  cat("\nSUGGESTED CLASSIFICATION FOR UNCLASSIFIED VARIABLES:\n")
  for(var in unclassified) {
    sample_values <- head(unique(leaf_data[[var]][!is.na(leaf_data[[var]]) & leaf_data[[var]] != ""]), 5)
    cat("Variable:", var, "| Sample values:", paste(sample_values, collapse = ", "), "\n")
    
    # Try to guess the type
    non_empty <- leaf_data[[var]][!is.na(leaf_data[[var]]) & leaf_data[[var]] != ""]
    if(length(non_empty) > 0) {
      # Check if it looks numeric (allowing for comma as decimal separator)
      numeric_test <- gsub(",", ".", non_empty)
      if(all(str_detect(numeric_test, "^-?[0-9]*\\.?[0-9]+$|^$"), na.rm = TRUE)) {
        cat("  -> Suggested: NUMERIC\n")
      } else if(length(unique(non_empty)) < 20 && length(non_empty) > 10) {
        cat("  -> Suggested: FACTOR\n")
      } else {
        cat("  -> Suggested: CHARACTER/OTHER\n")
      }
    }
  }
}

# ===============================
# 3. CONVERT DATA TYPES
# ===============================
cat("\n3. DATA TYPE CONVERSION\n")
cat("=======================\n")

# Create a copy for conversion
leaf_data_converted <- leaf_data

# Convert factors
for(var in existing_factors) {
  if(var %in% names(leaf_data_converted)) {
    leaf_data_converted[[var]] <- as.factor(leaf_data_converted[[var]])
    cat("Converted", var, "to factor -", nlevels(leaf_data_converted[[var]]), "levels\n")
  }
}

# Convert dates (handle different date formats)
for(var in existing_dates) {
  if(var %in% names(leaf_data_converted)) {
    # Try different date formats
    original_values <- leaf_data_converted[[var]]
    
    # Check for Excel numeric dates first (like 40542)
    if(all(str_detect(original_values[!is.na(original_values) & original_values != ""], "^[0-9]+$"), na.rm = TRUE)) {
      # Excel date format (days since 1900-01-01, but Excel incorrectly treats 1900 as leap year)
      leaf_data_converted[[var]] <- as.Date(as.numeric(original_values), origin = "1899-12-30")
      cat("Converted", var, "from Excel numeric format to Date\n")
    } else {
      # Try different date formats
      date_converted <- as.Date(original_values, format = "%d/%m/%Y")
      if(all(is.na(date_converted))) {
        date_converted <- as.Date(original_values, format = "%Y-%m-%d")
      }
      if(all(is.na(date_converted))) {
        date_converted <- as.Date(original_values, format = "%m/%d/%Y")
      }
      
      leaf_data_converted[[var]] <- date_converted
      cat("Converted", var, "to Date format\n")
    }
    
    # Show date range
    valid_dates <- leaf_data_converted[[var]][!is.na(leaf_data_converted[[var]])]
    if(length(valid_dates) > 0) {
      date_range <- range(valid_dates)
      cat("  Date range:", as.character(date_range[1]), "to", as.character(date_range[2]), "\n")
    }
  }
}

# Convert numerics (handle comma decimal separator)
for(var in existing_numerics) {
  if(var %in% names(leaf_data_converted)) {
    # Replace empty strings and spaces with NA first
    leaf_data_converted[[var]][leaf_data_converted[[var]] == "" | leaf_data_converted[[var]] == " "] <- NA
    
    # If it's character type, it might have comma decimals - convert them to dots
    if(is.character(leaf_data_converted[[var]])) {
      leaf_data_converted[[var]] <- gsub(",", ".", leaf_data_converted[[var]])
    }
    
    # Store original NA count
    original_nas <- sum(is.na(leaf_data[[var]]))
    
    # Convert to numeric
    leaf_data_converted[[var]] <- as.numeric(leaf_data_converted[[var]])
    
    # Check conversion success
    new_nas <- sum(is.na(leaf_data_converted[[var]]))
    conversion_nas <- new_nas - original_nas
    
    if(conversion_nas > 0) {
      cat("WARNING:", conversion_nas, "values became NA during conversion of", var, "\n")
      # Show some examples of problematic values
      original_values <- leaf_data[[var]]
      problematic <- original_values[is.na(leaf_data_converted[[var]]) & !is.na(original_values)]
      if(length(problematic) > 0) {
        cat("  Examples of unconverted values:", paste(head(problematic, 3), collapse = ", "), "\n")
      }
    } else {
      cat("Successfully converted", var, "to numeric\n")
    }
  }
}


str(leaf_data_converted)
```

## I.4. Campaign Analysis

```{r}
# ===============================
# 3. CAMPAIGN ANALYSIS
# ===============================

leaf_data_converted <- leaf_data_converted %>%
  mutate(Campagne = as.character(Campagne))
cat("3. CAMPAIGN ANALYSIS\n")
cat("====================\n")
campaign_summary <- leaf_data_converted %>%
  group_by(Campagne) %>%
  summarise(
    n_observations = n(),
    n_parcelles = n_distinct(Parc.),
    n_fermes = n_distinct(Ferme),
    .groups = 'drop'
  ) %>%
  arrange(Campagne)

print("Observations per Campaign:")
print(campaign_summary)
cat("\nCampaign period coverage:")
cat("From", min(leaf_data_converted$Campagne, na.rm = TRUE), "to", max(leaf_data_converted$Campagne, na.rm = TRUE), "\n")
cat("Total campaigns:", n_distinct(leaf_data_converted$Campagne, na.rm = TRUE), "\n\n")

# ===============================
# 4. FARM AND PARCELLE STRUCTURE ANALYSIS
# ===============================
cat("4. FARM AND PARCELLE STRUCTURE ANALYSIS\n")
cat("========================================\n")

# Parcelles per farm
parcelle_per_farm <- leaf_data_converted %>%
  group_by(Ferme) %>%
  summarise(
    n_parcelles = n_distinct(Parc.),
    campaigns = paste(sort(unique(Campagne)), collapse = ", "),
    varieties = n_distinct(Variete, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(n_parcelles))

print("Parcelles per Farm:")
print(parcelle_per_farm)

# Farm consistency across campaigns
farm_campaign <- leaf_data_converted %>%
  group_by(Campagne) %>%
  summarise(
    farms = list(sort(unique(Ferme))),
    n_farms = n_distinct(Ferme),
    .groups = 'drop'
  )

print("\nFarms per Campaign:")
print(farm_campaign %>% select(-farms))

# Check unique farms
all_farms <- unique(leaf_data_converted$Ferme)
cat("\nTotal unique farms across all campaigns:", length(all_farms), "\n")
print("Unique farms:")
print(all_farms)
```

## I.5. Farm and Parcels structure analysis

```{r}
# ===============================
# 4. FARM AND PARCELLE STRUCTURE ANALYSIS
# ===============================
cat("4. FARM AND PARCELLE STRUCTURE ANALYSIS\n")
cat("========================================\n")

# Parcelles per farm
parcelle_per_farm <- leaf_data_converted %>%
  group_by(Ferme) %>%
  summarise(
    n_parcelles = n_distinct(Parc.),
    campaigns = paste(sort(as.character(unique(Campagne))), collapse = ", "),
    varieties = n_distinct(Variete, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(n_parcelles))

print("Parcelles per Farm:")
print(parcelle_per_farm)

# Farm consistency across campaigns
farm_campaign <- leaf_data_converted %>%
  group_by(Campagne) %>%
  summarise(
    farms = list(sort(unique(Ferme))),
    n_farms = n_distinct(Ferme),
    .groups = 'drop'
  )

print("\nFarms per Campaign:")
print(farm_campaign %>% select(-farms))

# Check unique farms
all_farms <- unique(leaf_data_converted$Ferme)
cat("\nTotal unique farms across all campaigns:", length(all_farms), "\n")
print("Unique farms:")
print(all_farms)

# ===============================
# 5. PARCELLE NAMING PATTERN ANALYSIS
# ===============================
cat("\n5. PARCELLE NAMING PATTERN ANALYSIS\n")
cat("===================================\n")

# Extract parcelle patterns
parcelle_pattern <- leaf_data_converted %>%
  mutate(
    parcelle_prefix = str_extract(Parc., "^[A-Z]+"),
    parcelle_number = str_extract(Parc., "[0-9]+[A-Z]*$")
  ) %>%
  group_by(parcelle_prefix) %>%
  summarise(
    count = n(),
    unique_parcelles = n_distinct(Parc.),
    farms = paste(sort(unique(Ferme)), collapse = ", "),
    .groups = 'drop'
  ) %>%
  arrange(desc(count))

print("Parcelle naming patterns:")
print(parcelle_pattern)

# Evolution of parcelles over time
parcelle_evolution <- leaf_data_converted %>%
  group_by(Campagne) %>%
  summarise(
    parcelles = list(sort(unique(Parc.))),
    n_parcelles = n_distinct(Parc.),
    .groups = 'drop'
  ) %>%
  arrange(Campagne)

print("\nParcelle count evolution over campaigns:")
print(parcelle_evolution %>% select(-parcelles))

# ===============================
# 6. VARIETY ANALYSIS
# ===============================
cat("\n6. VARIETY ANALYSIS\n")
cat("===================\n")

variety_summary <- leaf_data_converted %>%
  group_by(Variete) %>%
  summarise(
    n_observations = n(),
    n_parcelles = n_distinct(Parc.),
    n_farms = n_distinct(Ferme),
    campaigns = paste(sort(as.character(unique(Campagne))), collapse = ", "),
    avg_surface = mean(Surf, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(desc(n_observations))

print("Variety distribution:")
print(variety_summary)
```

## I.6. Missing Data Analysis

```{r}
#4. DETAILED MISSING DATA ANALYSIS
# ===============================
cat("\n4. DETAILED MISSING DATA ANALYSIS\n")
cat("==================================\n")

# Overall missing data summary
missing_summary <- leaf_data_converted %>%
  summarise_all(~sum(is.na(.))) %>%
  gather(key = "variable", value = "missing_count") %>%
  mutate(
    missing_percentage = round((missing_count / nrow(leaf_data_converted)) * 100, 2),
    variable_type = case_when(
      variable %in% existing_factors ~ "Factor",
      variable %in% existing_dates ~ "Date", 
      variable %in% existing_numerics ~ "Numeric",
      TRUE ~ "Unclassified"
    )
  ) %>%
  arrange(desc(missing_count))

print("COMPLETE MISSING DATA SUMMARY:")
print(missing_summary)

# Missing data by variable type
cat("\nMISSING DATA BY VARIABLE TYPE:\n")
missing_by_type <- missing_summary %>%
  group_by(variable_type) %>%
  summarise(
    n_variables = n(),
    total_missing = sum(missing_count),
    avg_missing_pct = round(mean(missing_percentage), 2),
    max_missing_pct = max(missing_percentage),
    .groups = 'drop'
  )
print(missing_by_type)

# Variables with high missing rates (>50%)
high_missing <- missing_summary %>%
  filter(missing_percentage > 50) %>%
  arrange(desc(missing_percentage))

if(nrow(high_missing) > 0) {
  cat("\nVARIABLES WITH HIGH MISSING RATES (>50%):\n")
  print(high_missing)
} else {
  cat("\nNo variables with >50% missing values found.\n")
}

# ===============================
# 5. FACTOR VARIABLES ANALYSIS
# ===============================
cat("\n5. FACTOR VARIABLES ANALYSIS\n")
cat("============================\n")

for(var in existing_factors) {
  if(var %in% names(leaf_data_converted)) {
    cat("\n--- FACTOR:", var, "---\n")
    var_summary <- leaf_data_converted %>%
      count(.data[[var]], name = "count") %>%
      mutate(percentage = round(count/sum(count)*100, 2)) %>%
      arrange(desc(count))
    
    cat("Levels:", nlevels(leaf_data_converted[[var]]), "\n")
    cat("Missing values:", sum(is.na(leaf_data_converted[[var]])), "\n")
    print("Distribution (top 10):")
    print(head(var_summary, 10))
  }
}

# ===============================
# 6. NUMERICAL VARIABLES ANALYSIS
# ===============================
cat("\n6. NUMERICAL VARIABLES ANALYSIS\n")
cat("===============================\n")

if(length(existing_numerics) > 0) {
  # Comprehensive numeric summary
  numeric_data <- leaf_data_converted %>% select(all_of(existing_numerics))
  
  cat("SUMMARY STATISTICS FOR ALL NUMERICAL VARIABLES:\n")
  print(summary(numeric_data))
  
  # Detailed analysis for each numeric variable
  cat("\nDETAILED NUMERICAL ANALYSIS:\n")
  for(var in existing_numerics) {
    if(var %in% names(leaf_data_converted)) {
      cat("\n--- NUMERIC VARIABLE:", var, "---\n")
      
      var_data <- leaf_data_converted[[var]]
      non_na_data <- var_data[!is.na(var_data)]
      
      if(length(non_na_data) > 0) {
        cat("Missing values:", sum(is.na(var_data)), "(", 
            round(sum(is.na(var_data))/length(var_data)*100, 2), "%)\n")
        cat("Non-missing values:", length(non_na_data), "\n")
        cat("Min:", min(non_na_data), "\n")
        cat("Q1:", quantile(non_na_data, 0.25), "\n")
        cat("Median:", median(non_na_data), "\n")
        cat("Mean:", round(mean(non_na_data), 3), "\n")
        cat("Q3:", quantile(non_na_data, 0.75), "\n")
        cat("Max:", max(non_na_data), "\n")
        cat("Std Dev:", round(sd(non_na_data), 3), "\n")
        
        # Check for outliers (values beyond 1.5 * IQR)
        Q1 <- quantile(non_na_data, 0.25)
        Q3 <- quantile(non_na_data, 0.75)
        IQR <- Q3 - Q1
        outliers <- non_na_data[non_na_data < (Q1 - 1.5*IQR) | non_na_data > (Q3 + 1.5*IQR)]
        cat("Potential outliers:", length(outliers), "\n")
        if(length(outliers) > 0 && length(outliers) <= 10) {
          cat("Outlier values:", paste(round(outliers, 3), collapse = ", "), "\n")
        }
        
        # Zero values
        zero_count <- sum(non_na_data == 0)
        if(zero_count > 0) {
          cat("Zero values:", zero_count, "(", round(zero_count/length(non_na_data)*100, 2), "%)\n")
        }
        
        # Negative values
        neg_count <- sum(non_na_data < 0)
        if(neg_count > 0) {
          cat("Negative values:", neg_count, "(", round(neg_count/length(non_na_data)*100, 2), "%)\n")
        }
      } else {
        cat("No non-missing values found for", var, "\n")
      }
    }
  }
}
```

## I.7. Date variables analysis

```{r}
# ===============================
# 7. DATE VARIABLES ANALYSIS
# ===============================
cat("\n7. DATE VARIABLES ANALYSIS\n")
cat("==========================\n")

for(var in existing_dates) {
  if(var %in% names(leaf_data_converted)) {
    cat("\n--- DATE VARIABLE:", var, "---\n")
    
    date_data <- leaf_data_converted[[var]]
    non_na_dates <- date_data[!is.na(date_data)]
    
    if(length(non_na_dates) > 0) {
      cat("Missing values:", sum(is.na(date_data)), "(", 
          round(sum(is.na(date_data))/length(date_data)*100, 2), "%)\n")
      cat("Date range:", as.character(min(non_na_dates)), "to", as.character(max(non_na_dates)), "\n")
      cat("Time span:", as.numeric(max(non_na_dates) - min(non_na_dates)), "days\n")
      
      # Year distribution
      years <- year(non_na_dates)
      year_dist <- table(years)
      cat("Year distribution:\n")
      print(year_dist)
      
      # Month distribution
      months <- month(non_na_dates)
      month_dist <- table(months)
      cat("Month distribution:\n")
      print(month_dist)
    } else {
      cat("No valid dates found for", var, "\n")
    }
  }
}

# ===============================
# 8. CAMPAIGN ANALYSIS
# ===============================
cat("\n8. CAMPAIGN ANALYSIS\n")
cat("====================\n")

if("Campagne" %in% names(leaf_data_converted)) {
  campaign_analysis <- leaf_data_converted %>%
    group_by(Campagne) %>%
    summarise(
      n_observations = n(),
      n_parcelles = n_distinct(Parc., na.rm = TRUE),
      n_fermes = n_distinct(Ferme, na.rm = TRUE),
      n_varieties = n_distinct(Variete, na.rm = TRUE),
      avg_surface = round(mean(Surf, na.rm = TRUE), 2),
      total_surface = round(sum(Surf, na.rm = TRUE), 2),
      .groups = 'drop'
    )
  
  print("CAMPAIGN SUMMARY:")
  print(campaign_analysis)
}

# ===============================
# 9. VARIETY AND FARM ANALYSIS
# ===============================
cat("\n9. VARIETY AND FARM ANALYSIS\n")
cat("============================\n")

if("Variete" %in% names(leaf_data_converted)) {
  variety_analysis <- leaf_data_converted %>%
    group_by(Variete) %>%
    summarise(
      n_observations = n(),
      n_campaigns = n_distinct(Campagne, na.rm = TRUE),
      n_farms = n_distinct(Ferme, na.rm = TRUE),
      avg_yield = round(mean(Tc.Ha_A, na.rm = TRUE), 2),
      .groups = 'drop'
    ) %>%
    arrange(desc(n_observations))
  
  print("VARIETY ANALYSIS:")
  print(variety_analysis)
}

# ===============================
# 10. DATA QUALITY ASSESSMENT
# ===============================
cat("\n10. DATA QUALITY ASSESSMENT\n")
cat("============================\n")

# Completeness by row
row_completeness <- leaf_data_converted %>%
  mutate(row_id = row_number()) %>%
  rowwise() %>%
  mutate(
    missing_count = sum(is.na(c_across(-row_id))),
    completeness_pct = round((ncol(leaf_data_converted)-1-missing_count)/(ncol(leaf_data_converted)-1)*100, 2)
  ) %>%
  ungroup() %>%
  select(row_id, missing_count, completeness_pct)

cat("ROW COMPLETENESS SUMMARY:\n")
cat("Average completeness per row:", round(mean(row_completeness$completeness_pct), 2), "%\n")
cat("Rows with 100% completeness:", sum(row_completeness$completeness_pct == 100), "\n")
cat("Rows with <50% completeness:", sum(row_completeness$completeness_pct < 50), "\n")

# Most incomplete rows
cat("\nMOST INCOMPLETE ROWS (top 5):\n")
print(head(row_completeness %>% arrange(completeness_pct), 5))

# Most complete rows
cat("\nMOST COMPLETE ROWS (top 5):\n")
print(head(row_completeness %>% arrange(desc(completeness_pct)), 5))

```

# II. Dataset Integration (Three-Dataset Parcel-Year LinkageThree-Dataset Parcel-Year Linkage)

Creating normalized parcel-year keys by converting parcel names to uppercase and extracting campaign years (e.g., "2011-2012" → 2011 for leaf data, "03-04" → 2003 for yield data). Performing three-way joins to identify: (1) complete matches present in all datasets, (2) two-way overlaps (yield+soil, yield+leaf, soil+leaf), and (3) unique records per dataset.

```{r}
head(leaf_data_converted) 
# Unique and sorted
unique_parc. <- sort(unique(leaf_data_converted$Parc.))
unique_parcelle <- sort(unique(data$Parcelle))
unique_parc_detaillee <- sort(unique(soil_data$Parc.Detaillee))

# Display with numbering
cat("📌 Unique Parc. values:\n")
for (i in seq_along(unique_parc.)) {
  cat(paste0(i, ". ", unique_parc.[i], "\n"))
}

cat("\n📌 Unique Parcelle values:\n")
for (i in seq_along(unique_parcelle)) {
  cat(paste0(i, ". ", unique_parcelle[i], "\n"))
}

cat("📌 Unique Parc.Detaillee values:\n")
for (i in seq_along(unique_parc_detaillee)) {
  cat(paste0(i, ". ", unique_parc_detaillee[i], "\n"))
}
```

## II.1. Data Integration (All Uppercase)

```{r}
# Step 1: Convert to uppercase
soil_data$Parc.Detaillee <- toupper(soil_data$Parc.Detaillee)
data$Parcelle <- toupper(data$Parcelle)
leaf_data_converted$Parc. <- toupper(leaf_data_converted$Parc.)

# Step 2: Get unique sorted values
unique_parc_detaillee <- sort(unique(soil_data$Parc.Detaillee))
unique_parcelle <- sort(unique(data$Parcelle))
unique_parc <- sort(unique(leaf_data_converted$Parc.))

# Step 3: Display with count and numbered list
cat("📌 Unique values in soil_data$Parc.Detaillee:\n")
cat(paste0("Total: ", length(unique_parc_detaillee), " unique entries\n\n"))
for (i in seq_along(unique_parc_detaillee)) {
  cat(paste0(i, ". ", unique_parc_detaillee[i], "\n"))
}

cat("\n📌 Unique values in data$Parcelle:\n")
cat(paste0("Total: ", length(unique_parcelle), " unique entries\n\n"))
for (i in seq_along(unique_parcelle)) {
  cat(paste0(i, ". ", unique_parcelle[i], "\n"))
}

cat("\n📌 Unique values in leaf_data_converted$Parc.:\n")
cat(paste0("Total: ", length(unique_parc), " unique entries\n\n"))
for (i in seq_along(unique_parc)) {
  cat(paste0(i, ". ", unique_parc[i], "\n"))
}

```

```{r}
# Step 1: Convert to uppercase
soil_data$Parc.Detaillee <- toupper(soil_data$Parc.Detaillee)
data$Parcelle <- toupper(data$Parcelle)
leaf_data_converted$Parc. <- toupper(leaf_data_converted$Parc.)

# Step 2: Count frequencies
soil_freq <- sort(table(soil_data$Parc.Detaillee), decreasing = TRUE)
data_freq <- sort(table(data$Parcelle), decreasing = TRUE)
leaf_freq <- sort(table(leaf_data_converted$Parc.), decreasing = TRUE)

# Step 3: Display the frequencies nicely
cat("📊 Frequency of each unique 'Parc.Detaillee' in soil_data:\n")
cat(paste0("Total unique values: ", length(soil_freq), "\n\n"))
for (i in seq_along(soil_freq)) {
  cat(paste0(i, ". ", names(soil_freq)[i], " – ", soil_freq[i], " sample(s)\n"))
}

cat("\n📊 Frequency of each unique 'Parcelle' in data:\n")
cat(paste0("Total unique values: ", length(data_freq), "\n\n"))
for (i in seq_along(data_freq)) {
  cat(paste0(i, ". ", names(data_freq)[i], " – ", data_freq[i], " sample(s)\n"))
}

cat("\n📊 Frequency of each unique 'Parc.' in leaf_data_converted:\n")
cat(paste0("Total unique values: ", length(leaf_freq), "\n\n"))
for (i in seq_along(leaf_freq)) {
  cat(paste0(i, ". ", names(leaf_freq)[i], " – ", leaf_freq[i], " sample(s)\n"))
}

# Step 4: Standardize inputs to uppercase for matching
soil_parcels <- toupper(unique(soil_data$Parc.Detaillee))
yield_parcels <- toupper(unique(data$Parcelle))
leaf_parcels <- toupper(unique(leaf_data_converted$Parc.))

# Step 5: Matching statistics
common_all_three <- Reduce(intersect, list(soil_parcels, yield_parcels, leaf_parcels))
common_soil_yield <- intersect(soil_parcels, yield_parcels)
common_soil_leaf <- intersect(soil_parcels, leaf_parcels)
common_yield_leaf <- intersect(yield_parcels, leaf_parcels)

only_soil <- setdiff(soil_parcels, union(yield_parcels, leaf_parcels))
only_yield <- setdiff(yield_parcels, union(soil_parcels, leaf_parcels))
only_leaf <- setdiff(leaf_parcels, union(soil_parcels, yield_parcels))

# Step 6: Display results
cat("\n🔍 Parcel Matching Summary:\n")
cat("• Total unique parcels in SOIL data  :", length(soil_parcels), "\n")
cat("• Total unique parcels in YIELD data :", length(yield_parcels), "\n")
cat("• Total unique parcels in LEAF data  :", length(leaf_parcels), "\n")
cat("• ✅ Common to ALL THREE datasets    :", length(common_all_three), "\n")
cat("• ✅ Common to SOIL & YIELD only     :", length(common_soil_yield), "\n")
cat("• ✅ Common to SOIL & LEAF only      :", length(common_soil_leaf), "\n")
cat("• ✅ Common to YIELD & LEAF only     :", length(common_yield_leaf), "\n")
cat("• ❌ Parcels only in SOIL data       :", length(only_soil), "\n")
cat("• ❌ Parcels only in YIELD data      :", length(only_yield), "\n")
cat("• ❌ Parcels only in LEAF data       :", length(only_leaf), "\n")

# Step 7: Display samples of mismatches
cat("\n📂 Sample of parcels only in SOIL data:\n")
if (length(only_soil) > 0) {
  print(head(only_soil, 10))
} else {
  cat("✔️ None. All soil parcels are found in other datasets.\n")
}

cat("\n📂 Sample of parcels only in YIELD data:\n")
if (length(only_yield) > 0) {
  print(head(only_yield, 10))
} else {
  cat("✔️ None. All yield parcels are found in other datasets.\n")
}

cat("\n📂 Sample of parcels only in LEAF data:\n")
if (length(only_leaf) > 0) {
  print(head(only_leaf, 10))
} else {
  cat("✔️ None. All leaf parcels are found in other datasets.\n")
}

cat("\n📂 Parcels common to ALL THREE datasets:\n")
if (length(common_all_three) > 0) {
  print(head(common_all_three, 10))
} else {
  cat("❌ No parcels are common to all three datasets.\n")
}

View(leaf_data_converted)
```

```{r}
library(dplyr)
library(stringr)

# === YIELD DATA ===
cat("Yield data dimensions:", dim(data), "\n")
cat("Yield data columns:\n")
print(colnames(data))

# Extract start year from Camp, e.g. "03-04" -> 2003
data <- data %>%
  mutate(year_start = 2000 + as.numeric(substr(Camp, 1, 2)))

# Normalize parcel names: uppercase only, keep spaces
data <- data %>%
  mutate(Parcelle_norm = str_to_upper(Parcelle))

# === SOIL DATA ===
cat("\nSoil data dimensions:", dim(soil_data), "\n")
cat("Soil data columns:\n")
print(colnames(soil_data))

# Normalize parcel names in soil data: uppercase only, keep spaces + ensure Annee is numeric
soil_data <- soil_data %>%
  mutate(
    Parc.Detaillee_norm = str_to_upper(Parc.Detaillee),
    Annee = as.numeric(Annee)
  )

# === LEAF DATA ===
cat("\nLeaf data dimensions:", dim(leaf_data_converted), "\n")
cat("Leaf data columns:\n")
print(colnames(leaf_data_converted))

# Extract start year from Campagne, e.g. "2011-2012" -> 2011
leaf_data_converted <- leaf_data_converted %>%
  mutate(
    year_start = as.numeric(str_extract(Campagne, "^\\d{4}")),
    Parc_norm = str_to_upper(Parc.)
  )

# Check the campaign conversion
cat("\nSample of leaf campaign conversion:\n")
leaf_sample <- leaf_data_converted %>%
  select(Campagne, year_start) %>%
  distinct() %>%
  head(10)
print(leaf_sample)

# --- Create unique parcel-year keys ---
yield_keys <- data %>%
  distinct(Parcelle_norm, year_start) %>%
  rename(Year = year_start, Parcel = Parcelle_norm)

soil_keys <- soil_data %>%
  distinct(Parc.Detaillee_norm, Annee) %>%
  rename(Year = Annee, Parcel = Parc.Detaillee_norm)

leaf_keys <- leaf_data_converted %>%
  distinct(Parc_norm, year_start) %>%
  rename(Year = year_start, Parcel = Parc_norm)

# --- Find common and unique keys ---
# Three-way comparisons
common_all_three <- yield_keys %>%
  inner_join(soil_keys, by = c("Parcel", "Year")) %>%
  inner_join(leaf_keys, by = c("Parcel", "Year"))

# Pairwise comparisons
common_yield_soil <- inner_join(yield_keys, soil_keys, by = c("Parcel", "Year"))
common_yield_leaf <- inner_join(yield_keys, leaf_keys, by = c("Parcel", "Year"))
common_soil_leaf <- inner_join(soil_keys, leaf_keys, by = c("Parcel", "Year"))

# Unique to each dataset
only_in_yield <- yield_keys %>%
  anti_join(soil_keys, by = c("Parcel", "Year")) %>%
  anti_join(leaf_keys, by = c("Parcel", "Year"))

only_in_soil <- soil_keys %>%
  anti_join(yield_keys, by = c("Parcel", "Year")) %>%
  anti_join(leaf_keys, by = c("Parcel", "Year"))

only_in_leaf <- leaf_keys %>%
  anti_join(yield_keys, by = c("Parcel", "Year")) %>%
  anti_join(soil_keys, by = c("Parcel", "Year"))

# --- Summary output ---
cat("\n=== Three-Dataset Parcel-Year Linkage Summary ===\n")
cat("Total unique parcel-year in yield data: ", nrow(yield_keys), "\n")
cat("Total unique parcel-year in soil data : ", nrow(soil_keys), "\n")
cat("Total unique parcel-year in leaf data : ", nrow(leaf_keys), "\n")
cat("Common to ALL THREE datasets          : ", nrow(common_all_three), "\n")
cat("Common to YIELD & SOIL only           : ", nrow(common_yield_soil), "\n")
cat("Common to YIELD & LEAF only           : ", nrow(common_yield_leaf), "\n")
cat("Common to SOIL & LEAF only            : ", nrow(common_soil_leaf), "\n")
cat("Parcel-years only in yield data       : ", nrow(only_in_yield), "\n")
cat("Parcel-years only in soil data        : ", nrow(only_in_soil), "\n")
cat("Parcel-years only in leaf data        : ", nrow(only_in_leaf), "\n")

# --- Show samples ---
cat("\nSample parcel-years common to ALL THREE datasets:\n")
print(head(common_all_three, 10))

cat("\nSample parcel-years only in yield data:\n")
print(head(only_in_yield, 10))

cat("\nSample parcel-years only in soil data:\n")
print(head(only_in_soil, 10))

cat("\nSample parcel-years only in leaf data:\n")
print(head(only_in_leaf, 10))
```

## II.2. Data Integration (Trimmed and All uppercase)

```{r}
library(dplyr)
library(stringr)

# === YIELD DATA ===
cat("Yield data dimensions:", dim(data), "\n")
cat("Yield data columns:\n")
print(colnames(data))

# Extract start year from Camp, e.g. "03-04" -> 2003
data <- data %>%
  mutate(year_start = 2000 + as.numeric(substr(Camp, 1, 2)))

# Normalize parcel names: uppercase and trim all spaces
data <- data %>%
  mutate(Parcelle_norm = str_to_upper(str_replace_all(Parcelle, "\\s+", "")))

# === SOIL DATA ===
cat("\nSoil data dimensions:", dim(soil_data), "\n")
cat("Soil data columns:\n")
print(colnames(soil_data))

# Normalize parcel names in soil data: uppercase only, keep spaces + ensure Annee is numeric
soil_data <- soil_data %>%
  mutate(
    Parc.Detaillee_norm = str_to_upper(Parc.Detaillee),
    Annee = as.numeric(Annee)
  )

# === LEAF DATA ===
cat("\nLeaf data dimensions:", dim(leaf_data_converted), "\n")
cat("Leaf data columns:\n")
print(colnames(leaf_data_converted))

# Extract start year from Campagne, e.g. "2011-2012" -> 2011
leaf_data_converted <- leaf_data_converted %>%
  mutate(
    year_start = as.numeric(str_extract(Campagne, "^\\d{4}")),
    Parc_norm = str_to_upper(Parc.)
  )

# Check the campaign conversion
cat("\nSample of leaf campaign conversion:\n")
leaf_sample <- leaf_data_converted %>%
  select(Campagne, year_start) %>%
  distinct() %>%
  head(10)
print(leaf_sample)

# --- Create unique parcel-year keys ---
yield_keys <- data %>%
  distinct(Parcelle_norm, year_start) %>%
  rename(Year = year_start, Parcel = Parcelle_norm)

soil_keys <- soil_data %>%
  distinct(Parc.Detaillee_norm, Annee) %>%
  rename(Year = Annee, Parcel = Parc.Detaillee_norm)

leaf_keys <- leaf_data_converted %>%
  distinct(Parc_norm, year_start) %>%
  rename(Year = year_start, Parcel = Parc_norm)

# --- Find common and unique keys ---
# Three-way comparisons
common_all_three <- yield_keys %>%
  inner_join(soil_keys, by = c("Parcel", "Year")) %>%
  inner_join(leaf_keys, by = c("Parcel", "Year"))

# Pairwise comparisons
common_yield_soil <- inner_join(yield_keys, soil_keys, by = c("Parcel", "Year"))
common_yield_leaf <- inner_join(yield_keys, leaf_keys, by = c("Parcel", "Year"))
common_soil_leaf <- inner_join(soil_keys, leaf_keys, by = c("Parcel", "Year"))

# Unique to each dataset
only_in_yield <- yield_keys %>%
  anti_join(soil_keys, by = c("Parcel", "Year")) %>%
  anti_join(leaf_keys, by = c("Parcel", "Year"))

only_in_soil <- soil_keys %>%
  anti_join(yield_keys, by = c("Parcel", "Year")) %>%
  anti_join(leaf_keys, by = c("Parcel", "Year"))

only_in_leaf <- leaf_keys %>%
  anti_join(yield_keys, by = c("Parcel", "Year")) %>%
  anti_join(soil_keys, by = c("Parcel", "Year"))

# --- Summary output ---
cat("\n=== Three-Dataset Parcel-Year Linkage Summary ===\n")
cat("Total unique parcel-year in yield data: ", nrow(yield_keys), "\n")
cat("Total unique parcel-year in soil data : ", nrow(soil_keys), "\n")
cat("Total unique parcel-year in leaf data : ", nrow(leaf_keys), "\n")
cat("Common to ALL THREE datasets          : ", nrow(common_all_three), "\n")
cat("Common to YIELD & SOIL only           : ", nrow(common_yield_soil), "\n")
cat("Common to YIELD & LEAF only           : ", nrow(common_yield_leaf), "\n")
cat("Common to SOIL & LEAF only            : ", nrow(common_soil_leaf), "\n")
cat("Parcel-years only in yield data       : ", nrow(only_in_yield), "\n")
cat("Parcel-years only in soil data        : ", nrow(only_in_soil), "\n")
cat("Parcel-years only in leaf data        : ", nrow(only_in_leaf), "\n")

# --- Show samples ---
cat("\nSample parcel-years common to ALL THREE datasets:\n")
print(head(common_all_three, 10))

cat("\nSample parcel-years only in yield data:\n")
print(head(only_in_yield, 10))

cat("\nSample parcel-years only in soil data:\n")
print(head(only_in_soil, 10))

cat("\nSample parcel-years only in leaf data:\n")
print(head(only_in_leaf, 10))
```

## II.3. Not trimmed (space preserved)

```{r}
library(ggplot2)
library(patchwork)
library(scales)
library(dplyr)

# === VERSION 1: TWO-WAY PLOTS (Uppercase, Spaces Preserved) ===

# First, create normalized versions with uppercase but spaces preserved
data_twoway <- data %>%
  mutate(Parcelle_norm = str_to_upper(Parcelle))

soil_data_twoway <- soil_data %>%
  mutate(
    Parc.Detaillee_norm = str_to_upper(Parc.Detaillee),
    Annee = as.numeric(Annee)
  )

# Create keys for two-way analysis
yield_keys_twoway <- data_twoway %>%
  distinct(Parcelle_norm, year_start) %>%
  rename(Year = year_start, Parcel = Parcelle_norm)

soil_keys_twoway <- soil_data_twoway %>%
  distinct(Parc.Detaillee_norm, Annee) %>%
  rename(Year = Annee, Parcel = Parc.Detaillee_norm)

# Two-way matching
common_keys_twoway <- inner_join(yield_keys_twoway, soil_keys_twoway, by = c("Parcel", "Year"))
only_in_yield_twoway <- anti_join(yield_keys_twoway, soil_keys_twoway, by = c("Parcel", "Year"))
only_in_soil_twoway <- anti_join(soil_keys_twoway, yield_keys_twoway, by = c("Parcel", "Year"))

# Prepare data for two-way plotting
linkage_counts_twoway <- tibble(
  Category = c("Yield Only", "Soil Only", "Common"),
  Count = c(nrow(only_in_yield_twoway), nrow(only_in_soil_twoway), nrow(common_keys_twoway))
)

# ---- 1. Bar plot: Two-way Parcel-Year linkage summary ----
plot_linkage_twoway <- ggplot(linkage_counts_twoway, aes(x = Category, y = Count, fill = Category)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  geom_text(aes(label = scales::comma(Count)), vjust = -0.5, size = 5) +
  scale_fill_manual(values = c("Yield Only" = "#1f78b4", "Soil Only" = "#33a02c", "Common" = "#6a3d9a")) +
  labs(title = "Two-Way Parcel-Year Linkage Summary",
       subtitle = paste0("Yield unique: ", scales::comma(nrow(yield_keys_twoway)),
                         " | Soil unique: ", scales::comma(nrow(soil_keys_twoway))),
       x = NULL,
       y = "Number of Unique Parcel-Year Pairs") +
  theme_minimal(base_size = 15) +
  theme(axis.text.x = element_text(face = "bold"))

# ---- 2. Two-way Year distribution plots ----
yield_year_dist_twoway <- yield_keys_twoway %>%
  group_by(Year) %>%
  summarise(Parcel_Count = n()) %>%
  ungroup()

plot_yield_year_twoway <- ggplot(yield_year_dist_twoway, aes(x = Year, y = Parcel_Count)) +
  geom_col(fill = "#1f78b4") +
  scale_x_continuous(breaks = pretty(yield_year_dist_twoway$Year, 10)) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Yield Data: Parcel-Year Counts by Year",
       x = "Year",
       y = "Number of Unique Parcels") +
  theme_minimal(base_size = 13)

soil_year_dist_twoway <- soil_keys_twoway %>%
  group_by(Year) %>%
  summarise(Parcel_Count = n()) %>%
  ungroup()

plot_soil_year_twoway <- ggplot(soil_year_dist_twoway, aes(x = Year, y = Parcel_Count)) +
  geom_col(fill = "#33a02c") +
  scale_x_continuous(breaks = pretty(soil_year_dist_twoway$Year, 10)) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Soil Data: Parcel-Year Counts by Year",
       x = "Year",
       y = "Number of Unique Parcels") +
  theme_minimal(base_size = 13)

# ---- 3. Two-way Pie chart ----
plot_pie_twoway <- ggplot(linkage_counts_twoway, aes(x = "", y = Count, fill = Category)) +
  geom_col(width = 1, color = "white") +
  coord_polar("y") +
  scale_fill_manual(values = c("Yield Only" = "#1f78b4", "Soil Only" = "#33a02c", "Common" = "#6a3d9a")) +
  geom_text(aes(label = paste0(Category, "\n", scales::percent(Count / sum(Count), accuracy = 0.1))),
            position = position_stack(vjust = 0.5), size = 4) +
  labs(title = "Two-Way Linkage Proportions") +
  theme_void() +
  theme(plot.title = element_text(size = 14, face = "bold"))

# === VERSION 2: THREE-WAY PLOTS (Uppercase, Spaces Preserved) ===

# Create normalized leaf data with spaces preserved
leaf_data_twoway <- leaf_data_converted %>%
  mutate(
    year_start = as.numeric(str_extract(Campagne, "^\\d{4}")),
    Parc_norm = str_to_upper(Parc.)
  )

# Create keys for three-way analysis
leaf_keys_twoway <- leaf_data_twoway %>%
  distinct(Parc_norm, year_start) %>%
  rename(Year = year_start, Parcel = Parc_norm)

# === DETAILED THREE-WAY ANALYSIS ===
cat("=== Detailed Three-Way Overlap Analysis ===\n")

# Print unique key counts
cat("Total unique parcel-year pairs per dataset:\n")
cat("• Yield dataset:", nrow(yield_keys_twoway), "unique parcel-year pairs\n")
cat("• Soil dataset: ", nrow(soil_keys_twoway), "unique parcel-year pairs\n")
cat("• Leaf dataset: ", nrow(leaf_keys_twoway), "unique parcel-year pairs\n\n")

# Find all possible overlaps using Venn diagram logic
# 1. Common to all three datasets (intersection of all three)
common_all_three_twoway <- yield_keys_twoway %>%
  inner_join(soil_keys_twoway, by = c("Parcel", "Year")) %>%
  inner_join(leaf_keys_twoway, by = c("Parcel", "Year"))

# 2. Common to exactly two datasets (but NOT in the third)
yield_soil_only <- yield_keys_twoway %>%
  inner_join(soil_keys_twoway, by = c("Parcel", "Year")) %>%
  anti_join(leaf_keys_twoway, by = c("Parcel", "Year"))

yield_leaf_only <- yield_keys_twoway %>%
  inner_join(leaf_keys_twoway, by = c("Parcel", "Year")) %>%
  anti_join(soil_keys_twoway, by = c("Parcel", "Year"))

soil_leaf_only <- soil_keys_twoway %>%
  inner_join(leaf_keys_twoway, by = c("Parcel", "Year")) %>%
  anti_join(yield_keys_twoway, by = c("Parcel", "Year"))

# 3. Unique to exactly one dataset (not in any other)
yield_only <- yield_keys_twoway %>%
  anti_join(soil_keys_twoway, by = c("Parcel", "Year")) %>%
  anti_join(leaf_keys_twoway, by = c("Parcel", "Year"))

soil_only <- soil_keys_twoway %>%
  anti_join(yield_keys_twoway, by = c("Parcel", "Year")) %>%
  anti_join(leaf_keys_twoway, by = c("Parcel", "Year"))

leaf_only <- leaf_keys_twoway %>%
  anti_join(yield_keys_twoway, by = c("Parcel", "Year")) %>%
  anti_join(soil_keys_twoway, by = c("Parcel", "Year"))

# Print detailed results
cat("OVERLAP ANALYSIS:\n")
cat("• Common to ALL THREE datasets     :", nrow(common_all_three_twoway), "pairs\n")
cat("• Common to Yield+Soil only        :", nrow(yield_soil_only), "pairs\n")
cat("• Common to Yield+Leaf only        :", nrow(yield_leaf_only), "pairs\n")
cat("• Common to Soil+Leaf only         :", nrow(soil_leaf_only), "pairs\n")
cat("• Unique to Yield only             :", nrow(yield_only), "pairs\n")
cat("• Unique to Soil only              :", nrow(soil_only), "pairs\n")
cat("• Unique to Leaf only              :", nrow(leaf_only), "pairs\n\n")

# Calculate linkable data (parcel-years that appear in all three datasets)
total_potential_links <- nrow(common_all_three_twoway)
total_unlinkable <- nrow(yield_only) + nrow(soil_only) + nrow(leaf_only) + 
                   nrow(yield_soil_only) + nrow(yield_leaf_only) + nrow(soil_leaf_only)

cat("LINKAGE SUMMARY:\n")
cat("• ✅ FULLY LINKABLE (all 3 datasets):", total_potential_links, "parcel-year pairs\n")
cat("• ❌ NOT FULLY LINKABLE              :", total_unlinkable, "parcel-year pairs\n")
cat("• 📊 Linkage success rate           :", 
    round(total_potential_links / (total_potential_links + total_unlinkable) * 100, 1), "%\n\n")

# Show samples of fully linkable data
cat("Sample of FULLY LINKABLE parcel-year pairs:\n")
print(head(common_all_three_twoway, 10))

# Prepare data for three-way plotting
linkage_counts_threeway <- tibble(
  Category = c("All Three\n(Linkable)", "Yield+Soil Only", "Yield+Leaf Only", "Soil+Leaf Only", 
               "Yield Only", "Soil Only", "Leaf Only"),
  Count = c(nrow(common_all_three_twoway), nrow(yield_soil_only), 
            nrow(yield_leaf_only), nrow(soil_leaf_only),
            nrow(yield_only), nrow(soil_only), nrow(leaf_only)),
  Linkable = c("Yes", "No", "No", "No", "No", "No", "No")
)

# ---- 4. Three-way Bar plot ----
plot_linkage_threeway <- ggplot(linkage_counts_threeway, aes(x = reorder(Category, Count), y = Count, fill = Linkable)) +
  geom_col(width = 0.6, show.legend = TRUE) +
  geom_text(aes(label = scales::comma(Count)), hjust = -0.1, size = 4) +
  scale_fill_manual(values = c("Yes" = "#2E8B57", "No" = "#DC143C"), 
                    name = "Fully Linkable", 
                    labels = c("No" = "Not Linkable", "Yes" = "Linkable")) +
  coord_flip() +
  labs(title = "Three-Way Parcel-Year Linkage Analysis",
       subtitle = paste0("Linkable pairs: ", scales::comma(total_potential_links), " | ",
                         "Non-linkable pairs: ", scales::comma(total_unlinkable), " | ",
                         "Success rate: ", round(total_potential_links / (total_potential_links + total_unlinkable) * 100, 1), "%"),
       x = NULL,
       y = "Number of Unique Parcel-Year Pairs") +
  theme_minimal(base_size = 13) +
  theme(axis.text.y = element_text(face = "bold"),
        legend.position = "bottom")

# ---- 5. Three-way Year distributions ----
leaf_year_dist_twoway <- leaf_keys_twoway %>%
  group_by(Year) %>%
  summarise(Parcel_Count = n()) %>%
  ungroup()

plot_leaf_year_twoway <- ggplot(leaf_year_dist_twoway, aes(x = Year, y = Parcel_Count)) +
  geom_col(fill = "#fdbf6f") +
  scale_x_continuous(breaks = pretty(leaf_year_dist_twoway$Year, 10)) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Leaf Data: Parcel-Year Counts by Year",
       x = "Year",
       y = "Number of Unique Parcels") +
  theme_minimal(base_size = 13)

# Combined three-way year plot
plot_year_combined_threeway <- plot_yield_year_twoway / plot_soil_year_twoway / plot_leaf_year_twoway +
  plot_annotation(title = "Yearly Distribution of Unique Parcel-Year Combinations (Three Datasets)",
                  theme = theme(plot.title = element_text(size = 16, face = "bold")))

# ---- 6. Create final layouts ----
cat("=== Creating Two-Way Analysis Plots ===\n")
plot_year_combined_twoway <- plot_yield_year_twoway | plot_soil_year_twoway + 
  plot_annotation(title = "Two-Way Yearly Distribution",
                  theme = theme(plot.title = element_text(size = 16, face = "bold")))

final_plot_twoway <- plot_linkage_twoway | plot_pie_twoway
combined_plot_twoway <- final_plot_twoway / plot_year_combined_twoway + 
  plot_layout(heights = c(1, 2)) +
  plot_annotation(caption = "Two-way analysis: Uppercase parcel names, spaces preserved.",
                  theme = theme(plot.caption = element_text(size = 11, face = "italic")))

cat("=== Creating Three-Way Analysis Plots ===\n")
final_plot_threeway <- plot_linkage_threeway | plot_year_combined_threeway + 
  plot_layout(widths = c(1.2, 1)) +
  plot_annotation(caption = "Three-way analysis: Uppercase parcel names, spaces preserved.",
                  theme = theme(plot.caption = element_text(size = 11, face = "italic")))

# ---- Print plots ----
print("Two-Way Analysis:")
print(combined_plot_twoway)

print("\nThree-Way Analysis:")
print(final_plot_threeway)
```

## II.4. Correspondence Summary

i\. Complete Matches Identifying parcel-year pairs present in all three datasets for integrated analysis.

ii. Partial and Unique Records Documenting dataset-specific observations for selective analyses.

```{r}
library(ggplot2)
library(patchwork)
library(scales)
library(dplyr)
library(stringr)

# === PRÉPARATION DES DONNÉES ===

# Normalisation des données avec majuscules et espaces préservés
data_norm <- data %>%
  mutate(Parcelle_norm = str_to_upper(Parcelle))

soil_data_norm <- soil_data %>%
  mutate(
    Parc.Detaillee_norm = str_to_upper(Parc.Detaillee),
    Annee = as.numeric(Annee)
  )

leaf_data_norm <- leaf_data_converted %>%
  mutate(
    year_start = as.numeric(str_extract(Campagne, "^\\d{4}")),
    Parc_norm = str_to_upper(Parc.)
  )

# Création des clés uniques pour chaque jeu de données
yield_keys <- data_norm %>%
  distinct(Parcelle_norm, year_start) %>%
  rename(Annee = year_start, Parcelle = Parcelle_norm)

soil_keys <- soil_data_norm %>%
  distinct(Parc.Detaillee_norm, Annee) %>%
  rename(Parcelle = Parc.Detaillee_norm)

leaf_keys <- leaf_data_norm %>%
  distinct(Parc_norm, year_start) %>%
  rename(Annee = year_start, Parcelle = Parc_norm)

# === ANALYSE DES CORRESPONDANCES TOTALES ===

# Correspondances complètes (présentes dans les trois jeux de données)
correspondance_complete <- yield_keys %>%
  inner_join(soil_keys, by = c("Parcelle", "Annee")) %>%
  inner_join(leaf_keys, by = c("Parcelle", "Annee"))

# Correspondances à deux voies (incluant celles qui sont aussi dans le 3ème)
rendement_sol_total <- yield_keys %>%
  inner_join(soil_keys, by = c("Parcelle", "Annee"))

rendement_feuille_total <- yield_keys %>%
  inner_join(leaf_keys, by = c("Parcelle", "Annee"))

sol_feuille_total <- soil_keys %>%
  inner_join(leaf_keys, by = c("Parcelle", "Annee"))

# Données uniques (présentes dans un seul jeu de données)
unique_rendement <- yield_keys %>%
  anti_join(soil_keys, by = c("Parcelle", "Annee")) %>%
  anti_join(leaf_keys, by = c("Parcelle", "Annee"))

unique_sol <- soil_keys %>%
  anti_join(yield_keys, by = c("Parcelle", "Annee")) %>%
  anti_join(leaf_keys, by = c("Parcelle", "Annee"))

unique_feuille <- leaf_keys %>%
  anti_join(yield_keys, by = c("Parcelle", "Annee")) %>%
  anti_join(soil_keys, by = c("Parcelle", "Annee"))

# Configuration du thème académique français
theme_academique_fr <- theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5),
    axis.title = element_text(size = 11, face = "bold"),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 10),
    legend.position = "bottom",
    panel.grid.major = element_line(color = "grey90", size = 0.5),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    panel.background = element_rect(fill = "white", color = NA),
    strip.text = element_text(size = 11, face = "bold"),
    plot.caption = element_text(size = 9, hjust = 0, face = "italic")
  )

# === GRAPHIQUE 1: CORRESPONDANCES TOTALES ===

donnees_correspondance_totales <- tibble(
  Categorie = c("Trois jeux\n(Correspondance complète)", 
                "Rendement + Sol", 
                "Rendement + Feuille", 
                "Sol + Feuille"),
  Nombre = c(nrow(correspondance_complete), 
             nrow(rendement_sol_total), 
             nrow(rendement_feuille_total), 
             nrow(sol_feuille_total)),
  Type = c("Complète", "Partielle", "Partielle", "Partielle")
)

graphique_correspondances_totales <- ggplot(donnees_correspondance_totales, 
                                           aes(x = reorder(Categorie, Nombre), y = Nombre, fill = Type)) +
  geom_col(width = 0.7, color = "white", size = 0.5) +
  geom_text(aes(label = paste0("n = ", format(Nombre, big.mark = " "))), 
            hjust = -0.1, size = 3.5, fontface = "bold") +
  scale_fill_manual(values = c("Complète" = "#2E8B57", "Partielle" = "#4682B4"),
                    name = "Type de correspondance") +
  coord_flip() +
  labs(
    title = "Correspondances totales entre jeux de données",
    subtitle = "Nombre total de couples parcelle-année présents dans chaque combinaison",
    x = "Catégories de correspondance",
    y = "Nombre de couples parcelle-année uniques",
    caption = "Note : Les correspondances à 2 jeux incluent celles présentes aussi dans le 3ème jeu"
  ) +
  theme_academique_fr +
  scale_y_continuous(labels = function(x) format(x, big.mark = " "))

# === GRAPHIQUE 2: RÉPARTITION PAR JEU DE DONNÉES ===

donnees_jeux <- tibble(
  Jeu_donnees = c("Rendement", "Sol", "Feuilles"),
  Nombre_total = c(nrow(yield_keys), nrow(soil_keys), nrow(leaf_keys)),
  Dans_correspondance_complete = c(
    sum(yield_keys %>% semi_join(correspondance_complete, by = c("Parcelle", "Annee")) %>% nrow()),
    sum(soil_keys %>% semi_join(correspondance_complete, by = c("Parcelle", "Annee")) %>% nrow()),
    sum(leaf_keys %>% semi_join(correspondance_complete, by = c("Parcelle", "Annee")) %>% nrow())
  )
) %>%
  mutate(
    Taux_correspondance = Dans_correspondance_complete / Nombre_total * 100,
    Uniquement_jeu = Nombre_total - Dans_correspondance_complete
  )

# Reformater pour ggplot
donnees_jeux_long <- donnees_jeux %>%
  select(Jeu_donnees, Dans_correspondance_complete, Uniquement_jeu) %>%
  pivot_longer(cols = c(Dans_correspondance_complete, Uniquement_jeu),
               names_to = "Statut", values_to = "Nombre") %>%
  mutate(
    Statut = case_when(
      Statut == "Dans_correspondance_complete" ~ "Dans correspondance\ncomplète",
      Statut == "Uniquement_jeu" ~ "Unique à ce jeu"
    )
  )

graphique_jeux <- ggplot(donnees_jeux_long, aes(x = Jeu_donnees, y = Nombre, fill = Statut)) +
  geom_col(position = "stack", width = 0.7, color = "white", size = 0.5) +
  geom_text(data = donnees_jeux, 
            aes(x = Jeu_donnees, y = Nombre_total + 50, 
                label = paste0(round(Taux_correspondance, 1), "%")),
            inherit.aes = FALSE, size = 3.5, fontface = "bold") +
  scale_fill_manual(values = c("Dans correspondance\ncomplète" = "#2E8B57", 
                              "Unique à ce jeu" = "#DC143C"),
                    name = "Statut des données") +
  labs(
    title = "Répartition des données par jeu",
    subtitle = "Proportion des données présentes dans la correspondance complète",
    x = "Jeux de données",
    y = "Nombre de couples parcelle-année",
    caption = "Pourcentages indiqués : taux de présence dans la correspondance complète"
  ) +
  theme_academique_fr +
  scale_y_continuous(labels = function(x) format(x, big.mark = " "))

# === GRAPHIQUE 3: RÉPARTITION PROPORTIONNELLE ===

donnees_proportion <- tibble(
  Categorie = c("Correspondance\ncomplète", "Données\nuniques"),
  Nombre = c(nrow(correspondance_complete),
             nrow(unique_rendement) + nrow(unique_sol) + nrow(unique_feuille)),
  Couleur = c("#2E8B57", "#DC143C")
)

graphique_proportion <- ggplot(donnees_proportion, aes(x = "", y = Nombre, fill = Categorie)) +
  geom_col(width = 1, color = "white", size = 0.8) +
  coord_polar("y") +
  scale_fill_manual(values = setNames(donnees_proportion$Couleur, donnees_proportion$Categorie)) +
  geom_text(aes(label = paste0(format(Nombre, big.mark = " "), "\n(", 
                               round(Nombre/sum(Nombre)*100, 1), "%)")),
            position = position_stack(vjust = 0.5), 
            size = 3.5, fontface = "bold", color = "white") +
  labs(
    title = "Répartition générale",
    fill = "Type de données"
  ) +
  theme_academique_fr +
  theme(
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank(),
    axis.title = element_blank()
  )

# === GRAPHIQUE 4: ÉVOLUTION TEMPORELLE DES CORRESPONDANCES ===

# Analyse par année des correspondances complètes
correspondances_annuelles <- correspondance_complete %>%
  count(Annee, name = "Correspondances_completes") %>%
  full_join(
    yield_keys %>% count(Annee, name = "Total_rendement"),
    by = "Annee"
  ) %>%
  full_join(
    soil_keys %>% count(Annee, name = "Total_sol"),
    by = "Annee"
  ) %>%
  full_join(
    leaf_keys %>% count(Annee, name = "Total_feuille"),
    by = "Annee"
  ) %>%
  replace_na(list(Correspondances_completes = 0, Total_rendement = 0, 
                  Total_sol = 0, Total_feuille = 0)) %>%
  mutate(
    Taux_correspondance = ifelse(pmax(Total_rendement, Total_sol, Total_feuille, na.rm = TRUE) > 0,
                                Correspondances_completes / pmax(Total_rendement, Total_sol, Total_feuille, na.rm = TRUE) * 100,
                                0)
  )

graphique_temporel <- ggplot(correspondances_annuelles, aes(x = Annee)) +
  geom_col(aes(y = Correspondances_completes), fill = "#2E8B57", alpha = 0.8, width = 0.8) +
  geom_line(aes(y = Taux_correspondance * max(Correspondances_completes, na.rm = TRUE) / 100), 
            color = "#DC143C", size = 1.2, alpha = 0.8) +
  geom_point(aes(y = Taux_correspondance * max(Correspondances_completes, na.rm = TRUE) / 100), 
             color = "#DC143C", size = 2) +
  scale_x_continuous(breaks = seq(min(correspondances_annuelles$Annee, na.rm = TRUE), 
                                  max(correspondances_annuelles$Annee, na.rm = TRUE), by = 2)) +
  scale_y_continuous(
    name = "Nombre de correspondances complètes",
    labels = function(x) format(x, big.mark = " "),
    sec.axis = sec_axis(~ . * 100 / max(correspondances_annuelles$Correspondances_completes, na.rm = TRUE), 
                       name = "Taux de correspondance (%)")
  ) +
  labs(
    title = "Évolution temporelle des correspondances complètes",
    subtitle = "Nombre de couples parcelle-année présents dans les trois jeux de données",
    x = "Année",
    caption = "Barres : nombre de correspondances | Ligne : taux de correspondance"
  ) +
  theme_academique_fr +
  theme(
    axis.title.y.right = element_text(color = "#DC143C", face = "bold"),
    axis.text.y.right = element_text(color = "#DC143C")
  )

# === CRÉATION DU GRAPHIQUE COMBINÉ ===

graphique_combine <- (graphique_correspondances_totales | graphique_jeux) / 
                     (graphique_proportion | graphique_temporel) +
  plot_layout(heights = c(1, 1)) +
  plot_annotation(
    title = "ANALYSE DES CORRESPONDANCES TOTALES ENTRE JEUX DE DONNÉES AGRICOLES",
    subtitle = paste0("Analyse comparative des données de rendement, sol et feuilles | ",
                     "Correspondances complètes : ", format(nrow(correspondance_complete), big.mark = " "), 
                     " couples parcelle-année"),
    caption = paste0("Note : Correspondance complète = présence dans les trois jeux de données (rendement, sol, feuilles)\n",
                    "Les correspondances partielles incluent les chevauchements avec le troisième jeu\n",
                    "Période d'analyse : ", min(c(yield_keys$Annee, soil_keys$Annee, leaf_keys$Annee), na.rm = TRUE),
                    " - ", max(c(yield_keys$Annee, soil_keys$Annee, leaf_keys$Annee), na.rm = TRUE)),
    theme = theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      plot.caption = element_text(size = 9, hjust = 0, face = "italic")
    )
  )

# === SAUVEGARDE DES GRAPHIQUES ===

# Créer le dossier de sortie s'il n'existe pas
if (!dir.exists("graphiques_correspondances_totales")) {
  dir.create("graphiques_correspondances_totales")
}

# Sauvegarde du graphique combiné principal
ggsave(
  filename = "graphiques_correspondances_totales/analyse_correspondances_totales.png",
  plot = graphique_combine,
  width = 16, height = 12, dpi = 300, units = "in",
  bg = "white"
)

ggsave(
  filename = "graphiques_correspondances_totales/analyse_correspondances_totales.pdf",
  plot = graphique_combine,
  width = 16, height = 12, units = "in",
  bg = "white"
)

# Sauvegarde des graphiques individuels
ggsave(
  filename = "graphiques_correspondances_totales/correspondances_totales.png",
  plot = graphique_correspondances_totales,
  width = 10, height = 6, dpi = 300, units = "in",
  bg = "white"
)

ggsave(
  filename = "graphiques_correspondances_totales/repartition_jeux.png",
  plot = graphique_jeux,
  width = 10, height = 6, dpi = 300, units = "in",
  bg = "white"
)

ggsave(
  filename = "graphiques_correspondances_totales/repartition_proportionnelle.png",
  plot = graphique_proportion,
  width = 8, height = 6, dpi = 300, units = "in",
  bg = "white"
)

ggsave(
  filename = "graphiques_correspondances_totales/evolution_temporelle.png",
  plot = graphique_temporel,
  width = 12, height = 6, dpi = 300, units = "in",
  bg = "white"
)

# === STATISTIQUES DESCRIPTIVES ===

cat("=== RÉSUMÉ DE L'ANALYSE DES CORRESPONDANCES TOTALES ===\n\n")

cat("JEUX DE DONNÉES UNIQUES :\n")
cat("• Rendement :", format(nrow(yield_keys), big.mark = " "), "couples parcelle-année\n")
cat("• Sol       :", format(nrow(soil_keys), big.mark = " "), "couples parcelle-année\n")
cat("• Feuilles  :", format(nrow(leaf_keys), big.mark = " "), "couples parcelle-année\n\n")

cat("CORRESPONDANCES TOTALES (avec chevauchements inclus) :\n")
cat("• Rendement + Sol                      :", format(nrow(rendement_sol_total), big.mark = " "), "couples\n")
cat("• Rendement + Feuilles                 :", format(nrow(rendement_feuille_total), big.mark = " "), "couples\n")
cat("• Sol + Feuilles                       :", format(nrow(sol_feuille_total), big.mark = " "), "couples\n")
cat("• Correspondance complète (3 jeux)     :", format(nrow(correspondance_complete), big.mark = " "), "couples\n\n")

cat("DONNÉES UNIQUES (présentes dans un seul jeu) :\n")
cat("• Rendement uniquement                 :", format(nrow(unique_rendement), big.mark = " "), "couples\n")
cat("• Sol uniquement                       :", format(nrow(unique_sol), big.mark = " "), "couples\n")
cat("• Feuilles uniquement                  :", format(nrow(unique_feuille), big.mark = " "), "couples\n\n")

# Calcul des taux pour chaque jeu
taux_rendement <- round(nrow(correspondance_complete) / nrow(yield_keys) * 100, 2)
taux_sol <- round(nrow(correspondance_complete) / nrow(soil_keys) * 100, 2)
taux_feuilles <- round(nrow(correspondance_complete) / nrow(leaf_keys) * 100, 2)

cat("TAUX DE CORRESPONDANCE COMPLÈTE PAR JEU :\n")
cat("• Rendement  :", taux_rendement, "% de ses données dans la correspondance complète\n")
cat("• Sol        :", taux_sol, "% de ses données dans la correspondance complète\n")
cat("• Feuilles   :", taux_feuilles, "% de ses données dans la correspondance complète\n\n")
# Affichage du graphique principal
print(graphique_combine)
```

```{r}
library(dplyr)
library(stringr)

# === PRÉPARATION DES DONNÉES ===

# Normalisation des données avec majuscules et espaces préservés
data_norm <- data %>%
  mutate(Parcelle_norm = str_to_upper(Parcelle))

soil_data_norm <- soil_data %>%
  mutate(
    Parc.Detaillee_norm = str_to_upper(Parc.Detaillee),
    Annee = as.numeric(Annee)
  )

leaf_data_norm <- leaf_data_converted %>%
  mutate(
    year_start = as.numeric(str_extract(Campagne, "^\\d{4}")),
    Parc_norm = str_to_upper(Parc.)
  )

# Création des clés uniques pour chaque jeu de données
yield_keys <- data_norm %>%
  distinct(Parcelle_norm, year_start) %>%
  rename(Annee = year_start, Parcelle = Parcelle_norm)

soil_keys <- soil_data_norm %>%
  distinct(Parc.Detaillee_norm, Annee) %>%
  rename(Parcelle = Parc.Detaillee_norm)

leaf_keys <- leaf_data_norm %>%
  distinct(Parc_norm, year_start) %>%
  rename(Annee = year_start, Parcelle = Parc_norm)

# === FONCTION POUR GÉNÉRER DES ÉCHANTILLONS ===
generer_echantillons <- function(data, nom_dataset, n_sample = 10) {
  if (nrow(data) == 0) {
    return(paste0("Aucune donnée disponible pour ", nom_dataset))
  }
  
  echantillon <- data %>%
    sample_n(min(n_sample, nrow(data))) %>%
    arrange(Annee, Parcelle) %>%
    mutate(ligne = paste0("  • ", Parcelle, " (", Annee, ")"))
  
  return(echantillon$ligne)
}

# === CORRESPONDANCES POUR IDENTIFIER LES DONNÉES UNIQUES ===

# Correspondances complètes (présentes dans les trois jeux de données)
correspondance_complete <- yield_keys %>%
  inner_join(soil_keys, by = c("Parcelle", "Annee")) %>%
  inner_join(leaf_keys, by = c("Parcelle", "Annee"))

# Données uniques (présentes dans un seul jeu de données)
unique_rendement <- yield_keys %>%
  anti_join(soil_keys, by = c("Parcelle", "Annee")) %>%
  anti_join(leaf_keys, by = c("Parcelle", "Annee"))

unique_sol <- soil_keys %>%
  anti_join(yield_keys, by = c("Parcelle", "Annee")) %>%
  anti_join(leaf_keys, by = c("Parcelle", "Annee"))

unique_feuille <- leaf_keys %>%
  anti_join(yield_keys, by = c("Parcelle", "Annee")) %>%
  anti_join(soil_keys, by = c("Parcelle", "Annee"))

# === ANALYSE DES PLAGES TEMPORELLES ===

analyser_plage_temporelle <- function(data, nom_dataset) {
  if (nrow(data) == 0) {
    return(list(
      min_annee = NA,
      max_annee = NA,
      nb_annees = 0,
      annees_manquantes = NA,
      resume = paste0("Aucune donnée pour ", nom_dataset)
    ))
  }
  
  min_annee <- min(data$Annee, na.rm = TRUE)
  max_annee <- max(data$Annee, na.rm = TRUE)
  annees_presentes <- sort(unique(data$Annee))
  annees_completes <- seq(min_annee, max_annee)
  annees_manquantes <- setdiff(annees_completes, annees_presentes)
  
  return(list(
    min_annee = min_annee,
    max_annee = max_annee,
    nb_annees = length(annees_presentes),
    nb_annees_possibles = length(annees_completes),
    annees_presentes = annees_presentes,
    annees_manquantes = annees_manquantes,
    taux_completude = round(length(annees_presentes) / length(annees_completes) * 100, 1)
  ))
}

# Analyse des plages temporelles pour chaque jeu de données
plage_rendement <- analyser_plage_temporelle(yield_keys, "Rendement")
plage_sol <- analyser_plage_temporelle(soil_keys, "Sol")
plage_feuille <- analyser_plage_temporelle(leaf_keys, "Feuilles")
plage_correspondance_complete <- analyser_plage_temporelle(correspondance_complete, "Correspondance complète")

# === ANALYSE DES PARCELLES UNIQUES ===

analyser_parcelles <- function(data, nom_dataset) {
  if (nrow(data) == 0) {
    return(list(
      nb_parcelles_uniques = 0,
      parcelles_les_plus_frequentes = character(0),
      resume = paste0("Aucune parcelle pour ", nom_dataset)
    ))
  }
  
  parcelles_freq <- data %>%
    count(Parcelle, sort = TRUE)
  
  return(list(
    nb_parcelles_uniques = nrow(parcelles_freq),
    parcelles_les_plus_frequentes = head(parcelles_freq, 5),
    nb_observations_total = nrow(data)
  ))
}

parcelles_rendement <- analyser_parcelles(yield_keys, "Rendement")
parcelles_sol <- analyser_parcelles(soil_keys, "Sol")
parcelles_feuille <- analyser_parcelles(leaf_keys, "Feuilles")

# === 1. STATISTIQUES GÉNÉRALES ===
cat("📊 STATISTIQUES GÉNÉRALES\n")
cat("═══════════════════════════\n\n")

cat("TOTAL DES OBSERVATIONS PAR JEU DE DONNÉES :\n")
cat("• Rendement :", format(nrow(yield_keys), big.mark = " "), "couples parcelle-année\n")
cat("• Sol       :", format(nrow(soil_keys), big.mark = " "), "couples parcelle-année\n")
cat("• Feuilles  :", format(nrow(leaf_keys), big.mark = " "), "couples parcelle-année\n")
cat("• Correspondance complète :", format(nrow(correspondance_complete), big.mark = " "), "couples parcelle-année\n\n")

cat("NOMBRE DE PARCELLES UNIQUES PAR JEU :\n")
cat("• Rendement :", format(parcelles_rendement$nb_parcelles_uniques, big.mark = " "), "parcelles\n")
cat("• Sol       :", format(parcelles_sol$nb_parcelles_uniques, big.mark = " "), "parcelles\n")
cat("• Feuilles  :", format(parcelles_feuille$nb_parcelles_uniques, big.mark = " "), "parcelles\n\n")

# === 2. PLAGES TEMPORELLES ===
cat("📅 PLAGES TEMPORELLES\n")
cat("══════════════════════\n\n")

afficher_plage <- function(plage, nom) {
  cat(paste0("▶ ", nom, " :\n"))
  if (is.na(plage$min_annee)) {
    cat("  Aucune donnée disponible\n\n")
    return()
  }
  
  cat("  • Période :", plage$min_annee, "-", plage$max_annee, 
      "(", plage$max_annee - plage$min_annee + 1, "ans de couverture possible)\n")
  cat("  • Années avec données :", plage$nb_annees, "/", plage$nb_annees_possibles,
      "(", plage$taux_completude, "% de complétude temporelle)\n")
  
  if (length(plage$annees_manquantes) > 0 && length(plage$annees_manquantes) <= 10) {
    cat("  • Années manquantes :", paste(plage$annees_manquantes, collapse = ", "), "\n")
  } else if (length(plage$annees_manquantes) > 10) {
    cat("  • Années manquantes :", length(plage$annees_manquantes), "années (",
        paste(head(plage$annees_manquantes, 3), collapse = ", "), "...)\n")
  } else {
    cat("  • Couverture temporelle complète ✓\n")
  }
  cat("\n")
}

afficher_plage(plage_rendement, "DONNÉES DE RENDEMENT")
afficher_plage(plage_sol, "DONNÉES DE SOL")
afficher_plage(plage_feuille, "DONNÉES DE FEUILLES")
afficher_plage(plage_correspondance_complete, "CORRESPONDANCE COMPLÈTE")

# === 3. DONNÉES UNIQUES (présentes dans un seul jeu) ===
cat("🔍 DONNÉES UNIQUES (présentes dans un seul jeu de données)\n")
cat("══════════════════════════════════════════════════════════\n\n")

cat("▶ RENDEMENT UNIQUEMENT :\n")
cat("  Nombre :", format(nrow(unique_rendement), big.mark = " "), "couples parcelle-année\n")
if (nrow(unique_rendement) > 0) {
  cat("  Échantillons :\n")
  echantillons <- generer_echantillons(unique_rendement, "Rendement")
  cat(paste(echantillons, collapse = "\n"), "\n")
} else {
  cat("  Aucune donnée unique au jeu de rendement\n")
}
cat("\n")

cat("▶ SOL UNIQUEMENT :\n")
cat("  Nombre :", format(nrow(unique_sol), big.mark = " "), "couples parcelle-année\n")
if (nrow(unique_sol) > 0) {
  cat("  Échantillons :\n")
  echantillons <- generer_echantillons(unique_sol, "Sol")
  cat(paste(echantillons, collapse = "\n"), "\n")
} else {
  cat("  Aucune donnée unique au jeu de sol\n")
}
cat("\n")

cat("▶ FEUILLES UNIQUEMENT :\n")
cat("  Nombre :", format(nrow(unique_feuille), big.mark = " "), "couples parcelle-année\n")
if (nrow(unique_feuille) > 0) {
  cat("  Échantillons :\n")
  echantillons <- generer_echantillons(unique_feuille, "Feuilles")
  cat(paste(echantillons, collapse = "\n"), "\n")
} else {
  cat("  Aucune donnée unique au jeu de feuilles\n")
}
cat("\n")

# === 4. PARCELLES LES PLUS REPRÉSENTÉES ===
cat("📍 PARCELLES LES PLUS REPRÉSENTÉES\n")
cat("═══════════════════════════════════\n\n")

afficher_top_parcelles <- function(parcelles_info, nom) {
  cat(paste0("▶ ", nom, " :\n"))
  if (nrow(parcelles_info$parcelles_les_plus_frequentes) > 0) {
    for (i in 1:nrow(parcelles_info$parcelles_les_plus_frequentes)) {
      parcelle <- parcelles_info$parcelles_les_plus_frequentes[i, ]
      cat("  ", i, ". ", parcelle$Parcelle, " (", parcelle$n, " observations)\n")
    }
  } else {
    cat("  Aucune donnée disponible\n")
  }
  cat("\n")
}

afficher_top_parcelles(parcelles_rendement, "RENDEMENT")
afficher_top_parcelles(parcelles_sol, "SOL")
afficher_top_parcelles(parcelles_feuille, "FEUILLES")

# === 5. RÉSUMÉ GLOBAL ===
cat("📋 RÉSUMÉ GLOBAL\n")
cat("═════════════════\n\n")

total_observations <- nrow(yield_keys) + nrow(soil_keys) + nrow(leaf_keys)
total_unique <- nrow(unique_rendement) + nrow(unique_sol) + nrow(unique_feuille)
taux_correspondance_complete <- round(nrow(correspondance_complete) / total_observations * 100, 2)
taux_donnees_uniques <- round(total_unique / total_observations * 100, 2)

# Période globale
all_years <- c(yield_keys$Annee, soil_keys$Annee, leaf_keys$Annee)
periode_globale_min <- min(all_years, na.rm = TRUE)
periode_globale_max <- max(all_years, na.rm = TRUE)

cat("• Période d'étude globale :", periode_globale_min, "-", periode_globale_max, 
    "(", periode_globale_max - periode_globale_min + 1, "années)\n")
cat("• Total des observations :", format(total_observations, big.mark = " "), "couples parcelle-année\n")
cat("• Correspondance complète :", format(nrow(correspondance_complete), big.mark = " "), 
    "couples (", round(nrow(correspondance_complete) / total_observations * 100, 1), "% du total)\n")
cat("• Données uniques :", format(total_unique, big.mark = " "), 
    "couples (", taux_donnees_uniques, "% du total)\n")

# Efficacité de la correspondance
efficacite_max_possible <- min(nrow(yield_keys), nrow(soil_keys), nrow(leaf_keys))
efficacite_reelle <- round(nrow(correspondance_complete) / efficacite_max_possible * 100, 1)
cat("• Efficacité de correspondance :", efficacite_reelle, "% (vs maximum théorique possible)\n\n")

```

```{r}
# Convert key variables
leaf_data_converted$Tc_Ha_A <- as.numeric(gsub(",", ".", leaf_data_converted$Tc.Ha_A))
leaf_data_converted$T_pol_canne <- as.numeric(gsub(",", ".", leaf_data_converted$T_pol_canne))
leaf_data_converted$T_canne_reelle <- as.numeric(gsub(",", ".", leaf_data_converted$T_canne_reelle))

# Extract year from campaign
leaf_data_converted$Year <- as.numeric(str_extract(leaf_data_converted$Campagne, "\\d{4}"))
leaf_data_converted$Variete <- as.factor(leaf_data_converted$Variete)

# Calculate sugar efficiency
leaf_data_converted$Sugar_efficiency <- (leaf_data_converted$T_pol_canne / leaf_data_converted$T_canne_reelle) * 100

cat("Data prepared successfully!\n")
print(paste("Years available:", paste(unique(leaf_data_converted$Year), collapse = ", ")))
print(paste("Varieties:", paste(unique(leaf_data_converted$Variete), collapse = ", ")))


# ===============================
# 1. TREND EVOLUTION WITH SE + LOESS - COMBINED PLOT
# ===============================

create_combined_trends <- function(data) {
  
  # Prepare data for trends
  trend_data <- data %>%
    filter(!is.na(Year) & !is.na(Tc_Ha_A) & !is.na(T_pol_canne)) %>%
    group_by(Year) %>%
    summarise(
      Tc_mean = mean(Tc_Ha_A, na.rm = TRUE),
      Tc_se = sd(Tc_Ha_A, na.rm = TRUE) / sqrt(n()),
      Tc_sd = sd(Tc_Ha_A, na.rm = TRUE),
      Pol_mean = mean(T_pol_canne, na.rm = TRUE),
      Pol_se = sd(T_pol_canne, na.rm = TRUE) / sqrt(n()),
      Pol_sd = sd(T_pol_canne, na.rm = TRUE),
      n = n(),
      .groups = 'drop'
    )
  
  # Create dual-axis plot
  p1 <- ggplot(trend_data, aes(x = Year)) +
    geom_line(aes(y = Tc_mean), color = "#e74c3c", size = 1.5, alpha = 0.8) +
    geom_point(aes(y = Tc_mean), color = "#e74c3c", size = 4, alpha = 0.9) +
    geom_errorbar(aes(ymin = Tc_mean - Tc_se, ymax = Tc_mean + Tc_se), 
                  color = "#e74c3c", width = 0.2, size = 1, alpha = 0.7) +
    geom_smooth(aes(y = Tc_mean), method = "loess", se = TRUE, 
                color = "#c0392b", fill = "#e74c3c", alpha = 0.3, span = 0.8) +
    labs(title = "Sugarcane Yield Evolution - Senegal Sahel Region",
         subtitle = "Tc/Ha with Standard Error bars and LOESS smoothing",
         x = "Year", 
         y = "Tc/Ha (Tons per Hectare)",
         caption = "Error bars represent ±1 SE") +
    scale_x_continuous(breaks = pretty_breaks(n = 6)) +
    scale_y_continuous(labels = comma_format()) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray60"),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  p2 <- ggplot(trend_data, aes(x = Year)) +
    geom_line(aes(y = Pol_mean), color = "#3498db", size = 1.5, alpha = 0.8) +
    geom_point(aes(y = Pol_mean), color = "#3498db", size = 4, alpha = 0.9) +
    geom_errorbar(aes(ymin = Pol_mean - Pol_se, ymax = Pol_mean + Pol_se), 
                  color = "#3498db", width = 0.2, size = 1, alpha = 0.7) +
    geom_smooth(aes(y = Pol_mean), method = "loess", se = TRUE, 
                color = "#2980b9", fill = "#3498db", alpha = 0.3, span = 0.8) +
    labs(title = "Sugar Content (Pol) Evolution",
         subtitle = "T_pol/canne with Standard Error bars and LOESS smoothing",
         x = "Year", 
         y = "T_pol/canne (Pol tons per cane)",
         caption = "Error bars represent ±1 SE") +
    scale_x_continuous(breaks = pretty_breaks(n = 6)) +
    scale_y_continuous(labels = comma_format()) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray60"),
      axis.title = element_text(size = 12, face = "bold"),
      axis.text = element_text(size = 10),
      panel.grid.minor = element_blank(),
      panel.background = element_rect(fill = "white", color = NA),
      plot.background = element_rect(fill = "white", color = NA)
    )
  
  return(list(tc_plot = p1, pol_plot = p2, trend_data = trend_data))
}

trend_results <- create_combined_trends(leaf_data_converted)
print(trend_results$tc_plot)
print(trend_results$pol_plot)


# ===============================
# 2. COMBINED DUAL-AXIS PLOT
# ===============================
create_dual_axis_plot <- function(data) {
  
  trend_data <- data %>%
    filter(!is.na(Year) & !is.na(Tc_Ha_A) & !is.na(T_pol_canne)) %>%
    group_by(Year) %>%
    summarise(
      Tc_mean = mean(Tc_Ha_A, na.rm = TRUE),
      Tc_se = sd(Tc_Ha_A, na.rm = TRUE) / sqrt(n()),
      Pol_mean = mean(T_pol_canne, na.rm = TRUE),
      Pol_se = sd(T_pol_canne, na.rm = TRUE) / sqrt(n()),
      .groups = 'drop'
    )
  
  # Scale factor for dual axis
  scale_factor <- max(trend_data$Tc_mean, na.rm = TRUE) / max(trend_data$Pol_mean, na.rm = TRUE)
  
  p_dual <- ggplot(trend_data, aes(x = Year)) +
    # Primary axis (Tc/Ha_A)
    geom_line(aes(y = Tc_mean, color = "Tc/Ha"), size = 1.5) +
    geom_point(aes(y = Tc_mean, color = "Tc/Ha"), size = 4) +
    geom_errorbar(aes(ymin = Tc_mean - Tc_se, ymax = Tc_mean + Tc_se, color = "Tc/Ha"), 
                  width = 0.2, alpha = 0.7) +
    geom_smooth(aes(y = Tc_mean), method = "loess", se = TRUE, 
                color = "#e74c3c", fill = "#e74c3c", alpha = 0.2) +
    
    # Secondary axis (T_pol_canne scaled)
    geom_line(aes(y = Pol_mean * scale_factor, color = "T_pol/canne"), size = 1.5) +
    geom_point(aes(y = Pol_mean * scale_factor, color = "T_pol/canne"), size = 4) +
    geom_errorbar(aes(ymin = (Pol_mean - Pol_se) * scale_factor, 
                     ymax = (Pol_mean + Pol_se) * scale_factor, color = "T_pol/canne"), 
                  width = 0.2, alpha = 0.7) +
    geom_smooth(aes(y = Pol_mean * scale_factor), method = "loess", se = TRUE, 
                color = "#3498db", fill = "#3498db", alpha = 0.2) +
    
    scale_color_manual(values = c("Tc/Ha" = "#e74c3c", "T_pol/canne" = "#3498db")) +
    
    scale_y_continuous(
      name = "Tc/Ha (Tons per Hectare)",
      sec.axis = sec_axis(~./scale_factor, name = "T_pol/canne (Pol tons)")
    ) +
    
    labs(title = "Combined Sugarcane Yield Evolution",
         subtitle = "Tc/Ha and T_pol/canne trends with LOESS smoothing and SE bars",
         x = "Year",
         color = "Parameter",
         caption = "Dual-axis plot with independent LOESS trends") +
    
    theme_minimal() +
    theme(
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
      plot.subtitle = element_text(size = 12, hjust = 0.5),
      axis.title.y = element_text(color = "#e74c3c", face = "bold"),
      axis.title.y.right = element_text(color = "#3498db", face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    )
  
  return(p_dual)
}

dual_plot <- create_dual_axis_plot(leaf_data_converted)
print(dual_plot)

# ===============================
# 3. BOXPLOTS BY YEAR AND VARIETY
# ===============================

create_comprehensive_boxplots <- function(data) {
  
  clean_data <- data %>%
    filter(!is.na(Year) & !is.na(Tc_Ha_A) & !is.na(T_pol_canne) & !is.na(Variete))
  
  # Tc/Ha_A boxplots by year
  p1 <- ggplot(clean_data, aes(x = factor(Year), y = Tc_Ha_A)) +
    geom_violin(alpha = 0.6, fill = "#e74c3c") +
    geom_boxplot(width = 0.3, alpha = 0.8, outlier.color = "#c0392b") +
    stat_summary(fun = mean, geom = "point", size = 3, color = "white", shape = 21, fill = "red") +
    labs(title = "Tc/Ha Distribution by Year",
         subtitle = "Violin plots with boxplots and mean points",
         x = "Year", y = "Tc/Ha (Tons per Hectare)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # T_pol_canne boxplots by year
  p2 <- ggplot(clean_data, aes(x = factor(Year), y = T_pol_canne)) +
    geom_violin(alpha = 0.6, fill = "#3498db") +
    geom_boxplot(width = 0.3, alpha = 0.8, outlier.color = "#2980b9") +
    stat_summary(fun = mean, geom = "point", size = 3, color = "white", shape = 21, fill = "blue") +
    labs(title = "T_pol/canne Distribution by Year",
         subtitle = "Violin plots with boxplots and mean points", 
         x = "Year", y = "T_pol/canne (Pol tons)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
  
  # Variety comparison
  p3 <- ggplot(clean_data, aes(x = Variete, y = Tc_Ha_A, fill = Variete)) +
    geom_boxplot(alpha = 0.8, outlier.alpha = 0.6) +
    geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
    stat_summary(fun = mean, geom = "point", size = 3, color = "red", shape = 18) +
    facet_wrap(~Year, scales = "free_x") +
    scale_fill_viridis_d() +
    labs(title = "Tc/Ha by Variety Across Years",
         x = "Variety", y = "Tc/Ha (Tons per Hectare)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none")
  
  # Sugar efficiency comparison
  p4 <- ggplot(clean_data, aes(x = Variete, y = Sugar_efficiency, fill = Variete)) +
    geom_boxplot(alpha = 0.8, outlier.alpha = 0.6) +
    geom_jitter(width = 0.2, alpha = 0.4, size = 1) +
    stat_summary(fun = mean, geom = "point", size = 3, color = "red", shape = 18) +
    facet_wrap(~Year, scales = "free_x") +
    scale_fill_viridis_d() +
    labs(title = "Sugar Efficiency by Variety Across Years",
         x = "Variety", y = "Sugar Efficiency (%)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = "none")
  
  return(list(tc_year = p1, pol_year = p2, tc_variety = p3, sugar_variety = p4))
}

boxplot_results <- create_comprehensive_boxplots(leaf_data_converted)
grid.arrange(boxplot_results$tc_year, boxplot_results$pol_year, ncol = 2)
print(boxplot_results$tc_variety)
print(boxplot_results$sugar_variety)

# ===============================
# 4. STATISTICAL SUMMARY AND TRENDS
# ===============================

yearly_summary <- leaf_data_converted %>%
  filter(!is.na(Year) & !is.na(Tc_Ha_A) & !is.na(T_pol_canne)) %>%
  group_by(Year) %>%
  summarise(
    n = n(),
    Tc_mean = round(mean(Tc_Ha_A, na.rm = TRUE), 2),
    Tc_median = round(median(Tc_Ha_A, na.rm = TRUE), 2),
    Tc_sd = round(sd(Tc_Ha_A, na.rm = TRUE), 2),
    Tc_se = round(sd(Tc_Ha_A, na.rm = TRUE) / sqrt(n()), 2),
    Pol_mean = round(mean(T_pol_canne, na.rm = TRUE), 2),
    Pol_median = round(median(T_pol_canne, na.rm = TRUE), 2), 
    Pol_sd = round(sd(T_pol_canne, na.rm = TRUE), 2),
    Pol_se = round(sd(T_pol_canne, na.rm = TRUE) / sqrt(n()), 2),
    Sugar_eff_mean = round(mean(Sugar_efficiency, na.rm = TRUE), 2),
    .groups = 'drop'
  )

cat("\n=== YEARLY SUMMARY STATISTICS ===\n")
print(yearly_summary)

variety_summary <- leaf_data_converted %>%
  filter(!is.na(Variete) & !is.na(Tc_Ha_A) & !is.na(T_pol_canne)) %>%
  group_by(Variete) %>%
  summarise(
    n = n(),
    years = paste(unique(Year), collapse = ", "),
    Tc_mean = round(mean(Tc_Ha_A, na.rm = TRUE), 2),
    Tc_sd = round(sd(Tc_Ha_A, na.rm = TRUE), 2),
    Pol_mean = round(mean(T_pol_canne, na.rm = TRUE), 2),
    Pol_sd = round(sd(T_pol_canne, na.rm = TRUE), 2),
    Sugar_eff_mean = round(mean(Sugar_efficiency, na.rm = TRUE), 2),
    .groups = 'drop'
  )

cat("\n=== VARIETY SUMMARY STATISTICS ===\n")
print(variety_summary)

# ===============================
# 5. COMBINED GRID LAYOUT
# ===============================

create_dashboard <- function() {
  trends <- create_combined_trends(leaf_data_converted)
  boxplots <- create_comprehensive_boxplots(leaf_data_converted)
  combined_grid <- grid.arrange(
    trends$tc_plot, trends$pol_plot,
    boxplots$tc_year, boxplots$pol_year,
    ncol = 2, nrow = 2,
    top = "Sugarcane Production Analysis Dashboard - Senegal"
  )
  return(combined_grid)
}

dashboard <- create_dashboard()
```

# III. SUGARCANE NPK ANALYSIS (Visualization and Trend Analysis)

```{r}
# ===============================
# SUGARCANE NPK ANALYSIS 
# Visualization and Trend Analysis
# ===============================

library(tidyverse)
library(ggplot2)
library(plotly)
library(corrplot)
library(pheatmap)
library(viridis)
library(RColorBrewer)
library(gridExtra)
library(scales)
library(broom)
library(GGally)
library(factoextra)
library(FactoMineR)



cat("SUGARCANE DATASET ANALYSIS - SENEGAL\n")
cat("====================================\n")
print(dim(leaf_data_converted))
print(head(leaf_data_converted, 3))

# ===============================
# 1. DATA PREPROCESSING
# ===============================

cat("Actual column names:\n")
print(names(leaf_data_converted))

original_names <- names(leaf_data_converted)

leaf_data_converted$N_pct <- as.numeric(gsub(",", ".", leaf_data_converted$N.))
leaf_data_converted$P_pct <- as.numeric(gsub(",", ".", leaf_data_converted$X.P))
leaf_data_converted$K_pct <- as.numeric(gsub(",", ".", leaf_data_converted$X.K))
leaf_data_converted$Tc_Ha_A_1 <- as.numeric(gsub(",", ".", leaf_data_converted$Tc.Ha_A.1))
leaf_data_converted$RH_pctC_A_1 <- as.numeric(gsub(",", ".", leaf_data_converted$RH.C_A.1))
leaf_data_converted$Tc_Ha_A <- as.numeric(gsub(",", ".", leaf_data_converted$Tc.Ha_A))
leaf_data_converted$RH_pctC_A <- as.numeric(gsub(",", ".", leaf_data_converted$RH.C_A))
leaf_data_converted$T_canne_reelle <- as.numeric(gsub(",", ".", leaf_data_converted$T_canne_reelle))
leaf_data_converted$T_pol_canne <- as.numeric(gsub(",", ".", leaf_data_converted$T_pol_canne))
leaf_data_converted$H2O_feuilles <- as.numeric(gsub(",", ".", leaf_data_converted$H2O_feuilles))
leaf_data_converted$H20_gaines <- as.numeric(gsub(",", ".", leaf_data_converted$H20_gaines))
leaf_data_converted$Age_prel_ <- as.numeric(gsub(",", ".", leaf_data_converted$Age_prel))

leaf_data_converted$Variete <- as.factor(leaf_data_converted$Variete)
leaf_data_converted$Ferme <- as.factor(leaf_data_converted$Ferme)
leaf_data_converted$Campagne <- as.factor(leaf_data_converted$Campagne)

leaf_data_converted$NPK_ratio <- leaf_data_converted$N_pct / (leaf_data_converted$P_pct + leaf_data_converted$K_pct)
leaf_data_converted$Sugar_efficiency <- leaf_data_converted$T_pol_canne / leaf_data_converted$T_canne_reelle * 100
leaf_data_converted$Biomass_index <- leaf_data_converted$T_canne_reelle / leaf_data_converted$Surf

# ===============================
# 2. NPK TREND ANALYSIS WITH LOESS AND CONFIDENCE INTERVALS
# ===============================

create_npk_trends <- function(data) {
  npk_long <- data %>%
    select(Age_prel_, N_pct, P_pct, K_pct, Variete, Campagne) %>%
    filter(!is.na(Age_prel_) & !is.na(N_pct) & !is.na(P_pct) & !is.na(K_pct)) %>%
    pivot_longer(cols = c(N_pct, P_pct, K_pct), names_to = "Nutrient", values_to = "Concentration")

  p1 <- ggplot(npk_long, aes(x = Age_prel_, y = Concentration, color = Nutrient)) +
    geom_point(alpha = 0.6, size = 2) +
    geom_smooth(method = "loess", se = TRUE, span = 0.7, size = 1.2) +
    facet_wrap(~Nutrient, scales = "free_y", ncol = 3) +
    scale_color_manual(values = c("N_pct" = "#e74c3c", "P_pct" = "#3498db", "K_pct" = "#2ecc71")) +
    labs(title = "NPK Concentration Trends Over Plant Age",
         subtitle = "LOESS smoothing with 95% confidence intervals",
         x = "Plant Age (days)", y = "Nutrient Concentration (%)",
         caption = "Sugarcane - Senegal Sahel Region") +
    theme_minimal()
  return(p1)
}

npk_plot <- create_npk_trends(leaf_data_converted)
print(npk_plot)

# ===============================
# 3. NPK vs BIOMASS AND SUGAR CONTENT ANALYSIS
# ===============================

create_npk_biomass_plots <- function(data) {
  clean_data <- data %>%
    filter(!is.na(N_pct) & !is.na(P_pct) & !is.na(K_pct) &
           !is.na(T_canne_reelle) & !is.na(T_pol_canne))

  clean_data$Sugar_efficiency <- clean_data$T_pol_canne / clean_data$T_canne_reelle * 100

  p1 <- ggplot(clean_data, aes(x = N_pct, y = T_canne_reelle)) +
    geom_point(aes(color = Variete), size = 3, alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE, color = "darkred", size = 1.2)

  p2 <- ggplot(clean_data, aes(x = P_pct, y = T_pol_canne)) +
    geom_point(aes(color = Variete), size = 3, alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE, color = "darkblue", size = 1.2)

  p3 <- ggplot(clean_data, aes(x = K_pct, y = Sugar_efficiency)) +
    geom_point(aes(color = Variete), size = 3, alpha = 0.7) +
    geom_smooth(method = "lm", se = TRUE, color = "darkgreen", size = 1.2)

  clean_data$NP_ratio <- clean_data$N_pct / clean_data$P_pct
  p4 <- ggplot(clean_data, aes(x = NP_ratio, y = T_canne_reelle)) +
    geom_point(aes(color = Sugar_efficiency, size = T_pol_canne), alpha = 0.7) +
    geom_smooth(method = "loess", se = TRUE, color = "purple", size = 1.2)

  return(list(p1, p2, p3, p4))
}

biomass_plots <- create_npk_biomass_plots(leaf_data_converted)
grid.arrange(grobs = biomass_plots, ncol = 2)

# ===============================
# 4. ADVANCED CORRELATION MATRIX AND HEATMAPS
# ===============================

numerical_vars <- leaf_data_converted %>%
  select(N_pct, X.P, X.K, H2O_feuilles, H20_gaines,
         T_canne_reelle, T_pol_canne, Tc_Ha_A_1, RH_pctC_A_1,
         Age_prel_, Surf, Sugar_efficiency, NPK_ratio) %>%
  select_if(is.numeric)

cor_data <- numerical_vars[complete.cases(numerical_vars), ]
cor_matrix <- cor(cor_data)

corrplot(cor_matrix, method = "color", type = "upper")

pheatmap(cor_matrix,
         clustering_distance_rows = "euclidean",
         clustering_distance_cols = "euclidean",
         clustering_method = "complete")

```

# IV. Descriptive Statistics

```{r}
colnames(leaf_data) <- c(
  "Campaign", "Note", "Plot", "Farm", "Variety",
  "Surface_ha", "Crop_Cycle", "Planting_date", "A_ce_jour",
  "Yield_tonnes_ha_year_minus_1", "Cane_Moisture_percent_year_minus_1",
  "Sampling_date", "Sampling_Age", "Leaf_Moisture_percent",
  "Sheath_Moisture_percent", "Analysis_Date", "Nitrogen_percent",
  "Phosphorus_percent", "Potassium_percent", "Harvest_Date",
  "Yield_tonnes_ha_year_actual", "Cane_Moisture_percent_year_actual", "Growth_Period",
  "Period_Validation", "Plot_Campaign", "Harvest_Validation",
  "Real_T_canne", "T_Pol_canne",
  "Irrigation_Type", "Age_weeks", "Harvest_Age_year",
  "Agricultural_Surface_ha", "Surface_Age_days", "Harvest_Month",
  "Week_Number", "Common_Harvest_Date", "Data_Day",
  "Sampling_Month"
)
```

## IV.1. Dataset basic information

```{r}
# Display basic information about the dataset
cat("Dataset dimensions:", dim(leaf_data), "\n")
cat("Number of rows:", nrow(leaf_data), "\n")
cat("Number of columns:", ncol(leaf_data), "\n")

# Display column names
cat("\nColumn names:\n")
print(colnames(leaf_data))
```

## IV.2. Data Type Identification and Conversion

```{r}
# ===============================
cat("2. DATA TYPE IDENTIFICATION AND CONVERSION\n")
cat("==========================================\n")

# Define data types based on domain knowledge
factor_variables <- c(
  "Campaign",                  # Campaign
  "Note",              # Note
  "Plot",                       # Plot
  "Farm",                       # Farm
  "Variety",                    # Variety
  "Crop_Cycle",                 # Cycle
  "Growth_Period",              # Period
  "Period_Validation",          # Period validation
  "Plot_Campaign",              # Campaign plot
  "Harvest_Validation",         # Harvest validation
  "Irrigation_Type",            # Irrigation type
  "Harvest_Month",              # Harvest month
  "Sampling_Month",
  "Week_Number",
  "Data_Day"
)

# Date variables
date_variables <- c(
  "Planting_date",             # Planting date
  "Sampling_date",                   # Sampling date
  "Analysis_Date",               # Analysis date
  "Harvest_Date",                # Harvest date
  "Common_Harvest_Date"          # Common harvest date
)

# Numerical variables
numerical_variables <- c(
  "Surface_ha",  
  "A_ce_jour",
  "Sampling_Age",                 
  "Yield_tonnes_ha_year_minus_1",               
  "Cane_Moisture_percent_year_minus_1",         
  "Leaf_Moisture_percent",               # Leaf water content
  "Sheath_Moisture_percent",             # Sheath water content
  "Nitrogen_percent",                    # Nitrogen %
  "Phosphorus_percent",                  # Phosphorus %
  "Potassium_percent",                   # Potassium %
  "Yield_tonnes_ha_year_actual",         # Actual cane yield
  "Cane_Moisture_percent_year_actual",           # Actual sugar content
  "Real_T_canne",   
  "T_Pol_canne",
  "Age_weeks",
  "Harvest_Age_year",                    # Age at harvest
  "Agricultural_Surface_ha",             # Surface area actual
  "Surface_Age_days"
)

# Check which variables actually exist in the dataset
existing_factors <- intersect(factor_variables, names(leaf_data))
existing_dates <- intersect(date_variables, names(leaf_data))
existing_numerics <- intersect(numerical_variables, names(leaf_data))

cat("FACTOR VARIABLES (", length(existing_factors), " found):\n")
print(existing_factors)
cat("\nDATE VARIABLES (", length(existing_dates), " found):\n")
print(existing_dates)
cat("\nNUMERICAL VARIABLES (", length(existing_numerics), " found):\n")
print(existing_numerics)

# Variables not classified
all_classified <- c(existing_factors, existing_dates, existing_numerics)
unclassified <- setdiff(names(leaf_data), all_classified)
if(length(unclassified) > 0) {
  cat("\nUNCLASSIFIED VARIABLES:\n")
  print(unclassified)
  
  cat("\nSUGGESTED CLASSIFICATION FOR UNCLASSIFIED VARIABLES:\n")
  for(var in unclassified) {
    sample_values <- head(unique(leaf_data[[var]][!is.na(leaf_data[[var]]) & leaf_data[[var]] != ""]), 5)
    cat("Variable:", var, "| Sample values:", paste(sample_values, collapse = ", "), "\n")
    
    non_empty <- leaf_data[[var]][!is.na(leaf_data[[var]]) & leaf_data[[var]] != ""]
    if(length(non_empty) > 0) {
      numeric_test <- gsub(",", ".", non_empty)
      if(all(str_detect(numeric_test, "^-?[0-9]*\\.?[0-9]+$|^$"), na.rm = TRUE)) {
        cat("  -> Suggested: NUMERIC\n")
      } else if(length(unique(non_empty)) < 20 && length(non_empty) > 10) {
        cat("  -> Suggested: FACTOR\n")
      } else {
        cat("  -> Suggested: CHARACTER/OTHER\n")
      }
    }
  }
}

# ===============================
# 3. CONVERT DATA TYPES
# ===============================
cat("\n3. DATA TYPE CONVERSION\n")
cat("=======================\n")

# Create a copy for conversion
leaf_data_converted <- leaf_data

# Convert factors
for(var in existing_factors) {
  if(var %in% names(leaf_data_converted)) {
    leaf_data_converted[[var]] <- as.factor(leaf_data_converted[[var]])
    cat("Converted", var, "to factor -", nlevels(leaf_data_converted[[var]]), "levels\n")
  }
}

# Convert dates
for(var in existing_dates) {
  if(var %in% names(leaf_data_converted)) {
    original_values <- leaf_data_converted[[var]]
    
    if(all(str_detect(original_values[!is.na(original_values) & original_values != ""], "^[0-9]+$"), na.rm = TRUE)) {
      leaf_data_converted[[var]] <- as.Date(as.numeric(original_values), origin = "1899-12-30")
      cat("Converted", var, "from Excel numeric format to Date\n")
    } else {
      date_converted <- as.Date(original_values, format = "%d/%m/%Y")
      if(all(is.na(date_converted))) {
        date_converted <- as.Date(original_values, format = "%Y-%m-%d")
      }
      if(all(is.na(date_converted))) {
        date_converted <- as.Date(original_values, format = "%m/%d/%Y")
      }
      leaf_data_converted[[var]] <- date_converted
      cat("Converted", var, "to Date format\n")
    }
    
    valid_dates <- leaf_data_converted[[var]][!is.na(leaf_data_converted[[var]])]
    if(length(valid_dates) > 0) {
      date_range <- range(valid_dates)
      cat("  Date range:", as.character(date_range[1]), "to", as.character(date_range[2]), "\n")
    }
  }
}

# Convert numerics
for(var in existing_numerics) {
  if(var %in% names(leaf_data_converted)) {
    leaf_data_converted[[var]][leaf_data_converted[[var]] == "" | leaf_data_converted[[var]] == " "] <- NA
    if(is.character(leaf_data_converted[[var]])) {
      leaf_data_converted[[var]] <- gsub(",", ".", leaf_data_converted[[var]])
    }
    original_nas <- sum(is.na(leaf_data[[var]]))
    leaf_data_converted[[var]] <- as.numeric(leaf_data_converted[[var]])
    new_nas <- sum(is.na(leaf_data_converted[[var]]))
    conversion_nas <- new_nas - original_nas
    
    if(conversion_nas > 0) {
      cat("WARNING:", conversion_nas, "values became NA during conversion of", var, "\n")
      problematic <- leaf_data[[var]][is.na(leaf_data_converted[[var]]) & !is.na(leaf_data[[var]])]
      if(length(problematic) > 0) {
        cat("  Examples of unconverted values:", paste(head(problematic, 3), collapse = ", "), "\n")
      }
    } else {
      cat("Successfully converted", var, "to numeric\n")
    }
  }
}

str(leaf_data_converted)

```

## IV.3. Data Visualization by ggplot2

```{r}
# Load necessary packages
library(pacman)
p_load(
  dplyr, readr, lubridate, tidyr, stringr, skimr,
  VIM, summarytools, corrplot, janitor, purrr,
  ggplot2, gridExtra, RColorBrewer, viridis, scales,
  cowplot, ggthemes, Hmisc
)

# Set working directory and create output folder
# setwd("your/working/directory")  # Uncomment and set your directory

# Create folder for plots
if (!dir.exists("Enhanced_Leaf_Boxplots_Publication")) {
  dir.create("Enhanced_Leaf_Boxplots_Publication")
}

# Define enhanced color palettes for publication
publication_colors <- list(
  yield_colors = c("#006400", "#228B22", "#32CD32", "#7CFC00", "#98FB98", "#90EE90", "#F0FFF0", "#ADFF2F"),
  nutrient_colors = c("#8B0000", "#B22222", "#DC143C", "#FF4500", "#FF6347", "#FFA500", "#FFD700", "#FFFF00"),
  moisture_colors = c("#191970", "#4169E1", "#4682B4", "#5F9EA0", "#87CEEB", "#B0E0E6", "#E0F6FF", "#F0F8FF"),
  physical_colors = c("#8B4513", "#A0522D", "#CD853F", "#D2B48C", "#F4A460", "#FFDAB9", "#FFE4B5", "#F5DEB3"),
  age_colors = c("#800080", "#9932CC", "#9370DB", "#8A2BE2", "#7B68EE", "#6495ED", "#87CEEB", "#B0C4DE"),
  general_colors = c("#2F4F4F", "#696969", "#808080", "#A9A9A9", "#C0C0C0", "#D3D3D3", "#DCDCDC", "#F5F5F5"),
  comparison_colors = c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FECA57", "#FF9FF3", "#54A0FF", "#5F27CD")
)

# Enhanced publication-ready theme with TIGHT layout and narrow boxplots
publication_theme <- theme_minimal() +
  theme(
    text = element_text(size = 12, family = "Arial", color = "#1C1C1C"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5, color = "#000000", margin = margin(b = 8)),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "#333333", margin = margin(b = 10)),
    axis.title.x = element_text(size = 11, face = "bold", color = "#000000", margin = margin(t = 8)),
    axis.title.y = element_text(size = 11, face = "bold", color = "#000000", margin = margin(r = 8)),
    axis.text.x = element_text(size = 9, color = "#000000", angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 9, color = "#000000"),
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "#CCCCCC", size = 0.3, linetype = "solid"),
    panel.background = element_rect(fill = "white", color = "#000000", size = 0.8),
    plot.background = element_rect(fill = "white", color = "#000000", size = 1),
    strip.text = element_text(size = 11, face = "bold", color = "#000000"),
    plot.margin = margin(10, 10, 10, 10),  # Much tighter margins
    panel.spacing = unit(0.4, "lines"),    # Tighter panel spacing
    axis.line = element_line(color = "#000000", size = 0.6)
  )

# Function to detect outliers using IQR method
detect_outliers <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower_bound <- q1 - 1.5 * iqr
  upper_bound <- q3 + 1.5 * iqr
  return(x < lower_bound | x > upper_bound)
}

# Enhanced function to create narrow boxplots with red outlier detection
create_enhanced_boxplot <- function(data_long, title, subtitle, colors, x_label, y_label, 
                                  variable_col = "variable", value_col = "value", remove_outliers = FALSE) {
  
  # Detect outliers for each variable
  data_with_outliers <- data_long %>%
    group_by(!!sym(variable_col)) %>%
    mutate(
      is_outlier = detect_outliers(!!sym(value_col)),
      outlier_label = ifelse(is_outlier, "Outlier", "Normal")
    ) %>%
    ungroup()
  
  # Remove outliers if requested
  if (remove_outliers) {
    plot_data <- data_with_outliers %>% filter(!is_outlier)
    title <- paste(title, "(Without Outliers)")
  } else {
    plot_data <- data_with_outliers
  }
  
  p <- ggplot(plot_data, aes_string(x = variable_col, y = value_col, fill = variable_col)) +
    geom_boxplot(
      alpha = 0.85, 
      outlier.shape = NA,  # Hide default outliers
      width = 0.18,  # EVEN TIGHTER boxplots
      color = "#000000",  # Clear black contour
      size = 0.6,  # Slightly thinner for tighter look
      show.legend = FALSE
    ) +
    # Add red outlier points manually
    {if (!remove_outliers) {
      geom_point(
        data = plot_data %>% filter(is_outlier),
        aes_string(x = variable_col, y = value_col),
        color = "#DC143C",  # Bright red for outliers
        size = 1.2,  # Slightly smaller for tighter layout
        alpha = 0.8,
        shape = 16,
        inherit.aes = FALSE
      )
    }} +
    stat_summary(
      fun = median, 
      geom = "text", 
      aes(label = sprintf("%.2f", after_stat(y))),
      size = 2.5,  # Slightly smaller text for tighter layout
      color = "white", 
      fontface = "bold",
      vjust = 0.5
    ) +
    scale_fill_manual(values = colors) +
    labs(
      title = title,
      subtitle = subtitle,
      x = x_label,
      y = y_label
    ) +
    publication_theme +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
    coord_cartesian(clip = "off")
  
  return(p)
}

# Assume leaf_data_converted is already available from your data processing
# If not, you'll need to load and process your data first

# THEMATIC GROUP 1: YIELD VARIABLES
cat("Creating yield comparison boxplots...\n")

# Yield comparison (current vs previous year)
yield_comparison <- leaf_data_converted %>%
  select(Yield_tonnes_ha_year_minus_1, Yield_tonnes_ha_year_actual) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable, 
                          levels = c("Yield_tonnes_ha_year_minus_1", "Yield_tonnes_ha_year_actual"),
                          labels = c("Previous Year Yield", "Current Year Yield")))

p1_yield_comparison <- create_enhanced_boxplot(
  yield_comparison, 
  "Yield Comparison - Current vs Previous Year",
  "Sugar cane yield comparison (tonnes/ha)",
  publication_colors$yield_colors[1:2],
  "Year", "Yield (tonnes/ha)"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/01_Yield_Comparison.png", p1_yield_comparison,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# Create version without outliers
p1_yield_comparison_no_outliers <- create_enhanced_boxplot(
  yield_comparison, 
  "Yield Comparison - Current vs Previous Year",
  "Sugar cane yield comparison (tonnes/ha)",
  publication_colors$yield_colors[1:2],
  "Year", "Yield (tonnes/ha)",
  remove_outliers = TRUE
)

ggsave("Enhanced_Leaf_Boxplots_Publication/01_Yield_Comparison_NoOutliers.png", p1_yield_comparison_no_outliers,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 2: MOISTURE CONTENT VARIABLES
cat("Creating moisture content boxplots...\n")

# Moisture content variables
moisture_variables <- leaf_data_converted %>%
  select(Leaf_Moisture_percent, Sheath_Moisture_percent, 
         Cane_Moisture_percent_year_minus_1, Cane_Moisture_percent_year_actual) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Leaf_Moisture_percent", "Sheath_Moisture_percent", 
                                   "Cane_Moisture_percent_year_minus_1", "Cane_Moisture_percent_year_actual"),
                          labels = c("Leaf Moisture", "Sheath Moisture", "Cane Moisture (Prev)", "Cane Moisture (Curr)")))

p2_moisture <- create_enhanced_boxplot(
  moisture_variables,
  "Plant Moisture Content",
  "Moisture content in different plant parts (%)",
  publication_colors$moisture_colors[1:4],
  "Plant Part", "Moisture Content (%)"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/02_Moisture_Content.png", p2_moisture,
       width = 7, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# Create version without outliers
p2_moisture_no_outliers <- create_enhanced_boxplot(
  moisture_variables,
  "Plant Moisture Content",
  "Moisture content in different plant parts (%)",
  publication_colors$moisture_colors[1:4],
  "Plant Part", "Moisture Content (%)",
  remove_outliers = TRUE
)

ggsave("Enhanced_Leaf_Boxplots_Publication/02_Moisture_Content_NoOutliers.png", p2_moisture_no_outliers,
       width = 7, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 3: NUTRIENT CONTENT VARIABLES
cat("Creating nutrient content boxplots...\n")

# Nutrient content variables
nutrient_variables <- leaf_data_converted %>%
  select(Nitrogen_percent, Phosphorus_percent, Potassium_percent) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Nitrogen_percent", "Phosphorus_percent", "Potassium_percent"),
                          labels = c("Nitrogen (N)", "Phosphorus (P)", "Potassium (K)")))

p3_nutrients <- create_enhanced_boxplot(
  nutrient_variables,
  "Plant Nutrient Content",
  "Essential nutrient content in plant tissue (%)",
  publication_colors$nutrient_colors[1:3],
  "Nutrient", "Content (%)"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/03_Nutrient_Content.png", p3_nutrients,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# Create version without outliers
p3_nutrients_no_outliers <- create_enhanced_boxplot(
  nutrient_variables,
  "Plant Nutrient Content",
  "Essential nutrient content in plant tissue (%)",
  publication_colors$nutrient_colors[1:3],
  "Nutrient", "Content (%)",
  remove_outliers = TRUE
)

ggsave("Enhanced_Leaf_Boxplots_Publication/03_Nutrient_Content_NoOutliers.png", p3_nutrients_no_outliers,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 4: AGE VARIABLES
cat("Creating age-related boxplots...\n")

# Age-related variables
age_variables <- leaf_data_converted %>%
  select(Sampling_Age, Age_weeks, Harvest_Age_year) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Sampling_Age", "Age_weeks", "Harvest_Age_year"),
                          labels = c("Sampling Age", "Age (weeks)", "Harvest Age (years)")))

p4_age <- create_enhanced_boxplot(
  age_variables,
  "Plant Age Variables",
  "Different age measurements for sugar cane",
  publication_colors$age_colors[1:3],
  "Age Measurement", "Age Value"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/04_Age_Variables.png", p4_age,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 5: SURFACE AND PHYSICAL VARIABLES
cat("Creating surface and physical variable boxplots...\n")

# Surface and physical variables
physical_variables <- leaf_data_converted %>%
  select(Surface_ha, A_ce_jour, Agricultural_Surface_ha) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Surface_ha", "A_ce_jour", "Agricultural_Surface_ha"),
                          labels = c("Surface (ha)", "A ce jour", "Agricultural Surface (ha)")))

p5_physical <- create_enhanced_boxplot(
  physical_variables,
  "Surface and Physical Variables",
  "Land area and physical measurements",
  publication_colors$physical_colors[1:3],
  "Physical Measurement", "Value"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/05_Physical_Variables.png", p5_physical,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 6: SUGAR CONTENT VARIABLES
cat("Creating sugar content boxplots...\n")

# Sugar content variables (if they exist in your data)
if(all(c("Real_T_canne", "T_Pol_canne") %in% names(leaf_data_converted))) {
  sugar_variables <- leaf_data_converted %>%
    select(T_Pol_canne) %>%  # Excluding Real_T_canne as requested
    pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
    filter(!is.na(value)) %>%
    mutate(variable = factor(variable,
                            levels = c("T_Pol_canne"),
                            labels = c("T Pol Cane")))
  
  p6_sugar <- create_enhanced_boxplot(
    sugar_variables,
    "Sugar Content Variables",
    "Sugar content measurements",
    publication_colors$general_colors[1],
    "Sugar Measurement", "Value"
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/06_Sugar_Content.png", p6_sugar,
         width = 4, height = 4, dpi = 300, bg = "white")  # Tighter dimensions
}

# ENHANCED CORRELATION MATRIX
cat("Creating enhanced correlation matrix...\n")

# Select all numeric variables (excluding Real_T_canne and Surface_Age_days as requested)
excluded_vars <- c("Real_T_canne", "Surface_Age_days")

numeric_vars <- leaf_data_converted %>%
  select_if(is.numeric) %>%
  select(-any_of(excluded_vars))

# Remove variables that are entirely NA or have insufficient data for correlation
correlation_data <- numeric_vars %>%
  select_if(~ sum(!is.na(.)) > 10)  # Keep variables with at least 10 non-NA values

# Calculate correlation matrix with significance testing
cor_result <- rcorr(as.matrix(correlation_data), type = "pearson")
cor_matrix <- cor_result$r
p_matrix <- cor_result$P

# Handle any remaining NA/NaN/Inf values in correlation matrix
cor_matrix[is.na(cor_matrix)] <- 0
cor_matrix[is.infinite(cor_matrix)] <- 0

# Create enhanced correlation plot with BIGGER and BOLD fonts
png("Enhanced_Leaf_Boxplots_Publication/07_Enhanced_Correlation_Matrix.png", 
    width = 18, height = 16, units = "in", res = 300, bg = "white")

# Set up margins for title
par(mar = c(2, 2, 6, 3))

# Use standard blue to red color palette
corr_colors <- colorRampPalette(c("#053061", "#2166AC", "#4393C3", "#92C5DE", 
                                 "#D1E5F0", "#F7F7F7", 
                                 "#FDDBC7", "#F4A582", "#D6604D", "#B2182B", "#67001F"))(100)

# Create the correlation plot with BIGGER and BOLD text
corrplot(cor_matrix,
         method = "color",
         type = "upper",
         order = "original",
         col = corr_colors,
         tl.col = "#000000",
         tl.srt = 45,
         tl.cex = 1.2,         # BIGGER variable names
         tl.font = 2,          # BOLD variable names
         cl.cex = 1.3,         # BIGGER color legend text
         title = "Correlation Matrix of Leaf Dataset Variables\n(Excluding Real_T_canne and Surface_Age_days)",
         mar = c(0,0,4,0),
         diag = FALSE,
         addCoef.col = "#000000",
         number.cex = 0.7,     # BIGGER correlation values
         number.digits = 2,
         number.font = 2,      # BOLD correlation values
         outline = TRUE,
         cl.ratio = 0.2,
         cl.align.text = "c",
         cl.offset = 0.5)

# Add BIGGER and BOLDER significance stars
n_vars <- ncol(cor_matrix)
for(i in 1:(n_vars-1)) {
  for(j in (i+1):n_vars) {
    if(!is.na(p_matrix[i,j])) {
      if(p_matrix[i,j] < 0.001) {
        sig_symbol <- "***"
        sig_color <- "#000000"
        sig_cex <- 1.0        # BIGGER significance stars
      } else if(p_matrix[i,j] < 0.01) {
        sig_symbol <- "**"
        sig_color <- "#333333"
        sig_cex <- 1.0        # BIGGER significance stars
      } else if(p_matrix[i,j] < 0.05) {
        sig_symbol <- "*"
        sig_color <- "#666666"
        sig_cex <- 1.0        # BIGGER significance stars
      } else {
        sig_symbol <- ""
        sig_color <- "black"
        sig_cex <- 1.0
      }
      
      if(sig_symbol != "") {
        x_center <- j
        y_center <- n_vars - i + 1
        text(x_center, y_center + 0.35, sig_symbol, 
             cex = sig_cex, col = sig_color, font = 2)  # BOLD significance stars
      }
    }
  }
}

# Enhanced legend with BIGGER text
legend_x <- n_vars * 0.7
legend_y <- n_vars * 0.9

legend(x = legend_x, y = legend_y,
       legend = c(
         "Statistical Significance:",
         "*** p < 0.001", 
         "**  p < 0.01", 
         "*   p < 0.05",
         "",
         "Correlation Strength:",
         "Red: Strong Positive (r > 0.7)",
         "Light Red: Moderate Positive (0.3 < r ≤ 0.7)", 
         "White: Weak (-0.3 ≤ r ≤ 0.3)",
         "Light Blue: Moderate Negative (-0.7 ≤ r < -0.3)",
         "Blue: Strong Negative (r < -0.7)"
       ), 
       text.col = c("#000000", "#000000", "#333333", "#666666", "white",
                   "#000000", "#B2182B", "#F4A582", "#000000", "#92C5DE", "#2166AC"), 
       bty = "n", 
       cex = 1.0,            # BIGGER legend text
       text.font = 2,        # BOLD legend text
       title = "Legend",
       title.col = "#000000",
       title.cex = 1.2,      # BIGGER legend title
       title.font = 2)       # BOLD legend title

dev.off()

# YIELD COMPARISON ANALYSIS
cat("Creating detailed yield comparison plots...\n")

# Scatter plot comparing yields with trend line
p_yield_scatter <- ggplot(leaf_data_converted, 
                         aes(x = Yield_tonnes_ha_year_minus_1, y = Yield_tonnes_ha_year_actual)) +
  geom_point(alpha = 0.7, size = 2, color = "#2F4F4F") +
  geom_smooth(method = "lm", se = TRUE, color = "#DC143C", fill = "#FFB6C1", alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000", size = 1) +
  labs(
    title = "Yield Correlation: Current vs Previous Year",
    subtitle = "Red line shows trend, black dashed line shows perfect correlation",
    x = "Previous Year Yield (tonnes/ha)",
    y = "Current Year Yield (tonnes/ha)"
  ) +
  publication_theme +
  theme(legend.position = "none")

ggsave("Enhanced_Leaf_Boxplots_Publication/08_Yield_Correlation_Scatter.png", p_yield_scatter,
       width = 7, height = 5, dpi = 300, bg = "white")  # Tighter dimensions

# Difference analysis
yield_diff_data <- leaf_data_converted %>%
  filter(!is.na(Yield_tonnes_ha_year_minus_1) & !is.na(Yield_tonnes_ha_year_actual)) %>%
  mutate(
    Yield_Difference = Yield_tonnes_ha_year_actual - Yield_tonnes_ha_year_minus_1,
    Yield_Percent_Change = (Yield_Difference / Yield_tonnes_ha_year_minus_1) * 100,
    Change_Category = case_when(
      Yield_Percent_Change > 10 ~ "Strong Increase (>10%)",
      Yield_Percent_Change > 0 ~ "Moderate Increase (0-10%)",
      Yield_Percent_Change >= -10 ~ "Moderate Decrease (0 to -10%)",
      TRUE ~ "Strong Decrease (<-10%)"
    )
  )

# Yield difference histogram
p_yield_diff <- ggplot(yield_diff_data, aes(x = Yield_Difference)) +
  geom_histogram(bins = 30, fill = "#4ECDC4", color = "#000000", alpha = 0.8, size = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#DC143C", size = 1) +
  geom_vline(xintercept = mean(yield_diff_data$Yield_Difference, na.rm = TRUE), 
             color = "#2F4F4F", size = 1) +
  labs(
    title = "Distribution of Yield Changes",
    subtitle = "Red line: no change, Dark line: mean change",
    x = "Yield Change (tonnes/ha)",
    y = "Frequency"
  ) +
  publication_theme

ggsave("Enhanced_Leaf_Boxplots_Publication/09_Yield_Difference_Distribution.png", p_yield_diff,
       width = 7, height = 5, dpi = 300, bg = "white")  # Tighter dimensions

# Summary statistics
cat("\n=== YIELD COMPARISON SUMMARY ===\n")
summary_stats <- yield_diff_data %>%
  summarise(
    n_observations = n(),
    mean_prev_yield = mean(Yield_tonnes_ha_year_minus_1, na.rm = TRUE),
    mean_curr_yield = mean(Yield_tonnes_ha_year_actual, na.rm = TRUE),
    mean_difference = mean(Yield_Difference, na.rm = TRUE),
    mean_percent_change = mean(Yield_Percent_Change, na.rm = TRUE),
    correlation = cor(Yield_tonnes_ha_year_minus_1, Yield_tonnes_ha_year_actual, use = "complete.obs")
  )

print(summary_stats)

# Category distribution
cat("\n=== YIELD CHANGE CATEGORIES ===\n")
category_summary <- yield_diff_data %>%
  count(Change_Category) %>%
  mutate(percentage = n / sum(n) * 100)

print(category_summary)
```

```{r}


# Create folder for plots
if (!dir.exists("Enhanced_Leaf_Boxplots_Publication")) {
  dir.create("Enhanced_Leaf_Boxplots_Publication")
}

# Define enhanced color palettes for publication with MORE colors
publication_colors <- list(
  yield_colors = c("#006400", "#228B22", "#32CD32", "#7CFC00", "#98FB98", "#90EE90", "#F0FFF0", "#ADFF2F", 
                   "#9ACD32", "#6B8E23", "#556B2F", "#8FBC8F", "#20B2AA", "#008B8B", "#5F9EA0", "#4682B4"),
  nutrient_colors = c("#8B0000", "#B22222", "#DC143C", "#FF4500", "#FF6347", "#FFA500", "#FFD700", "#FFFF00",
                      "#FF1493", "#FF69B4", "#FFB6C1", "#FFA07A", "#FA8072", "#E9967A", "#CD5C5C", "#F08080"),
  moisture_colors = c("#191970", "#4169E1", "#4682B4", "#5F9EA0", "#87CEEB", "#B0E0E6", "#E0F6FF", "#F0F8FF",
                      "#00CED1", "#48D1CC", "#40E0D0", "#AFEEEE", "#B0E0E6", "#87CEFA", "#6495ED", "#4169E1"),
  physical_colors = c("#8B4513", "#A0522D", "#CD853F", "#D2B48C", "#F4A460", "#FFDAB9", "#FFE4B5", "#F5DEB3",
                      "#DEB887", "#BC8F8F", "#F5DEB3", "#FFEFD5", "#FFEBCD", "#FFDEAD", "#NAVAJOWHITE", "#MOCCASIN"),
  age_colors = c("#800080", "#9932CC", "#9370DB", "#8A2BE2", "#7B68EE", "#6495ED", "#87CEEB", "#B0C4DE",
                 "#DA70D6", "#EE82EE", "#DDA0DD", "#D8BFD8", "#THISTLE", "#PLUM", "#VIOLET", "#ORCHID"),
  general_colors = c("#2F4F4F", "#696969", "#808080", "#A9A9A9", "#C0C0C0", "#D3D3D3", "#DCDCDC", "#F5F5F5",
                     "#708090", "#778899", "#B0C4DE", "#E6E6FA", "#F8F8FF", "#F0F8FF", "#F5FFFA", "#HONEYDEW"),
  comparison_colors = c("#FF6B6B", "#4ECDC4", "#45B7D1", "#96CEB4", "#FECA57", "#FF9FF3", "#54A0FF", "#5F27CD",
                        "#FF7675", "#74B9FF", "#0984E3", "#00B894", "#FDCB6E", "#E17055", "#A29BFE", "#6C5CE7")
)

# Enhanced publication-ready theme with TIGHT layout and narrow boxplots
publication_theme <- theme_minimal() +
  theme(
    text = element_text(size = 12, family = "Arial", color = "#1C1C1C"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5, color = "#000000", margin = margin(b = 8)),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "#333333", margin = margin(b = 10)),
    axis.title.x = element_text(size = 11, face = "bold", color = "#000000", margin = margin(t = 8)),
    axis.title.y = element_text(size = 11, face = "bold", color = "#000000", margin = margin(r = 8)),
    axis.text.x = element_text(size = 9, color = "#000000", angle = 45, hjust = 1, vjust = 1),
    axis.text.y = element_text(size = 9, color = "#000000"),
    legend.position = "none",
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_line(color = "#CCCCCC", size = 0.3, linetype = "solid"),
    panel.background = element_rect(fill = "white", color = "#000000", size = 0.8),
    plot.background = element_rect(fill = "white", color = "#000000", size = 1),
    strip.text = element_text(size = 11, face = "bold", color = "#000000"),
    plot.margin = margin(10, 10, 10, 10),  # Much tighter margins
    panel.spacing = unit(0.4, "lines"),    # Tighter panel spacing
    axis.line = element_line(color = "#000000", size = 0.6)
  )

# Function to detect outliers using IQR method
detect_outliers <- function(x) {
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr <- q3 - q1
  lower_bound <- q1 - 1.5 * iqr
  upper_bound <- q3 + 1.5 * iqr
  return(x < lower_bound | x > upper_bound)
}

# Enhanced function to create narrow boxplots with red outlier detection
create_enhanced_boxplot <- function(data_long, title, subtitle, colors, x_label, y_label, 
                                  variable_col = "variable", value_col = "value", remove_outliers = FALSE) {
  
  # Detect outliers for each variable
  data_with_outliers <- data_long %>%
    group_by(!!sym(variable_col)) %>%
    mutate(
      is_outlier = detect_outliers(!!sym(value_col)),
      outlier_label = ifelse(is_outlier, "Outlier", "Normal")
    ) %>%
    ungroup()
  
  # Remove outliers if requested
  if (remove_outliers) {
    plot_data <- data_with_outliers %>% filter(!is_outlier)
    title <- paste(title, "(Without Outliers)")
  } else {
    plot_data <- data_with_outliers
  }
  
  p <- ggplot(plot_data, aes_string(x = variable_col, y = value_col, fill = variable_col)) +
    geom_boxplot(
      alpha = 0.85, 
      outlier.shape = NA,  # Hide default outliers
      width = 0.18,  # EVEN TIGHTER boxplots
      color = "#000000",  # Clear black contour
      size = 0.6,  # Slightly thinner for tighter look
      show.legend = FALSE
    ) +
    # Add red outlier points manually
    {if (!remove_outliers) {
      geom_point(
        data = plot_data %>% filter(is_outlier),
        aes_string(x = variable_col, y = value_col),
        color = "#DC143C",  # Bright red for outliers
        size = 1.2,  # Slightly smaller for tighter layout
        alpha = 0.8,
        shape = 16,
        inherit.aes = FALSE
      )
    }} +
    stat_summary(
      fun = median, 
      geom = "text", 
      aes(label = sprintf("%.2f", after_stat(y))),
      size = 2.5,  # Slightly smaller text for tighter layout
      color = "white", 
      fontface = "bold",
      vjust = 0.5
    ) +
    scale_fill_manual(values = colors) +
    labs(
      title = title,
      subtitle = subtitle,
      x = x_label,
      y = y_label
    ) +
    publication_theme +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
    coord_cartesian(clip = "off")
  
  return(p)
}

# Assume leaf_data_converted is already available from your data processing
# If not, you'll need to load and process your data first

# THEMATIC GROUP 1: YIELD VARIABLES
cat("Creating yield comparison boxplots...\n")

# Yield comparison (current vs previous year)
yield_comparison <- leaf_data_converted %>%
  select(Yield_tonnes_ha_year_minus_1, Yield_tonnes_ha_year_actual) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable, 
                          levels = c("Yield_tonnes_ha_year_minus_1", "Yield_tonnes_ha_year_actual"),
                          labels = c("Previous Year Yield", "Current Year Yield")))

p1_yield_comparison <- create_enhanced_boxplot(
  yield_comparison, 
  "Yield Comparison - Current vs Previous Year",
  "Sugar cane yield comparison (tonnes/ha)",
  publication_colors$yield_colors[1:2],
  "Year", "Yield (tonnes/ha)"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/01_Yield_Comparison.png", p1_yield_comparison,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# Create version without outliers
p1_yield_comparison_no_outliers <- create_enhanced_boxplot(
  yield_comparison, 
  "Yield Comparison - Current vs Previous Year",
  "Sugar cane yield comparison (tonnes/ha)",
  publication_colors$yield_colors[1:2],
  "Year", "Yield (tonnes/ha)",
  remove_outliers = TRUE
)

ggsave("Enhanced_Leaf_Boxplots_Publication/01_Yield_Comparison_NoOutliers.png", p1_yield_comparison_no_outliers,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 2: MOISTURE CONTENT VARIABLES
cat("Creating moisture content boxplots...\n")

# Moisture content variables
moisture_variables <- leaf_data_converted %>%
  select(Leaf_Moisture_percent, Sheath_Moisture_percent, 
         Cane_Moisture_percent_year_minus_1, Cane_Moisture_percent_year_actual) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Leaf_Moisture_percent", "Sheath_Moisture_percent", 
                                   "Cane_Moisture_percent_year_minus_1", "Cane_Moisture_percent_year_actual"),
                          labels = c("Leaf Moisture", "Sheath Moisture", "Cane Moisture (Prev)", "Cane Moisture (Curr)")))

p2_moisture <- create_enhanced_boxplot(
  moisture_variables,
  "Plant Moisture Content",
  "Moisture content in different plant parts (%)",
  publication_colors$moisture_colors[1:4],
  "Plant Part", "Moisture Content (%)"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/02_Moisture_Content.png", p2_moisture,
       width = 7, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# Create version without outliers
p2_moisture_no_outliers <- create_enhanced_boxplot(
  moisture_variables,
  "Plant Moisture Content",
  "Moisture content in different plant parts (%)",
  publication_colors$moisture_colors[1:4],
  "Plant Part", "Moisture Content (%)",
  remove_outliers = TRUE
)

ggsave("Enhanced_Leaf_Boxplots_Publication/02_Moisture_Content_NoOutliers.png", p2_moisture_no_outliers,
       width = 7, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 3: NUTRIENT CONTENT VARIABLES
cat("Creating nutrient content boxplots...\n")

# Nutrient content variables
nutrient_variables <- leaf_data_converted %>%
  select(Nitrogen_percent, Phosphorus_percent, Potassium_percent) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Nitrogen_percent", "Phosphorus_percent", "Potassium_percent"),
                          labels = c("Nitrogen (N)", "Phosphorus (P)", "Potassium (K)")))

p3_nutrients <- create_enhanced_boxplot(
  nutrient_variables,
  "Plant Nutrient Content",
  "Essential nutrient content in plant tissue (%)",
  publication_colors$nutrient_colors[1:3],
  "Nutrient", "Content (%)"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/03_Nutrient_Content.png", p3_nutrients,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# Create version without outliers
p3_nutrients_no_outliers <- create_enhanced_boxplot(
  nutrient_variables,
  "Plant Nutrient Content",
  "Essential nutrient content in plant tissue (%)",
  publication_colors$nutrient_colors[1:3],
  "Nutrient", "Content (%)",
  remove_outliers = TRUE
)

ggsave("Enhanced_Leaf_Boxplots_Publication/03_Nutrient_Content_NoOutliers.png", p3_nutrients_no_outliers,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 4: AGE VARIABLES
cat("Creating age-related boxplots...\n")

# Age-related variables
age_variables <- leaf_data_converted %>%
  select(Sampling_Age, Age_weeks, Harvest_Age_year) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Sampling_Age", "Age_weeks", "Harvest_Age_year"),
                          labels = c("Sampling Age", "Age (weeks)", "Harvest Age (years)")))

p4_age <- create_enhanced_boxplot(
  age_variables,
  "Plant Age Variables",
  "Different age measurements for sugar cane",
  publication_colors$age_colors[1:3],
  "Age Measurement", "Age Value"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/04_Age_Variables.png", p4_age,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 5: SURFACE AND PHYSICAL VARIABLES
cat("Creating surface and physical variable boxplots...\n")

# Surface and physical variables
physical_variables <- leaf_data_converted %>%
  select(Surface_ha, A_ce_jour, Agricultural_Surface_ha) %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  filter(!is.na(value)) %>%
  mutate(variable = factor(variable,
                          levels = c("Surface_ha", "A_ce_jour", "Agricultural_Surface_ha"),
                          labels = c("Surface (ha)", "A ce jour", "Agricultural Surface (ha)")))

p5_physical <- create_enhanced_boxplot(
  physical_variables,
  "Surface and Physical Variables",
  "Land area and physical measurements",
  publication_colors$physical_colors[1:3],
  "Physical Measurement", "Value"
)

ggsave("Enhanced_Leaf_Boxplots_Publication/05_Physical_Variables.png", p5_physical,
       width = 5, height = 4, dpi = 300, bg = "white")  # Tighter dimensions

# THEMATIC GROUP 6: SUGAR CONTENT VARIABLES
cat("Creating sugar content boxplots...\n")

# Sugar content variables (if they exist in your data)
if(all(c("Real_T_canne", "T_Pol_canne") %in% names(leaf_data_converted))) {
  sugar_variables <- leaf_data_converted %>%
    select(T_Pol_canne) %>%  # Excluding Real_T_canne as requested
    pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
    filter(!is.na(value)) %>%
    mutate(variable = factor(variable,
                            levels = c("T_Pol_canne"),
                            labels = c("T Pol Cane")))
  
  p6_sugar <- create_enhanced_boxplot(
    sugar_variables,
    "Sugar Content Variables",
    "Sugar content measurements",
    publication_colors$general_colors[1],
    "Sugar Measurement", "Value"
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/06_Sugar_Content.png", p6_sugar,
         width = 4, height = 4, dpi = 300, bg = "white")  # Tighter dimensions
}

# ENHANCED CORRELATION MATRIX
cat("Creating enhanced correlation matrix...\n")

# Select all numeric variables (excluding Real_T_canne and Surface_Age_days as requested)
excluded_vars <- c("Real_T_canne", "Surface_Age_days")

numeric_vars <- leaf_data_converted %>%
  select_if(is.numeric) %>%
  select(-any_of(excluded_vars))

# Remove variables that are entirely NA or have insufficient data for correlation
correlation_data <- numeric_vars %>%
  select_if(~ sum(!is.na(.)) > 10)  # Keep variables with at least 10 non-NA values

# Calculate correlation matrix with significance testing
cor_result <- rcorr(as.matrix(correlation_data), type = "pearson")
cor_matrix <- cor_result$r
p_matrix <- cor_result$P

# Handle any remaining NA/NaN/Inf values in correlation matrix
cor_matrix[is.na(cor_matrix)] <- 0
cor_matrix[is.infinite(cor_matrix)] <- 0

# Create enhanced correlation plot with BIGGER and BOLD fonts
png("Enhanced_Leaf_Boxplots_Publication/07_Enhanced_Correlation_Matrix.png", 
    width = 16, height = 14, units = "in", res = 300, bg = "white")

# Set up margins for title
par(mar = c(2, 2, 6, 3))

# Use standard blue to red color palette
corr_colors <- colorRampPalette(c("#053061", "#2166AC", "#4393C3", "#92C5DE", 
                                 "#D1E5F0", "#F7F7F7", 
                                 "#FDDBC7", "#F4A582", "#D6604D", "#B2182B", "#67001F"))(100)

# Create the correlation plot with BIGGER and BOLD text
corrplot(cor_matrix,
         method = "color",
         type = "upper",
         order = "original",
         col = corr_colors,
         tl.col = "#000000",
         tl.srt = 45,
         tl.cex = 1.2,         # BIGGER variable names
         tl.font = 2,          # BOLD variable names
         cl.cex = 1.3,         # BIGGER color legend text
         title = "Correlation Matrix of Leaf Dataset Variables\n(Excluding Real_T_canne and Surface_Age_days)",
         mar = c(0,0,4,0),
         diag = FALSE,
         addCoef.col = "#000000",
         number.cex = 0.7,     # BIGGER correlation values
         number.digits = 2,
         number.font = 2,      # BOLD correlation values
         outline = TRUE,
         cl.ratio = 0.2,
         cl.align.text = "c",
         cl.offset = 0.5)

# Add BIGGER and BOLDER significance stars
n_vars <- ncol(cor_matrix)
for(i in 1:(n_vars-1)) {
  for(j in (i+1):n_vars) {
    if(!is.na(p_matrix[i,j])) {
      if(p_matrix[i,j] < 0.001) {
        sig_symbol <- "***"
        sig_color <- "#000000"
        sig_cex <- 1.0        # BIGGER significance stars
      } else if(p_matrix[i,j] < 0.01) {
        sig_symbol <- "**"
        sig_color <- "#333333"
        sig_cex <- 1.0        # BIGGER significance stars
      } else if(p_matrix[i,j] < 0.05) {
        sig_symbol <- "*"
        sig_color <- "#666666"
        sig_cex <- 1.0        # BIGGER significance stars
      } else {
        sig_symbol <- ""
        sig_color <- "black"
        sig_cex <- 1.0
      }
      
      if(sig_symbol != "") {
        x_center <- j
        y_center <- n_vars - i + 1
        text(x_center, y_center + 0.35, sig_symbol, 
             cex = sig_cex, col = sig_color, font = 2)  # BOLD significance stars
      }
    }
  }
}

# Enhanced legend with BIGGER text
legend_x <- n_vars * 0.7
legend_y <- n_vars * 0.9

legend(
  "bottomleft",
  inset = c(0.12, 0.1), # nudges right and up from the very corner
  legend = c(
    "Statistical Significance:",
    "*** p < 0.001", 
    "**  p < 0.01", 
    "*   p < 0.05",
    "",
    "Correlation Strength:",
    "Red: Strong Positive (r > 0.7)",
    "Light Red: Moderate Positive (0.3 < r ≤ 0.7)", 
    "White: Weak (-0.3 ≤ r ≤ 0.3)",
    "Light Blue: Moderate Negative (-0.7 ≤ r < -0.3)",
    "Blue: Strong Negative (r < -0.7)"
  ),
  text.col = c("#000000", "#000000", "#333333", "#666666", "white",
               "#000000", "#B2182B", "#F4A582", "#000000", "#92C5DE", "#2166AC"),
  bty = "n",
  cex = 1.0,
  text.font = 2,
  title = "Legend",
  title.col = "#000000",
  title.cex = 1.2,
  title.font = 2
)

dev.off()

# YIELD COMPARISON ANALYSIS
cat("Creating detailed yield comparison plots...\n")

# Scatter plot comparing yields with trend line
p_yield_scatter <- ggplot(leaf_data_converted, 
                         aes(x = Yield_tonnes_ha_year_minus_1, y = Yield_tonnes_ha_year_actual)) +
  geom_point(alpha = 0.7, size = 2, color = "#2F4F4F") +
  geom_smooth(method = "lm", se = TRUE, color = "#DC143C", fill = "#FFB6C1", alpha = 0.3) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "#000000", size = 1) +
  labs(
    title = "Yield Correlation: Current vs Previous Year",
    subtitle = "Red line shows trend, black dashed line shows perfect correlation",
    x = "Previous Year Yield (tonnes/ha)",
    y = "Current Year Yield (tonnes/ha)"
  ) +
  publication_theme +
  theme(legend.position = "none")

ggsave("Enhanced_Leaf_Boxplots_Publication/08_Yield_Correlation_Scatter.png", p_yield_scatter,
       width = 7, height = 5, dpi = 300, bg = "white")  # Tighter dimensions

# Difference analysis
yield_diff_data <- leaf_data_converted %>%
  filter(!is.na(Yield_tonnes_ha_year_minus_1) & !is.na(Yield_tonnes_ha_year_actual)) %>%
  mutate(
    Yield_Difference = Yield_tonnes_ha_year_actual - Yield_tonnes_ha_year_minus_1,
    Yield_Percent_Change = (Yield_Difference / Yield_tonnes_ha_year_minus_1) * 100,
    Change_Category = case_when(
      Yield_Percent_Change > 10 ~ "Strong Increase (>10%)",
      Yield_Percent_Change > 0 ~ "Moderate Increase (0-10%)",
      Yield_Percent_Change >= -10 ~ "Moderate Decrease (0 to -10%)",
      TRUE ~ "Strong Decrease (<-10%)"
    )
  )

# Yield difference histogram
p_yield_diff <- ggplot(yield_diff_data, aes(x = Yield_Difference)) +
  geom_histogram(bins = 30, fill = "#4ECDC4", color = "#000000", alpha = 0.8, size = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", color = "#DC143C", size = 1) +
  geom_vline(xintercept = mean(yield_diff_data$Yield_Difference, na.rm = TRUE), 
             color = "#2F4F4F", size = 1) +
  labs(
    title = "Distribution of Yield Changes",
    subtitle = "Red line: no change, Dark line: mean change",
    x = "Yield Change (tonnes/ha)",
    y = "Frequency"
  ) +
  publication_theme

ggsave("Enhanced_Leaf_Boxplots_Publication/09_Yield_Difference_Distribution.png", p_yield_diff,
       width = 7, height = 5, dpi = 300, bg = "white")  # Tighter dimensions

# Summary statistics
cat("\n=== YIELD COMPARISON SUMMARY ===\n")
summary_stats <- yield_diff_data %>%
  summarise(
    n_observations = n(),
    mean_prev_yield = mean(Yield_tonnes_ha_year_minus_1, na.rm = TRUE),
    mean_curr_yield = mean(Yield_tonnes_ha_year_actual, na.rm = TRUE),
    mean_difference = mean(Yield_Difference, na.rm = TRUE),
    mean_percent_change = mean(Yield_Percent_Change, na.rm = TRUE),
    correlation = cor(Yield_tonnes_ha_year_minus_1, Yield_tonnes_ha_year_actual, use = "complete.obs")
  )

print(summary_stats)

# Category distribution
cat("\n=== YIELD CHANGE CATEGORIES ===\n")
category_summary <- yield_diff_data %>%
  count(Change_Category) %>%
  mutate(percentage = n / sum(n) * 100)

print(category_summary)
print(category_summary)

# ===================================================================
# CATEGORICAL COMPARISONS - MEANINGFUL GROUPINGS
# ===================================================================

cat("\n=== CREATING CATEGORICAL COMPARISON PLOTS ===\n")

# Enhanced theme for categorical comparisons with legend
categorical_theme <- publication_theme +
  theme(
    legend.position = "bottom",
    legend.title = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 9),
    legend.margin = margin(t = 10),
    axis.text.x = element_text(size = 9, color = "#000000", angle = 0, hjust = 0.5),  # No angle for categories
    strip.text = element_text(size = 10, face = "bold", color = "#000000")
  )

# Function for categorical boxplots with AUTOMATIC color generation
create_categorical_boxplot <- function(data, y_var, group_var, title, subtitle, y_label, group_label, colors = NULL) {
  
  # Filter out NA values
  plot_data <- data %>%
    filter(!is.na(!!sym(y_var)) & !is.na(!!sym(group_var)) & !!sym(group_var) != "")
  
  # Get unique groups and count them
  unique_groups <- unique(plot_data[[group_var]])
  n_groups <- length(unique_groups)
  
  # Create color palette automatically based on number of groups
  if(is.null(colors) || length(colors) < n_groups) {
    if(n_groups <= 16) {
      # Use predefined colors for smaller groups
      colors <- publication_colors$comparison_colors[1:n_groups]
    } else {
      # Generate colors automatically for larger groups
      if(n_groups <= 30) {
        # Use RColorBrewer for medium groups
        colors <- RColorBrewer::brewer.pal(min(n_groups, 11), "Spectral")
        if(n_groups > 11) {
          colors <- colorRampPalette(colors)(n_groups)
        }
      } else {
        # Use viridis for very large groups
        colors <- viridis::viridis(n_groups, alpha = 0.8)
      }
    }
  } else {
    # Extend provided colors if needed
    if(length(colors) < n_groups) {
      colors <- colorRampPalette(colors)(n_groups)
    }
  }
  
  # Ensure we have exactly the right number of colors
  colors <- colors[1:n_groups]
  
  cat("Creating plot for", n_groups, "groups in", group_var, "\n")
  
  p <- ggplot(plot_data, aes_string(x = group_var, y = y_var, fill = group_var)) +
    geom_boxplot(
      alpha = 0.85,
      outlier.shape = NA,
      width = 0.6,  # Wider for categorical comparisons
      color = "#000000",
      size = 0.6,
      show.legend = ifelse(n_groups > 10, FALSE, TRUE)  # Hide legend if too many groups
    ) +
    geom_point(
      position = position_jitter(width = 0.2, height = 0),
      alpha = 0.4,
      size = 0.8,
      color = "#333333",
      show.legend = FALSE
    ) +
    stat_summary(
      fun = median,
      geom = "text",
      aes(label = sprintf("%.2f", after_stat(y))),
      size = ifelse(n_groups > 15, 2.0, 2.5),  # Smaller text for many groups
      color = "white",
      fontface = "bold",
      vjust = 0.5
    ) +
    scale_fill_manual(values = colors, name = group_label) +
    labs(
      title = title,
      subtitle = paste(subtitle, "(", n_groups, "groups)"),
      x = group_label,
      y = y_label
    ) +
    categorical_theme +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.01)) +
    theme(
      axis.text.x = element_text(
        angle = ifelse(n_groups > 8, 45, 0), 
        hjust = ifelse(n_groups > 8, 1, 0.5),
        size = ifelse(n_groups > 15, 7, 9)  # Smaller text for many groups
      ),
      legend.position = ifelse(n_groups > 10, "none", "bottom")
    )
  
  return(p)
}

# 1. YIELD BY VARIETY
cat("Creating yield by variety comparison...\n")

if("Variety" %in% names(leaf_data_converted)) {
  
  # Current year yield by variety
  p_yield_variety_current <- create_categorical_boxplot(
    leaf_data_converted,
    "Yield_tonnes_ha_year_actual",
    "Variety",
    "Current Year Yield by Variety",
    "Sugar cane yield comparison across different varieties",
    "Current Year Yield (tonnes/ha)",
    "Variety"
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/10_Yield_by_Variety_Current.png", p_yield_variety_current,
         width = 10, height = 6, dpi = 300, bg = "white")
  
  # Previous year yield by variety
  p_yield_variety_previous <- create_categorical_boxplot(
    leaf_data_converted,
    "Yield_tonnes_ha_year_minus_1",
    "Variety",
    "Previous Year Yield by Variety",
    "Sugar cane yield comparison across different varieties",
    "Previous Year Yield (tonnes/ha)",
    "Variety"
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/11_Yield_by_Variety_Previous.png", p_yield_variety_previous,
         width = 10, height = 6, dpi = 300, bg = "white")
}

# 2. NUTRIENT CONTENT BY VARIETY
cat("Creating nutrient content by variety comparison...\n")

if("Variety" %in% names(leaf_data_converted)) {
  
  # Nitrogen by variety
  p_nitrogen_variety <- create_categorical_boxplot(
    leaf_data_converted,
    "Nitrogen_percent",
    "Variety",
    "Nitrogen Content by Variety",
    "Nitrogen content comparison across sugar cane varieties",
    "Nitrogen Content (%)",
    "Variety"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/12_Nitrogen_by_Variety.png", p_nitrogen_variety,
         width = 10, height = 6, dpi = 300, bg = "white")
  
  # Phosphorus by variety
  p_phosphorus_variety <- create_categorical_boxplot(
    leaf_data_converted,
    "Phosphorus_percent",
    "Variety",
    "Phosphorus Content by Variety",
    "Phosphorus content comparison across sugar cane varieties",
    "Phosphorus Content (%)",
    "Variety"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/13_Phosphorus_by_Variety.png", p_phosphorus_variety,
         width = 10, height = 6, dpi = 300, bg = "white")
  
  # Potassium by variety
  p_potassium_variety <- create_categorical_boxplot(
    leaf_data_converted,
    "Potassium_percent",
    "Variety",
    "Potassium Content by Variety",
    "Potassium content comparison across sugar cane varieties",
    "Potassium Content (%)",
    "Variety"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/14_Potassium_by_Variety.png", p_potassium_variety,
         width = 10, height = 6, dpi = 300, bg = "white")
}

# 3. MOISTURE CONTENT BY VARIETY
cat("Creating moisture content by variety comparison...\n")

if("Variety" %in% names(leaf_data_converted)) {
  
  # Leaf moisture by variety
  p_leaf_moisture_variety <- create_categorical_boxplot(
    leaf_data_converted,
    "Leaf_Moisture_percent",
    "Variety",
    "Leaf Moisture Content by Variety",
    "Leaf moisture content comparison across sugar cane varieties",
    "Leaf Moisture Content (%)",
    "Variety"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/15_Leaf_Moisture_by_Variety.png", p_leaf_moisture_variety,
         width = 10, height = 6, dpi = 300, bg = "white")
  
  # Sheath moisture by variety
  p_sheath_moisture_variety <- create_categorical_boxplot(
    leaf_data_converted,
    "Sheath_Moisture_percent",
    "Variety",
    "Sheath Moisture Content by Variety",
    "Sheath moisture content comparison across sugar cane varieties",
    "Sheath Moisture Content (%)",
    "Variety"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/16_Sheath_Moisture_by_Variety.png", p_sheath_moisture_variety,
         width = 10, height = 6, dpi = 300, bg = "white")
}

# 4. YIELD BY FARM
cat("Creating yield by farm comparison...\n")

if("Farm" %in% names(leaf_data_converted)) {
  
  # Current year yield by farm
  p_yield_farm <- create_categorical_boxplot(
    leaf_data_converted,
    "Yield_tonnes_ha_year_actual",
    "Farm",
    "Current Year Yield by Farm",
    "Sugar cane yield comparison across different farms",
    "Current Year Yield (tonnes/ha)",
    "Farm"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/17_Yield_by_Farm.png", p_yield_farm,
         width = 12, height = 6, dpi = 300, bg = "white")
}

# 5. YIELD BY IRRIGATION TYPE
cat("Creating yield by irrigation type comparison...\n")

if("Irrigation_Type" %in% names(leaf_data_converted)) {
  
  # Current year yield by irrigation type
  p_yield_irrigation <- create_categorical_boxplot(
    leaf_data_converted,
    "Yield_tonnes_ha_year_actual",
    "Irrigation_Type",
    "Current Year Yield by Irrigation Type",
    "Sugar cane yield comparison across different irrigation methods",
    "Current Year Yield (tonnes/ha)",
    "Irrigation Type"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/18_Yield_by_Irrigation.png", p_yield_irrigation,
         width = 8, height = 6, dpi = 300, bg = "white")
  
  # Nutrient content by irrigation type
  nutrient_irrigation_data <- leaf_data_converted %>%
    select(Irrigation_Type, Nitrogen_percent, Phosphorus_percent, Potassium_percent) %>%
    filter(!is.na(Irrigation_Type) & Irrigation_Type != "") %>%
    pivot_longer(cols = c(Nitrogen_percent, Phosphorus_percent, Potassium_percent),
                 names_to = "Nutrient", values_to = "Content") %>%
    filter(!is.na(Content)) %>%
    mutate(Nutrient = factor(Nutrient,
                            levels = c("Nitrogen_percent", "Phosphorus_percent", "Potassium_percent"),
                            labels = c("Nitrogen (N)", "Phosphorus (P)", "Potassium (K)")))
  
  p_nutrients_irrigation <- ggplot(nutrient_irrigation_data, 
                                  aes(x = Irrigation_Type, y = Content, fill = Nutrient)) +
    geom_boxplot(alpha = 0.85, width = 0.6, color = "#000000", size = 0.6,
                 position = position_dodge(width = 0.8)) +
    geom_point(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
               alpha = 0.4, size = 0.8, color = "#333333") +
    scale_fill_manual(values = publication_colors$nutrient_colors[1:3], name = "Nutrient") +
    labs(
      title = "Nutrient Content by Irrigation Type",
      subtitle = "NPK content comparison across irrigation methods",
      x = "Irrigation Type",
      y = "Nutrient Content (%)"
    ) +
    categorical_theme
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/19_Nutrients_by_Irrigation.png", p_nutrients_irrigation,
         width = 10, height = 6, dpi = 300, bg = "white")
}

# 6. YIELD BY GROWTH PERIOD
cat("Creating yield by growth period comparison...\n")

if("Growth_Period" %in% names(leaf_data_converted)) {
  
  # Current year yield by growth period
  p_yield_period <- create_categorical_boxplot(
    leaf_data_converted,
    "Yield_tonnes_ha_year_actual",
    "Growth_Period",
    "Current Year Yield by Growth Period",
    "Sugar cane yield comparison across different growth periods",
    "Current Year Yield (tonnes/ha)",
    "Growth Period"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/20_Yield_by_Growth_Period.png", p_yield_period,
         width = 8, height = 6, dpi = 300, bg = "white")
}

# 7. YIELD BY CROP CYCLE
cat("Creating yield by crop cycle comparison...\n")

if("Crop_Cycle" %in% names(leaf_data_converted)) {
  
  # Current year yield by crop cycle
  p_yield_cycle <- create_categorical_boxplot(
    leaf_data_converted,
    "Yield_tonnes_ha_year_actual",
    "Crop_Cycle",
    "Current Year Yield by Crop Cycle",
    "Sugar cane yield comparison across different crop cycles",
    "Current Year Yield (tonnes/ha)",
    "Crop Cycle"  # Removed colors parameter to use automatic generation
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/21_Yield_by_Crop_Cycle.png", p_yield_cycle,
         width = 8, height = 6, dpi = 300, bg = "white")
}

# 8. HARVEST AGE BY VARIETY
cat("Creating harvest age by variety comparison...\n")

if(all(c("Variety", "Harvest_Age_year") %in% names(leaf_data_converted))) {
  
  p_age_variety <- create_categorical_boxplot(
    leaf_data_converted,
    "Harvest_Age_year",
    "Variety",
    "Harvest Age by Variety",
    "Harvest age comparison across sugar cane varieties",
    "Harvest Age (years)",
    "Variety",
    publication_colors$age_colors
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/22_Harvest_Age_by_Variety.png", p_age_variety,
         width = 10, height = 6, dpi = 300, bg = "white")
}

# 9. SAMPLING AGE BY GROWTH PERIOD
cat("Creating sampling age by growth period comparison...\n")

if(all(c("Growth_Period", "Sampling_Age") %in% names(leaf_data_converted))) {
  
  p_sampling_age_period <- create_categorical_boxplot(
    leaf_data_converted,
    "Sampling_Age",
    "Growth_Period",
    "Sampling Age by Growth Period",
    "Sampling age comparison across different growth periods",
    "Sampling Age",
    "Growth Period",
    publication_colors$age_colors
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/23_Sampling_Age_by_Period.png", p_sampling_age_period,
         width = 8, height = 6, dpi = 300, bg = "white")
}

# 10. SURFACE AREA BY FARM
cat("Creating surface area by farm comparison...\n")

if(all(c("Farm", "Surface_ha") %in% names(leaf_data_converted))) {
  
  p_surface_farm <- create_categorical_boxplot(
    leaf_data_converted,
    "Surface_ha",
    "Farm",
    "Surface Area by Farm",
    "Plot surface area comparison across different farms",
    "Surface Area (ha)",
    "Farm",
    publication_colors$physical_colors
  )
  
  ggsave("Enhanced_Leaf_Boxplots_Publication/24_Surface_by_Farm.png", p_surface_farm,
         width = 12, height = 6, dpi = 300, bg = "white")
}

# 11. COMPREHENSIVE MULTI-FACETED ANALYSIS
cat("Creating comprehensive multi-faceted analysis...\n")

# Yield performance by variety and irrigation type (if both exist)
if(all(c("Variety", "Irrigation_Type") %in% names(leaf_data_converted))) {
  
  multi_comparison_data <- leaf_data_converted %>%
    filter(!is.na(Variety) & !is.na(Irrigation_Type) & 
           !is.na(Yield_tonnes_ha_year_actual) &
           Variety != "" & Irrigation_Type != "") %>%
    select(Variety, Irrigation_Type, Yield_tonnes_ha_year_actual)
  
  if(nrow(multi_comparison_data) > 10) {  # Only if sufficient data
    
    p_multi_yield <- ggplot(multi_comparison_data, 
                           aes(x = Variety, y = Yield_tonnes_ha_year_actual, fill = Irrigation_Type)) +
      geom_boxplot(alpha = 0.85, width = 0.7, color = "#000000", size = 0.6,
                   position = position_dodge(width = 0.8)) +
      geom_point(position = position_jitterdodge(jitter.width = 0.2, dodge.width = 0.8),
                 alpha = 0.4, size = 0.8, color = "#333333") +
      scale_fill_manual(values = publication_colors$comparison_colors, 
                        name = "Irrigation Type") +
      labs(
        title = "Yield Performance: Variety × Irrigation Type Interaction",
        subtitle = "Sugar cane yield comparison across varieties and irrigation methods",
        x = "Variety",
        y = "Current Year Yield (tonnes/ha)"
      ) +
      categorical_theme +
      theme(axis.text.x = element_text(angle = 45, hjust = 1))
    
    ggsave("Enhanced_Leaf_Boxplots_Publication/25_Yield_Variety_Irrigation_Interaction.png", p_multi_yield,
           width = 14, height = 7, dpi = 300, bg = "white")
  }
}

# 12. STATISTICAL SUMMARY TABLE FOR CATEGORICAL COMPARISONS
cat("Creating statistical summary for categorical comparisons...\n")

# Function to create statistical summary
create_categorical_summary <- function(data, numeric_var, category_var) {
  if(all(c(numeric_var, category_var) %in% names(data))) {
    summary_table <- data %>%
      filter(!is.na(!!sym(numeric_var)) & !is.na(!!sym(category_var)) & !!sym(category_var) != "") %>%
      group_by(!!sym(category_var)) %>%
      summarise(
        n = n(),
        mean = round(mean(!!sym(numeric_var), na.rm = TRUE), 3),
        median = round(median(!!sym(numeric_var), na.rm = TRUE), 3),
        sd = round(sd(!!sym(numeric_var), na.rm = TRUE), 3),
        min = round(min(!!sym(numeric_var), na.rm = TRUE), 3),
        max = round(max(!!sym(numeric_var), na.rm = TRUE), 3),
        .groups = 'drop'
      ) %>%
      arrange(desc(mean))
    
    cat(paste("\n=== SUMMARY:", numeric_var, "BY", category_var, "===\n"))
    print(summary_table)
    return(summary_table)
  }
}

# Generate summaries for key comparisons
if("Variety" %in% names(leaf_data_converted)) {
  create_categorical_summary(leaf_data_converted, "Yield_tonnes_ha_year_actual", "Variety")
  create_categorical_summary(leaf_data_converted, "Nitrogen_percent", "Variety")
  create_categorical_summary(leaf_data_converted, "Leaf_Moisture_percent", "Variety")
}

if("Farm" %in% names(leaf_data_converted)) {
  create_categorical_summary(leaf_data_converted, "Yield_tonnes_ha_year_actual", "Farm")
}

if("Irrigation_Type" %in% names(leaf_data_converted)) {
  create_categorical_summary(leaf_data_converted, "Yield_tonnes_ha_year_actual", "Irrigation_Type")
}
```

# V. Multivariate Data Analysis - PCA

```{r}
# Multivariate Data Analysis for Sugar Cane Leaf Dataset using explor package
# Load required libraries
library(explor)
library(FactoMineR)
library(factoextra)
library(corrplot)
library(dplyr)
library(tidyr)
library(ggplot2)
library(plotly)
library(DT)
library(RColorBrewer)

cat("================================================================\n")
cat("MULTIVARIATE DATA ANALYSIS - SUGAR CANE LEAF DATASET\n")
cat("================================================================\n")

# Prepare the data for multivariate analysis
cat("Preparing leaf data for multivariate analysis...\n")

cat("• Total rows in leaf_data_converted:", nrow(leaf_data_converted), "\n")
cat("• Total columns:", ncol(leaf_data_converted), "\n")
cat("• Column names:", paste(names(leaf_data_converted)[1:min(10, ncol(leaf_data_converted))], collapse = ", "), "...\n\n")

# Get all numeric variables, excluding ID-like variables and previously excluded ones
excluded_vars <- c("Real_T_canne", "Surface_Age_days", "Plot", "Campaign", 
                   "Farm", "Variety", "Crop_Cycle", "Growth_Period",
                   "Irrigation_Type", "Harvest_Month", "Sampling_Month", 
                   "Week_Number", "Data_Day", "Period_Validation", 
                   "Plot_Campaign", "Harvest_Validation", "Note")

numeric_vars <- leaf_data_converted %>%
  select_if(is.numeric) %>%
  select(-any_of(excluded_vars)) %>%
  names()

cat("Identified numeric variables:", length(numeric_vars), "\n")
if(length(numeric_vars) > 0) {
  cat("Variables for PCA:\n")
  for(i in 1:length(numeric_vars)) {
    cat("  ", i, ". ", numeric_vars[i], "\n")
  }
} else {
  cat("ERROR: No numeric variables found!\n")
  stop("Please check your data structure")
}

# Check for missing values in each variable
missing_summary <- leaf_data_converted %>%
  select(all_of(numeric_vars)) %>%
  summarise_all(~sum(is.na(.))) %>%
  pivot_longer(everything(), names_to = "variable", values_to = "missing_count") %>%
  mutate(missing_percent = round(missing_count / nrow(leaf_data_converted) * 100, 2)) %>%
  arrange(desc(missing_count))

cat("\nMissing values summary:\n")
print(missing_summary)

# Remove variables with too many missing values (>70% for leaf data)
vars_to_keep <- missing_summary %>%
  filter(missing_percent <= 70) %>%
  pull(variable)

cat("\nKeeping", length(vars_to_keep), "variables with ≤70% missing values\n")
cat("Variables kept:", paste(vars_to_keep, collapse = ", "), "\n")

# Variables removed due to high missingness
vars_removed <- missing_summary %>%
  filter(missing_percent > 70) %>%
  pull(variable)

if(length(vars_removed) > 0) {
  cat("Variables removed due to >70% missing:", paste(vars_removed, collapse = ", "), "\n")
}

# Create analysis dataset with categorical info for supplementary analysis
categorical_vars <- c("Campaign", "Farm", "Variety", "Crop_Cycle", "Growth_Period", "Irrigation_Type")
existing_categorical <- intersect(categorical_vars, names(leaf_data_converted))

analysis_data <- leaf_data_converted %>%
  select(all_of(vars_to_keep), any_of(existing_categorical))

# Remove rows with too many missing values (>50% of numeric variables)
numeric_data_for_filtering <- analysis_data %>% select(all_of(vars_to_keep))
rows_to_keep <- rowSums(is.na(numeric_data_for_filtering)) <= length(vars_to_keep) * 0.5

analysis_data <- analysis_data[rows_to_keep, ]

cat("\nAfter filtering:\n")
cat("• Numeric variables:", length(vars_to_keep), "\n")
cat("• Categorical variables:", length(existing_categorical), "\n")
cat("• Observations:", nrow(analysis_data), "\n\n")

if(nrow(analysis_data) == 0) {
  stop("ERROR: No observations left after filtering!")
}

# Extract numeric data for PCA and handle remaining missing values
pca_data <- analysis_data %>%
  select(all_of(vars_to_keep))

# Replace remaining missing values with column medians (more robust for agricultural data)
pca_data <- pca_data %>%
  mutate_all(~ifelse(is.na(.), median(., na.rm = TRUE), .))

# Check for and remove variables with zero variance
var_check <- apply(pca_data, 2, function(x) {
  if(all(is.na(x))) {
    return(NA)
  } else {
    return(var(x, na.rm = TRUE))
  }
})

# Identify problematic columns
zero_var_cols <- names(var_check)[var_check == 0 | is.na(var_check) | is.infinite(var_check)]

if(length(zero_var_cols) > 0) {
  cat("Removing", length(zero_var_cols), "variables with zero/NA/infinite variance:\n")
  cat(paste(zero_var_cols, collapse = ", "), "\n\n")
  pca_data <- pca_data %>% select(-all_of(zero_var_cols))
  vars_to_keep <- setdiff(vars_to_keep, zero_var_cols)
} else {
  cat("All variables have valid variance.\n\n")
}

# Scale the data (important for PCA when variables have different units)
pca_data_scaled <- scale(pca_data)

# Add meaningful rownames
if("Campaign" %in% names(analysis_data) && "Plot" %in% names(leaf_data_converted)) {
  rownames(pca_data_scaled) <- paste0("Obs_", 1:nrow(pca_data_scaled))
} else {
  rownames(pca_data_scaled) <- paste0("Sample_", 1:nrow(pca_data_scaled))
}

# Update final variables
numeric_vars_final <- vars_to_keep

cat("Final dataset for PCA:\n")
cat("• Variables:", ncol(pca_data_scaled), "\n")
cat("• Observations:", nrow(pca_data_scaled), "\n")
cat("• Final variables:", paste(numeric_vars_final, collapse = ", "), "\n\n")

# Create output directory
if(!dir.exists("Enhanced_Leaf_PCA_Analysis")) {
  dir.create("Enhanced_Leaf_PCA_Analysis", recursive = TRUE)
}

cat("================================================================\n")
cat("1. PRINCIPAL COMPONENT ANALYSIS (PCA)\n")
cat("================================================================\n")

# Perform PCA using FactoMineR
# Include supplementary categorical variables if available
if(length(existing_categorical) > 0) {
  # Prepare supplementary qualitative variables
  sup_data <- analysis_data %>% 
    select(all_of(existing_categorical)) %>%
    mutate_all(as.factor)
  
  # Remove rows with all NA in supplementary variables
  sup_data[is.na(sup_data)] <- "Unknown"
  
  # Find columns in supplementary data
  sup_quali <- which(names(analysis_data) %in% existing_categorical)
  
  cat("Including supplementary categorical variables:", paste(existing_categorical, collapse = ", "), "\n")
  
  pca_result <- PCA(
    cbind(pca_data_scaled, sup_data), 
    scale.unit = FALSE,  # Already scaled
    quali.sup = sup_quali,  # Supplementary qualitative variables
    graph = FALSE,       # Don't show default plots
    ncp = min(10, ncol(pca_data_scaled))  # Max 10 components
  )
} else {
  cat("No supplementary categorical variables found\n")
  pca_result <- PCA(pca_data_scaled, 
                    scale.unit = FALSE,
                    graph = FALSE,
                    ncp = min(10, ncol(pca_data_scaled)))
}

cat("================================================================\n")
cat("2. INTERACTIVE PCA EXPLORATION\n")
cat("================================================================\n")

# Note: This will open an interactive Shiny app
# Uncomment the line below to launch the interactive explorer
explor(pca_result)

cat("================================================================\n")
cat("3. CORRELATION ANALYSIS\n")
cat("================================================================\n")

# Create correlation matrix
cor_matrix <- cor(pca_data, use = "complete.obs")

# Enhanced correlation plot
png("Enhanced_Leaf_PCA_Analysis/correlation_matrix.png", 
    width = 14, height = 12, units = "in", res = 300, bg = "white")

par(mar = c(2, 2, 4, 2))
corrplot(cor_matrix, 
         method = "color",
         type = "upper",
         order = "hclust",
         tl.col = "#000000",
         tl.srt = 45,
         tl.cex = 1.0,
         tl.font = 2,
         cl.cex = 1.1,
         col = colorRampPalette(c("#053061", "#2166AC", "#4393C3", "#92C5DE", 
                                 "#D1E5F0", "#F7F7F7", 
                                 "#FDDBC7", "#F4A582", "#D6604D", "#B2182B", "#67001F"))(100),
         title = "Sugar Cane Leaf Dataset - Correlation Matrix\n",
         mar = c(0,0,3,0),
         diag = FALSE,
         addCoef.col = "#000000",
         number.cex = 0.6,
         number.digits = 2,
         outline = TRUE)

dev.off()

cat("Correlation matrix saved as correlation_matrix.png\n\n")

cat("================================================================\n")
cat("4. PCA SUMMARY STATISTICS\n")
cat("================================================================\n")

# Print PCA summary
cat("Eigenvalues and variance explained:\n")
eigen_summary <- round(pca_result$eig, 3)
print(eigen_summary)

# Calculate how many components explain 80% of variance
cum_var <- pca_result$eig[,3]
components_80 <- which(cum_var >= 80)[1]
cat("\nComponents needed to explain 80% variance:", components_80, "\n")

cat("\nVariables with highest contributions to PC1:\n")
contrib_pc1 <- sort(pca_result$var$contrib[,1], decreasing = TRUE)[1:min(10, nrow(pca_result$var$contrib))]
print(round(contrib_pc1, 2))

cat("\nVariables with highest contributions to PC2:\n")
contrib_pc2 <- sort(pca_result$var$contrib[,2], decreasing = TRUE)[1:min(10, nrow(pca_result$var$contrib))]
print(round(contrib_pc2, 2))

cat("\nVariables with highest quality of representation (Cos2) on PC1-PC2:\n")
cos2_pc12 <- pca_result$var$cos2[,1] + pca_result$var$cos2[,2]
best_represented <- sort(cos2_pc12, decreasing = TRUE)[1:min(10, length(cos2_pc12))]
print(round(best_represented, 3))

cat("================================================================\n")
cat("5. DETAILED PCA VISUALIZATIONS\n")
cat("================================================================\n")

# 1. Enhanced scree plot
scree_plot <- fviz_eig(pca_result, 
                       addlabels = TRUE, 
                       ylim = c(0, max(pca_result$eig[,2]) * 1.1),
                       barfill = "#4ECDC4",
                       barcolor = "#2F4F4F",
                       linecolor = "#DC143C") +
  ggtitle("PCA Scree Plot - Sugar Cane Leaf Dataset") +
  labs(subtitle = paste("First", components_80, "components explain 80% of variance")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 16, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 12))

ggsave("Enhanced_Leaf_PCA_Analysis/pca_scree_plot.png", scree_plot,
       width = 10, height = 6, dpi = 300, bg = "white")

# 2. Enhanced biplot
biplot <- fviz_pca_biplot(pca_result, 
                         repel = TRUE,
                         col.var = "#2E9FDF",
                         col.ind = "#696969",
                         alpha.ind = 0.3,
                         alpha.var = 0.8,
                         title = "PCA Biplot - Sugar Cane Leaf Variables and Observations",
                         subtitle = paste("PC1 (", round(pca_result$eig[1,2], 1), "%) vs PC2 (", 
                                        round(pca_result$eig[2,2], 1), "%)", sep="")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11))

ggsave("Enhanced_Leaf_PCA_Analysis/pca_biplot.png", biplot,
       width = 12, height = 9, dpi = 300, bg = "white")

# 3. Variables contribution plot
var_plot <- fviz_pca_var(pca_result,
                        col.var = "contrib",
                        gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                        repel = TRUE,
                        title = "Variables Contribution to PC1-PC2",
                        subtitle = "Color intensity = Contribution level") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11))

ggsave("Enhanced_Leaf_PCA_Analysis/pca_variables_plot.png", var_plot,
       width = 10, height = 8, dpi = 300, bg = "white")

# 4. Variables quality plot (cos2)
cos2_plot <- fviz_pca_var(pca_result,
                         col.var = "cos2",
                         gradient.cols = c("#BB4444", "#EE9988", "#FFFFFF", "#77AADD", "#4477AA"),
                         repel = TRUE,
                         title = "Variables Quality of Representation (Cos2)",
                         subtitle = "How well variables are represented by PC1-PC2") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11))

ggsave("Enhanced_Leaf_PCA_Analysis/pca_variables_cos2_plot.png", cos2_plot,
       width = 10, height = 8, dpi = 300, bg = "white")

# 5. Individuals plot colored by contribution
ind_plot <- fviz_pca_ind(pca_result,
                        col.ind = "contrib",
                        gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
                        repel = FALSE,
                        alpha.ind = 0.6,
                        title = "Individuals Plot - Colored by Contribution",
                        subtitle = "Each point represents one observation") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
        plot.subtitle = element_text(hjust = 0.5, size = 11))

ggsave("Enhanced_Leaf_PCA_Analysis/pca_individuals_plot.png", ind_plot,
       width = 10, height = 8, dpi = 300, bg = "white")

cat("================================================================\n")
cat("6. SUPPLEMENTARY CATEGORICAL ANALYSIS\n")
cat("================================================================\n")

# Analyze supplementary categorical variables if available
if(length(existing_categorical) > 0) {
  for(cat_var in existing_categorical) {
    if(cat_var %in% names(analysis_data)) {
      cat("Analyzing", cat_var, "...\n")
      
      # Create plot colored by categorical variable
      cat_levels <- unique(analysis_data[[cat_var]])
      cat_levels <- cat_levels[!is.na(cat_levels) & cat_levels != ""]
      
      if(length(cat_levels) > 1 && length(cat_levels) <= 20) {
        # Create color palette
        if(length(cat_levels) <= 8) {
          colors <- RColorBrewer::brewer.pal(max(3, length(cat_levels)), "Set2")[1:length(cat_levels)]
        } else {
          colors <- rainbow(length(cat_levels))
        }
        
        # Plot individuals colored by categorical variable
        tryCatch({
          sup_plot <- fviz_pca_ind(pca_result,
                                  habillage = cat_var,
                                  palette = colors,
                                  addEllipses = TRUE,
                                  ellipse.level = 0.68,
                                  repel = FALSE,
                                  alpha.ind = 0.7,
                                  title = paste("PCA - Individuals colored by", cat_var),
                                  subtitle = paste("Ellipses show 68% confidence intervals")) +
            theme_minimal() +
            theme(plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
                  plot.subtitle = element_text(hjust = 0.5, size = 11),
                  legend.title = element_text(face = "bold"))
          
          filename <- paste0("Enhanced_Leaf_PCA_Analysis/pca_by_", tolower(cat_var), ".png")
          ggsave(filename, sup_plot, width = 12, height = 8, dpi = 300, bg = "white")
          
          cat("  • Saved:", filename, "\n")
        }, error = function(e) {
          cat("  • Could not create plot for", cat_var, ":", e$message, "\n")
        })
      } else {
        cat("  • Skipping", cat_var, "- too many levels or insufficient data\n")
      }
    }
  }
}

cat("================================================================\n")
cat("7. AGRICULTURAL INTERPRETATION ANALYSIS\n")
cat("================================================================\n")

# Group variables by agricultural meaning
agricultural_groups <- list(
  Yield = grep("yield|Yield", numeric_vars_final, value = TRUE, ignore.case = TRUE),
  Moisture = grep("moisture|Moisture", numeric_vars_final, value = TRUE, ignore.case = TRUE),
  Nutrients = grep("nitrogen|phosphorus|potassium|Nitrogen|Phosphorus|Potassium", numeric_vars_final, value = TRUE, ignore.case = TRUE),
  Age = grep("age|Age|weeks|Weeks", numeric_vars_final, value = TRUE, ignore.case = TRUE),
  Surface = grep("surface|Surface|ha", numeric_vars_final, value = TRUE, ignore.case = TRUE),
  Sugar = grep("pol|sugar|Sugar|T_", numeric_vars_final, value = TRUE, ignore.case = TRUE)
)

# Remove empty groups
agricultural_groups <- agricultural_groups[sapply(agricultural_groups, length) > 0]

cat("Variable groups identified:\n")
for(group_name in names(agricultural_groups)) {
  cat("•", group_name, ":", length(agricultural_groups[[group_name]]), "variables\n")
  cat("  Variables:", paste(agricultural_groups[[group_name]], collapse = ", "), "\n")
}

# Create contribution analysis by groups
cat("\nContribution of variable groups to first 3 PCs:\n")
for(i in 1:min(3, ncol(pca_result$var$contrib))) {
  cat("\n--- PC", i, "(", round(pca_result$eig[i,2], 1), "% variance) ---\n")
  
  for(group_name in names(agricultural_groups)) {
    group_vars <- agricultural_groups[[group_name]]
    group_contrib <- sum(pca_result$var$contrib[group_vars, i], na.rm = TRUE)
    cat(group_name, "group total contribution:", round(group_contrib, 2), "%\n")
  }
}

cat("================================================================\n")
cat("8. SAVE COMPREHENSIVE RESULTS\n")
cat("================================================================\n")

# Save all PCA results
save(pca_result, pca_data, pca_data_scaled, cor_matrix, 
     agricultural_groups, analysis_data, numeric_vars_final,
     file = "Enhanced_Leaf_PCA_Analysis/leaf_pca_complete_results.RData")

# Create comprehensive summary table
n_components <- min(10, nrow(pca_result$eig))
summary_table <- data.frame(
  Component = paste0("PC", 1:n_components),
  Eigenvalue = round(pca_result$eig[1:n_components, 1], 3),
  Variance_Percent = round(pca_result$eig[1:n_components, 2], 2),
  Cumulative_Percent = round(pca_result$eig[1:n_components, 3], 2)
)

write.csv(summary_table, "Enhanced_Leaf_PCA_Analysis/pca_summary.csv", row.names = FALSE)

# Create variable contributions table
contrib_table <- data.frame(
  Variable = rownames(pca_result$var$contrib),
  PC1_Contrib = round(pca_result$var$contrib[,1], 3),
  PC2_Contrib = round(pca_result$var$contrib[,2], 3),
  PC3_Contrib = round(pca_result$var$contrib[,3], 3),
  PC1_Cos2 = round(pca_result$var$cos2[,1], 3),
  PC2_Cos2 = round(pca_result$var$cos2[,2], 3),
  PC3_Cos2 = round(pca_result$var$cos2[,3], 3)
)

write.csv(contrib_table, "Enhanced_Leaf_PCA_Analysis/variables_contributions.csv", row.names = FALSE)

# Create coordinates table for individuals
coord_table <- data.frame(
  Sample = rownames(pca_result$ind$coord),
  PC1 = round(pca_result$ind$coord[,1], 3),
  PC2 = round(pca_result$ind$coord[,2], 3),
  PC3 = round(pca_result$ind$coord[,3], 3)
)

# Add categorical information if available
if(length(existing_categorical) > 0) {
  for(cat_var in existing_categorical) {
    if(cat_var %in% names(analysis_data)) {
      coord_table[[cat_var]] <- analysis_data[[cat_var]]
    }
  }
}

write.csv(coord_table, "Enhanced_Leaf_PCA_Analysis/sample_coordinates.csv", row.names = FALSE)


cat("explor(pca_result)\n\n")

# Summary statistics
cat("FINAL DATASET SUMMARY:\n")
cat("• Original variables:", length(numeric_vars), "\n")
cat("• Variables after filtering:", length(numeric_vars_final), "\n")
cat("• Final observations:", nrow(pca_data_scaled), "\n")
cat("• Categorical variables included:", length(existing_categorical), "\n")
cat("• Components explaining 80% variance:", components_80, "\n")

```

```{r}
# 3x3 PCA Variables Plot Grid 
# Load required libraries
library(FactoMineR)
library(factoextra)
library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(viridis)
library(cowplot)
library(dplyr)

cat("================================================================\n")
cat("3x3 PCA VARIABLES PLOT GRID - ALL VARIABLES VISIBLE\n")
cat("SUGAR CANE LEAF DATASET - UNIFORM SIZING\n")
cat("================================================================\n")

# Check if PCA result exists
if(!exists("pca_result")) {
  stop("ERROR: pca_result not found. Please run the PCA analysis first!")
}

# Define sophisticated color palettes for agricultural data
agricultural_colors <- c("#228B22", "#32CD32", "#FF6347", "#FFD700", "#87CEEB", "#DDA0DD", "#F4A460", "#98FB98", "#FF69B4")

# Create custom theme for bold, professional plots with UNIFORM sizing
theme_pca_agricultural <- function() {
  theme_minimal() +
  theme(
    # UNIFORM TEXT SIZING FOR AGRICULTURAL DATA
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5, color = "#1B4332", margin = margin(b = 8)),
    plot.subtitle = element_text(size = 10, hjust = 0.5, color = "#52796F", margin = margin(b = 6)),
    axis.title = element_text(size = 10, face = "bold", color = "#2D6A4F"),
    axis.text = element_text(size = 8.5, color = "#1B4332"),
    
    # UNIFORM LEGEND STYLING
    legend.title = element_text(size = 9, face = "bold", color = "#1B4332", hjust = 0.5),
    legend.text = element_text(size = 8, color = "#1B4332"),
    legend.key.size = unit(0.45, "cm"),
    legend.position = "right",
    legend.box = "vertical",
    legend.margin = margin(0, 0, 0, 8),
    legend.box.margin = margin(0, 0, 0, 0),
    
    # AGRICULTURAL-THEMED GRID AND BACKGROUND
    panel.grid.major = element_line(color = "#E8F5E8", size = 0.4, linetype = "dashed"),
    panel.grid.minor = element_line(color = "#F0FFF0", size = 0.25, linetype = "dotted"),
    panel.background = element_rect(fill = "#FEFFFE", color = NA),
    plot.background = element_rect(fill = "#FFFFFF", color = NA),
    
    # UNIFORM LEGEND BACKGROUND
    legend.background = element_rect(fill = "#F8FFF8", color = "#95D5B2", size = 0.4),
    legend.key = element_rect(fill = "transparent", color = NA),
    
    # UNIFORM MARGINS
    plot.margin = margin(10, 10, 10, 10),
    
    # STRIP TEXT FOR MULTI-PANEL CONSISTENCY
    strip.text = element_text(face = "bold", size = 9, color = "#1B4332"),
    
    # ENSURE SQUARE ASPECT RATIO
    aspect.ratio = 1
  )
}

# Enhanced function to create PCA variable plot showing ALL variables
create_agricultural_pca_plot <- function(pca_result, axes = c(1, 2), plot_title = "", 
                                         color_type = "contrib") {
  
  # Check if axes exist
  if (max(axes) > ncol(pca_result$var$coord)) {
    return(create_placeholder_plot(plot_title, axes))
  }
  
  # Variance explained for axis labels
  var_exp_1 <- round(pca_result$eig[axes[1], 2], 1)
  var_exp_2 <- round(pca_result$eig[axes[2], 2], 1)
  
  # --- CASE 1: Contribution ---
  if (color_type == "contrib") {
    p <- fviz_pca_var(pca_result,
                      axes = axes,
                      col.var = "contrib",
                      alpha.var = 0.85,
                      repel = TRUE,
                      labelsize = 3.0,
                      pointsize = 2.2,
                      arrowsize = 0.7,
                      select.var = list(contrib = Inf)) +  # Show ALL variables
      scale_color_gradient2(low = "#2D6A4F", mid = "#F77F00", high = "#D62828",
                            midpoint = 50, name = "Contrib\n%",
                            guide = guide_colorbar(title.position = "top", 
                                                   title.hjust = 0.5,
                                                   barwidth = 1.3,
                                                   barheight = 5.0,
                                                   ticks.colour = "#1B4332",
                                                   ticks.linewidth = 0.6,
                                                   frame.colour = "#1B4332",
                                                   frame.linewidth = 0.8)) +
      labs(title = plot_title,
           subtitle = paste0("PC", axes[1], " & PC", axes[2], " • All Variables"),
           x = paste0("PC", axes[1], " (", var_exp_1, "% variance)"),
           y = paste0("PC", axes[2], " (", var_exp_2, "% variance)"))
    
  # --- CASE 2: Cos² quality ---
  } else if (color_type == "cos2") {
    p <- fviz_pca_var(pca_result,
                      axes = axes,
                      col.var = "cos2",
                      alpha.var = 0.85,
                      repel = TRUE,
                      labelsize = 3.0,
                      pointsize = 2.2,
                      arrowsize = 0.7,
                      select.var = list(cos2 = 0)) +  # Show ALL variables
      scale_color_viridis_c(option = "viridis", name = "Quality\n(Cos²)",
                            guide = guide_colorbar(title.position = "top", 
                                                   title.hjust = 0.5,
                                                   barwidth = 1.3,
                                                   barheight = 5.0,
                                                   ticks.colour = "#1B4332",
                                                   ticks.linewidth = 0.6,
                                                   frame.colour = "#1B4332",
                                                   frame.linewidth = 0.8)) +
      labs(title = plot_title,
           subtitle = paste0("PC", axes[1], " & PC", axes[2], " • All Variables"),
           x = paste0("PC", axes[1], " (", var_exp_1, "% variance)"),
           y = paste0("PC", axes[2], " (", var_exp_2, "% variance)"))
  }
  
  # Apply agricultural theme
  p <- p + theme_pca_agricultural()
  
  # Add reference lines, border, and fixed square ratio
  p <- p +
    geom_hline(yintercept = 0, linetype = "solid", color = "#52796F", size = 0.9, alpha = 0.8) +
    geom_vline(xintercept = 0, linetype = "solid", color = "#52796F", size = 0.9, alpha = 0.8) +
    annotate("rect", xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = Inf, 
             fill = NA, color = "#1B4332", size = 1.3) +
    coord_fixed(ratio = 1,
                xlim = c(-1.1, 1.1),
                ylim = c(-1.1, 1.1),
                expand = FALSE,
                clip = "on")
  
  return(p)
}

# Function to create placeholder plot for missing dimensions
create_placeholder_plot <- function(title, axes) {
  ggplot() + 
    annotate("text", x = 0, y = 0, 
             label = paste("PC", max(axes), "\nnot available\nin dataset"), 
             size = 5, color = "#52796F", fontface = "bold") +
    annotate("text", x = 0, y = -0.3,
             label = "Insufficient components\nfor this combination",
             size = 3.5, color = "#7F8C8D") +
    labs(title = title, 
         subtitle = paste0("PC", axes[1], " & PC", axes[2], " • Not Available")) +
    theme_pca_agricultural() +
    theme(axis.text = element_blank(), 
          axis.ticks = element_blank(),
          panel.grid = element_blank()) +
    coord_fixed(ratio = 1, xlim = c(-1, 1), ylim = c(-1, 1))
}

# Define the 9 different dimension combinations
plot_combinations <- list(
  list(axes = c(1, 2), title = "Primary Components", color = "contrib"),
  list(axes = c(1, 3), title = "Main vs Third", color = "cos2"),
  list(axes = c(2, 3), title = "Secondary vs Third", color = "contrib"),
  list(axes = c(1, 4), title = "Primary Extended", color = "cos2"),
  list(axes = c(2, 4), title = "Secondary Extended", color = "contrib"),
  list(axes = c(3, 4), title = "Middle Components", color = "cos2"),
  list(axes = c(1, 5), title = "Primary vs Fifth", color = "contrib"),
  list(axes = c(2, 5), title = "Secondary vs Fifth", color = "cos2"),
  list(axes = c(3, 5), title = "Tertiary vs Fifth", color = "contrib")
)

# Create individual plots showing ALL variables
cat("Creating individual PCA variable plots showing ALL variables...\n")
plot_list <- list()

for(i in 1:9) {
  combo <- plot_combinations[[i]]
  
  cat("Creating plot", i, ":", combo$title, "(", paste(combo$axes, collapse=" vs "), ") with", combo$color, "coloring...\n")
  
  plot_list[[i]] <- create_agricultural_pca_plot(pca_result, 
                                               axes = combo$axes, 
                                               plot_title = combo$title,
                                               color_type = combo$color)
}

cat("All 9 plots created successfully!\n\n")

# Create the 3x3 grid layout
cat("Arranging plots in 3x3 grid...\n")

# Create main title
main_title <- ggdraw() + 
  draw_label("Sugar Cane Leaf Dataset: PCA Variables Grid - All Variables Shown", 
             fontface = 'bold', size = 18, color = "#1B4332") +
  draw_label("Complete visualization of all variable loadings across multiple PC dimensions", 
             fontface = 'plain', size = 12, color = "#52796F", 
             x = 0.5, y = 0.25) +
  draw_label("No variables filtered - all displayed in each plot", 
             fontface = 'italic', size = 10, color = "#74A57F", 
             x = 0.5, y = 0.05) +
  theme(plot.background = element_rect(fill = "#FFFFFF", color = NA))

# Arrange plots in 3x3 grid
variables_grid <- plot_grid(
  plot_list[[1]], plot_list[[2]], plot_list[[3]],
  plot_list[[4]], plot_list[[5]], plot_list[[6]],
  plot_list[[7]], plot_list[[8]], plot_list[[9]],
  ncol = 3, nrow = 3,
  align = "hv",
  axis = "tblr",
  rel_widths = c(1, 1, 1),
  rel_heights = c(1, 1, 1),
  hjust = 0.5,
  vjust = 0.5,
  scale = 0.96
)

# Combine title with grid
final_plot <- plot_grid(
  main_title, 
  variables_grid,
  ncol = 1,
  rel_heights = c(0.08, 0.92),
  align = "hv"
)

# Create output directory
if(!dir.exists("PCA_Analysis_All_Variables")) {
  dir.create("PCA_Analysis_All_Variables", recursive = TRUE)
}

# Save the comprehensive plot
cat("Saving comprehensive 3x3 PCA variables grid (all variables)...\n")

# High-resolution version
ggsave("PCA_Analysis_All_Variables/pca_variables_3x3_grid_all_variables.png", 
       final_plot,
       width = 20, height = 20,
       dpi = 300, 
       bg = "white",
       device = "png")

# Medium resolution version
ggsave("PCA_Analysis_All_Variables/pca_variables_3x3_grid_medium.png", 
       final_plot,
       width = 16, height = 16,
       dpi = 200, 
       bg = "white",
       device = "png")


# Print variable information
if(exists("pca_result")) {
  cat("VARIABLES INCLUDED IN ANALYSIS:\n")
  var_names <- rownames(pca_result$var$coord)
  for(i in 1:length(var_names)) {
    cat(sprintf("%2d. %s\n", i, var_names[i]))
  }
  
  cat("\nPCA DIMENSIONS AVAILABLE:", ncol(pca_result$var$coord), "\n")
  cat("VARIANCE EXPLAINED BY FIRST 5 PCs:\n")
  first_5 <- min(5, nrow(pca_result$eig))
  for(i in 1:first_5) {
    cat(sprintf("PC%d: %.1f%% (cumulative: %.1f%%)\n", 
                i, pca_result$eig[i,2], pca_result$eig[i,3]))
  }
}


```
