---
title: "15-09-2025_R-INLA"
format: revealjs
editor: visual
---

# R-INLA SPATIAL ANALYSIS: PREPROCESSING

# I. Shapefile Importation

This section imports the PARCELLES22324 shapefile and transforms it to UTM Zone 28N coordinates for accurate spatial analysis. The code validates geometries, filters columns to keep id, Bloc2022, and Var, calculates parcel centroids, and creates an irrigation type variable based on block classifications. Multiple visualizations verify data quality and spatial structure.

```{r}
# Load required libraries
library(sf)
library(ggplot2)
library(dplyr)

# Set file paths
shp_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Spatial_Data/Bandes et Parcelles22324/PARCELLES22324.shp"
output_folder <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy"

# Create output folder if it doesn't exist
if (!dir.exists(output_folder)) {
  dir.create(output_folder, recursive = TRUE)
  cat("Created output folder:", output_folder, "\n")
}

# Import the shapefile
cat("Importing shapefile...\n")
parcelles <- st_read(shp_path)

# Display basic information about the shapefile
cat("\n=== SHAPEFILE OVERVIEW ===\n")
cat("Number of features:", nrow(parcelles), "\n")
cat("Geometry type:", unique(st_geometry_type(parcelles)), "\n")

# Display CURRENT CRS information
cat("\n=== CURRENT CRS INFORMATION ===\n")
current_crs <- st_crs(parcelles)
cat("Current CRS:", current_crs$input, "\n")
cat("CRS Name:", current_crs$name, "\n")
if (!is.null(current_crs$epsg)) {
  cat("EPSG Code:", current_crs$epsg, "\n")
}
cat("Units:", current_crs$units_gdal, "\n")
cat("Is geographic (lat/lon)?", st_is_longlat(parcelles), "\n")

# Transform to UTM Zone 28N (EPSG:32628) - Senegal zone
cat("\n=== TRANSFORMING TO UTM ZONE 28N ===\n")
parcelles_utm <- st_transform(parcelles, crs = 32628)

# Verify transformation
new_crs <- st_crs(parcelles_utm)
cat("New CRS:", new_crs$input, "\n")
cat("New CRS Name:", new_crs$name, "\n")
cat("New EPSG Code:", new_crs$epsg, "\n")
cat("New Units:", new_crs$units_gdal, "\n")
cat("Is geographic (lat/lon)?", st_is_longlat(parcelles_utm), "\n")

# Display column names and types
cat("\n=== ATTRIBUTE TABLE STRUCTURE ===\n")
attr_data <- st_drop_geometry(parcelles_utm)
col_info <- sapply(attr_data, class)
for (i in 1:length(col_info)) {
  cat(sprintf("%-20s: %s\n", names(col_info)[i], paste(col_info[[i]], collapse = ", ")))
}

# Display comprehensive summary of attribute table
cat("\n=== ATTRIBUTE TABLE SUMMARY ===\n")
cat("Number of columns (excluding geometry):", ncol(attr_data), "\n")
cat("Number of rows:", nrow(attr_data), "\n")

# Display first few rows
cat("\n=== FIRST 6 ROWS OF ATTRIBUTE DATA ===\n")
print(head(attr_data, 6))

# Summary for each column
cat("\n=== DETAILED COLUMN SUMMARIES ===\n")
for (col_name in names(attr_data)) {
  cat("\n--- Column:", col_name, "---\n")
  col_data <- attr_data[[col_name]]
  
  if (is.numeric(col_data)) {
    cat("Type: Numeric\n")
    cat("Min:", min(col_data, na.rm = TRUE), "\n")
    cat("Max:", max(col_data, na.rm = TRUE), "\n")
    cat("Mean:", round(mean(col_data, na.rm = TRUE), 3), "\n")
    cat("Median:", median(col_data, na.rm = TRUE), "\n")
    cat("NA count:", sum(is.na(col_data)), "\n")
  } else if (is.character(col_data) || is.factor(col_data)) {
    cat("Type: Character/Factor\n")
    unique_vals <- unique(col_data)
    cat("Unique values:", length(unique_vals), "\n")
    cat("NA count:", sum(is.na(col_data)), "\n")
    if (length(unique_vals) <= 10) {
      cat("Values:", paste(unique_vals, collapse = ", "), "\n")
    } else {
      cat("Sample values:", paste(head(unique_vals, 5), collapse = ", "), "...\n")
    }
    # Show frequency table for categorical data
    if (length(unique_vals) <= 20) {
      cat("Frequency table:\n")
      freq_table <- table(col_data, useNA = "ifany")
      print(freq_table)
    }
  } else {
    cat("Type:", class(col_data), "\n")
    cat("Unique values:", length(unique(col_data)), "\n")
    cat("NA count:", sum(is.na(col_data)), "\n")
  }
}

# Define custom theme for plotting
custom_theme <- theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5, margin = margin(b = 15)),
    plot.subtitle = element_text(size = 12, hjust = 0.5, color = "gray40", margin = margin(b = 10)),
    plot.caption = element_text(size = 10, color = "gray50", hjust = 0, margin = margin(t = 10)),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 10),
    panel.grid.major = element_line(color = "gray90", linewidth = 0.3),
    panel.grid.minor = element_line(color = "gray95", linewidth = 0.2)
  )

# Filter and keep only specific columns: id, Bloc2022, Var
cat("\n=== FILTERING ATTRIBUTE TABLE ===\n")
cat("Original number of columns:", ncol(st_drop_geometry(parcelles_utm)), "\n")

# Check which columns exist in the dataset
available_cols <- names(st_drop_geometry(parcelles_utm))
cat("Available columns:", paste(available_cols, collapse = ", "), "\n")

# Define the columns we want to keep
desired_cols <- c("id", "Bloc2022", "Var")
existing_cols <- desired_cols[desired_cols %in% available_cols]
missing_cols <- desired_cols[!desired_cols %in% available_cols]

if (length(missing_cols) > 0) {
  cat("Warning: The following columns were not found:", paste(missing_cols, collapse = ", "), "\n")
}

cat("Keeping columns:", paste(existing_cols, collapse = ", "), "\n")

# Keep only the specified columns (plus geometry)
parcelles_filtered <- parcelles_utm[, existing_cols]

cat("New number of columns:", ncol(st_drop_geometry(parcelles_filtered)), "\n")

# Calculate centroids for each parcel
cat("\n=== CALCULATING CENTROIDS ===\n")
cat("Calculating centroids for each parcel...\n")

# First, check for geometry validity and fix any issues
cat("Checking geometry validity...\n")
valid_geoms <- st_is_valid(parcelles_filtered)
invalid_count <- sum(!valid_geoms, na.rm = TRUE)
cat("Invalid geometries found:", invalid_count, "\n")

# Fix invalid geometries if any
if (invalid_count > 0) {
  cat("Fixing invalid geometries...\n")
  parcelles_filtered <- st_make_valid(parcelles_filtered)
}

# Check for empty geometries
empty_geoms <- st_is_empty(parcelles_filtered)
empty_count <- sum(empty_geoms, na.rm = TRUE)
cat("Empty geometries found:", empty_count, "\n")

# Remove empty geometries if any
if (empty_count > 0) {
  cat("Removing empty geometries...\n")
  parcelles_filtered <- parcelles_filtered[!empty_geoms, ]
}

# Calculate centroids using st_point_on_surface for better results
# st_point_on_surface guarantees the point is inside the polygon
cat("Calculating centroids using point-on-surface method...\n")

# Try both methods and use the most appropriate one
tryCatch({
  # Method 1: Use st_point_on_surface (guarantees point inside polygon)
  centroids_pos <- st_point_on_surface(parcelles_filtered)
  centroid_coords <- st_coordinates(centroids_pos)
  
  # Add centroid coordinates to the filtered dataset
  parcelles_filtered$X.Centroid <- centroid_coords[, "X"]
  parcelles_filtered$Y.Centroid <- centroid_coords[, "Y"]
  
  cat("Used point-on-surface method (points guaranteed inside polygons)\n")
  
}, error = function(e) {
  cat("Point-on-surface failed, using geometric centroid...\n")
  
  # Method 2: Fallback to regular centroid
  centroids <- st_centroid(parcelles_filtered)
  centroid_coords <- st_coordinates(centroids)
  
  parcelles_filtered$X.Centroid <- centroid_coords[, "X"]
  parcelles_filtered$Y.Centroid <- centroid_coords[, "Y"]
  
  cat("Used geometric centroid method\n")
})

# Verify all parcels have centroids
na_centroids_x <- sum(is.na(parcelles_filtered$X.Centroid))
na_centroids_y <- sum(is.na(parcelles_filtered$Y.Centroid))
cat("Parcels with missing X centroids:", na_centroids_x, "\n")
cat("Parcels with missing Y centroids:", na_centroids_y, "\n")

# Check if we have unique IDs and centroids match
if ("id" %in% existing_cols) {
  unique_ids <- length(unique(parcelles_filtered[[existing_cols[existing_cols == "id"]]]))
  total_parcels <- nrow(parcelles_filtered)
  cat("Unique parcel IDs:", unique_ids, "\n")
  cat("Total parcels with centroids:", total_parcels, "\n")
  
  if (unique_ids != total_parcels) {
    cat("WARNING: Number of unique IDs doesn't match total parcels!\n")
  }
}

cat("Centroids calculated successfully!\n")
cat("X.Centroid range:", round(min(parcelles_filtered$X.Centroid, na.rm = TRUE), 2), "to", round(max(parcelles_filtered$X.Centroid, na.rm = TRUE), 2), "meters\n")
cat("Y.Centroid range:", round(min(parcelles_filtered$Y.Centroid, na.rm = TRUE), 2), "to", round(max(parcelles_filtered$Y.Centroid, na.rm = TRUE), 2), "meters\n")

# Add irrigation type variable based on Bloc2022 values
cat("\n=== ADDING IRRIGATION TYPE VARIABLE ===\n")

# Check if Bloc2022 column exists
if ("Bloc2022" %in% names(parcelles_filtered)) {
  # Create irrig_type variable
  parcelles_filtered$irrig_type <- ifelse(
    parcelles_filtered$Bloc2022 %in% c("P", "S", "V", "W"),
    "GAG",
    "GRV"
  )
  
  
  # Show frequency table
  if (any(!is.na(parcelles_filtered$Bloc2022))) {
    irrig_summary <- table(parcelles_filtered$Bloc2022, parcelles_filtered$irrig_type, useNA = "ifany")
    cat("\nCross-tabulation of Bloc2022 vs irrig_type:\n")
    print(irrig_summary)
    
    # Overall irrigation type distribution
    irrig_freq <- table(parcelles_filtered$irrig_type, useNA = "ifany")
    cat("\nOverall irrigation type distribution:\n")
    print(irrig_freq)
  } else {
    cat("Warning: All Bloc2022 values are NA\n")
    parcelles_filtered$irrig_type <- "GRV"  # Default for all NA
  }
  
} else {
  cat("Warning: Bloc2022 column not found. Adding default irrigation type.\n")
  parcelles_filtered$irrig_type <- "GRV"  # Default when Bloc2022 doesn't exist
}

cat("Successfully added irrig_type variable!\n")

# Display the new attribute table structure
cat("\n=== NEW FILTERED ATTRIBUTE TABLE STRUCTURE ===\n")
attr_data_filtered <- st_drop_geometry(parcelles_filtered)
col_info_filtered <- sapply(attr_data_filtered, class)
for (i in 1:length(col_info_filtered)) {
  cat(sprintf("%-15s: %s\n", names(col_info_filtered)[i], paste(col_info_filtered[[i]], collapse = ", ")))
}

# Display first few rows of the filtered data
cat("\n=== FIRST 6 ROWS OF FILTERED ATTRIBUTE DATA ===\n")
print(head(attr_data_filtered, 6))

# Summary of the filtered columns (including new irrig_type)
cat("\n=== SUMMARY OF FILTERED COLUMNS ===\n")
for (col_name in names(attr_data_filtered)) {
  cat("\n--- Column:", col_name, "---\n")
  col_data <- attr_data_filtered[[col_name]]
  
  if (is.numeric(col_data)) {
    cat("Type: Numeric\n")
    cat("Min:", round(min(col_data, na.rm = TRUE), 3), "\n")
    cat("Max:", round(max(col_data, na.rm = TRUE), 3), "\n")
    cat("Mean:", round(mean(col_data, na.rm = TRUE), 3), "\n")
    cat("Median:", round(median(col_data, na.rm = TRUE), 3), "\n")
    cat("NA count:", sum(is.na(col_data)), "\n")
  } else if (is.character(col_data) || is.factor(col_data)) {
    cat("Type: Character/Factor\n")
    unique_vals <- unique(col_data)
    cat("Unique values:", length(unique_vals), "\n")
    cat("NA count:", sum(is.na(col_data)), "\n")
    if (length(unique_vals) <= 10) {
      cat("Values:", paste(unique_vals, collapse = ", "), "\n")
    } else {
      cat("Sample values:", paste(head(unique_vals, 5), collapse = ", "), "...\n")
    }
    # Show frequency table for categorical data
    if (length(unique_vals) <= 20) {
      cat("Frequency table:\n")
      freq_table <- table(col_data, useNA = "ifany")
      print(freq_table)
    }
  }
}

# Create plot with transformed data (UTM Zone 28N) using filtered data
cat("\n=== CREATING PLOTS WITH UTM COORDINATES ===\n")

# Calculate area in hectares (since UTM is in meters) 
parcelles_filtered$area_ha <- as.numeric(st_area(parcelles_filtered)) / 10000

# Basic plot with UTM coordinates using filtered data
p1 <- ggplot(parcelles_filtered) +
  geom_sf(fill = "lightblue", color = "darkblue", alpha = 0.7, linewidth = 0.2) +
  labs(
    title = "PARCELLES22324 - UTM Zone 28N Projection (Filtered Data)",
    subtitle = paste("Filtered attributes with irrigation types | Total parcels:", nrow(parcelles_filtered)),
    caption = paste("CRS: UTM Zone 28N | Units: meters | Total area:", 
                   round(sum(parcelles_filtered$area_ha, na.rm = TRUE), 1), "hectares")
  ) +
  xlab("Easting (m)") +
  ylab("Northing (m)") +
  custom_theme



# Optional: Create a second plot showing centroids with validation
p2 <- ggplot(parcelles_filtered) +
  geom_sf(fill = "lightblue", color = "darkblue", alpha = 0.5, linewidth = 0.2) +
  geom_point(aes(x = X.Centroid, y = Y.Centroid), color = "red", size = 0.8, alpha = 0.9) +
  labs(
    title = "PARCELLES22324 - Parcels with Validated Centroids",
    subtitle = paste("Red points show parcel centroids (point-on-surface method) | Total parcels:", nrow(parcelles_filtered)),
    caption = "Blue polygons = parcels, Red points = centroids (guaranteed inside polygons)"
  ) +
  xlab("Easting (m)") +
  ylab("Northing (m)") +
  custom_theme

# Create a diagnostic plot to check centroid placement
p3 <- ggplot(parcelles_filtered) +
  geom_sf(aes(fill = !is.na(X.Centroid)), color = "white", linewidth = 0.1) +
  scale_fill_manual(values = c("TRUE" = "lightgreen", "FALSE" = "red"), 
                   name = "Has Centroid", 
                   labels = c("FALSE" = "Missing", "TRUE" = "Present")) +
  geom_point(aes(x = X.Centroid, y = Y.Centroid), color = "darkred", size = 0.5, alpha = 0.8) +
  labs(
    title = "Centroid Validation Check",
    subtitle = "Green = parcels with centroids, Red = parcels without centroids",
    caption = paste("Total parcels:", nrow(parcelles_filtered), 
                   "| With centroids:", sum(!is.na(parcelles_filtered$X.Centroid)))
  ) +
  xlab("Easting (m)") +
  ylab("Northing (m)") +
  custom_theme

# Save the plots
ggsave(
  filename = file.path(output_folder, "parcelles_filtered_utm28n.png"),
  plot = p1,
  width = 12, height = 9, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(output_folder, "parcelles_with_centroids.png"),
  plot = p2,
  width = 12, height = 9, dpi = 300, bg = "white"
)

ggsave(
  filename = file.path(output_folder, "centroid_validation_check.png"),
  plot = p3,
  width = 12, height = 9, dpi = 300, bg = "white"
)


# Display all plots
print(p1)
print(p2)
print(p3)


# Additional validation: Check for centroids outside parcels (if any)
cat("\n=== CENTROID VALIDATION ===\n")

# Create point geometries from centroid coordinates
centroid_points <- st_as_sf(
  data.frame(
    id = if("id" %in% existing_cols) parcelles_filtered[[existing_cols[existing_cols == "id"]]] else seq_len(nrow(parcelles_filtered)),
    X = parcelles_filtered$X.Centroid,
    Y = parcelles_filtered$Y.Centroid
  ),
  coords = c("X", "Y"),
  crs = st_crs(parcelles_filtered)
)

# Check which centroids fall within their corresponding parcels
if (nrow(centroid_points) == nrow(parcelles_filtered)) {
  within_check <- st_within(centroid_points, parcelles_filtered, sparse = FALSE)
  within_own_parcel <- sapply(1:nrow(parcelles_filtered), function(i) within_check[i, i])
  
  centroids_outside <- sum(!within_own_parcel, na.rm = TRUE)
  cat("Centroids outside their own parcel:", centroids_outside, "\n")
  
  if (centroids_outside > 0) {
    cat("These parcels have centroids outside (using point-on-surface should fix this):\n")
    outside_indices <- which(!within_own_parcel)
    cat("Parcel indices:", paste(head(outside_indices, 10), collapse = ", "), 
        if(length(outside_indices) > 10) "..." else "", "\n")
  }
} else {
  cat("Cannot perform within-parcel validation due to geometry mismatch\n")
}

# Save the filtered dataset as a new shapefile
cat("\n=== SAVING FILTERED DATASET ===\n")
output_shp_path <- file.path(output_folder, "parcelles_filtered.shp")
st_write(parcelles_filtered, output_shp_path, delete_dsn = TRUE)
cat("Filtered shapefile saved to:", output_shp_path, "\n")

# Export attribute table as CSV for easy viewing
csv_path <- file.path(output_folder, "parcelles_attributes.csv")
write.csv(attr_data_filtered, csv_path, row.names = FALSE)
cat("Attribute table saved to:", csv_path, "\n")
```

# II. SPATIAL NEIGHBORHOOD ANALYSIS

```{r}
# ===== SPATIAL NEIGHBORHOOD ANALYSIS =====
# Load required libraries
library(sf)
library(ggplot2)
library(dplyr)

# Install and load spatial analysis packages if needed
if (!requireNamespace("spdep", quietly = TRUE)) {
  cat("Installing spdep package...\n")
  install.packages("spdep")
}
library(spdep)

if (!requireNamespace("units", quietly = TRUE)) {
  install.packages("units")
}
library(units)

cat("========================================\n")
cat("      SPATIAL NEIGHBORHOOD ANALYSIS\n") 
cat("========================================\n")

# Assuming your filtered shapefile is called 'parcelles_filtered'
# If you have a different name, change it here:
shapefile <- parcelles_filtered

cat("\n=== STEP 1: PREPARING SHAPEFILE FOR SPATIAL ANALYSIS ===\n")

# Ensure geometries are valid
shapefile <- st_make_valid(shapefile)
cat("Geometries validated\n")

# Add ID.Field (sequential numeric identifier for INLA)
shapefile$ID.Field <- 1:nrow(shapefile)
cat("Added ID.Field: 1 to", nrow(shapefile), "\n")

# Check if 'id' column exists, rename to 'field' if it does
if ("id" %in% names(shapefile)) {
  shapefile <- shapefile %>% rename(field = id)
  cat("Renamed 'id' column to 'field'\n")
} else {
  # If no 'id' column, create 'field' from ID.Field
  shapefile$field <- shapefile$ID.Field
  cat("Created 'field' column from ID.Field\n")
}

# Display updated structure
cat("\nUpdated shapefile structure:\n")
attr_data_updated <- st_drop_geometry(shapefile)
print(names(attr_data_updated))
cat("Total parcels:", nrow(shapefile), "\n")

cat("\n=== STEP 2: CALCULATING BOUNDARY-TO-BOUNDARY DISTANCES ===\n")

# Calculate boundary-to-boundary distances between all parcels
cat("Computing pairwise distances between parcel boundaries...\n")
cat("This may take a while for large datasets...\n")

# Get parcel boundaries (exterior rings only)
boundaries <- st_boundary(shapefile)

# Calculate distance matrix (boundary to boundary)
dist_matrix <- st_distance(boundaries, boundaries)

# Convert to numeric matrix and remove units
dist_matrix <- units::drop_units(dist_matrix)

# Set diagonal to NA (parcel cannot be neighbor to itself)
diag(dist_matrix) <- NA

cat("Distance matrix calculated: ", nrow(dist_matrix), "x", ncol(dist_matrix), "\n")
cat("Min boundary distance:", round(min(dist_matrix, na.rm = TRUE), 2), "meters\n")
cat("Max boundary distance:", round(max(dist_matrix, na.rm = TRUE), 2), "meters\n")
cat("Mean boundary distance:", round(mean(dist_matrix, na.rm = TRUE), 2), "meters\n")

cat("\n=== STEP 3: DISTANCE DISTRIBUTION ANALYSIS ===\n")

# Extract all distances for analysis (excluding diagonal)
all_distances <- as.vector(dist_matrix)
all_distances <- all_distances[!is.na(all_distances)]

# Calculate statistics
dist_stats <- list(
  min = min(all_distances),
  q25 = quantile(all_distances, 0.25),
  median = median(all_distances),
  mean = mean(all_distances),
  q75 = quantile(all_distances, 0.75),
  q90 = quantile(all_distances, 0.90),
  q95 = quantile(all_distances, 0.95),
  max = max(all_distances)
)

cat("Distance Statistics (meters):\n")
cat("Min:", round(dist_stats$min, 2), "\n")
cat("25th percentile:", round(dist_stats$q25, 2), "\n")
cat("Median:", round(dist_stats$median, 2), "\n")
cat("Mean:", round(dist_stats$mean, 2), "\n")
cat("75th percentile:", round(dist_stats$q75, 2), "\n")
cat("90th percentile:", round(dist_stats$q90, 2), "\n")
cat("95th percentile:", round(dist_stats$q95, 2), "\n")
cat("Max:", round(dist_stats$max, 2), "\n")

cat("\n=== STEP 4: DETERMINING OPTIMAL THRESHOLD ===\n")

# Suggest multiple threshold options
threshold_options <- c(
  dist_stats$min + 1,                    # Just above minimum (touching parcels)
  dist_stats$q25,                        # 25th percentile
  dist_stats$median * 0.1,               # 10% of median
  dist_stats$median * 0.2,               # 20% of median
  50,                                    # Fixed 50m threshold
  100,                                   # Fixed 100m threshold
  200                                    # Fixed 200m threshold
)

cat("Suggested threshold options:\n")
for (i in 1:length(threshold_options)) {
  thresh <- threshold_options[i]
  n_neighbors <- sum(dist_matrix <= thresh & !is.na(dist_matrix))
  avg_neighbors <- n_neighbors / nrow(shapefile)
  cat(sprintf("Option %d: %.1f meters (avg %.1f neighbors per parcel)\n", 
              i, thresh, avg_neighbors))
}

# Auto-select threshold: aim for 3-8 neighbors per parcel on average
target_neighbors <- 5
optimal_threshold <- quantile(all_distances, 0.1)  # Start with 10th percentile

# Iteratively find good threshold
for (test_thresh in seq(10, 500, by = 10)) {
  avg_neighbors <- sum(dist_matrix <= test_thresh & !is.na(dist_matrix)) / nrow(shapefile)
  if (avg_neighbors >= target_neighbors) {
    optimal_threshold <- test_thresh
    break
  }
}

cat("\nRecommended threshold:", round(optimal_threshold, 1), "meters\n")
cat("(Aiming for ~", target_neighbors, "neighbors per parcel on average)\n")

# Allow user to choose or use default
DISTANCE_THRESHOLD <- optimal_threshold
cat("Using threshold:", round(DISTANCE_THRESHOLD, 1), "meters\n")

cat("\n=== STEP 5: CREATING ADJACENCY MATRIX ===\n")

# Create adjacency matrix based on distance threshold
W.P1 <- matrix(0, nrow = nrow(shapefile), ncol = nrow(shapefile))

# Fill adjacency matrix: 1 if distance <= threshold, 0 otherwise
W.P1[dist_matrix <= DISTANCE_THRESHOLD & !is.na(dist_matrix)] <- 1

# Ensure diagonal is 0 (parcel not neighbor to itself)
diag(W.P1) <- 0

# Add row and column names using field values
if ("field" %in% names(shapefile)) {
  dimnames(W.P1)[[1]] <- dimnames(W.P1)[[2]] <- shapefile$field
} else {
  dimnames(W.P1)[[1]] <- dimnames(W.P1)[[2]] <- shapefile$ID.Field
}

# Calculate adjacency statistics
n_connections <- sum(W.P1)
avg_neighbors <- n_connections / nrow(shapefile)
max_neighbors <- max(rowSums(W.P1))
min_neighbors <- min(rowSums(W.P1))

cat("Adjacency Matrix created:", nrow(W.P1), "x", ncol(W.P1), "\n")
cat("Total connections:", n_connections, "\n")
cat("Average neighbors per parcel:", round(avg_neighbors, 2), "\n")
cat("Min neighbors:", min_neighbors, "\n")
cat("Max neighbors:", max_neighbors, "\n")
cat("Parcels with no neighbors:", sum(rowSums(W.P1) == 0), "\n")

cat("\n=== STEP 6: CREATING REFERENCE TABLE FOR INLA ===\n")

# Create reference table (map_sp equivalent)
map_sp <- data.frame(
  field = if("field" %in% names(shapefile)) shapefile$field else shapefile$ID.Field,
  ID.Field = shapefile$ID.Field
)

# Ensure ID.Field is numeric
map_sp$ID.Field <- as.numeric(map_sp$ID.Field)

cat("Reference table created:\n")
print(head(map_sp, 10))
cat("Dimensions:", nrow(map_sp), "rows x", ncol(map_sp), "columns\n")

cat("\n=== STEP 7: CREATING VISUALIZATIONS ===\n")

# Set output folder (same as before or create new one)
output_folder <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy"

# Plot 1: Distance distribution histogram
p_dist_hist <- ggplot(data.frame(distance = all_distances), aes(x = distance)) +
  geom_histogram(bins = 50, fill = "skyblue", alpha = 0.7, color = "black") +
  geom_vline(xintercept = DISTANCE_THRESHOLD, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Distribution of Boundary-to-Boundary Distances",
    subtitle = paste("Red line = Selected threshold (", round(DISTANCE_THRESHOLD, 1), "m)"),
    x = "Distance (meters)",
    y = "Frequency",
    caption = paste("Total parcel pairs:", length(all_distances))
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Plot 2: Distance distribution (zoomed on lower values)
p_dist_zoom <- ggplot(data.frame(distance = all_distances[all_distances <= quantile(all_distances, 0.95)]), 
                      aes(x = distance)) +
  geom_histogram(bins = 50, fill = "lightgreen", alpha = 0.7, color = "black") +
  geom_vline(xintercept = DISTANCE_THRESHOLD, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Distance Distribution (Zoomed - 95th Percentile)",
    subtitle = paste("Red line = Selected threshold (", round(DISTANCE_THRESHOLD, 1), "m)"),
    x = "Distance (meters)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Plot 3: Number of neighbors per parcel
neighbor_counts <- rowSums(W.P1)
p_neighbors <- ggplot(data.frame(neighbors = neighbor_counts), aes(x = neighbors)) +
  geom_bar(fill = "orange", alpha = 0.7, color = "black") +
  labs(
    title = "Distribution of Neighbor Counts",
    subtitle = paste("Threshold:", round(DISTANCE_THRESHOLD, 1), "m | Avg neighbors:", round(avg_neighbors, 2)),
    x = "Number of Neighbors",
    y = "Number of Parcels"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Plot 4: Spatial plot showing neighbor connections
# Sample some connections to avoid overcrowding
set.seed(123)
sample_parcels <- sample(1:nrow(shapefile), min(50, nrow(shapefile)))

# Create connection lines
connection_lines <- data.frame()
for (i in sample_parcels) {
  neighbors <- which(W.P1[i, ] == 1)
  if (length(neighbors) > 0) {
    for (j in neighbors) {
      connection_lines <- rbind(connection_lines, data.frame(
        x1 = shapefile$X.Centroid[i],
        y1 = shapefile$Y.Centroid[i],
        x2 = shapefile$X.Centroid[j],
        y2 = shapefile$Y.Centroid[j]
      ))
    }
  }
}

p_spatial <- ggplot(shapefile) +
  geom_sf(aes(fill = rowSums(W.P1)), color = "white", size = 0.1) +
  scale_fill_viridis_c(name = "Neighbors", option = "plasma") +
  {if(nrow(connection_lines) > 0) geom_segment(data = connection_lines, 
                                               aes(x = x1, y = y1, xend = x2, yend = y2), 
                                               color = "red", alpha = 0.3, size = 0.2)} +
  labs(
    title = "Spatial Neighborhood Network",
    subtitle = paste("Color = Number of neighbors | Red lines = Sample connections (", 
                    nrow(connection_lines), "shown)"),
    caption = paste("Threshold:", round(DISTANCE_THRESHOLD, 1), "m")
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )

# Save plots
ggsave(file.path(output_folder, "distance_distribution.png"), p_dist_hist, 
       width = 10, height = 7, dpi = 300)
ggsave(file.path(output_folder, "distance_distribution_zoomed.png"), p_dist_zoom, 
       width = 10, height = 7, dpi = 300)
ggsave(file.path(output_folder, "neighbor_counts.png"), p_neighbors, 
       width = 10, height = 7, dpi = 300)
ggsave(file.path(output_folder, "spatial_network.png"), p_spatial, 
       width = 12, height = 10, dpi = 300)

# Display plots
print(p_dist_hist)
print(p_dist_zoom)
print(p_neighbors)
print(p_spatial)

cat("\n=== STEP 8: UPDATING MAIN SHAPEFILE ===\n")

# Add ID.Field to the main parcelles_filtered object
parcelles_filtered$ID.Field <- shapefile$ID.Field

# Rename id to field if it exists
if ("id" %in% names(parcelles_filtered) && !"field" %in% names(parcelles_filtered)) {
  parcelles_filtered <- parcelles_filtered %>% rename(field = id)
}

cat("Updated main shapefile with ID.Field\n")
cat("Final attribute structure:\n")
print(names(st_drop_geometry(parcelles_filtered)))
```

## II.1. Spatial Neighborhood Analysis (Fixed 110m Threshold)

This section defines spatial neighborhoods by calculating boundary-to-boundary distances between all parcels and applying a 110-meter threshold to create the adjacency matrix W.P1 required for INLA. The code generates a reference table mapping parcel IDs to sequential numbers and produces visualizations showing distance distributions and neighbor connections.

```{r}
# ===== SPATIAL NEIGHBORHOOD ANALYSIS =====
# Load required libraries
library(sf)
library(ggplot2)
library(dplyr)

# Install and load spatial analysis packages if needed
if (!requireNamespace("spdep", quietly = TRUE)) {
  cat("Installing spdep package...\n")
  install.packages("spdep")
}
library(spdep)

if (!requireNamespace("units", quietly = TRUE)) {
  install.packages("units")
}
library(units)

cat("========================================\n")
cat("      SPATIAL NEIGHBORHOOD ANALYSIS\n") 
cat("========================================\n")

# Assuming your filtered shapefile is called 'parcelles_filtered'
# If you have a different name, change it here:
shapefile <- parcelles_filtered

cat("\n=== STEP 1: PREPARING SHAPEFILE FOR SPATIAL ANALYSIS ===\n")

# Ensure geometries are valid
shapefile <- st_make_valid(shapefile)
cat("Geometries validated\n")

# Add ID.Field (sequential numeric identifier for INLA)
shapefile$ID.Field <- 1:nrow(shapefile)
cat("Added ID.Field: 1 to", nrow(shapefile), "\n")

# Check if 'id' column exists, rename to 'field' if it does
if ("id" %in% names(shapefile)) {
  shapefile <- shapefile %>% rename(field = id)
  cat("Renamed 'id' column to 'field'\n")
} else {
  # If no 'id' column, create 'field' from ID.Field
  shapefile$field <- shapefile$ID.Field
  cat("Created 'field' column from ID.Field\n")
}

# Display updated structure
cat("\nUpdated shapefile structure:\n")
attr_data_updated <- st_drop_geometry(shapefile)
print(names(attr_data_updated))
cat("Total parcels:", nrow(shapefile), "\n")

cat("\n=== STEP 2: CALCULATING BOUNDARY-TO-BOUNDARY DISTANCES ===\n")

# Calculate boundary-to-boundary distances between all parcels
cat("Computing pairwise distances between parcel boundaries...\n")
cat("This may take a while for large datasets...\n")

# Get parcel boundaries (exterior rings only)
boundaries <- st_boundary(shapefile)

# Calculate distance matrix (boundary to boundary)
dist_matrix <- st_distance(boundaries, boundaries)

# Convert to numeric matrix and remove units
dist_matrix <- units::drop_units(dist_matrix)

# Set diagonal to NA (parcel cannot be neighbor to itself)
diag(dist_matrix) <- NA

cat("Distance matrix calculated: ", nrow(dist_matrix), "x", ncol(dist_matrix), "\n")
cat("Min boundary distance:", round(min(dist_matrix, na.rm = TRUE), 2), "meters\n")
cat("Max boundary distance:", round(max(dist_matrix, na.rm = TRUE), 2), "meters\n")
cat("Mean boundary distance:", round(mean(dist_matrix, na.rm = TRUE), 2), "meters\n")

cat("\n=== STEP 3: DISTANCE DISTRIBUTION ANALYSIS ===\n")

# Extract all distances for analysis (excluding diagonal)
all_distances <- as.vector(dist_matrix)
all_distances <- all_distances[!is.na(all_distances)]

# Calculate statistics
dist_stats <- list(
  min = min(all_distances),
  q25 = quantile(all_distances, 0.25),
  median = median(all_distances),
  mean = mean(all_distances),
  q75 = quantile(all_distances, 0.75),
  q90 = quantile(all_distances, 0.90),
  q95 = quantile(all_distances, 0.95),
  max = max(all_distances)
)

cat("Distance Statistics (meters):\n")
cat("Min:", round(dist_stats$min, 2), "\n")
cat("25th percentile:", round(dist_stats$q25, 2), "\n")
cat("Median:", round(dist_stats$median, 2), "\n")
cat("Mean:", round(dist_stats$mean, 2), "\n")
cat("75th percentile:", round(dist_stats$q75, 2), "\n")
cat("90th percentile:", round(dist_stats$q90, 2), "\n")
cat("95th percentile:", round(dist_stats$q95, 2), "\n")
cat("Max:", round(dist_stats$max, 2), "\n")

cat("\n=== STEP 4: DETERMINING OPTIMAL THRESHOLD ===\n")

# Suggest multiple threshold options
threshold_options <- c(
  dist_stats$min + 1,                    # Just above minimum (touching parcels)
  dist_stats$q25,                        # 25th percentile
  dist_stats$median * 0.1,               # 10% of median
  dist_stats$median * 0.2,               # 20% of median
  50,                                    # Fixed 50m threshold
  100,                                   # Fixed 100m threshold
  200                                    # Fixed 200m threshold
)

cat("Suggested threshold options:\n")
for (i in 1:length(threshold_options)) {
  thresh <- threshold_options[i]
  n_neighbors <- sum(dist_matrix <= thresh & !is.na(dist_matrix))
  avg_neighbors <- n_neighbors / nrow(shapefile)
  cat(sprintf("Option %d: %.1f meters (avg %.1f neighbors per parcel)\n", 
              i, thresh, avg_neighbors))
}

# Allow user to choose or use default
DISTANCE_THRESHOLD <- 110  # Fixed threshold of 110 meters
cat("Using fixed threshold:", DISTANCE_THRESHOLD, "meters\n")

# Show what this threshold will give us
n_neighbors_110 <- sum(dist_matrix <= DISTANCE_THRESHOLD & !is.na(dist_matrix))
avg_neighbors_110 <- n_neighbors_110 / nrow(shapefile)
cat("With 110m threshold:\n")
cat("- Total connections:", n_neighbors_110, "\n")
cat("- Average neighbors per parcel:", round(avg_neighbors_110, 2), "\n")

cat("\n=== STEP 5: CREATING ADJACENCY MATRIX ===\n")

# Create adjacency matrix based on distance threshold
W.P1 <- matrix(0, nrow = nrow(shapefile), ncol = nrow(shapefile))

# Fill adjacency matrix: 1 if distance <= threshold, 0 otherwise
W.P1[dist_matrix <= DISTANCE_THRESHOLD & !is.na(dist_matrix)] <- 1

# Ensure diagonal is 0 (parcel not neighbor to itself)
diag(W.P1) <- 0

# Add row and column names using field values
if ("field" %in% names(shapefile)) {
  dimnames(W.P1)[[1]] <- dimnames(W.P1)[[2]] <- shapefile$field
} else {
  dimnames(W.P1)[[1]] <- dimnames(W.P1)[[2]] <- shapefile$ID.Field
}

# Calculate adjacency statistics
n_connections <- sum(W.P1)
avg_neighbors <- n_connections / nrow(shapefile)
max_neighbors <- max(rowSums(W.P1))
min_neighbors <- min(rowSums(W.P1))

cat("Adjacency Matrix created:", nrow(W.P1), "x", ncol(W.P1), "\n")
cat("Total connections:", n_connections, "\n")
cat("Average neighbors per parcel:", round(avg_neighbors, 2), "\n")
cat("Min neighbors:", min_neighbors, "\n")
cat("Max neighbors:", max_neighbors, "\n")
cat("Parcels with no neighbors:", sum(rowSums(W.P1) == 0), "\n")

cat("\n=== STEP 6: CREATING REFERENCE TABLE FOR INLA ===\n")

# Create reference table (map_sp equivalent)
map_sp <- data.frame(
  field = if("field" %in% names(shapefile)) shapefile$field else shapefile$ID.Field,
  ID.Field = shapefile$ID.Field
)

# Ensure ID.Field is numeric
map_sp$ID.Field <- as.numeric(map_sp$ID.Field)

cat("Reference table created:\n")
print(head(map_sp, 10))
cat("Dimensions:", nrow(map_sp), "rows x", ncol(map_sp), "columns\n")

cat("\n=== STEP 7: CREATING VISUALIZATIONS ===\n")

# Set output folder (same as before or create new one)
output_folder <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy"

# Plot 1: Distance distribution histogram
p_dist_hist <- ggplot(data.frame(distance = all_distances), aes(x = distance)) +
  geom_histogram(bins = 50, fill = "skyblue", alpha = 0.7, color = "black") +
  geom_vline(xintercept = 110, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Distribution of Boundary-to-Boundary Distances",
    subtitle = "Red line = Selected threshold (110m)",
    x = "Distance (meters)",
    y = "Frequency",
    caption = paste("Total parcel pairs:", length(all_distances))
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Plot 2: Distance distribution (zoomed on lower values)
p_dist_zoom <- ggplot(data.frame(distance = all_distances[all_distances <= quantile(all_distances, 0.95)]), 
                      aes(x = distance)) +
  geom_histogram(bins = 50, fill = "lightgreen", alpha = 0.7, color = "black") +
  geom_vline(xintercept = 110, color = "red", linetype = "dashed", size = 1) +
  labs(
    title = "Distance Distribution (Zoomed - 95th Percentile)",
    subtitle = "Red line = Selected threshold (110m)",
    x = "Distance (meters)",
    y = "Frequency"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Plot 3: Number of neighbors per parcel
neighbor_counts <- rowSums(W.P1)
p_neighbors <- ggplot(data.frame(neighbors = neighbor_counts), aes(x = neighbors)) +
  geom_bar(fill = "orange", alpha = 0.7, color = "black") +
  labs(
    title = "Distribution of Neighbor Counts",
    subtitle = paste("Threshold: 110m | Avg neighbors:", round(avg_neighbors, 2)),
    x = "Number of Neighbors",
    y = "Number of Parcels"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    axis.title = element_text(size = 12)
  )

# Plot 4: Spatial plot showing neighbor connections
# Sample some connections to avoid overcrowding
set.seed(123)
sample_parcels <- sample(1:nrow(shapefile), min(50, nrow(shapefile)))

# Create connection lines
connection_lines <- data.frame()
for (i in sample_parcels) {
  neighbors <- which(W.P1[i, ] == 1)
  if (length(neighbors) > 0) {
    for (j in neighbors) {
      connection_lines <- rbind(connection_lines, data.frame(
        x1 = shapefile$X.Centroid[i],
        y1 = shapefile$Y.Centroid[i],
        x2 = shapefile$X.Centroid[j],
        y2 = shapefile$Y.Centroid[j]
      ))
    }
  }
}

p_spatial <- ggplot(shapefile) +
  geom_sf(aes(fill = rowSums(W.P1)), color = "white", size = 0.1) +
  scale_fill_viridis_c(name = "Neighbors", option = "plasma") +
  {if(nrow(connection_lines) > 0) geom_segment(data = connection_lines, 
                                               aes(x = x1, y = y1, xend = x2, yend = y2), 
                                               color = "red", alpha = 0.3, size = 0.2)} +
  labs(
    title = "Spatial Neighborhood Network",
    subtitle = paste("Color = Number of neighbors | Red lines = Sample connections (", 
                    nrow(connection_lines), "shown)"),
    caption = "Threshold: 110m"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.subtitle = element_text(size = 12),
    legend.position = "bottom"
  )

# Save plots
ggsave(file.path(output_folder, "distance_distribution.png"), p_dist_hist, 
       width = 10, height = 7, dpi = 300)
ggsave(file.path(output_folder, "distance_distribution_zoomed.png"), p_dist_zoom, 
       width = 10, height = 7, dpi = 300)
ggsave(file.path(output_folder, "neighbor_counts.png"), p_neighbors, 
       width = 10, height = 7, dpi = 300)
ggsave(file.path(output_folder, "spatial_network.png"), p_spatial, 
       width = 12, height = 10, dpi = 300)

# Display plots
print(p_dist_hist)
print(p_dist_zoom)
print(p_neighbors)
print(p_spatial)

cat("\n=== STEP 8: UPDATING MAIN SHAPEFILE ===\n")

# Add ID.Field to the main parcelles_filtered object
parcelles_filtered$ID.Field <- shapefile$ID.Field

# Rename id to field if it exists
if ("id" %in% names(parcelles_filtered) && !"field" %in% names(parcelles_filtered)) {
  parcelles_filtered <- parcelles_filtered %>% rename(field = id)
}

cat("Updated main shapefile with ID.Field\n")
cat("Final attribute structure:\n")
print(names(st_drop_geometry(parcelles_filtered)))
```

# III. Soil Data Importation

This section loads the soil database, standardizes parcel names through automated cleaning and manual correspondence rules, and matches soil samples to shapefile parcels. The code generates detailed reports of matched and unmatched parcels, analyzes temporal coverage across years, and documents the matching process with comprehensive statistics.

## III.1. PARCEL CORRESPONDENCE ANALYSIS

```{r}
# ============================================================================
# PARCEL CORRESPONDENCE ANALYSIS WITH DETAILED UNMATCHED REPORTS
# ============================================================================
# This script analyzes correspondence between soil data and shapefile parcels
# and creates detailed CSV reports of unmatched parcels for further investigation

# Load required libraries
library(sf)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(viridis)

# Set working directory (adjust path as needed)
setwd("C:/Users/U078-N800/Desktop/Ml_Internship/Copy")

# ============================================================================
# LOAD AND PROCESS SOIL DATA
# ============================================================================

cat("=== LOADING SOIL DATA ===\n")

# Load the soil data
soil_data <- read.csv("BD_soil.csv", sep = ";", dec = ",", header = TRUE, 
                      stringsAsFactors = FALSE, fileEncoding = "latin1")

# Display basic soil data information
cat("Soil dataset dimensions:", dim(soil_data), "\n")
cat("Number of rows:", nrow(soil_data), "\n")
cat("Number of columns:", ncol(soil_data), "\n")

# Show column names
cat("\nSoil data column names:\n")
print(colnames(soil_data))

# Process Parc.Detaillee column - convert to uppercase
cat("\n=== PROCESSING SOIL DATA PARC.DETAILLEE ===\n")

# Convert to uppercase
soil_data$Parc.Detaillee_UPPER <- toupper(soil_data$Parc.Detaillee)

# Get unique values and create detailed soil parcel info
unique_soil_parcels <- sort(unique(soil_data$Parc.Detaillee_UPPER))
n_unique_soil <- length(unique_soil_parcels)

# Create detailed soil parcel information
soil_parcel_details <- soil_data %>%
  group_by(Parc.Detaillee_UPPER) %>%
  summarise(
    Original_Names = paste(unique(Parc.Detaillee), collapse = "; "),
    Sample_Count = n(),
    Years_Sampled = paste(sort(unique(Annee[!is.na(Annee)])), collapse = ", "),
    First_Sample_Year = min(Annee, na.rm = TRUE),
    Last_Sample_Year = max(Annee, na.rm = TRUE),
    .groups = 'drop'
  ) %>%
  arrange(Parc.Detaillee_UPPER)

cat("\n📌 SOIL DATA ANALYSIS:\n")
cat("Number of unique parcels in soil data:", n_unique_soil, "\n")
cat("Total soil samples:", nrow(soil_data), "\n")

# ============================================================================
# USE EXISTING PROCESSED SHAPEFILE DATA
# ============================================================================

cat("\n=== USING EXISTING PROCESSED SHAPEFILE DATA ===\n")

# Use the parcelles_filtered object from previous analysis
# Assuming parcelles_filtered is already loaded in the environment
parcelles <- parcelles_filtered

cat("Shapefile dimensions:", dim(parcelles), "\n")
cat("Number of features:", nrow(parcelles), "\n")

# Show updated shapefile column names
cat("\nShapefile column names:\n")
print(names(st_drop_geometry(parcelles)))

# Process field column - convert to uppercase
parcelles$field_UPPER <- toupper(parcelles$field)

# Get unique values and create detailed shapefile parcel info
unique_shp_parcels <- sort(unique(parcelles$field_UPPER))
n_unique_shp <- length(unique_shp_parcels)

# Create detailed shapefile parcel information
shp_parcel_details <- parcelles %>%
  st_drop_geometry() %>%
  group_by(field_UPPER) %>%
  summarise(
    Original_Names = paste(unique(field), collapse = "; "),
    Feature_Count = n(),
    .groups = 'drop'
  ) %>%
  arrange(field_UPPER)

cat("\n📌 SHAPEFILE ANALYSIS:\n")
cat("Number of unique parcel IDs in shapefile:", n_unique_shp, "\n")
cat("Total shapefile features:", nrow(parcelles), "\n")

# ============================================================================
# ADVANCED NAME CLEANING AND MATCHING
# ============================================================================

cat("\n=== ADVANCED NAME CLEANING ===\n")

# Define enhanced cleaning function
clean_parcel_name <- function(name) {
  if (is.na(name) || name == "") return("")
  
  # Convert to uppercase and remove extra spaces
  cleaned <- toupper(trimws(name))
  
  # Remove special characters and standardize
  cleaned <- gsub("[^A-Z0-9]", "", cleaned)
  
  # Handle common variations
  cleaned <- gsub("PARCELLE", "PARC", cleaned)
  cleaned <- gsub("PARCEL", "PARC", cleaned)
  cleaned <- gsub("BLOC", "BL", cleaned)
  cleaned <- gsub("SECTEUR", "SECT", cleaned)
  
  return(cleaned)
}

# Apply cleaning to soil data
cat("Applying advanced cleaning to soil parcel names...\n")
soil_data$Parc_CLEANED <- sapply(soil_data$Parc.Detaillee_UPPER, clean_parcel_name)

# Update soil parcel details with cleaned names
soil_parcel_details$Parc_CLEANED <- sapply(soil_parcel_details$Parc.Detaillee_UPPER, clean_parcel_name)

# Get unique cleaned soil parcels
unique_soil_cleaned <- sort(unique(soil_parcel_details$Parc_CLEANED))
n_unique_soil_cleaned <- length(unique_soil_cleaned)

# Apply cleaning to shapefile data  
cat("Applying advanced cleaning to shapefile parcel names...\n")
parcelles$field_CLEANED <- sapply(parcelles$field_UPPER, clean_parcel_name)

# Update shapefile parcel details with cleaned names
shp_parcel_details$field_CLEANED <- sapply(shp_parcel_details$field_UPPER, clean_parcel_name)

# Get unique cleaned shapefile parcels
unique_shp_cleaned <- sort(unique(shp_parcel_details$field_CLEANED))
n_unique_shp_cleaned <- length(unique_shp_cleaned)

# ============================================================================
# DETAILED CORRESPONDENCE ANALYSIS
# ============================================================================

cat("\n=== DETAILED CORRESPONDENCE ANALYSIS ===\n")

# Find exact matches (before cleaning)
exact_matches <- intersect(unique_soil_parcels, unique_shp_parcels)
n_exact_matches <- length(exact_matches)

# Find matches after cleaning
cleaned_matches <- intersect(unique_soil_cleaned, unique_shp_cleaned)
n_cleaned_matches <- length(cleaned_matches)

# Find additional matches gained from cleaning
additional_matches <- setdiff(cleaned_matches, exact_matches)
n_additional_matches <- length(additional_matches)

# Find unmatched parcels after cleaning
unmatched_soil_cleaned <- setdiff(unique_soil_cleaned, unique_shp_cleaned)
unmatched_shp_cleaned <- setdiff(unique_shp_cleaned, unique_soil_cleaned)

# Calculate percentages
pct_soil_sampled_updated <- round((n_cleaned_matches / n_unique_shp_cleaned) * 100, 1)
pct_soil_matched_updated <- round((n_cleaned_matches / n_unique_soil_cleaned) * 100, 1)

cat("\n🔍 DETAILED CORRESPONDENCE RESULTS:\n")
cat("=" , rep("=", 60), "\n", sep="")
cat("📊 SUMMARY STATISTICS:\n")
cat("• Unique parcels in SOIL data (cleaned):   ", sprintf("%4d", n_unique_soil_cleaned), "\n")
cat("• Unique parcels in SHAPEFILE (cleaned):   ", sprintf("%4d", n_unique_shp_cleaned), "\n")
cat("• EXACT matches (before cleaning):         ", sprintf("%4d", n_exact_matches), "\n")
cat("• ADDITIONAL matches (from cleaning):      ", sprintf("%4d", n_additional_matches), "\n")
cat("• TOTAL MATCHES (after cleaning):          ", sprintf("%4d", n_cleaned_matches), "\n")
cat("• Unmatched SOIL parcels:                  ", sprintf("%4d", length(unmatched_soil_cleaned)), "\n")  
cat("• Unmatched SHAPEFILE parcels:             ", sprintf("%4d", length(unmatched_shp_cleaned)), "\n")

cat("\n📈 COVERAGE ANALYSIS:\n")
cat("• Percentage of shapefile parcels with soil data: ", sprintf("%5.1f%%", pct_soil_sampled_updated), "\n")
cat("• Percentage of soil samples with parcel match:   ", sprintf("%5.1f%%", pct_soil_matched_updated), "\n")

# ============================================================================
# CREATE DETAILED UNMATCHED PARCEL REPORTS
# ============================================================================

cat("\n=== CREATING DETAILED UNMATCHED PARCEL REPORTS ===\n")

# Create output folder for reports
output_folder <- "parcel_analysis_reports"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
  cat("Created output folder:", output_folder, "\n")
}

# 1. UNMATCHED SOIL PARCELS REPORT
cat("\nCreating detailed unmatched soil parcels report...\n")

unmatched_soil_report <- soil_parcel_details %>%
  filter(Parc_CLEANED %in% unmatched_soil_cleaned) %>%
  mutate(
    Cleaned_Name = Parc_CLEANED,
    Original_Uppercase = Parc.Detaillee_UPPER
  ) %>%
  arrange(Cleaned_Name)

# Add analysis columns
unmatched_soil_report <- unmatched_soil_report %>%
  mutate(
    Name_Length = nchar(Cleaned_Name),
    Contains_Numbers = grepl("[0-9]", Cleaned_Name),
    Potential_Issue = case_when(
      Cleaned_Name == "" ~ "Empty after cleaning",
      Name_Length < 3 ~ "Very short name",
      Name_Length > 20 ~ "Very long name",
      !Contains_Numbers ~ "No numbers (may be generic)",
      TRUE ~ "Check manually"
    )
  )

# Save unmatched soil parcels report
write.csv(unmatched_soil_report, 
          file.path(output_folder, "unmatched_soil_parcels_detailed.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✓ Saved detailed unmatched soil parcels report:", nrow(unmatched_soil_report), "records\n")

# 2. UNMATCHED SHAPEFILE PARCELS REPORT
cat("Creating detailed unmatched shapefile parcels report...\n")

unmatched_shp_report <- shp_parcel_details %>%
  filter(field_CLEANED %in% unmatched_shp_cleaned) %>%
  mutate(
    Cleaned_Name = field_CLEANED,
    Original_Uppercase = field_UPPER
  ) %>%
  arrange(Cleaned_Name)

# Add analysis columns
unmatched_shp_report <- unmatched_shp_report %>%
  mutate(
    Name_Length = nchar(Cleaned_Name),
    Contains_Numbers = grepl("[0-9]", Cleaned_Name),
    Potential_Issue = case_when(
      Cleaned_Name == "" ~ "Empty after cleaning",
      Name_Length < 3 ~ "Very short name",
      Name_Length > 20 ~ "Very long name",
      !Contains_Numbers ~ "No numbers (may be generic)",
      TRUE ~ "Check manually"
    )
  )

# Save unmatched shapefile parcels report
write.csv(unmatched_shp_report, 
          file.path(output_folder, "unmatched_shapefile_parcels_detailed.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✓ Saved detailed unmatched shapefile parcels report:", nrow(unmatched_shp_report), "records\n")
```

## III.2. Manual matching

```{r}
# ============================================================================
# MANUAL CORRESPONDENCE ENHANCEMENT FOR PARCEL MATCHING
# ============================================================================
# This script applies manual correspondences to improve parcel matching
# Run this AFTER the original analysis code

# NOTE: This script assumes you have already run the original analysis code
# and have the necessary objects and data loaded in your R environment

cat("=== MANUAL CORRESPONDENCE ENHANCEMENT ===\n")
cat("This script enhances the original analysis with manual correspondences\n\n")

# ============================================================================
# CREATE MANUAL CORRESPONDENCE LOOKUP TABLE
# ============================================================================

cat("=== CREATING MANUAL CORRESPONDENCE LOOKUP ===\n")

# Create the manual correspondence lookup table based on your provided data
manual_correspondences <- data.frame(
  Original_Names_Soil = c(
    "D1B1 0-60", "D1B1 60-120", "D 1B 3", "E10a 1", "E10B 1", "E 3a 1", 
    "E 3B 1", "E 8B 1A", "E 8B 1B", "E8B1C", "F 2A 3", "F 8A 1", "F 9A 1", 
    "F 9B 1", "G 1A 1", "G 1B 1", "G 2A 1", "G 2B 1", "G 3A 1", "H13A 1", 
    "H 9A 1", "J 5A 1", "J 5B 1", "J5B ext", "K 7A 1", "K 7A 2", "KH13D", 
    "KH13E", "KH 1B", "KH 1C", "KH 2D1", "KH 2D2", "KH 2E", "KH2 E1", 
    "KH 2F", "KH 3A", "KH 3F", "KH 4D", "KH 9B", "KH 9K 1", "KH 9K 2", 
    "L 3A 1", "MB 1F 1", "MB 3A 3", "MB 3F", "MB 3L", "MB 3N", "MB 3R", 
    "MB 3T", "MB 3Y 1", "MB 5S", "MB 7A10", "MB 7A11", "MB 7C 6", "P 2B 2", 
    "Q 7A 1", "Q 7A 2", "Q 7A 3", "Q 7A 4", "Q7B2", "Q7B3", "Q7B4", "Q7C1", 
    "Q7C2", "R 2A2", "R 2A 4", "S 1B 1", "S 1B 2", "S 1B 3", "S 1B 4", 
    "S 1C 1", "S 1C 2", "S 1C 3", "S 1C 4", "S 1F 1", "S 1F2", "S 1F 3", 
    "S 1F 4", "S 1G 1", "S 1G 2", "S 1G 3", "S 1G 4", "S 1H 1", "S 1H 3", 
    "S 2A 1", "S 2A 2", "S 2A 3", "S 2A 4", "S 2B 1", "S 2B 2", "S 2B 3", 
    "S 2B 4", "S 2C 1", "S 2C 2", "S 2C 3", "S 2C 4", "S 2E 1", "S 2E 2", 
    "S 2E 3", "S 2E3", "S 2E 4", "S 2F 1", "S 2F 3", "S 2G 1", "S 2G 2", 
    "S 2G 4", "S 6(EFGH)", "S6 NÂ°1", "S6 NÂ°2", "S6 NÂ°3", "S6 NÂ°4", 
    "T 2", "T30", "T33E 1", "T39c 1", "T 4A", "T 4B", "U 3B 1", "U 3B1", 
    "U 3B 2", "U 3B2", "U 3K", "U 5G", "U 6C", "U 8D1"
  ),
  Manual_Correspondence_Shp = c(
    "D 1B", "D 1B", "D 1B", "E10A", "E10B", "E 3A", "E 3B", "E 8B", "E 8B", 
    "E 8B", "F 2A", "F 8A", "F 9A", "F 9B", "G 1A", "G 1B", "G 2A", "G 2B", 
    "G 3A", "H13A", "H 9A", "J 5A", "J 5B", "J 5B", "K 7A", "K 7A", "KH13D", 
    "KH13E", "KH 1B", "KH 1C", "KH 2D", "KH 2D", "KH 2E", "KH2 E1", "KH 2F", 
    "KH 3A", "KH 3F", "KH 4D", "KH 9B", "KH 9K", "KH 9K", "L 3A", "MB 1F", 
    "MB 3A", "MB 3F", "MB 3L", "MB 3N", "MB 3R", "MB 3T", "MB 3Y", "MB 5S", 
    "MB 7A9", "MB 7A9", "MB 7C 5", "P 2B", "Q 7A", "Q 7A", "Q 7A", "Q 7A", 
    "Q 7B", "Q 7B", "Q 7B", "Q 7C", "Q 7C", "R 2A", "R 2A", "S 1B", "S 1B", 
    "S 1B", "S 1B", "S 1C", "S 1C", "S 1C", "S 1C", "S 1F", "S 1F", "S 1F", 
    "S 1F", "S 1G", "S 1G", "S 1G", "S 1G", "S 1H", "S 1H", "S 2A", "S 2A", 
    "S 2A", "S 2A", "S 2B", "S 2B", "S 2B", "S 2B", "S 2C", "S 2C", "S 2C", 
    "S 2C", "S 2E", "S 2E", "S 2E", "S 2E", "S 2E", "S 2F", "S 2F", "S 2G", 
    "S 2G", "S 2G", "S 6E", "S 6A", "S 6A", "S 6A", "S 6A", "T 2A", "T30", 
    "T33E", "T39C", "T 4", "T 4", "U 3B", "U 3B", "U 3B", "U 3B", "U 3K", 
    "U 5G", "U 6C", "U 8D"
  ),
  stringsAsFactors = FALSE
)

# Convert to uppercase for matching consistency
manual_correspondences$Original_Names_Soil_UPPER <- toupper(manual_correspondences$Original_Names_Soil)
manual_correspondences$Manual_Correspondence_Shp_UPPER <- toupper(manual_correspondences$Manual_Correspondence_Shp)

cat("Manual correspondence entries created:", nrow(manual_correspondences), "\n")

# ============================================================================
# APPLY MANUAL CORRESPONDENCES TO SOIL DATA
# ============================================================================

cat("\n=== APPLYING MANUAL CORRESPONDENCES ===\n")

# Create a copy of soil data for manual corrections
soil_data_manual <- soil_data

# Apply manual correspondences
cat("Applying manual correspondences to soil data...\n")
soil_data_manual$Parc_Manual_Corrected <- soil_data_manual$Parc.Detaillee_UPPER

# Apply manual correspondences
corrections_applied <- 0
for (i in 1:nrow(manual_correspondences)) {
  soil_original <- manual_correspondences$Original_Names_Soil_UPPER[i]
  shp_corrected <- manual_correspondences$Manual_Correspondence_Shp_UPPER[i]
  
  # Count records that will be updated
  matches <- soil_data_manual$Parc.Detaillee_UPPER == soil_original
  corrections_applied <- corrections_applied + sum(matches, na.rm = TRUE)
  
  # Update soil data where original name matches
  soil_data_manual$Parc_Manual_Corrected[matches] <- shp_corrected
}

cat("Manual corrections applied to", corrections_applied, "soil records\n")

# Apply cleaning function to manually corrected data
soil_data_manual$Parc_Manual_CLEANED <- sapply(soil_data_manual$Parc_Manual_Corrected, clean_parcel_name)

# Create updated soil parcel details with manual corrections
soil_parcel_details_manual <- soil_data_manual %>%
  group_by(Parc_Manual_Corrected) %>%
  summarise(
    Original_Names = paste(unique(Parc.Detaillee), collapse = "; "),
    Original_Names_Upper = paste(unique(Parc.Detaillee_UPPER), collapse = "; "),
    Sample_Count = n(),
    Years_Sampled = paste(sort(unique(Annee[!is.na(Annee)])), collapse = ", "),
    First_Sample_Year = min(Annee, na.rm = TRUE),
    Last_Sample_Year = max(Annee, na.rm = TRUE),
    Manual_Correction_Applied = any(Parc.Detaillee_UPPER != Parc_Manual_Corrected),
    .groups = 'drop'
  ) %>%
  arrange(Parc_Manual_Corrected)

# Add cleaned names
soil_parcel_details_manual$Parc_Manual_CLEANED <- sapply(soil_parcel_details_manual$Parc_Manual_Corrected, clean_parcel_name)

# Get unique manually corrected and cleaned parcels
unique_soil_manual_cleaned <- sort(unique(soil_parcel_details_manual$Parc_Manual_CLEANED))
n_unique_soil_manual <- length(unique_soil_manual_cleaned)

# ============================================================================
# ENHANCED CORRESPONDENCE ANALYSIS WITH MANUAL CORRECTIONS
# ============================================================================

cat("\n=== ENHANCED CORRESPONDENCE ANALYSIS ===\n")

# Find matches after manual corrections
manual_matches <- intersect(unique_soil_manual_cleaned, unique_shp_cleaned)
n_manual_matches <- length(manual_matches)

# Calculate improvement from manual correspondences
improvement_from_manual <- n_manual_matches - n_cleaned_matches

# Find remaining unmatched parcels
unmatched_soil_manual <- setdiff(unique_soil_manual_cleaned, unique_shp_cleaned)
unmatched_shp_manual <- setdiff(unique_shp_cleaned, unique_soil_manual_cleaned)

# Calculate updated percentages
pct_soil_sampled_manual <- round((n_manual_matches / n_unique_shp_cleaned) * 100, 1)
pct_soil_matched_manual <- round((n_manual_matches / n_unique_soil_manual) * 100, 1)

cat("\n🔍 ENHANCED CORRESPONDENCE RESULTS WITH MANUAL CORRECTIONS:\n")
cat("=" , rep("=", 70), "\n", sep="")
cat("📊 COMPARISON STATISTICS:\n")
cat("• ORIGINAL matches (before manual):            ", sprintf("%4d", n_cleaned_matches), "\n")
cat("• ENHANCED matches (after manual):             ", sprintf("%4d", n_manual_matches), "\n")
cat("• IMPROVEMENT from manual correspondences:     ", sprintf("%4d", improvement_from_manual), "\n")
cat("• Remaining unmatched SOIL parcels:            ", sprintf("%4d", length(unmatched_soil_manual)), "\n")
cat("• Remaining unmatched SHAPEFILE parcels:       ", sprintf("%4d", length(unmatched_shp_manual)), "\n")

cat("\n📈 ENHANCED COVERAGE ANALYSIS:\n")
cat("• Original coverage: ", sprintf("%5.1f%%", pct_soil_sampled_updated), " → Enhanced: ", sprintf("%5.1f%%", pct_soil_sampled_manual), "\n")
cat("• Coverage improvement: ", sprintf("+%.1f%%", pct_soil_sampled_manual - pct_soil_sampled_updated), "\n")

# ============================================================================
# CREATE ENHANCED REPORTS WITH MANUAL CORRECTIONS
# ============================================================================

cat("\n=== CREATING ENHANCED REPORTS ===\n")

# Create enhanced output folder
output_folder_manual <- "parcel_analysis_enhanced_with_manual"
if (!dir.exists(output_folder_manual)) {
  dir.create(output_folder_manual)
  cat("Created enhanced output folder:", output_folder_manual, "\n")
}

# 1. ENHANCED MATCHED PARCELS REPORT
cat("\nCreating enhanced matched parcels report...\n")

matched_soil_manual <- soil_parcel_details_manual %>%
  filter(Parc_Manual_CLEANED %in% manual_matches)

matched_shp_manual <- shp_parcel_details %>%
  filter(field_CLEANED %in% manual_matches)

enhanced_matched_report <- matched_soil_manual %>%
  left_join(matched_shp_manual, by = c("Parc_Manual_CLEANED" = "field_CLEANED")) %>%
  mutate(
    Cleaned_Name = Parc_Manual_CLEANED,
    Manual_Corrected_Name = Parc_Manual_Corrected,
    Soil_Original_Names = Original_Names.x,
    Soil_Original_Names_Upper = Original_Names_Upper,
    Shapefile_Original_Names = Original_Names.y,
    Shapefile_Feature_Count = Feature_Count
  ) %>%
  arrange(Cleaned_Name)

write.csv(enhanced_matched_report, 
          file.path(output_folder_manual, "enhanced_matched_parcels_with_manual.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✓ Saved enhanced matched parcels report:", nrow(enhanced_matched_report), "records\n")

# 2. REMAINING UNMATCHED SOIL PARCELS (AFTER MANUAL CORRECTIONS)
cat("Creating remaining unmatched soil parcels report...\n")

remaining_unmatched_soil <- soil_parcel_details_manual %>%
  filter(Parc_Manual_CLEANED %in% unmatched_soil_manual) %>%
  mutate(
    Cleaned_Name = Parc_Manual_CLEANED,
    Manual_Corrected_Name = Parc_Manual_Corrected,
    Name_Length = nchar(Parc_Manual_CLEANED),
    Contains_Numbers = grepl("[0-9]", Parc_Manual_CLEANED),
    Potential_Issue = case_when(
      Parc_Manual_CLEANED == "" ~ "Empty after cleaning",
      Name_Length < 3 ~ "Very short name",
      Name_Length > 20 ~ "Very long name",
      !Contains_Numbers ~ "No numbers (may be generic)",
      TRUE ~ "Requires further investigation"
    )
  ) %>%
  arrange(Cleaned_Name)

write.csv(remaining_unmatched_soil, 
          file.path(output_folder_manual, "remaining_unmatched_soil_after_manual.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✓ Saved remaining unmatched soil parcels:", nrow(remaining_unmatched_soil), "records\n")

# 3. REMAINING UNMATCHED SHAPEFILE PARCELS
remaining_unmatched_shp <- shp_parcel_details %>%
  filter(field_CLEANED %in% unmatched_shp_manual) %>%
  mutate(
    Cleaned_Name = field_CLEANED,
    Original_Uppercase = field_UPPER,
    Name_Length = nchar(field_CLEANED),
    Contains_Numbers = grepl("[0-9]", field_CLEANED),
    Potential_Issue = case_when(
      field_CLEANED == "" ~ "Empty after cleaning",
      Name_Length < 3 ~ "Very short name",
      Name_Length > 20 ~ "Very long name",
      !Contains_Numbers ~ "No numbers (may be generic)",
      TRUE ~ "Needs soil sampling"
    )
  ) %>%
  arrange(Cleaned_Name)

write.csv(remaining_unmatched_shp, 
          file.path(output_folder_manual, "remaining_unmatched_shapefile_after_manual.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✓ Saved remaining unmatched shapefile parcels:", nrow(remaining_unmatched_shp), "records\n")

# 4. MANUAL CORRESPONDENCE IMPACT ANALYSIS
cat("Creating manual correspondence impact analysis...\n")

manual_impact_analysis <- manual_correspondences %>%
  left_join(
    soil_data_manual %>% 
      group_by(Parc.Detaillee_UPPER) %>% 
      summarise(Sample_Count = n(), .groups = 'drop'),
    by = c("Original_Names_Soil_UPPER" = "Parc.Detaillee_UPPER")
  ) %>%
  mutate(
    Found_In_Shapefile = Manual_Correspondence_Shp_UPPER %in% unique_shp_parcels,
    # Check if the CLEANED version of the manual correspondence matches
    Manual_Cleaned = sapply(Manual_Correspondence_Shp_UPPER, clean_parcel_name),
    Correction_Successful = Manual_Cleaned %in% manual_matches,
    Sample_Count = ifelse(is.na(Sample_Count), 0, Sample_Count),
    Status = case_when(
      Sample_Count == 0 ~ "No samples found",
      !Found_In_Shapefile ~ "Target not in shapefile",
      Correction_Successful ~ "Successfully matched",
      TRUE ~ "Match failed (cleaning issue?)"
    )
  ) %>%
  arrange(Original_Names_Soil)

write.csv(manual_impact_analysis, 
          file.path(output_folder_manual, "manual_correspondence_impact_analysis.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

# Create summary of manual correspondence success
manual_success_summary <- manual_impact_analysis %>%
  count(Status) %>%
  mutate(Percentage = round(n / nrow(manual_impact_analysis) * 100, 1))

cat("✓ Saved manual correspondence impact analysis:", nrow(manual_impact_analysis), "records\n")

# 5. ENHANCED COMPREHENSIVE SUMMARY
cat("Creating enhanced comprehensive summary...\n")

enhanced_summary <- data.frame(
  Category = c(
    "=== ORIGINAL ANALYSIS ===",
    "Original Soil Parcels (Cleaned)",
    "Original Shapefile Parcels (Cleaned)", 
    "Original Matches",
    "Original Unmatched Soil",
    "Original Unmatched Shapefile",
    "Original Coverage Percentage",
    "",
    "=== MANUAL ENHANCEMENT ===",
    "Manual Correspondence Entries",
    "Soil Records Corrected",
    "Enhanced Matches",
    "Improvement from Manual Corrections",
    "Remaining Unmatched Soil",
    "Remaining Unmatched Shapefile", 
    "Enhanced Coverage Percentage",
    "Coverage Improvement",
    "",
    "=== MANUAL CORRESPONDENCE SUCCESS ===",
    "Successfully Matched",
    "Target Not in Shapefile",
    "Match Failed",
    "No Samples Found"
  ),
  Value = c(
    "",
    n_unique_soil_cleaned,
    n_unique_shp_cleaned,
    n_cleaned_matches,
    length(unmatched_soil_cleaned),
    length(unmatched_shp_cleaned),
    paste0(pct_soil_sampled_updated, "%"),
    "",
    "",
    nrow(manual_correspondences),
    corrections_applied,
    n_manual_matches,
    improvement_from_manual,
    length(unmatched_soil_manual),
    length(unmatched_shp_manual),
    paste0(pct_soil_sampled_manual, "%"),
    paste0("+", pct_soil_sampled_manual - pct_soil_sampled_updated, "%"),
    "",
    "",
    sum(manual_success_summary$n[manual_success_summary$Status == "Successfully matched"]),
    sum(manual_success_summary$n[manual_success_summary$Status == "Target not in shapefile"]),
    sum(manual_success_summary$n[manual_success_summary$Status == "Match failed"]),
    sum(manual_success_summary$n[manual_success_summary$Status == "No samples found"])
  )
)

write.csv(enhanced_summary, 
          file.path(output_folder_manual, "enhanced_comprehensive_summary.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✓ Saved enhanced comprehensive summary\n")

# 6. ENHANCED YEARLY ANALYSIS
cat("Creating enhanced yearly analysis...\n")

enhanced_yearly_analysis <- data.frame()

for (year in unique_years) {
  year_data <- soil_data_manual[soil_data_manual$Annee == year & !is.na(soil_data_manual$Annee), ]
  
  # Get unique parcels for this year (manually corrected and cleaned)
  year_parcels_cleaned <- unique(year_data$Parc_Manual_CLEANED)
  year_parcels_cleaned <- year_parcels_cleaned[year_parcels_cleaned != ""]
  
  # Find matches and unmatched
  year_matches <- intersect(year_parcels_cleaned, unique_shp_cleaned)
  year_unmatched <- setdiff(year_parcels_cleaned, unique_shp_cleaned)
  
  # Calculate improvement for this year
  year_data_original <- soil_data[soil_data$Annee == year & !is.na(soil_data$Annee), ]
  year_parcels_original <- unique(year_data_original$Parc_CLEANED)
  year_parcels_original <- year_parcels_original[year_parcels_original != ""]
  year_matches_original <- intersect(year_parcels_original, unique_shp_cleaned)
  year_improvement <- length(year_matches) - length(year_matches_original)
  
  enhanced_yearly_analysis <- rbind(enhanced_yearly_analysis, data.frame(
    Year = year,
    Total_Samples = nrow(year_data),
    Unique_Parcels_Sampled = length(year_parcels_cleaned),
    Original_Matched_Parcels = length(year_matches_original),
    Enhanced_Matched_Parcels = length(year_matches),
    Improvement_from_Manual = year_improvement,
    Remaining_Unmatched = length(year_unmatched),
    Original_Coverage_Pct = round((length(year_matches_original) / n_unique_shp_cleaned) * 100, 1),
    Enhanced_Coverage_Pct = round((length(year_matches) / n_unique_shp_cleaned) * 100, 1),
    Enhanced_Match_Rate = round((length(year_matches) / length(year_parcels_cleaned)) * 100, 1)
  ))
}

write.csv(enhanced_yearly_analysis, 
          file.path(output_folder_manual, "enhanced_yearly_analysis.csv"), 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✓ Saved enhanced yearly analysis\n")

# ============================================================================
# FINAL SUMMARY AND RECOMMENDATIONS
# ============================================================================

cat("\n" , rep("=", 80), "\n", sep="")
cat("🎯 MANUAL CORRESPONDENCE ENHANCEMENT COMPLETE!\n")
cat(rep("=", 80), "\n")

cat("\n📋 ENHANCED FILES CREATED IN '", output_folder_manual, "':\n", sep="")
cat("1. enhanced_matched_parcels_with_manual.csv          (", nrow(enhanced_matched_report), " records)\n")
cat("2. remaining_unmatched_soil_after_manual.csv         (", nrow(remaining_unmatched_soil), " records)\n")
cat("3. remaining_unmatched_shapefile_after_manual.csv    (", nrow(remaining_unmatched_shp), " records)\n")
cat("4. manual_correspondence_impact_analysis.csv         (", nrow(manual_impact_analysis), " records)\n")
cat("5. enhanced_comprehensive_summary.csv                (overview)\n")
cat("6. enhanced_yearly_analysis.csv                      (yearly breakdown)\n")

cat("\n📊 ENHANCEMENT IMPACT:\n")
cat("• Manual correspondences processed:    ", nrow(manual_correspondences), "\n")
cat("• Soil records corrected:              ", corrections_applied, "\n")
cat("• Matching improvement:                ", improvement_from_manual, " additional matches\n")
cat("• Original coverage (cleaned):         ", pct_soil_sampled_updated, "%\n")
cat("• Enhanced coverage (direct):          ", pct_soil_sampled_manual, "%\n")
cat("• Coverage improvement:                +", pct_soil_sampled_manual - pct_soil_sampled_updated, "%\n")

# Show manual correspondence success breakdown
cat("\n📈 MANUAL CORRESPONDENCE SUCCESS RATE:\n")
for (i in 1:nrow(manual_success_summary)) {
  status <- manual_success_summary$Status[i]
  count <- manual_success_summary$n[i]
  percentage <- manual_success_summary$Percentage[i]
  cat("• ", sprintf("%-25s", status), ": ", sprintf("%3d", count), " (", sprintf("%4.1f%%", percentage), ")\n", sep="")
}


improvement_rate <- round((improvement_from_manual / n_cleaned_matches) * 100, 1)
cat("• Manual corrections improved matching by ", improvement_rate, "%\n", sep="")
```

## III.3. Soil Data Variables Explorer

This section examines all variables in the soil database, classifying them by type and assessing data completeness. The code creates an inventory showing each variable's structure, missing data patterns, and sample values to guide variable selection for analysis.

```{r}
# ============================================================================
# SOIL DATA VARIABLES EXPLORER
# ============================================================================
# This script examines all variables in the soil dataset to help you decide
# which variables to keep for parcel linking analysis

cat("=== SOIL DATA VARIABLES EXPLORER ===\n")

# Ensure soil data is loaded (assuming from your previous analysis)
if (!exists("soil_data")) {
  cat("Loading soil data...\n")
  soil_data <- read.csv("BD_soil.csv", sep = ";", dec = ",", header = TRUE, 
                        stringsAsFactors = FALSE, fileEncoding = "latin1")
}

# ============================================================================
# BASIC DATASET INFORMATION
# ============================================================================

cat("📊 BASIC DATASET INFORMATION:\n")
cat("=" , rep("=", 50), "\n", sep="")
cat("Dataset dimensions:", nrow(soil_data), "rows ×", ncol(soil_data), "columns\n")
cat("Total observations:", nrow(soil_data), "\n")
cat("Number of variables:", ncol(soil_data), "\n\n")

# ============================================================================
# DETAILED VARIABLE ANALYSIS
# ============================================================================

cat("🔍 DETAILED VARIABLE ANALYSIS:\n")
cat("=" , rep("=", 80), "\n", sep="")

# Create a comprehensive variable summary
variable_summary <- data.frame(
  Variable_Number = 1:ncol(soil_data),
  Variable_Name = names(soil_data),
  Data_Type = sapply(soil_data, class),
  Non_Missing_Count = sapply(soil_data, function(x) sum(!is.na(x))),
  Missing_Count = sapply(soil_data, function(x) sum(is.na(x))),
  Unique_Values = sapply(soil_data, function(x) length(unique(x[!is.na(x)]))),
  stringsAsFactors = FALSE
)

# Calculate percentages
variable_summary$Missing_Percentage <- round((variable_summary$Missing_Count / nrow(soil_data)) * 100, 1)
variable_summary$Complete_Percentage <- round((variable_summary$Non_Missing_Count / nrow(soil_data)) * 100, 1)

# Add sample values for each variable
variable_summary$Sample_Values <- sapply(names(soil_data), function(var_name) {
  values <- soil_data[[var_name]][!is.na(soil_data[[var_name]])]
  if (length(values) == 0) {
    return("All NA")
  } else if (length(values) <= 3) {
    return(paste(values, collapse = ", "))
  } else {
    return(paste(values[1:3], collapse = ", "))
  }
})

# Add variable classification
variable_summary$Variable_Type <- sapply(1:nrow(variable_summary), function(i) {
  var_name <- variable_summary$Variable_Name[i]
  unique_vals <- variable_summary$Unique_Values[i]
  data_type <- variable_summary$Data_Type[i]
  
  if (grepl("Parc|parc", var_name, ignore.case = TRUE)) {
    return("Parcel_ID")
  } else if (grepl("Annee|Year|year", var_name, ignore.case = TRUE)) {
    return("Time")
  } else if (grepl("Date|date", var_name, ignore.case = TRUE)) {
    return("Date")
  } else if (data_type %in% c("character", "factor") && unique_vals <= 20) {
    return("Categorical")
  } else if (data_type %in% c("character", "factor")) {
    return("Text/ID")
  } else if (data_type %in% c("numeric", "integer")) {
    return("Numeric")
  } else {
    return("Other")
  }
})

# Sort by variable number for easy reference
variable_summary <- variable_summary[order(variable_summary$Variable_Number), ]

# ============================================================================
# DISPLAY VARIABLE SUMMARY TABLE
# ============================================================================

cat("📋 COMPLETE VARIABLE SUMMARY:\n")
cat(sprintf("%-3s %-25s %-12s %-8s %-8s %-8s %-8s %-12s %-20s\n", 
           "No.", "Variable_Name", "Data_Type", "Complete", "Missing", "Missing%", "Unique", "Category", "Sample_Values"))
cat(rep("-", 120), "\n")

for (i in 1:nrow(variable_summary)) {
  cat(sprintf("%-3d %-25s %-12s %-8d %-8d %-8.1f %-8d %-12s %-20s\n",
             variable_summary$Variable_Number[i],
             substr(variable_summary$Variable_Name[i], 1, 25),
             variable_summary$Data_Type[i],
             variable_summary$Non_Missing_Count[i],
             variable_summary$Missing_Count[i],
             variable_summary$Missing_Percentage[i],
             variable_summary$Unique_Values[i],
             variable_summary$Variable_Type[i],
             substr(variable_summary$Sample_Values[i], 1, 20)
             ))
}

# ============================================================================
# VARIABLE CATEGORIES BREAKDOWN
# ============================================================================

cat("\n📊 VARIABLES BY CATEGORY:\n")
cat("=" , rep("=", 40), "\n", sep="")

category_counts <- table(variable_summary$Variable_Type)
for (category in names(category_counts)) {
  cat(sprintf("%-15s: %d variables\n", category, category_counts[category]))
}

# ============================================================================
# DETAILED BREAKDOWN BY CATEGORY
# ============================================================================

cat("\n🔍 DETAILED BREAKDOWN BY CATEGORY:\n")

for (category in unique(variable_summary$Variable_Type)) {
  cat("\n", rep("-", 60), "\n", sep="")
  cat("📂 ", toupper(category), " VARIABLES:\n", sep="")
  cat(rep("-", 60), "\n")
  
  category_vars <- variable_summary[variable_summary$Variable_Type == category, ]
  
  for (i in 1:nrow(category_vars)) {
    var_info <- category_vars[i, ]
    cat(sprintf("%2d. %-25s [%s] - %.1f%% complete, %d unique values\n",
               var_info$Variable_Number,
               var_info$Variable_Name,
               var_info$Data_Type,
               var_info$Complete_Percentage,
               var_info$Unique_Values))
    
    # Show sample values for this variable
    cat(sprintf("    Sample values: %s\n", var_info$Sample_Values))
    
    # For numeric variables, show additional statistics
    if (var_info$Data_Type %in% c("numeric", "integer") && var_info$Non_Missing_Count > 0) {
      values <- soil_data[[var_info$Variable_Name]][!is.na(soil_data[[var_info$Variable_Name]])]
      if (length(values) > 0) {
        cat(sprintf("    Range: %.3f to %.3f, Mean: %.3f\n", 
                   min(values, na.rm = TRUE), 
                   max(values, na.rm = TRUE), 
                   mean(values, na.rm = TRUE)))
      }
    }
    cat("\n")
  }
}

# ============================================================================
# DATA QUALITY ASSESSMENT
# ============================================================================

cat("\n⚠️  DATA QUALITY ASSESSMENT:\n")
cat("=" , rep("=", 50), "\n", sep="")

# Variables with high missing data
high_missing <- variable_summary[variable_summary$Missing_Percentage > 50, ]
if (nrow(high_missing) > 0) {
  cat("🚨 Variables with >50% missing data:\n")
  for (i in 1:nrow(high_missing)) {
    cat(sprintf("• %s: %.1f%% missing\n", 
               high_missing$Variable_Name[i], 
               high_missing$Missing_Percentage[i]))
  }
} else {
  cat("✅ No variables have >50% missing data\n")
}

cat("\n")

# Variables with very few unique values
low_unique <- variable_summary[variable_summary$Variable_Type == "Numeric" & 
                              variable_summary$Unique_Values <= 5 & 
                              variable_summary$Non_Missing_Count > 0, ]
if (nrow(low_unique) > 0) {
  cat("ℹ️  Numeric variables with ≤5 unique values (potential categorical):\n")
  for (i in 1:nrow(low_unique)) {
    cat(sprintf("• %s: %d unique values\n", 
               low_unique$Variable_Name[i], 
               low_unique$Unique_Values[i]))
  }
} else {
  cat("ℹ️  All numeric variables have >5 unique values\n")
}

# ============================================================================
# EXPORT VARIABLE SUMMARY
# ============================================================================

cat("\n💾 EXPORTING VARIABLE SUMMARY:\n")
cat("=" , rep("=", 40), "\n", sep="")

# Create output folder if it doesn't exist
if (!dir.exists("variable_analysis")) {
  dir.create("variable_analysis")
}

# Export detailed variable summary
write.csv(variable_summary, "variable_analysis/soil_variables_detailed_summary.csv", 
          row.names = FALSE, fileEncoding = "UTF-8")

cat("✅ Exported detailed variable summary to: variable_analysis/soil_variables_detailed_summary.csv\n")

# Create a simple variable list for easy reference
simple_variable_list <- data.frame(
  Number = variable_summary$Variable_Number,
  Variable_Name = variable_summary$Variable_Name,
  Category = variable_summary$Variable_Type,
  Complete_Percentage = variable_summary$Complete_Percentage,
  Description = paste0(variable_summary$Data_Type, " - ", variable_summary$Unique_Values, " unique values")
)

write.csv(simple_variable_list, "variable_analysis/soil_variables_simple_list.csv", 
          row.names = FALSE, fileEncoding = "UTF-8")
```

## III.4. Soil Data Export

This section creates a clean soil dataset with selected variables including pH, nutrients, particle sizes, and organic matter measurements. The code adds three versions of parcel identifiers for traceability, generates a data dictionary, and exports the final dataset ready for spatial modeling.

```{r}
# ============================================================================
# CREATE CLEAN SOIL DATASET WITH MATCHED PARCEL IDS
# ============================================================================
# This script creates a clean soil data CSV with selected variables and 
# the parcel IDs that have been processed through original + manual matching

cat("=== CREATING CLEAN SOIL DATASET ===\n")

# ============================================================================
# VERIFY REQUIRED DATA IS AVAILABLE
# ============================================================================

cat("📋 VERIFYING REQUIRED DATA:\n")
cat("=" , rep("=", 50), "\n", sep="")

# Check if we have the manually corrected soil data from previous analysis
required_objects <- c("soil_data", "soil_data_manual", "clean_parcel_name")

missing_objects <- c()
for (obj in required_objects) {
  if (!exists(obj)) {
    missing_objects <- c(missing_objects, obj)
  }
}

if (length(missing_objects) > 0) {
  cat("⚠️  Missing required objects from previous analysis:\n")
  for (obj in missing_objects) {
    cat("• ", obj, "\n")
  }
  cat("\n🔄 Loading original soil data and applying corrections...\n")
  
  # Load original soil data if not available
  if (!exists("soil_data")) {
    soil_data <- read.csv("BD_soil.csv", sep = ";", dec = ",", header = TRUE, 
                          stringsAsFactors = FALSE, fileEncoding = "latin1")
    cat("✅ Loaded original soil data\n")
  }
  
  # Recreate cleaning function if not available
  if (!exists("clean_parcel_name")) {
    clean_parcel_name <- function(name) {
      if (is.na(name) || name == "") return("")
      cleaned <- toupper(trimws(name))
      cleaned <- gsub("[^A-Z0-9]", "", cleaned)
      cleaned <- gsub("PARCELLE", "PARC", cleaned)
      cleaned <- gsub("PARCEL", "PARC", cleaned)
      cleaned <- gsub("BLOC", "BL", cleaned)
      cleaned <- gsub("SECTEUR", "SECT", cleaned)
      return(cleaned)
    }
    cat("✅ Recreated cleaning function\n")
  }
  
  # If soil_data_manual doesn't exist, we need to recreate it
  if (!exists("soil_data_manual")) {
    cat("🔧 Recreating manual corrections...\n")
    
    # Apply basic processing to soil data
    soil_data$Parc.Detaillee_UPPER <- toupper(soil_data$Parc.Detaillee)
    soil_data$Parc_CLEANED <- sapply(soil_data$Parc.Detaillee_UPPER, clean_parcel_name)
    
    # Recreate manual correspondences
    manual_correspondences <- data.frame(
      Original_Names_Soil_UPPER = c(
        "D1B1 0-60", "D1B1 60-120", "D 1B 3", "E10A 1", "E10B 1", "E 3A 1", 
        "E 3B 1", "E 8B 1A", "E 8B 1B", "E8B1C", "F 2A 3", "F 8A 1", "F 9A 1", 
        "F 9B 1", "G 1A 1", "G 1B 1", "G 2A 1", "G 2B 1", "G 3A 1", "H13A 1", 
        "H 9A 1", "J 5A 1", "J 5B 1", "J5B EXT", "K 7A 1", "K 7A 2", "KH13D", 
        "KH13E", "KH 1B", "KH 1C", "KH 2D1", "KH 2D2", "KH 2E", "KH2 E1", 
        "KH 2F", "KH 3A", "KH 3F", "KH 4D", "KH 9B", "KH 9K 1", "KH 9K 2", 
        "L 3A 1", "MB 1F 1", "MB 3A 3", "MB 3F", "MB 3L", "MB 3N", "MB 3R", 
        "MB 3T", "MB 3Y 1", "MB 5S", "MB 7A10", "MB 7A11", "MB 7C 6", "P 2B 2", 
        "Q 7A 1", "Q 7A 2", "Q 7A 3", "Q 7A 4", "Q7B2", "Q7B3", "Q7B4", "Q7C1", 
        "Q7C2", "R 2A2", "R 2A 4", "S 1B 1", "S 1B 2", "S 1B 3", "S 1B 4", 
        "S 1C 1", "S 1C 2", "S 1C 3", "S 1C 4", "S 1F 1", "S 1F2", "S 1F 3", 
        "S 1F 4", "S 1G 1", "S 1G 2", "S 1G 3", "S 1G 4", "S 1H 1", "S 1H 3", 
        "S 2A 1", "S 2A 2", "S 2A 3", "S 2A 4", "S 2B 1", "S 2B 2", "S 2B 3", 
        "S 2B 4", "S 2C 1", "S 2C 2", "S 2C 3", "S 2C 4", "S 2E 1", "S 2E 2", 
        "S 2E 3", "S 2E3", "S 2E 4", "S 2F 1", "S 2F 3", "S 2G 1", "S 2G 2", 
        "S 2G 4", "S 6(EFGH)", "S6 NÂ°1", "S6 NÂ°2", "S6 NÂ°3", "S6 NÂ°4", 
        "T 2", "T30", "T33E 1", "T39C 1", "T 4A", "T 4B", "U 3B 1", "U 3B1", 
        "U 3B 2", "U 3B2", "U 3K", "U 5G", "U 6C", "U 8D1"
      ),
      Manual_Correspondence_Shp_UPPER = c(
        "D 1B", "D 1B", "D 1B", "E10A", "E10B", "E 3A", "E 3B", "E 8B", "E 8B", 
        "E 8B", "F 2A", "F 8A", "F 9A", "F 9B", "G 1A", "G 1B", "G 2A", "G 2B", 
        "G 3A", "H13A", "H 9A", "J 5A", "J 5B", "J 5B", "K 7A", "K 7A", "KH13D", 
        "KH13E", "KH 1B", "KH 1C", "KH 2D", "KH 2D", "KH 2E", "KH2 E1", "KH 2F", 
        "KH 3A", "KH 3F", "KH 4D", "KH 9B", "KH 9K", "KH 9K", "L 3A", "MB 1F", 
        "MB 3A", "MB 3F", "MB 3L", "MB 3N", "MB 3R", "MB 3T", "MB 3Y", "MB 5S", 
        "MB 7A9", "MB 7A9", "MB 7C 5", "P 2B", "Q 7A", "Q 7A", "Q 7A", "Q 7A", 
        "Q 7B", "Q 7B", "Q 7B", "Q 7C", "Q 7C", "R 2A", "R 2A", "S 1B", "S 1B", 
        "S 1B", "S 1B", "S 1C", "S 1C", "S 1C", "S 1C", "S 1F", "S 1F", "S 1F", 
        "S 1F", "S 1G", "S 1G", "S 1G", "S 1G", "S 1H", "S 1H", "S 2A", "S 2A", 
        "S 2A", "S 2A", "S 2B", "S 2B", "S 2B", "S 2B", "S 2C", "S 2C", "S 2C", 
        "S 2C", "S 2E", "S 2E", "S 2E", "S 2E", "S 2E", "S 2F", "S 2F", "S 2G", 
        "S 2G", "S 2G", "S 6E", "S 6A", "S 6A", "S 6A", "S 6A", "T 2A", "T30", 
        "T33E", "T39C", "T 4", "T 4", "U 3B", "U 3B", "U 3B", "U 3B", "U 3K", 
        "U 5G", "U 6C", "U 8D"
      ),
      stringsAsFactors = FALSE
    )
    
    # Apply manual correspondences
    soil_data_manual <- soil_data
    soil_data_manual$Parc_Manual_Corrected <- soil_data_manual$Parc.Detaillee_UPPER
    
    corrections_applied <- 0
    for (i in 1:nrow(manual_correspondences)) {
      soil_original <- manual_correspondences$Original_Names_Soil_UPPER[i]
      shp_corrected <- manual_correspondences$Manual_Correspondence_Shp_UPPER[i]
      
      matches <- soil_data_manual$Parc.Detaillee_UPPER == soil_original
      corrections_applied <- corrections_applied + sum(matches, na.rm = TRUE)
      soil_data_manual$Parc_Manual_Corrected[matches] <- shp_corrected
    }
    
    # Apply cleaning to manual corrections
    soil_data_manual$Parc_Manual_CLEANED <- sapply(soil_data_manual$Parc_Manual_Corrected, clean_parcel_name)
    
    cat("✅ Applied manual correspondences to", corrections_applied, "records\n")
  }
  
} else {
  cat("✅ All required data objects are available\n")
}

# ============================================================================
# DEFINE SELECTED VARIABLES
# ============================================================================

cat("\n📝 DEFINING SELECTED VARIABLES:\n")
cat("=" , rep("=", 50), "\n", sep="")

# Define the variables you want to keep
selected_variables <- c(
  # Parcel identification (we'll add the processed IDs)
  "Parc.Detaillee",           # Original parcel name
  "Annee",                    # Year
  
  # Your selected soil measurement variables
  "Date_prelevement", 
  "pH", 
  "C.N", 
  "CE_mmho_cm", 
  "Ca_ech_me_100g", 
  "Mg_ech_me_100g", 
  "Na_ech_me_100g", 
  "K_ech_me_100g", 
  "P2O5_assimilable_mg_kg", 
  "Argiles", 
  "Limons_fins", 
  "Limons_grossiers", 
  "Sables_fins", 
  "Sables_grossiers", 
  "Carbone_organique_pour_mille", 
  "Azote_total_pour_mille", 
  "Limons", 
  "Sables", 
  "Matiere_organique.", 
  "S.Ca.Mg.K.Na..me.100g", 
  "CaO_g.100g", 
  "Mg.Ca", 
  "Mg.K", 
  "Indice_de_batance", 
  "Texture"
)

# Check which variables exist in the soil data
available_vars <- selected_variables[selected_variables %in% names(soil_data_manual)]
missing_vars <- selected_variables[!selected_variables %in% names(soil_data_manual)]

cat("✅ Available variables:", length(available_vars), "\n")
cat("⚠️  Missing variables:", length(missing_vars), "\n")

if (length(missing_vars) > 0) {
  cat("\nMissing variables:\n")
  for (var in missing_vars) {
    cat("• ", var, "\n")
  }
  
  # Try to find similar variable names
  cat("\n🔍 Looking for similar variable names in dataset:\n")
  all_vars <- names(soil_data_manual)
  for (missing_var in missing_vars) {
    # Simple similarity check
    similar_vars <- c()
    for (existing_var in all_vars) {
      # Check if any part of the missing variable name appears in existing variable
      missing_parts <- unlist(strsplit(missing_var, "[._]"))
      for (part in missing_parts) {
        if (nchar(part) > 2 && grepl(part, existing_var, ignore.case = TRUE)) {
          similar_vars <- c(similar_vars, existing_var)
          break
        }
      }
    }
    if (length(similar_vars) > 0) {
      cat("  '", missing_var, "' might be: ", paste(unique(similar_vars), collapse = ", "), "\n")
    }
  }
}

# ============================================================================
# CREATE CLEAN SOIL DATASET
# ============================================================================

cat("\n🔧 CREATING CLEAN SOIL DATASET:\n")
cat("=" , rep("=", 50), "\n", sep="")

# Start with the manually corrected soil data
clean_soil_data <- soil_data_manual

# Add the processed parcel ID columns
clean_soil_data$Parcel_ID_Original <- clean_soil_data$Parc.Detaillee_UPPER
clean_soil_data$Parcel_ID_Manual_Corrected <- clean_soil_data$Parc_Manual_Corrected
clean_soil_data$Parcel_ID_Cleaned_Final <- clean_soil_data$Parc_Manual_CLEANED

# Select only the available variables plus our parcel ID columns
final_variables <- c(
  "Parcel_ID_Original",           # Original parcel name (uppercase)
  "Parcel_ID_Manual_Corrected",   # After manual corrections
  "Parcel_ID_Cleaned_Final",      # Final cleaned version for matching
  available_vars                  # Your selected variables that exist
)

# Create the final clean dataset
clean_soil_data <- clean_soil_data[, final_variables]

# Add some metadata columns
clean_soil_data$Manual_Correction_Applied <- clean_soil_data$Parcel_ID_Original != clean_soil_data$Parcel_ID_Manual_Corrected
clean_soil_data$Data_Processing_Date <- Sys.Date()

# ============================================================================
# DATA QUALITY SUMMARY
# ============================================================================

cat("📊 DATA QUALITY SUMMARY:\n")
cat("=" , rep("=", 50), "\n", sep="")

cat("Total records in clean dataset:", nrow(clean_soil_data), "\n")
cat("Total variables in clean dataset:", ncol(clean_soil_data), "\n")

# Count records with manual corrections
manual_corrections <- sum(clean_soil_data$Manual_Correction_Applied, na.rm = TRUE)
cat("Records with manual parcel ID corrections:", manual_corrections, "\n")

# Show unique parcels
unique_original <- length(unique(clean_soil_data$Parcel_ID_Original))
unique_final <- length(unique(clean_soil_data$Parcel_ID_Cleaned_Final))
cat("Unique parcels (original names):", unique_original, "\n")
cat("Unique parcels (final cleaned):", unique_final, "\n")

# Year range
if ("Annee" %in% names(clean_soil_data)) {
  years <- clean_soil_data$Annee[!is.na(clean_soil_data$Annee)]
  if (length(years) > 0) {
    cat("Year range:", min(years), "to", max(years), "\n")
    cat("Years with data:", paste(sort(unique(years)), collapse = ", "), "\n")
  }
}

# Missing data summary for selected variables
cat("\n📋 MISSING DATA SUMMARY FOR SELECTED VARIABLES:\n")
measurement_vars <- available_vars[!available_vars %in% c("Parc.Detaillee", "Annee", "Date_prelevement")]

for (var in measurement_vars) {
  missing_count <- sum(is.na(clean_soil_data[[var]]))
  missing_pct <- round((missing_count / nrow(clean_soil_data)) * 100, 1)
  cat(sprintf("%-30s: %4d missing (%5.1f%%)\n", substr(var, 1, 30), missing_count, missing_pct))
}

# ============================================================================
# EXPORT CLEAN DATASET
# ============================================================================

cat("\n💾 EXPORTING CLEAN DATASET:\n")
cat("=" , rep("=", 40), "\n", sep="")

# Create output folder
output_folder <- "clean_soil_dataset"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
  cat("✅ Created output folder:", output_folder, "\n")
}

# Export main clean dataset
output_filename <- file.path(output_folder, "clean_soil_data_with_matched_parcels.csv")
write.csv(clean_soil_data, output_filename, row.names = FALSE, fileEncoding = "UTF-8")

cat("✅ Exported clean soil dataset to:", output_filename, "\n")

# Create a data dictionary
data_dictionary <- data.frame(
  Variable_Name = names(clean_soil_data),
  Variable_Type = sapply(clean_soil_data, class),
  Description = c(
    "Original parcel name (uppercase)",
    "Parcel name after manual corrections",  
    "Final cleaned parcel name for GIS matching",
    "Original parcel name (mixed case)",
    "Year of sampling",
    if ("Date_prelevement" %in% available_vars) "Sampling date" else NULL,
    if ("pH" %in% available_vars) "Soil pH" else NULL,
    if ("C.N" %in% available_vars) "Carbon to Nitrogen ratio" else NULL,
    if ("CE_mmho_cm" %in% available_vars) "Electrical conductivity (mmho/cm)" else NULL,
    if ("Ca_ech_me_100g" %in% available_vars) "Exchangeable Calcium (me/100g)" else NULL,
    if ("Mg_ech_me_100g" %in% available_vars) "Exchangeable Magnesium (me/100g)" else NULL,
    if ("Na_ech_me_100g" %in% available_vars) "Exchangeable Sodium (me/100g)" else NULL,
    if ("K_ech_me_100g" %in% available_vars) "Exchangeable Potassium (me/100g)" else NULL,
    if ("P2O5_assimilable_mg_kg" %in% available_vars) "Assimilable P2O5 (mg/kg)" else NULL,
    if ("Argiles" %in% available_vars) "Clay content (%)" else NULL,
    if ("Limons_fins" %in% available_vars) "Fine silt content (%)" else NULL,
    if ("Limons_grossiers" %in% available_vars) "Coarse silt content (%)" else NULL,
    if ("Sables_fins" %in% available_vars) "Fine sand content (%)" else NULL,
    if ("Sables_grossiers" %in% available_vars) "Coarse sand content (%)" else NULL,
    if ("Carbone_organique_pour_mille" %in% available_vars) "Organic carbon (‰)" else NULL,
    if ("Azote_total_pour_mille" %in% available_vars) "Total nitrogen (‰)" else NULL,
    if ("Limons" %in% available_vars) "Total silt content (%)" else NULL,
    if ("Sables" %in% available_vars) "Total sand content (%)" else NULL,
    if ("Matiere_organique." %in% available_vars) "Organic matter (%)" else NULL,
    if ("S.Ca.Mg.K.Na..me.100g" %in% available_vars) "Sum of exchangeable cations (me/100g)" else NULL,
    if ("CaO_g.100g" %in% available_vars) "Calcium oxide (g/100g)" else NULL,
    if ("Mg.Ca" %in% available_vars) "Mg/Ca ratio" else NULL,
    if ("Mg.K" %in% available_vars) "Mg/K ratio" else NULL,
    if ("Indice_de_batance" %in% available_vars) "Balance index" else NULL,
    if ("Texture" %in% available_vars) "Soil texture class" else NULL,
    "Whether manual parcel name correction was applied",
    "Date when this dataset was processed"
  ),
  Non_Missing_Count = sapply(clean_soil_data, function(x) sum(!is.na(x))),
  Missing_Count = sapply(clean_soil_data, function(x) sum(is.na(x))),
  stringsAsFactors = FALSE
)

# Remove NULL entries from Description
data_dictionary$Description <- ifelse(is.null(data_dictionary$Description), "", data_dictionary$Description)

dict_filename <- file.path(output_folder, "data_dictionary.csv")
write.csv(data_dictionary, dict_filename, row.names = FALSE, fileEncoding = "UTF-8")

cat("✅ Exported data dictionary to:", dict_filename, "\n")

# Create a summary report
summary_report <- data.frame(
  Metric = c(
    "Total records",
    "Total variables", 
    "Variables requested",
    "Variables found", 
    "Variables missing",
    "Records with manual corrections",
    "Unique parcels (original)",
    "Unique parcels (final cleaned)",
    "Year range",
    "Processing date"
  ),
  Value = c(
    nrow(clean_soil_data),
    ncol(clean_soil_data),
    length(selected_variables),
    length(available_vars),
    length(missing_vars),
    manual_corrections,
    unique_original,
    unique_final,
    if ("Annee" %in% names(clean_soil_data) && length(years) > 0) paste(min(years), "to", max(years)) else "Not available",
    as.character(Sys.Date())
  ),
  stringsAsFactors = FALSE
)

summary_filename <- file.path(output_folder, "dataset_summary.csv")
write.csv(summary_report, summary_filename, row.names = FALSE, fileEncoding = "UTF-8")

cat("✅ Exported dataset summary to:", summary_filename, "\n")
```

## III.5. Shapefile Data Export

This section adds the final cleaned parcel identifier to the shapefile as the second column and exports it in ESRI Shapefile format with all component files, preserving the UTM coordinate system for seamless integration with the soil dataset.

```{r}
# ============================================================================
# EXPORT SHAPEFILE WITH FINAL CLEANED PARCEL ID
# ============================================================================
# This script exports your shapefile with only the final cleaned parcel ID 
# added as the second column

cat("=== EXPORTING SHAPEFILE WITH FINAL CLEANED PARCEL ID ===\n")

# Load required libraries
library(sf)
library(dplyr)

# ============================================================================
# VERIFY REQUIRED DATA IS AVAILABLE
# ============================================================================

cat("📋 VERIFYING REQUIRED SPATIAL DATA:\n")
cat("=" , rep("=", 50), "\n", sep="")

# Check if we have the processed shapefile from previous analysis
if (!exists("parcelles_filtered") && !exists("parcelles")) {
  cat("⚠️  Shapefile data not found in environment.\n")
  cat("🔄 Attempting to load shapefile...\n")
  
  # Try to load the shapefile (you may need to adjust the path)
  shp_files <- list.files(pattern = "*.shp$", recursive = TRUE)
  if (length(shp_files) > 0) {
    cat("Found shapefile(s):\n")
    for (i in 1:length(shp_files)) {
      cat(i, ". ", shp_files[i], "\n")
    }
    cat("\n⚠️  Please ensure your shapefile is loaded as 'parcelles' or 'parcelles_filtered'\n")
    cat("Example: parcelles <- st_read('your_shapefile.shp')\n\n")
  } else {
    cat("❌ No shapefile found in current directory\n")
    cat("Please load your shapefile first:\n")
    cat("parcelles <- st_read('path/to/your/shapefile.shp')\n\n")
  }
  stop("Shapefile data required. Please load your shapefile first.")
}

# Use the available shapefile data
if (exists("parcelles_filtered")) {
  parcelles <- parcelles_filtered
  cat("✅ Using 'parcelles_filtered' from previous analysis\n")
} else if (exists("parcelles")) {
  cat("✅ Using 'parcelles' shapefile\n")
} else {
  stop("No shapefile data available")
}

# Verify it's a spatial object
if (!inherits(parcelles, "sf")) {
  stop("❌ Parcelles object is not a valid sf spatial object")
}

cat("Shapefile loaded successfully:\n")
cat("• Features (polygons):", nrow(parcelles), "\n")
cat("• Columns:", ncol(parcelles), "\n")
cat("• CRS:", st_crs(parcelles)$input, "\n")

# ============================================================================
# RECREATE CLEANING FUNCTION IF NEEDED
# ============================================================================

if (!exists("clean_parcel_name")) {
  clean_parcel_name <- function(name) {
    if (is.na(name) || name == "") return("")
    cleaned <- toupper(trimws(name))
    cleaned <- gsub("[^A-Z0-9]", "", cleaned)
    cleaned <- gsub("PARCELLE", "PARC", cleaned)
    cleaned <- gsub("PARCEL", "PARC", cleaned)
    cleaned <- gsub("BLOC", "BL", cleaned)
    cleaned <- gsub("SECTEUR", "SECT", cleaned)
    return(cleaned)
  }
  cat("✅ Recreated parcel name cleaning function\n")
}

# ============================================================================
# PROCESS SHAPEFILE WITH FINAL CLEANED ID
# ============================================================================

cat("\n🔧 PROCESSING SHAPEFILE WITH FINAL CLEANED ID:\n")
cat("=" , rep("=", 50), "\n", sep="")

# Display current column names
cat("Current shapefile columns:\n")
shp_cols <- names(st_drop_geometry(parcelles))
for (i in 1:length(shp_cols)) {
  cat(sprintf("%2d. %-20s\n", i, shp_cols[i]))
}

# Identify the field column (parcel identifier)
field_column <- NULL
possible_field_cols <- c("field", "Field", "FIELD", "parcel", "Parcel", "PARCEL", 
                        "id", "ID", "Id", "nom", "NOM", "name", "NAME")

for (col in possible_field_cols) {
  if (col %in% names(parcelles)) {
    field_column <- col
    break
  }
}

if (is.null(field_column)) {
  cat("\n❌ Could not automatically identify parcel ID column.\n")
  cat("Available columns:\n")
  print(names(st_drop_geometry(parcelles)))
  cat("\nPlease specify the correct column name manually:\n")
  cat("Example: field_column <- 'your_parcel_id_column_name'\n")
  stop("Parcel ID column identification required")
}

cat("\n✅ Using parcel ID column:", field_column, "\n")

# Create final cleaned parcel ID
cat("\n🔧 Creating final cleaned parcel ID...\n")
final_cleaned_id <- sapply(toupper(parcelles[[field_column]]), clean_parcel_name)

# Create new shapefile with final cleaned ID in second position
cat("📝 Adding final cleaned ID as second column...\n")

# Get all columns except geometry
data_cols <- st_drop_geometry(parcelles)

# Create new data frame with final cleaned ID as second column
new_data <- data.frame(
  data_cols[1],  # First original column
  Final_Cleaned_ID = final_cleaned_id,  # Our new column in second position
  data_cols[2:ncol(data_cols)]  # Rest of original columns
)

# Create new sf object with the reordered columns
final_shapefile <- st_sf(new_data, geometry = st_geometry(parcelles))

# Verify the column order
cat("\n✅ Final shapefile columns:\n")
final_cols <- names(st_drop_geometry(final_shapefile))
for (i in 1:length(final_cols)) {
  cat(sprintf("%2d. %-20s", i, final_cols[i]))
  if (i == 2) cat(" ← Final cleaned ID")
  cat("\n")
}

cat("\n📊 SUMMARY:\n")
cat("• Total parcels:", nrow(final_shapefile), "\n")
cat("• Total columns:", ncol(final_shapefile), "\n")
cat("• Final cleaned ID added as column 2\n")

# ============================================================================
# EXPORT FINAL SHAPEFILE
# ============================================================================

cat("\n💾 EXPORTING FINAL SHAPEFILE:\n")
cat("=" , rep("=", 40), "\n", sep="")

# Create output folder
output_folder <- "clean_soil_dataset"
if (!dir.exists(output_folder)) {
  dir.create(output_folder)
}

# Export the final shapefile
shapefile_path <- file.path(output_folder, "parcels_with_final_id")

# Remove existing files if they exist
existing_files <- list.files(output_folder, pattern = "parcels_with_final_id.*", full.names = TRUE)
if (length(existing_files) > 0) {
  file.remove(existing_files)
}

# Write shapefile
st_write(final_shapefile, paste0(shapefile_path, ".shp"), 
         driver = "ESRI Shapefile", delete_dsn = TRUE, quiet = TRUE)

cat("✅ Exported final shapefile to:", paste0(shapefile_path, ".shp"), "\n")

# Export attribute table for reference
attr_filename <- file.path(output_folder, "final_shapefile_attributes.csv")
write.csv(st_drop_geometry(final_shapefile), attr_filename, row.names = FALSE, fileEncoding = "UTF-8")
```
