---
title: "INLA Cross-Validation"
author: "Antsa Sarobidy"
format: html
editor: visual
---

# I. Part 1 : MULTI-VARIABLE ENHANCED SPATIO-TEMPORAL IMPUTATION WITH 5-FOLD CROSS-VALIDATION

## I.1. Data Loading and Variable Definition + MULTI-VARIABLE ENHANCED SPATIO-TEMPORAL IMPUTATION WITH 5-FOLD CROSS-VALIDATION

This section loads the normalized soil dataset from the geopackage file and defines two categories of variables for analysis. The compositional variables consist of five granulometry fractions that must sum to 100% (clay, fine silt, coarse silt, fine sand, and coarse sand), while non-compositional variables include 18 independent soil properties such as pH, nutrient content, and various ratios. The code validates the completeness of granulometry data and calculates summary statistics for all variables including their availability rates across the dataset.

```{r}
# ============================================================================
# MULTI-VARIABLE ENHANCED SPATIO-TEMPORAL IMPUTATION WITH 5-FOLD CROSS-VALIDATION
# Compositional treatment for granulometry fractions
# ============================================================================

library(sf)
library(INLA)
library(dplyr)
library(ggplot2)
library(viridis)
library(spdep)
library(readr)
library(purrr)
library(tibble)

cat("========================================\n")
cat("  MULTI-VARIABLE SPATIO-TEMPORAL       \n")
cat("  IMPUTATION WITH 5-FOLD CV            \n")
cat("========================================\n")

# ============================================================================
# STEP 1: LOAD THE COMPLETE DATASET AND IDENTIFY SOIL VARIABLES
# ============================================================================
cat("📂 LOADING COMPLETE DATASET AND IDENTIFYING SOIL VARIABLES:\n")
cat("=", rep("=", 60), "\n", sep="")

gpkg_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/parcels_with_soil_complete.gpkg"

if (file.exists(gpkg_path)) {
  parcelles_complete <- st_read(gpkg_path, quiet = TRUE)
  cat("✅ Complete dataset loaded:", nrow(parcelles_complete), "observations\n")
} else {
  cat("❌ Complete dataset not found. Please run the enhanced join script first.\n")
  stop("Dataset not available")
}

# Define compositional and non-compositional variables
GRANULO_FRACTIONS <- c(
  "Clay_percent",
  "Fine_Silt_percent",
  "Coarse_Silt_percent",
  "Fine_Sand_percent",
  "Coarse_Sand_percent"
)

NON_COMPOSITIONAL_VARIABLES <- c(
  "pH",
  "CN_Ratio", 
  "Electrical_Conductivity_mmho_cm",
  "Exchangeable_Calcium_meq_100g",
  "Exchangeable_Magnesium_meq_100g", 
  "Exchangeable_Sodium_meq_100g",
  "Exchangeable_Potassium_meq_100g",
  "Available_P2O5_mg_kg",
  "Organic_Carbon_per_mille",
  "Total_Nitrogen_per_mille",
  "Total_Silt_percent",
  "Total_Sand_percent", 
  "Organic_Matter_percent",
  "Sum_of_Exchangeable_Bases_meq_100g",
  "CaO_g_100g",
  "Mg_Ca_Ratio",
  "Mg_K_Ratio",
  "Compaction_Index"
)

cat("Compositional variables (granulometry):", length(GRANULO_FRACTIONS), "\n")
cat("Non-compositional variables:", length(NON_COMPOSITIONAL_VARIABLES), "\n")

for (i in seq_along(GRANULO_FRACTIONS)) {
  var_name <- GRANULO_FRACTIONS[i]
  n_available <- sum(!is.na(parcelles_complete[[var_name]]))
  pct_available <- round(n_available/nrow(parcelles_complete)*100, 1)
  cat(sprintf("• %2d. %-35s: %4d available (%5.1f%%), %4d missing\n", 
              i, var_name, n_available, pct_available, sum(is.na(parcelles_complete[[var_name]]))))
}

for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  n_available <- sum(!is.na(parcelles_complete[[var_name]]))
  pct_available <- round(n_available/nrow(parcelles_complete)*100, 1)
  cat(sprintf("• %2d. %-35s: %4d available (%5.1f%%), %4d missing\n", 
              length(GRANULO_FRACTIONS) + i, var_name, n_available, pct_available, sum(is.na(parcelles_complete[[var_name]]))))
}

cat("\nDataset structure:\n")
cat("• Total observations:", nrow(parcelles_complete), "\n")
cat("• Unique parcels:", length(unique(parcelles_complete$Final_Cleaned_ID)), "\n")
cat("• Year range:", min(parcelles_complete$Year, na.rm = TRUE), "-", max(parcelles_complete$Year, na.rm = TRUE), "\n")

# Validate compositional data
parcelles_complete <- parcelles_complete %>%
  mutate(
    has_complete_granulo = !is.na(Clay_percent) & 
                          !is.na(Fine_Silt_percent) & 
                          !is.na(Coarse_Silt_percent) & 
                          !is.na(Fine_Sand_percent) & 
                          !is.na(Coarse_Sand_percent),
    granulo_sum = Clay_percent + Fine_Silt_percent + Coarse_Silt_percent + 
                  Fine_Sand_percent + Coarse_Sand_percent
  )

complete_granulo <- sum(parcelles_complete$has_complete_granulo, na.rm = TRUE)
cat("\n🔍 COMPOSITIONAL DATA VALIDATION:\n")
cat("• Observations with complete granulometry:", complete_granulo, "\n")

if(complete_granulo > 0) {
  sum_stats <- parcelles_complete %>%
    filter(has_complete_granulo) %>%
    pull(granulo_sum) %>%
    summary()
  cat("• Sum range:", round(sum_stats[1], 2), "to", round(sum_stats[6], 2), "%\n")
}

# ============================================================================
# STEP 2: PREPARE SHARED SPATIAL STRUCTURE
# ============================================================================
cat("\n🗺️ PREPARING SHARED SPATIAL STRUCTURE:\n")
cat("=", rep("=", 40), "\n", sep="")

unique_parcels_spatial <- parcelles_complete %>%
  group_by(Final_Cleaned_ID) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(Final_Cleaned_ID)

cat("Unique parcels for spatial analysis:", nrow(unique_parcels_spatial), "\n")

unique_parcels_spatial <- st_make_valid(unique_parcels_spatial)
cat("Geometries validated\n")

field_mapping <- data.frame(
  Original_ID = unique_parcels_spatial$Final_Cleaned_ID,
  Sequential_ID = 1:nrow(unique_parcels_spatial)
)

cat("Field mapping created: 1 to", nrow(field_mapping), "\n")

unique_parcels_spatial <- unique_parcels_spatial %>%
  left_join(field_mapping, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
  arrange(Sequential_ID)

# ============================================================================
# STEP 3: CREATE SPATIAL WEIGHTS MATRIX W.P1
# ============================================================================
cat("\n🔗 CREATING SHARED SPATIAL WEIGHTS MATRIX:\n")
cat("=", rep("=", 42), "\n", sep="")

cat("Computing boundary-to-boundary distances (this may take time)...\n")

boundaries <- st_boundary(unique_parcels_spatial)
dist_matrix <- st_distance(boundaries, boundaries)
dist_matrix <- units::drop_units(dist_matrix)
diag(dist_matrix) <- NA

cat("Distance matrix calculated:", nrow(dist_matrix), "x", ncol(dist_matrix), "\n")

DISTANCE_THRESHOLD <- 110
cat("Using distance threshold:", DISTANCE_THRESHOLD, "meters\n")

W.P1_complete <- matrix(0, nrow = nrow(unique_parcels_spatial), ncol = nrow(unique_parcels_spatial))
W.P1_complete[dist_matrix <= DISTANCE_THRESHOLD & !is.na(dist_matrix)] <- 1
diag(W.P1_complete) <- 0
rownames(W.P1_complete) <- colnames(W.P1_complete) <- 1:nrow(W.P1_complete)

n_connections <- sum(W.P1_complete)
avg_neighbors <- n_connections / nrow(unique_parcels_spatial)

cat("✅ W.P1 matrix created:\n")
cat("• Dimensions:", nrow(W.P1_complete), "x", ncol(W.P1_complete), "\n")
cat("• Total connections:", n_connections, "\n")
cat("• Average neighbors per parcel:", round(avg_neighbors, 2), "\n")

# ============================================================================
# STEP 4: PREPARE SHARED TEMPORAL STRUCTURE
# ============================================================================
years_unique <- sort(unique(parcelles_complete$Year))
n_years <- length(years_unique)
year_mapping <- data.frame(
  Year = years_unique,
  ID.Year = 1:n_years
)

cat("Years mapped:", min(years_unique), "to", max(years_unique), "(", n_years, "years)\n")

# ============================================================================
# ILR TRANSFORMATION FUNCTIONS
# ============================================================================

create_ilr_basis <- function(D) {
  V <- matrix(0, nrow = D - 1, ncol = D)
  for (i in 1:(D - 1)) {
    V[i, 1:i] <- 1/i
    V[i, i + 1] <- -1
    V[i, ] <- V[i, ] * sqrt(i / (i + 1))
  }
  return(V)
}

ilr_transform <- function(composition_matrix) {
  D <- ncol(composition_matrix)
  V <- create_ilr_basis(D)
  log_comp <- log(composition_matrix)
  ilr_coords <- log_comp %*% t(V)
  return(ilr_coords)
}

ilr_inverse <- function(ilr_matrix) {
  D <- ncol(ilr_matrix) + 1
  V <- create_ilr_basis(D)
  log_comp <- ilr_matrix %*% V
  composition <- exp(log_comp)
  row_sums <- rowSums(composition)
  composition <- composition / row_sums
  return(composition)
}

# ============================================================================
# STEP 5: CREATE 5-FOLD CV SPLITS
# ============================================================================

create_cv_folds <- function(unique_ids, n_folds = 5, seed = 123) {
  set.seed(seed)
  n <- length(unique_ids)
  fold_ids <- sample(rep(1:n_folds, length.out = n))
  
  fold_assignment <- data.frame(
    Final_Cleaned_ID = unique_ids,
    cv_fold = fold_ids
  )
  
  return(fold_assignment)
}

# ============================================================================
# STEP 6: DEFINE IMPUTATION FUNCTIONS WITH CV
# ============================================================================

# Non-compositional imputation with 5-fold CV
impute_single_variable_cv <- function(variable_name, parcelles_data, W_matrix, field_map, year_map, 
                                      cv_folds, n_folds = 5, verbose = TRUE) {
  
  if (verbose) {
    cat("\n", "=", rep("=", 60), "\n", sep="")
    cat("🔬 PROCESSING VARIABLE:", variable_name, "\n")
    cat("=", rep("=", 60), "\n", sep="")
  }
  
  var_values <- parcelles_data[[variable_name]]
  n_total <- length(var_values)
  n_available <- sum(!is.na(var_values))
  n_missing <- n_total - n_available
  pct_available <- round(n_available/n_total*100, 1)
  
  if (verbose) {
    cat("Variable statistics:\n")
    cat("• Total observations:", n_total, "\n")
    cat("• Available (non-NA):", n_available, "(", pct_available, "%)\n")
    cat("• Missing (NA):", n_missing, "(", round(n_missing/n_total*100, 1), "%)\n")
    
    if (n_available > 0) {
      cat("• Min value:", round(min(var_values, na.rm = TRUE), 3), "\n")
      cat("• Max value:", round(max(var_values, na.rm = TRUE), 3), "\n")
      cat("• Mean value:", round(mean(var_values, na.rm = TRUE), 3), "\n")
      cat("• Std deviation:", round(sd(var_values, na.rm = TRUE), 3), "\n")
    }
  }
  
  if (n_available < 20) {
    if (verbose) cat("❌ Insufficient data (< 20 observations). Skipping variable.\n")
    return(NULL)
  }
  
  # Initialize storage for all fold results
  all_fold_predictions <- list()
  fold_metrics <- list()
  
  # Loop through each fold
  for (fold in 1:n_folds) {
    if (verbose) cat("\n  📊 Processing Fold", fold, "/", n_folds, "...\n")
    
    # Assign train/validation based on current fold
    parcelles_data_fold <- parcelles_data %>%
      left_join(cv_folds, by = "Final_Cleaned_ID") %>%
      mutate(
        data_split = case_when(
          is.na(.data[[variable_name]]) ~ "predict",
          cv_fold == fold ~ "validation",
          !is.na(.data[[variable_name]]) ~ "train",
          TRUE ~ "predict"
        )
      )
    
    train_obs <- sum(parcelles_data_fold$data_split == "train")
    val_obs <- sum(parcelles_data_fold$data_split == "validation")
    pred_obs <- sum(parcelles_data_fold$data_split == "predict")
    
    if (verbose) {
      cat("    • Training:", train_obs, "\n")
      cat("    • Validation:", val_obs, "\n")
      cat("    • To predict:", pred_obs, "\n")
    }
    
    # Prepare INLA data
    df_inla <- parcelles_data_fold %>%
      st_drop_geometry() %>%
      left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
      left_join(year_map, by = "Year") %>%
      mutate(
        Y = case_when(
          data_split == "train" ~ .data[[variable_name]],
          TRUE ~ NA_real_
        ),
        Y_true = .data[[variable_name]],
        ID.Field = Sequential_ID,
        ID.Year = as.integer(ID.Year),
        Month = as.integer(ifelse(is.na(Month), 6, Month)),
        Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
        Var = as.factor(ifelse(is.na(Var), "Unknown", as.character(Var))),
        irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
      ) %>%
      filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
      arrange(ID.Field, ID.Year)
    
    if (max(df_inla$ID.Field) > nrow(W_matrix)) {
      if (verbose) cat("    ❌ Spatial structure incompatible. Skipping fold.\n")
      next
    }
    
    # Fit model
    fit_result <- tryCatch({
      prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
      
      inla(
        Y ~ 1 + Bloc + irrg_ty +
          f(ID.Field, model = "besag", graph = W_matrix,
            group = ID.Year, 
            control.group = list(model = "ar1"),
            hyper = list(prec = prec_prior),
            constr = TRUE),
        data = df_inla,
        family = "gaussian",
        control.predictor = list(compute = TRUE, link = 1),
        control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
        control.inla = list(strategy = "adaptive", int.strategy = "eb"),
        verbose = FALSE
      )
    }, error = function(e) {
      tryCatch({
        inla(
          Y ~ 1 + Bloc + f(ID.Field, model = "besag", graph = W_matrix, constr = TRUE),
          data = df_inla,
          family = "gaussian",
          control.predictor = list(compute = TRUE, link = 1),
          control.compute = list(dic = TRUE),
          verbose = FALSE
        )
      }, error = function(e2) NULL)
    })
    
    if (is.null(fit_result) || is.null(fit_result$summary.fitted.values)) {
      if (verbose) cat("    ❌ Model fitting failed for fold", fold, "\n")
      next
    }
    
    # Extract predictions
    predictions <- fit_result$summary.fitted.values
    
    df_predicted <- df_inla %>%
      mutate(
        fold = fold,
        predicted = predictions$mean,
        predicted_sd = predictions$sd,
        predicted_lower = predictions$`0.025quant`,
        predicted_upper = predictions$`0.975quant`
      )
    
    all_fold_predictions[[fold]] <- df_predicted
    
    # Calculate fold-specific validation metrics
    validation_data <- df_predicted %>%
      filter(data_split == "validation", !is.na(Y_true), !is.na(predicted))
    
    if (nrow(validation_data) > 5) {
      pred_vals <- validation_data$predicted
      true_vals <- validation_data$Y_true
      
      ss_res <- sum((true_vals - pred_vals)^2)
      ss_tot <- sum((true_vals - mean(true_vals))^2)
      r_squared <- 1 - (ss_res / ss_tot)
      
      fold_metrics[[fold]] <- list(
        fold = fold,
        r_squared = r_squared,
        correlation = cor(true_vals, pred_vals),
        rmse = sqrt(mean((true_vals - pred_vals)^2)),
        mae = mean(abs(true_vals - pred_vals)),
        n_validation = nrow(validation_data)
      )
      
      if (verbose) {
        cat("    ✅ Fold", fold, "metrics: R²=", round(r_squared, 4), 
            " RMSE=", round(fold_metrics[[fold]]$rmse, 4), "\n")
      }
    }
  }
  
  # Combine all fold predictions
  if (length(all_fold_predictions) == 0) {
    if (verbose) cat("❌ No successful folds\n")
    return(NULL)
  }
  
  combined_predictions <- bind_rows(all_fold_predictions)
  
  # Calculate aggregated CV metrics
  all_validation <- combined_predictions %>%
    filter(data_split == "validation", !is.na(Y_true), !is.na(predicted))
  
  cv_metrics <- list(variable = variable_name)
  
  if (nrow(all_validation) > 5) {
    pred_vals <- all_validation$predicted
    true_vals <- all_validation$Y_true
    
    ss_res <- sum((true_vals - pred_vals)^2)
    ss_tot <- sum((true_vals - mean(true_vals))^2)
    
    cv_metrics$cv_r_squared = 1 - (ss_res / ss_tot)
    cv_metrics$cv_correlation = cor(true_vals, pred_vals)
    cv_metrics$cv_rmse = sqrt(mean((true_vals - pred_vals)^2))
    cv_metrics$cv_mae = mean(abs(true_vals - pred_vals))
    cv_metrics$n_total_validation = nrow(all_validation)
    
    # Calculate per-fold statistics
    fold_r2_values <- sapply(fold_metrics, function(x) x$r_squared)
    cv_metrics$mean_fold_r_squared = mean(fold_r2_values, na.rm = TRUE)
    cv_metrics$sd_fold_r_squared = sd(fold_r2_values, na.rm = TRUE)
    
    if (verbose) {
      cat("\n  🎯 Cross-Validation Results (Aggregated):\n")
      cat("    • Total validation obs:", cv_metrics$n_total_validation, "\n")
      cat("    • CV R²:", round(cv_metrics$cv_r_squared, 4), "\n")
      cat("    • Mean fold R² (±SD):", round(cv_metrics$mean_fold_r_squared, 4), 
          "±", round(cv_metrics$sd_fold_r_squared, 4), "\n")
      cat("    • CV RMSE:", round(cv_metrics$cv_rmse, 4), "\n")
      cat("    • CV MAE:", round(cv_metrics$cv_mae, 4), "\n")
    }
  }
  
  # Fit final model on all data for predictions
  if (verbose) cat("\n  🚀 Fitting final model on all data...\n")
  
  parcelles_data_final <- parcelles_data %>%
    mutate(data_split = ifelse(is.na(.data[[variable_name]]), "predict", "train"))
  
  df_inla_final <- parcelles_data_final %>%
    st_drop_geometry() %>%
    left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
    left_join(year_map, by = "Year") %>%
    mutate(
      Y = case_when(
        data_split == "train" ~ .data[[variable_name]],
        TRUE ~ NA_real_
      ),
      ID.Field = Sequential_ID,
      ID.Year = as.integer(ID.Year),
      Month = as.integer(ifelse(is.na(Month), 6, Month)),
      Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
      Var = as.factor(ifelse(is.na(Var), "Unknown", as.character(Var))),
      irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
    ) %>%
    filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
    arrange(ID.Field, ID.Year)
  
  fit_final <- tryCatch({
    prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
    
    inla(
      Y ~ 1 + Bloc + irrg_ty +
        f(ID.Field, model = "besag", graph = W_matrix,
          group = ID.Year, 
          control.group = list(model = "ar1"),
          hyper = list(prec = prec_prior),
          constr = TRUE),
      data = df_inla_final,
      family = "gaussian",
      control.predictor = list(compute = TRUE, link = 1),
      control.compute = list(dic = TRUE, waic = TRUE),
      control.inla = list(strategy = "adaptive", int.strategy = "eb"),
      verbose = FALSE
    )
  }, error = function(e) NULL)
  
  if (is.null(fit_final)) {
    if (verbose) cat("  ❌ Final model fitting failed\n")
    return(NULL)
  }
  
  predictions_final <- fit_final$summary.fitted.values
  
  df_predicted_final <- df_inla_final %>%
    mutate(
      !!paste0(variable_name, "_predicted") := predictions_final$mean,
      !!paste0(variable_name, "_predicted_sd") := predictions_final$sd,
      !!paste0(variable_name, "_predicted_lower") := predictions_final$`0.025quant`,
      !!paste0(variable_name, "_predicted_upper") := predictions_final$`0.975quant`,
      !!paste0(variable_name, "_complete") := case_when(
        data_split == "train" ~ Y,
        TRUE ~ predictions_final$mean
      ),
      !!paste0(variable_name, "_prediction_type") := ifelse(data_split == "train", "Observed", "Predicted")
    )
  
  result <- list(
    variable_name = variable_name,
    model_fit = fit_final,
    predictions = df_predicted_final,
    cv_predictions = combined_predictions,
    cv_metrics = cv_metrics,
    fold_metrics = fold_metrics,
    n_total = n_total,
    n_available = n_available,
    n_missing = n_missing,
    pct_available = pct_available
  )
  
  return(result)
}

# Compositional imputation with 5-fold CV
impute_compositional_granulometry_cv <- function(parcelles_data, W_matrix, field_map, year_map, 
                                                 cv_folds, n_folds = 5, verbose = TRUE) {
  
  if (verbose) {
    cat("\n", "=", rep("=", 60), "\n", sep="")
    cat("🔬 PROCESSING COMPOSITIONAL GRANULOMETRY (ILR) WITH 5-FOLD CV\n")
    cat("=", rep("=", 60), "\n", sep="")
  }
  
  parcelles_data <- parcelles_data %>%
    mutate(obs_id = row_number())
  
  sampled_obs <- parcelles_data %>%
    filter(has_complete_granulo)
  
  unsampled_obs <- parcelles_data %>%
    filter(!has_complete_granulo)
  
  n_sampled <- nrow(sampled_obs)
  n_unsampled <- nrow(unsampled_obs)
  
  if (verbose) {
    cat("• Sampled observations:", n_sampled, "\n")
    cat("• Unsampled observations:", n_unsampled, "\n")
  }
  
  if (n_sampled < 30) {
    if (verbose) cat("❌ Insufficient sampled observations\n")
    return(NULL)
  }
  
  # Prepare compositional data and ILR transform
  comp_matrix <- sampled_obs %>%
    st_drop_geometry() %>%
    dplyr::select(all_of(GRANULO_FRACTIONS)) %>%
    as.matrix()
  
  comp_matrix[comp_matrix <= 0] <- 0.01
  row_sums <- rowSums(comp_matrix)
  comp_matrix <- comp_matrix / row_sums * 100
  
  ilr_coords <- ilr_transform(comp_matrix / 100)
  colnames(ilr_coords) <- paste0("ILR_", 1:4)
  
  if (verbose) {
    cat("✅ ILR transformation: 5 fractions -> 4 orthonormal coordinates\n")
  }
  
  sampled_obs_ilr <- sampled_obs %>%
    bind_cols(as.data.frame(ilr_coords))
  
  # Process each ILR coordinate with CV
  ilr_cv_results <- list()
  
  for (i in 1:4) {
    ilr_var <- paste0("ILR_", i)
    
    if (verbose) cat("\n  🔄 Processing", ilr_var, "with 5-fold CV...\n")
    
    all_fold_predictions <- list()
    fold_metrics <- list()
    
    # CV loop
    for (fold in 1:n_folds) {
      if (verbose) cat("    Fold", fold, "/", n_folds, "... ")
      
      sampled_obs_fold <- sampled_obs_ilr %>%
        left_join(cv_folds, by = "Final_Cleaned_ID") %>%
        mutate(
          data_split = case_when(
            cv_fold == fold ~ "validation",
            TRUE ~ "train"
          )
        )
      
      all_obs_fold <- bind_rows(
        sampled_obs_fold %>% mutate(obs_type = "sampled"),
        unsampled_obs %>% mutate(obs_type = "unsampled", data_split = "predict")
      )
      
      df_inla <- all_obs_fold %>%
        st_drop_geometry() %>%
        left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
        left_join(year_map, by = "Year") %>%
        mutate(
          Y = case_when(
            data_split == "train" ~ .data[[ilr_var]],
            TRUE ~ NA_real_
          ),
          Y_true = .data[[ilr_var]],
          ID.Field = Sequential_ID,
          ID.Year = as.integer(ID.Year),
          Month = as.integer(ifelse(is.na(Month), 6, Month)),
          Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
          irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
        ) %>%
        filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
        arrange(ID.Field, ID.Year)
      
      fit_result <- tryCatch({
        prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
        
        inla(
          Y ~ 1 + Bloc + irrg_ty +
            f(ID.Field, model = "besag", graph = W_matrix,
              group = ID.Year, 
              control.group = list(model = "ar1"),
              hyper = list(prec = prec_prior),
              constr = TRUE),
          data = df_inla,
          family = "gaussian",
          control.predictor = list(compute = TRUE, link = 1),
          control.compute = list(dic = TRUE, waic = TRUE),
          control.inla = list(strategy = "adaptive", int.strategy = "eb"),
          verbose = FALSE
        )
      }, error = function(e) NULL)
      
      if (!is.null(fit_result)) {
        predictions <- fit_result$summary.fitted.values
        
        df_predicted <- df_inla %>%
          mutate(
            fold = fold,
            predicted = predictions$mean,
            predicted_sd = predictions$sd
          )
        
        all_fold_predictions[[fold]] <- df_predicted
        
        # Validation metrics
        val_data <- df_predicted %>%
          filter(data_split == "validation", !is.na(Y_true))
        
        if (nrow(val_data) > 5) {
          pred_vals <- val_data$predicted
          true_vals <- val_data$Y_true
          
          ss_res <- sum((true_vals - pred_vals)^2)
          ss_tot <- sum((true_vals - mean(true_vals))^2)
          
          fold_metrics[[fold]] <- list(
            fold = fold,
            r_squared = 1 - (ss_res / ss_tot),
            rmse = sqrt(mean((true_vals - pred_vals)^2)),
            n_val = nrow(val_data)
          )
          
          if (verbose) cat("R²=", round(fold_metrics[[fold]]$r_squared, 3), "\n")
        } else {
          if (verbose) cat("insufficient data\n")
        }
      } else {
        if (verbose) cat("failed\n")
      }
    }
    
    # Aggregate CV results for this ILR coordinate
    if (length(all_fold_predictions) > 0) {
      combined_predictions <- bind_rows(all_fold_predictions)
      
      all_validation <- combined_predictions %>%
        filter(data_split == "validation", !is.na(Y_true))
      
      cv_metrics_ilr <- list()
      if (nrow(all_validation) > 5) {
        pred_vals <- all_validation$predicted
        true_vals <- all_validation$Y_true
        
        ss_res <- sum((true_vals - pred_vals)^2)
        ss_tot <- sum((true_vals - mean(true_vals))^2)
        
        cv_metrics_ilr$cv_r_squared = 1 - (ss_res / ss_tot)
        cv_metrics_ilr$cv_rmse = sqrt(mean((true_vals - pred_vals)^2))
        cv_metrics_ilr$n_val = nrow(all_validation)
        
        fold_r2 <- sapply(fold_metrics, function(x) x$r_squared)
        cv_metrics_ilr$mean_fold_r_squared = mean(fold_r2, na.rm = TRUE)
        cv_metrics_ilr$sd_fold_r_squared = sd(fold_r2, na.rm = TRUE)
      }
      
      # Fit final model on all data
      all_obs_final <- bind_rows(
        sampled_obs_ilr %>% mutate(obs_type = "sampled", data_split = "train"),
        unsampled_obs %>% mutate(obs_type = "unsampled", data_split = "predict")
      )
      
      df_inla_final <- all_obs_final %>%
        st_drop_geometry() %>%
        left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
        left_join(year_map, by = "Year") %>%
        mutate(
          Y = case_when(
            data_split == "train" ~ .data[[ilr_var]],
            TRUE ~ NA_real_
          ),
          ID.Field = Sequential_ID,
          ID.Year = as.integer(ID.Year),
          Month = as.integer(ifelse(is.na(Month), 6, Month)),
          Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
          irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
        ) %>%
        filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
        arrange(ID.Field, ID.Year)
      
      fit_final <- tryCatch({
        prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
        
        inla(
          Y ~ 1 + Bloc + irrg_ty +
            f(ID.Field, model = "besag", graph = W_matrix,
              group = ID.Year, 
              control.group = list(model = "ar1"),
              hyper = list(prec = prec_prior),
              constr = TRUE),
          data = df_inla_final,
          family = "gaussian",
          control.predictor = list(compute = TRUE, link = 1),
          control.compute = list(dic = TRUE),
          verbose = FALSE
        )
      }, error = function(e) NULL)
      
      if (!is.null(fit_final)) {
        predictions_final <- fit_final$summary.fitted.values
        
        df_predicted_final <- df_inla_final %>%
          mutate(
            !!paste0(ilr_var, "_predicted") := predictions_final$mean,
            !!paste0(ilr_var, "_predicted_sd") := predictions_final$sd
          )
        
        ilr_cv_results[[ilr_var]] <- list(
          model = fit_final,
          predictions = df_predicted_final,
          cv_predictions = combined_predictions,
          cv_metrics = cv_metrics_ilr,
          fold_metrics = fold_metrics
        )
        
        if (verbose && !is.null(cv_metrics_ilr$cv_r_squared)) {
          cat("    ✅ CV R²:", round(cv_metrics_ilr$cv_r_squared, 4), 
              "± SD:", round(cv_metrics_ilr$sd_fold_r_squared, 4), "\n")
        }
      }
    }
  }
  
  if (length(ilr_cv_results) != 4) {
    if (verbose) cat("❌ Not all ILR coordinates fitted successfully\n")
    return(NULL)
  }
  
  # Transform back to composition
  if (verbose) cat("\n🔄 Transforming back to composition...\n")
  
  all_predictions <- ilr_cv_results[[1]]$predictions %>%
    dplyr::select(obs_id, Final_Cleaned_ID, Year, data_split, obs_type,
                  ILR_1_predicted, ILR_1_predicted_sd)
  
  for (i in 2:4) {
    ilr_var <- paste0("ILR_", i)
    pred_cols <- ilr_cv_results[[ilr_var]]$predictions %>%
      dplyr::select(obs_id, !!paste0(ilr_var, "_predicted"), !!paste0(ilr_var, "_predicted_sd"))
    
    all_predictions <- all_predictions %>%
      left_join(pred_cols, by = "obs_id")
  }
  
  ilr_pred_matrix <- all_predictions %>%
    dplyr::select(ILR_1_predicted, ILR_2_predicted, ILR_3_predicted, ILR_4_predicted) %>%
    as.matrix()
  
  comp_predicted <- ilr_inverse(ilr_pred_matrix)
  comp_predicted_pct <- comp_predicted * 100
  colnames(comp_predicted_pct) <- GRANULO_FRACTIONS
  
  if (verbose) {
    cat("✅ ILR back-transformation complete\n")
    sums_check <- rowSums(comp_predicted_pct)
    cat("• Sum range:", round(min(sums_check), 2), "to", round(max(sums_check), 2), "%\n")
  }
  
  final_results <- all_predictions %>%
    bind_cols(as.data.frame(comp_predicted_pct) %>%
                rename_with(~paste0(., "_predicted"), all_of(GRANULO_FRACTIONS)))
  
  for (frac in GRANULO_FRACTIONS) {
    original_map <- bind_rows(sampled_obs_ilr, unsampled_obs) %>%
      st_drop_geometry() %>%
      dplyr::select(obs_id, !!frac) %>%
      rename(original_value = !!frac)
    
    final_results <- final_results %>%
      left_join(original_map, by = "obs_id")
    
    pred_col <- paste0(frac, "_predicted")
    
    final_results[[paste0(frac, "_complete")]] <- ifelse(
      final_results$data_split == "train",
      final_results$original_value,
      final_results[[pred_col]]
    )
    
    final_results[[paste0(frac, "_prediction_type")]] <- final_results$data_split
    
    final_results <- final_results %>% dplyr::select(-original_value)
  }
  
  # Calculate CV metrics for fractions
  cv_frac_metrics <- list()
  
  # Get CV predictions for all ILR coordinates
  all_ilr_cv <- ilr_cv_results[[1]]$cv_predictions %>%
    dplyr::select(obs_id, Final_Cleaned_ID, Year, fold, data_split, ILR_1 = predicted)
  
  for (i in 2:4) {
    ilr_var <- paste0("ILR_", i)
    cv_cols <- ilr_cv_results[[ilr_var]]$cv_predictions %>%
      dplyr::select(obs_id, fold, !!paste0("ILR_", i) := predicted)
    
    all_ilr_cv <- all_ilr_cv %>%
      left_join(cv_cols, by = c("obs_id", "fold"))
  }
  
  # Transform CV predictions back to composition
  ilr_cv_matrix <- all_ilr_cv %>%
    dplyr::select(ILR_1, ILR_2, ILR_3, ILR_4) %>%
    as.matrix()
  
  comp_cv_predicted <- ilr_inverse(ilr_cv_matrix)
  comp_cv_predicted_pct <- comp_cv_predicted * 100
  colnames(comp_cv_predicted_pct) <- GRANULO_FRACTIONS
  
  cv_predictions_fractions <- all_ilr_cv %>%
    bind_cols(as.data.frame(comp_cv_predicted_pct))
  
  # Calculate metrics for each fraction
  for (frac in GRANULO_FRACTIONS) {
    true_map <- sampled_obs_ilr %>%
      st_drop_geometry() %>%
      dplyr::select(obs_id, !!frac) %>%
      rename(true_val = !!frac)
    
    cv_comparison <- cv_predictions_fractions %>%
      filter(data_split == "validation") %>%
      left_join(true_map, by = "obs_id")
    
    if (sum(!is.na(cv_comparison$true_val)) > 5) {
      true_vals <- cv_comparison$true_val
      pred_vals <- cv_comparison[[frac]]
      
      valid_idx <- !is.na(true_vals) & !is.na(pred_vals)
      true_vals <- true_vals[valid_idx]
      pred_vals <- pred_vals[valid_idx]
      
      if (length(true_vals) > 0) {
        ss_res <- sum((true_vals - pred_vals)^2)
        ss_tot <- sum((true_vals - mean(true_vals))^2)
        
        cv_frac_metrics[[frac]] <- list(
          cv_r_squared = 1 - (ss_res / ss_tot),
          cv_rmse = sqrt(mean((true_vals - pred_vals)^2)),
          cv_mae = mean(abs(true_vals - pred_vals)),
          n = length(true_vals)
        )
      }
    }
  }
  
  overall_metrics <- list(
    fraction_metrics = cv_frac_metrics,
    n_sampled = n_sampled,
    n_unsampled = n_unsampled
  )
  
  if (verbose) {
    cat("\n🎯 Cross-Validation Results (Back-transformed Fractions):\n")
    for (frac in GRANULO_FRACTIONS) {
      if (!is.null(cv_frac_metrics[[frac]])) {
        cat("• ", frac, ": CV R²=", round(cv_frac_metrics[[frac]]$cv_r_squared, 4),
            " RMSE=", round(cv_frac_metrics[[frac]]$cv_rmse, 2), "% (n=", 
            cv_frac_metrics[[frac]]$n, ")\n", sep="")
      }
    }
  }
  
  return(list(
    predictions = final_results,
    cv_predictions = cv_predictions_fractions,
    ilr_results = ilr_cv_results,
    metrics = overall_metrics,
    n_sampled = n_sampled,
    n_unsampled = n_unsampled
  ))
}

# ============================================================================
# STEP 7: CREATE CV FOLDS
# ============================================================================
cat("\n🎲 CREATING 5-FOLD CROSS-VALIDATION SPLITS:\n")
cat("=", rep("=", 45), "\n", sep="")

# Create CV folds based on unique parcels
unique_parcels_with_data <- parcelles_complete %>%
  st_drop_geometry() %>%
  filter(!is.na(pH) | has_complete_granulo) %>%
  dplyr::select(Final_Cleaned_ID) %>%
  distinct() %>%
  pull(Final_Cleaned_ID)

cv_fold_assignment <- create_cv_folds(unique_parcels_with_data, n_folds = 5, seed = 123)

cat("• Total unique parcels with data:", length(unique_parcels_with_data), "\n")
cat("• Parcels per fold:\n")
for (fold in 1:5) {
  n_parcels <- sum(cv_fold_assignment$cv_fold == fold)
  cat("  Fold", fold, ":", n_parcels, "parcels\n")
}

# ============================================================================
# STEP 8: PROCESS ALL VARIABLES WITH CV
# ============================================================================
cat("\n🔄 PROCESSING ALL VARIABLES WITH 5-FOLD CV:\n")
cat("=", rep("=", 50), "\n", sep="")

all_results <- list()
all_metrics <- data.frame()
failed_variables <- character(0)

# Process compositional granulometry
cat("\n[COMPOSITIONAL] Processing granulometry fractions with CV...\n")

granulo_result <- impute_compositional_granulometry_cv(
  parcelles_data = parcelles_complete,
  W_matrix = W.P1_complete,
  field_map = field_mapping,
  year_map = year_mapping,
  cv_folds = cv_fold_assignment,
  n_folds = 5,
  verbose = TRUE
)

if (!is.null(granulo_result)) {
  all_results[["Granulometry_Compositional"]] <- granulo_result
  
  if (!is.null(granulo_result$metrics$fraction_metrics)) {
    for (frac in GRANULO_FRACTIONS) {
      if (!is.null(granulo_result$metrics$fraction_metrics[[frac]])) {
        metrics_row <- data.frame(
          Variable = frac,
          N_Total = nrow(parcelles_complete),
          N_Available = granulo_result$n_sampled,
          N_Missing = granulo_result$n_unsampled,
          Pct_Available = round(granulo_result$n_sampled / nrow(parcelles_complete) * 100, 1),
          CV_R_Squared = granulo_result$metrics$fraction_metrics[[frac]]$cv_r_squared,
          CV_RMSE = granulo_result$metrics$fraction_metrics[[frac]]$cv_rmse,
          CV_MAE = granulo_result$metrics$fraction_metrics[[frac]]$cv_mae,
          N_CV_Validation = granulo_result$metrics$fraction_metrics[[frac]]$n,
          stringsAsFactors = FALSE
        )
        
        all_metrics <- rbind(all_metrics, metrics_row)
      }
    }
  }
  cat("✅ SUCCESS for compositional granulometry\n")
} else {
  failed_variables <- c(failed_variables, GRANULO_FRACTIONS)
  cat("❌ FAILED for compositional granulometry\n")
}

# Process non-compositional variables
for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  
  cat("\n[", i, "/", length(NON_COMPOSITIONAL_VARIABLES), "] Processing:", var_name, "\n")
  
  result <- impute_single_variable_cv(
    variable_name = var_name,
    parcelles_data = parcelles_complete,
    W_matrix = W.P1_complete,
    field_map = field_mapping,
    year_map = year_mapping,
    cv_folds = cv_fold_assignment,
    n_folds = 5,
    verbose = TRUE
  )
  
  if (!is.null(result)) {
    all_results[[var_name]] <- result
    
    metrics_row <- data.frame(
      Variable = var_name,
      N_Total = result$n_total,
      N_Available = result$n_available,
      N_Missing = result$n_missing,
      Pct_Available = result$pct_available,
      CV_R_Squared = ifelse(is.null(result$cv_metrics$cv_r_squared), NA, result$cv_metrics$cv_r_squared),
      CV_RMSE = ifelse(is.null(result$cv_metrics$cv_rmse), NA, result$cv_metrics$cv_rmse),
      CV_MAE = ifelse(is.null(result$cv_metrics$cv_mae), NA, result$cv_metrics$cv_mae),
      N_CV_Validation = ifelse(is.null(result$cv_metrics$n_total_validation), 0, result$cv_metrics$n_total_validation),
      Mean_Fold_R_Squared = ifelse(is.null(result$cv_metrics$mean_fold_r_squared), NA, result$cv_metrics$mean_fold_r_squared),
      SD_Fold_R_Squared = ifelse(is.null(result$cv_metrics$sd_fold_r_squared), NA, result$cv_metrics$sd_fold_r_squared),
      stringsAsFactors = FALSE
    )
    
    all_metrics <- rbind(all_metrics, metrics_row)
    cat("✅ SUCCESS for", var_name, "\n")
  } else {
    failed_variables <- c(failed_variables, var_name)
    cat("❌ FAILED for", var_name, "\n")
  }
}

# ============================================================================
# STEP 9: COMBINE ALL PREDICTIONS
# ============================================================================
cat("\n🔗 COMBINING ALL PREDICTIONS:\n")
cat("=", rep("=", 50), "\n", sep="")

successful_variables <- names(all_results)
cat("Successfully processed:", length(successful_variables), "variable groups\n")

final_dataset <- parcelles_complete %>% 
  st_drop_geometry() %>%
  group_by(Final_Cleaned_ID, Year) %>%
  mutate(
    replicate_id = row_number(),
    unique_key = paste0(Final_Cleaned_ID, "_", Year, "_rep", replicate_id)
  ) %>%
  ungroup() %>%
  mutate(original_row_id = row_number())

final_dataset <- parcelles_complete %>%
  mutate(original_row_id = row_number()) %>%
  left_join(
    final_dataset %>% dplyr::select(original_row_id, replicate_id, unique_key),
    by = "original_row_id"
  )

# Add CV fold assignment
final_dataset <- final_dataset %>%
  left_join(cv_fold_assignment, by = "Final_Cleaned_ID")

# Add granulometry predictions
if ("Granulometry_Compositional" %in% names(all_results)) {
  cat("\n[COMPOSITIONAL] Adding granulometry predictions...\n")
  
  final_dataset <- final_dataset %>%
    mutate(obs_id = row_number())
  
  granulo_preds <- all_results[["Granulometry_Compositional"]]$predictions
  
  final_dataset <- final_dataset %>%
    left_join(
      granulo_preds %>% 
        dplyr::select(obs_id, matches("^Clay_|^Fine_Silt_|^Coarse_Silt_|^Fine_Sand_|^Coarse_Sand_|^ILR_")),
      by = "obs_id"
    )
  
  final_dataset <- final_dataset %>% dplyr::select(-obs_id)
  cat("  ✅ Compositional predictions added\n")
}

# Add non-compositional predictions
for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  
  if (var_name %in% names(all_results)) {
    cat("\n[", i, "/", length(NON_COMPOSITIONAL_VARIABLES), "] Adding:", var_name, "\n")
    
    result <- all_results[[var_name]]
    
    df_inla_reference <- parcelles_complete %>% 
      st_drop_geometry() %>% 
      left_join(field_mapping, by = c("Final_Cleaned_ID" = "Original_ID"), 
                relationship = "many-to-one") %>%  
      left_join(year_mapping, by = "Year", 
                relationship = "many-to-one") %>%  
      filter(!is.na(Sequential_ID), !is.na(ID.Year)) %>% 
      group_by(Final_Cleaned_ID, Year) %>%
      mutate(
        replicate_id = row_number(),
        unique_key = paste0(Final_Cleaned_ID, "_", Year, "_rep", replicate_id)
      ) %>%
      ungroup() %>%
      mutate(original_row_id = row_number()) %>% 
      arrange(Sequential_ID, ID.Year, replicate_id) %>%
      mutate(inla_row_id = row_number())
    
    if (nrow(df_inla_reference) == nrow(result$predictions)) {
      predictions_aligned <- df_inla_reference %>% 
        dplyr::select(Final_Cleaned_ID, Year, replicate_id, unique_key, original_row_id) %>% 
        bind_cols(
          result$predictions %>% 
            dplyr::select(matches(paste0("^", var_name, "_")))
        )
      
      final_dataset <- final_dataset %>% 
        left_join(
          predictions_aligned %>% dplyr::select(-Final_Cleaned_ID, -Year, -replicate_id, -original_row_id),
          by = "unique_key"
        )
      
      first_pred_col <- paste0(var_name, "_predicted")
      if (first_pred_col %in% colnames(final_dataset)) {
        non_null_count <- sum(!is.na(final_dataset[[first_pred_col]]))
        cat("  ✅ Successfully added", non_null_count, "predictions for", var_name, "\n")
      }
    }
    
    gc()
  }
}

final_dataset <- final_dataset %>% 
  dplyr::select(-original_row_id)

cat("\n✅ Final combined dataset created\n")
cat("• Total observations:", nrow(final_dataset), "\n")
cat("• Total columns:", ncol(final_dataset), "\n")

gc()

# ============================================================================
# STEP 10: SAVE COMPREHENSIVE OUTPUTS
# ============================================================================
cat("\n💾 SAVING COMPREHENSIVE OUTPUTS:\n")
cat("=", rep("=", 35), "\n", sep="")

base_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/"

gpkg_output <- paste0(base_path, "parcels_ALL_VARIABLES_CV_predictions.gpkg")
st_write(final_dataset, gpkg_output, delete_dsn = TRUE, quiet = TRUE)
cat("✅ Complete dataset saved:", basename(gpkg_output), "\n")

csv_output <- paste0(base_path, "parcels_ALL_VARIABLES_CV_attributes.csv")
st_drop_geometry(final_dataset) %>% write_csv(csv_output)
cat("✅ Attributes saved:", basename(csv_output), "\n")

metrics_output <- paste0(base_path, "ALL_VARIABLES_CV_metrics_summary.csv")
write_csv(all_metrics, metrics_output)
cat("✅ CV metrics summary saved:", basename(metrics_output), "\n")

# Save CV fold assignment
cv_folds_output <- paste0(base_path, "CV_fold_assignment.csv")
write_csv(cv_fold_assignment, cv_folds_output)
cat("✅ CV fold assignment saved:", basename(cv_folds_output), "\n")
```

## I.2. MULTI-VARIABLE ENHANCED SPATIO-TEMPORAL IMPUTATION WITH 5-FOLD CROSS-VALIDATION \[PARALLEL COMPUTING VERSION\]

```{r}
# ============================================================================
# MULTI-VARIABLE ENHANCED SPATIO-TEMPORAL IMPUTATION WITH 5-FOLD CROSS-VALIDATION
# PARALLEL COMPUTING VERSION
# Compositional treatment for granulometry fractions
# ============================================================================

library(sf)
library(INLA)
library(dplyr)
library(ggplot2)
library(viridis)
library(spdep)
library(readr)
library(purrr)
library(tibble)
library(foreach)
library(doParallel)
library(parallel)

cat("========================================\n")
cat("  MULTI-VARIABLE SPATIO-TEMPORAL       \n")
cat("  IMPUTATION WITH 5-FOLD CV            \n")
cat("  PARALLEL COMPUTING ENABLED           \n")
cat("========================================\n")

# ============================================================================
# STEP 0: SETUP PARALLEL COMPUTING
# ============================================================================
cat("\n🚀 SETTING UP PARALLEL COMPUTING:\n")
cat("=", rep("=", 40), "\n", sep="")

# Detect number of cores
n_cores_available <- detectCores()
# Use 75% of available cores (leave some for system)
n_cores_to_use <- max(1, floor(n_cores_available * 0.75))

cat("• Available CPU cores:", n_cores_available, "\n")
cat("• Cores to use for parallel processing:", n_cores_to_use, "\n")

# Register parallel backend
cl <- makeCluster(n_cores_to_use)
registerDoParallel(cl)

cat("✅ Parallel backend registered\n")

# ============================================================================
# STEP 1: LOAD THE COMPLETE DATASET AND IDENTIFY SOIL VARIABLES
# ============================================================================
cat("\n📂 LOADING COMPLETE DATASET AND IDENTIFYING SOIL VARIABLES:\n")
cat("=", rep("=", 60), "\n", sep="")

gpkg_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/parcels_with_soil_complete.gpkg"

if (file.exists(gpkg_path)) {
  parcelles_complete <- st_read(gpkg_path, quiet = TRUE)
  cat("✅ Complete dataset loaded:", nrow(parcelles_complete), "observations\n")
} else {
  cat("❌ Complete dataset not found. Please run the enhanced join script first.\n")
  stopCluster(cl)
  stop("Dataset not available")
}

# Define compositional and non-compositional variables
GRANULO_FRACTIONS <- c(
  "Clay_percent",
  "Fine_Silt_percent",
  "Coarse_Silt_percent",
  "Fine_Sand_percent",
  "Coarse_Sand_percent"
)

NON_COMPOSITIONAL_VARIABLES <- c(
  "pH",
  "CN_Ratio", 
  "Electrical_Conductivity_mmho_cm",
  "Exchangeable_Calcium_meq_100g",
  "Exchangeable_Magnesium_meq_100g", 
  "Exchangeable_Sodium_meq_100g",
  "Exchangeable_Potassium_meq_100g",
  "Available_P2O5_mg_kg",
  "Organic_Carbon_per_mille",
  "Total_Nitrogen_per_mille",
  "Total_Silt_percent",
  "Total_Sand_percent", 
  "Organic_Matter_percent",
  "Sum_of_Exchangeable_Bases_meq_100g",
  "CaO_g_100g",
  "Mg_Ca_Ratio",
  "Mg_K_Ratio",
  "Compaction_Index"
)

cat("Compositional variables (granulometry):", length(GRANULO_FRACTIONS), "\n")
cat("Non-compositional variables:", length(NON_COMPOSITIONAL_VARIABLES), "\n")

for (i in seq_along(GRANULO_FRACTIONS)) {
  var_name <- GRANULO_FRACTIONS[i]
  n_available <- sum(!is.na(parcelles_complete[[var_name]]))
  pct_available <- round(n_available/nrow(parcelles_complete)*100, 1)
  cat(sprintf("• %2d. %-35s: %4d available (%5.1f%%), %4d missing\n", 
              i, var_name, n_available, pct_available, sum(is.na(parcelles_complete[[var_name]]))))
}

for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  n_available <- sum(!is.na(parcelles_complete[[var_name]]))
  pct_available <- round(n_available/nrow(parcelles_complete)*100, 1)
  cat(sprintf("• %2d. %-35s: %4d available (%5.1f%%), %4d missing\n", 
              length(GRANULO_FRACTIONS) + i, var_name, n_available, pct_available, sum(is.na(parcelles_complete[[var_name]]))))
}

cat("\nDataset structure:\n")
cat("• Total observations:", nrow(parcelles_complete), "\n")
cat("• Unique parcels:", length(unique(parcelles_complete$Final_Cleaned_ID)), "\n")
cat("• Year range:", min(parcelles_complete$Year, na.rm = TRUE), "-", max(parcelles_complete$Year, na.rm = TRUE), "\n")

# Validate compositional data
parcelles_complete <- parcelles_complete %>%
  mutate(
    has_complete_granulo = !is.na(Clay_percent) & 
                          !is.na(Fine_Silt_percent) & 
                          !is.na(Coarse_Silt_percent) & 
                          !is.na(Fine_Sand_percent) & 
                          !is.na(Coarse_Sand_percent),
    granulo_sum = Clay_percent + Fine_Silt_percent + Coarse_Silt_percent + 
                  Fine_Sand_percent + Coarse_Sand_percent
  )

complete_granulo <- sum(parcelles_complete$has_complete_granulo, na.rm = TRUE)
cat("\n🔍 COMPOSITIONAL DATA VALIDATION:\n")
cat("• Observations with complete granulometry:", complete_granulo, "\n")

if(complete_granulo > 0) {
  sum_stats <- parcelles_complete %>%
    filter(has_complete_granulo) %>%
    pull(granulo_sum) %>%
    summary()
  cat("• Sum range:", round(sum_stats[1], 2), "to", round(sum_stats[6], 2), "%\n")
}

# ============================================================================
# STEP 2: PREPARE SHARED SPATIAL STRUCTURE
# ============================================================================
cat("\n🗺️ PREPARING SHARED SPATIAL STRUCTURE:\n")
cat("=", rep("=", 40), "\n", sep="")

unique_parcels_spatial <- parcelles_complete %>%
  group_by(Final_Cleaned_ID) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(Final_Cleaned_ID)

cat("Unique parcels for spatial analysis:", nrow(unique_parcels_spatial), "\n")

unique_parcels_spatial <- st_make_valid(unique_parcels_spatial)
cat("Geometries validated\n")

field_mapping <- data.frame(
  Original_ID = unique_parcels_spatial$Final_Cleaned_ID,
  Sequential_ID = 1:nrow(unique_parcels_spatial)
)

cat("Field mapping created: 1 to", nrow(field_mapping), "\n")

unique_parcels_spatial <- unique_parcels_spatial %>%
  left_join(field_mapping, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
  arrange(Sequential_ID)

# ============================================================================
# STEP 3: CREATE SPATIAL WEIGHTS MATRIX W.P1
# ============================================================================
cat("\n🔗 CREATING SHARED SPATIAL WEIGHTS MATRIX:\n")
cat("=", rep("=", 42), "\n", sep="")

cat("Computing boundary-to-boundary distances (this may take time)...\n")

boundaries <- st_boundary(unique_parcels_spatial)
dist_matrix <- st_distance(boundaries, boundaries)
dist_matrix <- units::drop_units(dist_matrix)
diag(dist_matrix) <- NA

cat("Distance matrix calculated:", nrow(dist_matrix), "x", ncol(dist_matrix), "\n")

DISTANCE_THRESHOLD <- 110
cat("Using distance threshold:", DISTANCE_THRESHOLD, "meters\n")

W.P1_complete <- matrix(0, nrow = nrow(unique_parcels_spatial), ncol = nrow(unique_parcels_spatial))
W.P1_complete[dist_matrix <= DISTANCE_THRESHOLD & !is.na(dist_matrix)] <- 1
diag(W.P1_complete) <- 0
rownames(W.P1_complete) <- colnames(W.P1_complete) <- 1:nrow(W.P1_complete)

n_connections <- sum(W.P1_complete)
avg_neighbors <- n_connections / nrow(unique_parcels_spatial)

cat("✅ W.P1 matrix created:\n")
cat("• Dimensions:", nrow(W.P1_complete), "x", ncol(W.P1_complete), "\n")
cat("• Total connections:", n_connections, "\n")
cat("• Average neighbors per parcel:", round(avg_neighbors, 2), "\n")

# ============================================================================
# STEP 4: PREPARE SHARED TEMPORAL STRUCTURE
# ============================================================================
years_unique <- sort(unique(parcelles_complete$Year))
n_years <- length(years_unique)
year_mapping <- data.frame(
  Year = years_unique,
  ID.Year = 1:n_years
)

cat("Years mapped:", min(years_unique), "to", max(years_unique), "(", n_years, "years)\n")

# ============================================================================
# ILR TRANSFORMATION FUNCTIONS
# ============================================================================

create_ilr_basis <- function(D) {
  V <- matrix(0, nrow = D - 1, ncol = D)
  for (i in 1:(D - 1)) {
    V[i, 1:i] <- 1/i
    V[i, i + 1] <- -1
    V[i, ] <- V[i, ] * sqrt(i / (i + 1))
  }
  return(V)
}

ilr_transform <- function(composition_matrix) {
  D <- ncol(composition_matrix)
  V <- create_ilr_basis(D)
  log_comp <- log(composition_matrix)
  ilr_coords <- log_comp %*% t(V)
  return(ilr_coords)
}

ilr_inverse <- function(ilr_matrix) {
  D <- ncol(ilr_matrix) + 1
  V <- create_ilr_basis(D)
  log_comp <- ilr_matrix %*% V
  composition <- exp(log_comp)
  row_sums <- rowSums(composition)
  composition <- composition / row_sums
  return(composition)
}

# ============================================================================
# STEP 5: CREATE 5-FOLD CV SPLITS
# ============================================================================

create_cv_folds <- function(unique_ids, n_folds = 5, seed = 123) {
  set.seed(seed)
  n <- length(unique_ids)
  fold_ids <- sample(rep(1:n_folds, length.out = n))
  
  fold_assignment <- data.frame(
    Final_Cleaned_ID = unique_ids,
    cv_fold = fold_ids
  )
  
  return(fold_assignment)
}

# ============================================================================
# STEP 6: DEFINE IMPUTATION FUNCTIONS WITH CV AND PARALLEL PROCESSING
# ============================================================================

# Non-compositional imputation with 5-fold CV (PARALLELIZED)
impute_single_variable_cv_parallel <- function(variable_name, parcelles_data, W_matrix, 
                                               field_map, year_map, cv_folds, n_folds = 5, 
                                               verbose = TRUE) {
  
  if (verbose) {
    cat("\n", "=", rep("=", 60), "\n", sep="")
    cat("🔬 PROCESSING VARIABLE:", variable_name, "(PARALLEL)\n")
    cat("=", rep("=", 60), "\n", sep="")
  }
  
  var_values <- parcelles_data[[variable_name]]
  n_total <- length(var_values)
  n_available <- sum(!is.na(var_values))
  n_missing <- n_total - n_available
  pct_available <- round(n_available/n_total*100, 1)
  
  if (verbose) {
    cat("Variable statistics:\n")
    cat("• Total observations:", n_total, "\n")
    cat("• Available (non-NA):", n_available, "(", pct_available, "%)\n")
    cat("• Missing (NA):", n_missing, "(", round(n_missing/n_total*100, 1), "%)\n")
    
    if (n_available > 0) {
      cat("• Min value:", round(min(var_values, na.rm = TRUE), 3), "\n")
      cat("• Max value:", round(max(var_values, na.rm = TRUE), 3), "\n")
      cat("• Mean value:", round(mean(var_values, na.rm = TRUE), 3), "\n")
      cat("• Std deviation:", round(sd(var_values, na.rm = TRUE), 3), "\n")
    }
  }
  
  if (n_available < 20) {
    if (verbose) cat("❌ Insufficient data (< 20 observations). Skipping variable.\n")
    return(NULL)
  }
  
  # Export necessary objects to cluster
  if (verbose) cat("  📤 Exporting data to parallel workers...\n")
  
  # PARALLEL LOOP THROUGH FOLDS
  fold_results <- foreach(
    fold = 1:n_folds,
    .packages = c("sf", "INLA", "dplyr"),
    .export = c("parcelles_data", "cv_folds", "variable_name", "field_map", "year_map", "W_matrix"),
    .errorhandling = "pass",
    .verbose = FALSE
  ) %dopar% {
    
    # Assign train/validation based on current fold
    parcelles_data_fold <- parcelles_data %>%
      left_join(cv_folds, by = "Final_Cleaned_ID") %>%
      mutate(
        data_split = case_when(
          is.na(.data[[variable_name]]) ~ "predict",
          cv_fold == fold ~ "validation",
          !is.na(.data[[variable_name]]) ~ "train",
          TRUE ~ "predict"
        )
      )
    
    train_obs <- sum(parcelles_data_fold$data_split == "train")
    val_obs <- sum(parcelles_data_fold$data_split == "validation")
    pred_obs <- sum(parcelles_data_fold$data_split == "predict")
    
    # Prepare INLA data
    df_inla <- parcelles_data_fold %>%
      st_drop_geometry() %>%
      left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
      left_join(year_map, by = "Year") %>%
      mutate(
        Y = case_when(
          data_split == "train" ~ .data[[variable_name]],
          TRUE ~ NA_real_
        ),
        Y_true = .data[[variable_name]],
        ID.Field = Sequential_ID,
        ID.Year = as.integer(ID.Year),
        Month = as.integer(ifelse(is.na(Month), 6, Month)),
        Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
        Var = as.factor(ifelse(is.na(Var), "Unknown", as.character(Var))),
        irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
      ) %>%
      filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
      arrange(ID.Field, ID.Year)
    
    if (max(df_inla$ID.Field) > nrow(W_matrix)) {
      return(list(success = FALSE, fold = fold, error = "Spatial structure incompatible"))
    }
    
    # Fit model
    fit_result <- tryCatch({
      prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
      
      inla(
        Y ~ 1 + Bloc + irrg_ty +
          f(ID.Field, model = "besag", graph = W_matrix,
            group = ID.Year, 
            control.group = list(model = "ar1"),
            hyper = list(prec = prec_prior),
            constr = TRUE),
        data = df_inla,
        family = "gaussian",
        control.predictor = list(compute = TRUE, link = 1),
        control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
        control.inla = list(strategy = "adaptive", int.strategy = "eb"),
        verbose = FALSE
      )
    }, error = function(e) {
      tryCatch({
        inla(
          Y ~ 1 + Bloc + f(ID.Field, model = "besag", graph = W_matrix, constr = TRUE),
          data = df_inla,
          family = "gaussian",
          control.predictor = list(compute = TRUE, link = 1),
          control.compute = list(dic = TRUE),
          verbose = FALSE
        )
      }, error = function(e2) NULL)
    })
    
    if (is.null(fit_result) || is.null(fit_result$summary.fitted.values)) {
      return(list(success = FALSE, fold = fold, error = "Model fitting failed"))
    }
    
    # Extract predictions
    predictions <- fit_result$summary.fitted.values
    
    df_predicted <- df_inla %>%
      mutate(
        fold = fold,
        predicted = predictions$mean,
        predicted_sd = predictions$sd,
        predicted_lower = predictions$`0.025quant`,
        predicted_upper = predictions$`0.975quant`
      )
    
    # Calculate fold-specific validation metrics
    validation_data <- df_predicted %>%
      filter(data_split == "validation", !is.na(Y_true), !is.na(predicted))
    
    fold_metric <- NULL
    if (nrow(validation_data) > 5) {
      pred_vals <- validation_data$predicted
      true_vals <- validation_data$Y_true
      
      ss_res <- sum((true_vals - pred_vals)^2)
      ss_tot <- sum((true_vals - mean(true_vals))^2)
      r_squared <- 1 - (ss_res / ss_tot)
      
      fold_metric <- list(
        fold = fold,
        r_squared = r_squared,
        correlation = cor(true_vals, pred_vals),
        rmse = sqrt(mean((true_vals - pred_vals)^2)),
        mae = mean(abs(true_vals - pred_vals)),
        n_validation = nrow(validation_data)
      )
    }
    
    return(list(
      success = TRUE,
      fold = fold,
      predictions = df_predicted,
      metrics = fold_metric,
      train_obs = train_obs,
      val_obs = val_obs,
      pred_obs = pred_obs
    ))
  }
  
  # Process parallel results
  if (verbose) cat("\n  📥 Processing parallel results...\n")
  
  all_fold_predictions <- list()
  fold_metrics <- list()
  
  for (result in fold_results) {
    if (is.list(result) && !is.null(result$success) && result$success) {
      fold <- result$fold
      all_fold_predictions[[fold]] <- result$predictions
      if (!is.null(result$metrics)) {
        fold_metrics[[fold]] <- result$metrics
        if (verbose) {
          cat("    ✅ Fold", fold, "metrics: R²=", round(result$metrics$r_squared, 4), 
              " RMSE=", round(result$metrics$rmse, 4), "\n")
        }
      }
    } else {
      if (verbose) cat("    ❌ Fold", result$fold, "failed:", result$error, "\n")
    }
  }
  
  # Combine all fold predictions
  if (length(all_fold_predictions) == 0) {
    if (verbose) cat("❌ No successful folds\n")
    return(NULL)
  }
  
  combined_predictions <- bind_rows(all_fold_predictions)
  
  # Calculate aggregated CV metrics
  all_validation <- combined_predictions %>%
    filter(data_split == "validation", !is.na(Y_true), !is.na(predicted))
  
  cv_metrics <- list(variable = variable_name)
  
  if (nrow(all_validation) > 5) {
    pred_vals <- all_validation$predicted
    true_vals <- all_validation$Y_true
    
    ss_res <- sum((true_vals - pred_vals)^2)
    ss_tot <- sum((true_vals - mean(true_vals))^2)
    
    cv_metrics$cv_r_squared = 1 - (ss_res / ss_tot)
    cv_metrics$cv_correlation = cor(true_vals, pred_vals)
    cv_metrics$cv_rmse = sqrt(mean((true_vals - pred_vals)^2))
    cv_metrics$cv_mae = mean(abs(true_vals - pred_vals))
    cv_metrics$n_total_validation = nrow(all_validation)
    
    # Calculate per-fold statistics
    fold_r2_values <- sapply(fold_metrics, function(x) x$r_squared)
    cv_metrics$mean_fold_r_squared = mean(fold_r2_values, na.rm = TRUE)
    cv_metrics$sd_fold_r_squared = sd(fold_r2_values, na.rm = TRUE)
    
    if (verbose) {
      cat("\n  🎯 Cross-Validation Results (Aggregated):\n")
      cat("    • Total validation obs:", cv_metrics$n_total_validation, "\n")
      cat("    • CV R²:", round(cv_metrics$cv_r_squared, 4), "\n")
      cat("    • Mean fold R² (±SD):", round(cv_metrics$mean_fold_r_squared, 4), 
          "±", round(cv_metrics$sd_fold_r_squared, 4), "\n")
      cat("    • CV RMSE:", round(cv_metrics$cv_rmse, 4), "\n")
      cat("    • CV MAE:", round(cv_metrics$cv_mae, 4), "\n")
    }
  }
  
  # Fit final model on all data for predictions
  if (verbose) cat("\n  🚀 Fitting final model on all data...\n")
  
  parcelles_data_final <- parcelles_data %>%
    mutate(data_split = ifelse(is.na(.data[[variable_name]]), "predict", "train"))
  
  df_inla_final <- parcelles_data_final %>%
    st_drop_geometry() %>%
    left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
    left_join(year_map, by = "Year") %>%
    mutate(
      Y = case_when(
        data_split == "train" ~ .data[[variable_name]],
        TRUE ~ NA_real_
      ),
      ID.Field = Sequential_ID,
      ID.Year = as.integer(ID.Year),
      Month = as.integer(ifelse(is.na(Month), 6, Month)),
      Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
      Var = as.factor(ifelse(is.na(Var), "Unknown", as.character(Var))),
      irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
    ) %>%
    filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
    arrange(ID.Field, ID.Year)
  
  fit_final <- tryCatch({
    prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
    
    inla(
      Y ~ 1 + Bloc + irrg_ty +
        f(ID.Field, model = "besag", graph = W_matrix,
          group = ID.Year, 
          control.group = list(model = "ar1"),
          hyper = list(prec = prec_prior),
          constr = TRUE),
      data = df_inla_final,
      family = "gaussian",
      control.predictor = list(compute = TRUE, link = 1),
      control.compute = list(dic = TRUE, waic = TRUE),
      control.inla = list(strategy = "adaptive", int.strategy = "eb"),
      verbose = FALSE
    )
  }, error = function(e) NULL)
  
  if (is.null(fit_final)) {
    if (verbose) cat("  ❌ Final model fitting failed\n")
    return(NULL)
  }
  
  predictions_final <- fit_final$summary.fitted.values
  
  df_predicted_final <- df_inla_final %>%
    mutate(
      !!paste0(variable_name, "_predicted") := predictions_final$mean,
      !!paste0(variable_name, "_predicted_sd") := predictions_final$sd,
      !!paste0(variable_name, "_predicted_lower") := predictions_final$`0.025quant`,
      !!paste0(variable_name, "_predicted_upper") := predictions_final$`0.975quant`,
      !!paste0(variable_name, "_complete") := case_when(
        data_split == "train" ~ Y,
        TRUE ~ predictions_final$mean
      ),
      !!paste0(variable_name, "_prediction_type") := ifelse(data_split == "train", "Observed", "Predicted")
    )
  
  result <- list(
    variable_name = variable_name,
    model_fit = fit_final,
    predictions = df_predicted_final,
    cv_predictions = combined_predictions,
    cv_metrics = cv_metrics,
    fold_metrics = fold_metrics,
    n_total = n_total,
    n_available = n_available,
    n_missing = n_missing,
    pct_available = pct_available
  )
  
  return(result)
}

# Compositional imputation with 5-fold CV (PARALLELIZED)
impute_compositional_granulometry_cv_parallel <- function(parcelles_data, W_matrix, field_map, 
                                                          year_map, cv_folds, n_folds = 5, 
                                                          verbose = TRUE, GRANULO_FRACTIONS) {
  
  if (verbose) {
    cat("\n", "=", rep("=", 60), "\n", sep="")
    cat("🔬 PROCESSING COMPOSITIONAL GRANULOMETRY (ILR) WITH 5-FOLD CV (PARALLEL)\n")
    cat("=", rep("=", 60), "\n", sep="")
  }
  
  parcelles_data <- parcelles_data %>%
    mutate(obs_id = row_number())
  
  sampled_obs <- parcelles_data %>%
    filter(has_complete_granulo)
  
  unsampled_obs <- parcelles_data %>%
    filter(!has_complete_granulo)
  
  n_sampled <- nrow(sampled_obs)
  n_unsampled <- nrow(unsampled_obs)
  
  if (verbose) {
    cat("• Sampled observations:", n_sampled, "\n")
    cat("• Unsampled observations:", n_unsampled, "\n")
  }
  
  if (n_sampled < 30) {
    if (verbose) cat("❌ Insufficient sampled observations\n")
    return(NULL)
  }
  
  # Prepare compositional data and ILR transform
  comp_matrix <- sampled_obs %>%
    st_drop_geometry() %>%
    dplyr::select(all_of(GRANULO_FRACTIONS)) %>%
    as.matrix()
  
  comp_matrix[comp_matrix <= 0] <- 0.01
  row_sums <- rowSums(comp_matrix)
  comp_matrix <- comp_matrix / row_sums * 100
  
  ilr_coords <- ilr_transform(comp_matrix / 100)
  colnames(ilr_coords) <- paste0("ILR_", 1:4)
  
  if (verbose) {
    cat("✅ ILR transformation: 5 fractions -> 4 orthonormal coordinates\n")
    cat("  📤 Preparing parallel processing of ILR coordinates...\n")
  }
  
  sampled_obs_ilr <- sampled_obs %>%
    bind_cols(as.data.frame(ilr_coords))
  
  # PARALLEL PROCESSING OF ILR COORDINATES
  ilr_results_list <- foreach(
    i = 1:4,
    .packages = c("sf", "INLA", "dplyr"),
    .export = c("sampled_obs_ilr", "unsampled_obs", "cv_folds", "field_map", "year_map", 
                "W_matrix", "n_folds"),
    .errorhandling = "pass",
    .verbose = FALSE
  ) %dopar% {
    
    ilr_var <- paste0("ILR_", i)
    
    all_fold_predictions <- list()
    fold_metrics <- list()
    
    # CV loop for this ILR coordinate
    for (fold in 1:n_folds) {
      
      sampled_obs_fold <- sampled_obs_ilr %>%
        left_join(cv_folds, by = "Final_Cleaned_ID") %>%
        mutate(
          data_split = case_when(
            cv_fold == fold ~ "validation",
            TRUE ~ "train"
          )
        )
      
      all_obs_fold <- bind_rows(
        sampled_obs_fold %>% mutate(obs_type = "sampled"),
        unsampled_obs %>% mutate(obs_type = "unsampled", data_split = "predict")
      )
      
      df_inla <- all_obs_fold %>%
        st_drop_geometry() %>%
        left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
        left_join(year_map, by = "Year") %>%
        mutate(
          Y = case_when(
            data_split == "train" ~ .data[[ilr_var]],
            TRUE ~ NA_real_
          ),
          Y_true = .data[[ilr_var]],
          ID.Field = Sequential_ID,
          ID.Year = as.integer(ID.Year),
          Month = as.integer(ifelse(is.na(Month), 6, Month)),
          Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
          irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
        ) %>%
        filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
        arrange(ID.Field, ID.Year)
      
      fit_result <- tryCatch({
        prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
        
        inla(
          Y ~ 1 + Bloc + irrg_ty +
            f(ID.Field, model = "besag", graph = W_matrix,
              group = ID.Year, 
              control.group = list(model = "ar1"),
              hyper = list(prec = prec_prior),
              constr = TRUE),
          data = df_inla,
          family = "gaussian",
          control.predictor = list(compute = TRUE, link = 1),
          control.compute = list(dic = TRUE, waic = TRUE),
          control.inla = list(strategy = "adaptive", int.strategy = "eb"),
          verbose = FALSE
        )
      }, error = function(e) NULL)
      
      if (!is.null(fit_result)) {
        predictions <- fit_result$summary.fitted.values
        
        df_predicted <- df_inla %>%
          mutate(
            fold = fold,
            predicted = predictions$mean,
            predicted_sd = predictions$sd
          )
        
        all_fold_predictions[[fold]] <- df_predicted
        
        # Validation metrics
        val_data <- df_predicted %>%
          filter(data_split == "validation", !is.na(Y_true))
        
        if (nrow(val_data) > 5) {
          pred_vals <- val_data$predicted
          true_vals <- val_data$Y_true
          
          ss_res <- sum((true_vals - pred_vals)^2)
          ss_tot <- sum((true_vals - mean(true_vals))^2)
          
          fold_metrics[[fold]] <- list(
            fold = fold,
            r_squared = 1 - (ss_res / ss_tot),
            rmse = sqrt(mean((true_vals - pred_vals)^2)),
            n_val = nrow(val_data)
          )
        }
      }
    }
    
    # Aggregate CV results for this ILR coordinate
    cv_metrics_ilr <- NULL
    combined_predictions <- NULL
    
    if (length(all_fold_predictions) > 0) {
      combined_predictions <- bind_rows(all_fold_predictions)
      
      all_validation <- combined_predictions %>%
        filter(data_split == "validation", !is.na(Y_true))
      
      if (nrow(all_validation) > 5) {
        pred_vals <- all_validation$predicted
        true_vals <- all_validation$Y_true
        
        ss_res <- sum((true_vals - pred_vals)^2)
        ss_tot <- sum((true_vals - mean(true_vals))^2)
        
        cv_metrics_ilr <- list(
          cv_r_squared = 1 - (ss_res / ss_tot),
          cv_rmse = sqrt(mean((true_vals - pred_vals)^2)),
          n_val = nrow(all_validation)
        )
        
        fold_r2 <- sapply(fold_metrics, function(x) x$r_squared)
        cv_metrics_ilr$mean_fold_r_squared = mean(fold_r2, na.rm = TRUE)
        cv_metrics_ilr$sd_fold_r_squared = sd(fold_r2, na.rm = TRUE)
      }
    }
    
    # Fit final model on all data
    all_obs_final <- bind_rows(
      sampled_obs_ilr %>% mutate(obs_type = "sampled", data_split = "train"),
      unsampled_obs %>% mutate(obs_type = "unsampled", data_split = "predict")
    )
    
    df_inla_final <- all_obs_final %>%
      st_drop_geometry() %>%
      left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
      left_join(year_map, by = "Year") %>%
      mutate(
        Y = case_when(
          data_split == "train" ~ .data[[ilr_var]],
          TRUE ~ NA_real_
        ),
        ID.Field = Sequential_ID,
        ID.Year = as.integer(ID.Year),
        Month = as.integer(ifelse(is.na(Month), 6, Month)),
        Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
        irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
      ) %>%
      filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
      arrange(ID.Field, ID.Year)
    
    fit_final <- tryCatch({
      prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
      
      inla(
        Y ~ 1 + Bloc + irrg_ty +
          f(ID.Field, model = "besag", graph = W_matrix,
            group = ID.Year, 
            control.group = list(model = "ar1"),
            hyper = list(prec = prec_prior),
            constr = TRUE),
        data = df_inla_final,
        family = "gaussian",
        control.predictor = list(compute = TRUE, link = 1),
        control.compute = list(dic = TRUE),
        verbose = FALSE
      )
    }, error = function(e) NULL)
    
    predictions_final <- NULL
    if (!is.null(fit_final)) {
      predictions_final <- fit_final$summary.fitted.values
      
      df_predicted_final <- df_inla_final %>%
        mutate(
          !!paste0(ilr_var, "_predicted") := predictions_final$mean,
          !!paste0(ilr_var, "_predicted_sd") := predictions_final$sd
        )
      
      return(list(
        success = TRUE,
        ilr_var = ilr_var,
        model = fit_final,
        predictions = df_predicted_final,
        cv_predictions = combined_predictions,
        cv_metrics = cv_metrics_ilr,
        fold_metrics = fold_metrics
      ))
    } else {
      return(list(success = FALSE, ilr_var = ilr_var))
    }
  }
  
  # Process parallel ILR results
  if (verbose) cat("\n  📥 Processing parallel ILR results...\n")
  
  ilr_cv_results <- list()
  for (result in ilr_results_list) {
    if (is.list(result) && !is.null(result$success) && result$success) {
      ilr_var <- result$ilr_var
      ilr_cv_results[[ilr_var]] <- result
      
      if (verbose && !is.null(result$cv_metrics$cv_r_squared)) {
        cat("    ✅", ilr_var, "- CV R²:", round(result$cv_metrics$cv_r_squared, 4), 
            "± SD:", round(result$cv_metrics$sd_fold_r_squared, 4), "\n")
      }
    } else {
      if (verbose) cat("    ❌", result$ilr_var, "failed\n")
    }
  }
  
  if (length(ilr_cv_results) != 4) {
    if (verbose) cat("❌ Not all ILR coordinates fitted successfully\n")
    return(NULL)
  }
  
  # Transform back to composition
  if (verbose) cat("\n🔄 Transforming back to composition...\n")
  
  all_predictions <- ilr_cv_results[[1]]$predictions %>%
    dplyr::select(obs_id, Final_Cleaned_ID, Year, data_split, obs_type,
                  ILR_1_predicted, ILR_1_predicted_sd)
  
  for (i in 2:4) {
    ilr_var <- paste0("ILR_", i)
    pred_cols <- ilr_cv_results[[ilr_var]]$predictions %>%
      dplyr::select(obs_id, !!paste0(ilr_var, "_predicted"), !!paste0(ilr_var, "_predicted_sd"))
    
    all_predictions <- all_predictions %>%
      left_join(pred_cols, by = "obs_id")
  }
  
  ilr_pred_matrix <- all_predictions %>%
    dplyr::select(ILR_1_predicted, ILR_2_predicted, ILR_3_predicted, ILR_4_predicted) %>%
    as.matrix()
  
  comp_predicted <- ilr_inverse(ilr_pred_matrix)
  comp_predicted_pct <- comp_predicted * 100
  colnames(comp_predicted_pct) <- GRANULO_FRACTIONS
  
  if (verbose) {
    cat("✅ ILR back-transformation complete\n")
    sums_check <- rowSums(comp_predicted_pct)
    cat("• Sum range:", round(min(sums_check), 2), "to", round(max(sums_check), 2), "%\n")
  }
  
  final_results <- all_predictions %>%
    bind_cols(as.data.frame(comp_predicted_pct) %>%
                rename_with(~paste0(., "_predicted"), all_of(GRANULO_FRACTIONS)))
  
  for (frac in GRANULO_FRACTIONS) {
    original_map <- bind_rows(sampled_obs_ilr, unsampled_obs) %>%
      st_drop_geometry() %>%
      dplyr::select(obs_id, !!frac) %>%
      rename(original_value = !!frac)
    
    final_results <- final_results %>%
      left_join(original_map, by = "obs_id")
    
    pred_col <- paste0(frac, "_predicted")
    
    final_results[[paste0(frac, "_complete")]] <- ifelse(
      final_results$data_split == "train",
      final_results$original_value,
      final_results[[pred_col]]
    )
    
    final_results[[paste0(frac, "_prediction_type")]] <- final_results$data_split
    
    final_results <- final_results %>% dplyr::select(-original_value)
  }
  
  # Calculate CV metrics for fractions
  cv_frac_metrics <- list()
  
  # Get CV predictions for all ILR coordinates
  all_ilr_cv <- ilr_cv_results[[1]]$cv_predictions %>%
    dplyr::select(obs_id, Final_Cleaned_ID, Year, fold, data_split, ILR_1 = predicted)
  
  for (i in 2:4) {
    ilr_var <- paste0("ILR_", i)
    cv_cols <- ilr_cv_results[[ilr_var]]$cv_predictions %>%
      dplyr::select(obs_id, fold, !!paste0("ILR_", i) := predicted)
    
    all_ilr_cv <- all_ilr_cv %>%
      left_join(cv_cols, by = c("obs_id", "fold"))
  }
  
  # Transform CV predictions back to composition
  ilr_cv_matrix <- all_ilr_cv %>%
    dplyr::select(ILR_1, ILR_2, ILR_3, ILR_4) %>%
    as.matrix()
  
  comp_cv_predicted <- ilr_inverse(ilr_cv_matrix)
  comp_cv_predicted_pct <- comp_cv_predicted * 100
  colnames(comp_cv_predicted_pct) <- GRANULO_FRACTIONS
  
  cv_predictions_fractions <- all_ilr_cv %>%
    bind_cols(as.data.frame(comp_cv_predicted_pct))
  
  # Calculate metrics for each fraction
  for (frac in GRANULO_FRACTIONS) {
    true_map <- sampled_obs_ilr %>%
      st_drop_geometry() %>%
      dplyr::select(obs_id, !!frac) %>%
      rename(true_val = !!frac)
    
    cv_comparison <- cv_predictions_fractions %>%
      filter(data_split == "validation") %>%
      left_join(true_map, by = "obs_id")
    
    if (sum(!is.na(cv_comparison$true_val)) > 5) {
      true_vals <- cv_comparison$true_val
      pred_vals <- cv_comparison[[frac]]
      
      valid_idx <- !is.na(true_vals) & !is.na(pred_vals)
      true_vals <- true_vals[valid_idx]
      pred_vals <- pred_vals[valid_idx]
      
      if (length(true_vals) > 0) {
        ss_res <- sum((true_vals - pred_vals)^2)
        ss_tot <- sum((true_vals - mean(true_vals))^2)
        
        cv_frac_metrics[[frac]] <- list(
          cv_r_squared = 1 - (ss_res / ss_tot),
          cv_rmse = sqrt(mean((true_vals - pred_vals)^2)),
          cv_mae = mean(abs(true_vals - pred_vals)),
          n = length(true_vals)
        )
      }
    }
  }
  
  overall_metrics <- list(
    fraction_metrics = cv_frac_metrics,
    n_sampled = n_sampled,
    n_unsampled = n_unsampled
  )
  
  if (verbose) {
    cat("\n🎯 Cross-Validation Results (Back-transformed Fractions):\n")
    for (frac in GRANULO_FRACTIONS) {
      if (!is.null(cv_frac_metrics[[frac]])) {
        cat("• ", frac, ": CV R²=", round(cv_frac_metrics[[frac]]$cv_r_squared, 4),
            " RMSE=", round(cv_frac_metrics[[frac]]$cv_rmse, 2), "% (n=", 
            cv_frac_metrics[[frac]]$n, ")\n", sep="")
      }
    }
  }
  
  return(list(
    predictions = final_results,
    cv_predictions = cv_predictions_fractions,
    ilr_results = ilr_cv_results,
    metrics = overall_metrics,
    n_sampled = n_sampled,
    n_unsampled = n_unsampled
  ))
}

# ============================================================================
# STEP 7: CREATE CV FOLDS
# ============================================================================
cat("\n🎲 CREATING 5-FOLD CROSS-VALIDATION SPLITS:\n")
cat("=", rep("=", 45), "\n", sep="")

# Create CV folds based on unique parcels
unique_parcels_with_data <- parcelles_complete %>%
  st_drop_geometry() %>%
  filter(!is.na(pH) | has_complete_granulo) %>%
  dplyr::select(Final_Cleaned_ID) %>%
  distinct() %>%
  pull(Final_Cleaned_ID)

cv_fold_assignment <- create_cv_folds(unique_parcels_with_data, n_folds = 5, seed = 123)

cat("• Total unique parcels with data:", length(unique_parcels_with_data), "\n")
cat("• Parcels per fold:\n")
for (fold in 1:5) {
  n_parcels <- sum(cv_fold_assignment$cv_fold == fold)
  cat("  Fold", fold, ":", n_parcels, "parcels\n")
}

# ============================================================================
# STEP 8: PROCESS ALL VARIABLES WITH CV (PARALLEL)
# ============================================================================
cat("\n🔄 PROCESSING ALL VARIABLES WITH 5-FOLD CV (PARALLEL):\n")
cat("=", rep("=", 50), "\n", sep="")

all_results <- list()
all_metrics <- data.frame()
failed_variables <- character(0)

# Process compositional granulometry
cat("\n[COMPOSITIONAL] Processing granulometry fractions with CV (PARALLEL)...\n")

granulo_result <- impute_compositional_granulometry_cv_parallel(
  parcelles_data = parcelles_complete,
  W_matrix = W.P1_complete,
  field_map = field_mapping,
  year_map = year_mapping,
  cv_folds = cv_fold_assignment,
  n_folds = 5,
  verbose = TRUE,
  GRANULO_FRACTIONS = GRANULO_FRACTIONS
)

if (!is.null(granulo_result)) {
  all_results[["Granulometry_Compositional"]] <- granulo_result
  
  if (!is.null(granulo_result$metrics$fraction_metrics)) {
    for (frac in GRANULO_FRACTIONS) {
      if (!is.null(granulo_result$metrics$fraction_metrics[[frac]])) {
        metrics_row <- data.frame(
          Variable = frac,
          N_Total = nrow(parcelles_complete),
          N_Available = granulo_result$n_sampled,
          N_Missing = granulo_result$n_unsampled,
          Pct_Available = round(granulo_result$n_sampled / nrow(parcelles_complete) * 100, 1),
          CV_R_Squared = granulo_result$metrics$fraction_metrics[[frac]]$cv_r_squared,
          CV_RMSE = granulo_result$metrics$fraction_metrics[[frac]]$cv_rmse,
          CV_MAE = granulo_result$metrics$fraction_metrics[[frac]]$cv_mae,
          N_CV_Validation = granulo_result$metrics$fraction_metrics[[frac]]$n,
          stringsAsFactors = FALSE
        )
        
        all_metrics <- rbind(all_metrics, metrics_row)
      }
    }
  }
  cat("✅ SUCCESS for compositional granulometry\n")
} else {
  failed_variables <- c(failed_variables, GRANULO_FRACTIONS)
  cat("❌ FAILED for compositional granulometry\n")
}

# Process non-compositional variables (SEQUENTIAL - one at a time due to memory)
cat("\n[NON-COMPOSITIONAL] Processing variables sequentially...\n")
cat("(Sequential processing to manage memory for INLA models)\n")

for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  
  cat("\n[", i, "/", length(NON_COMPOSITIONAL_VARIABLES), "] Processing:", var_name, "\n")
  
  result <- impute_single_variable_cv_parallel(
    variable_name = var_name,
    parcelles_data = parcelles_complete,
    W_matrix = W.P1_complete,
    field_map = field_mapping,
    year_map = year_mapping,
    cv_folds = cv_fold_assignment,
    n_folds = 5,
    verbose = TRUE
  )
  
  if (!is.null(result)) {
    all_results[[var_name]] <- result
    
    metrics_row <- data.frame(
      Variable = var_name,
      N_Total = result$n_total,
      N_Available = result$n_available,
      N_Missing = result$n_missing,
      Pct_Available = result$pct_available,
      CV_R_Squared = ifelse(is.null(result$cv_metrics$cv_r_squared), NA, result$cv_metrics$cv_r_squared),
      CV_RMSE = ifelse(is.null(result$cv_metrics$cv_rmse), NA, result$cv_metrics$cv_rmse),
      CV_MAE = ifelse(is.null(result$cv_metrics$cv_mae), NA, result$cv_metrics$cv_mae),
      N_CV_Validation = ifelse(is.null(result$cv_metrics$n_total_validation), 0, result$cv_metrics$n_total_validation),
      Mean_Fold_R_Squared = ifelse(is.null(result$cv_metrics$mean_fold_r_squared), NA, result$cv_metrics$mean_fold_r_squared),
      SD_Fold_R_Squared = ifelse(is.null(result$cv_metrics$sd_fold_r_squared), NA, result$cv_metrics$sd_fold_r_squared),
      stringsAsFactors = FALSE
    )
    
    all_metrics <- rbind(all_metrics, metrics_row)
    cat("✅ SUCCESS for", var_name, "\n")
  } else {
    failed_variables <- c(failed_variables, var_name)
    cat("❌ FAILED for", var_name, "\n")
  }
  
  # Force garbage collection after each variable
  gc()
}

# ============================================================================
# STEP 9: COMBINE ALL PREDICTIONS
# ============================================================================
cat("\n🔗 COMBINING ALL PREDICTIONS:\n")
cat("=", rep("=", 50), "\n", sep="")

successful_variables <- names(all_results)
cat("Successfully processed:", length(successful_variables), "variable groups\n")

final_dataset <- parcelles_complete %>% 
  st_drop_geometry() %>%
  group_by(Final_Cleaned_ID, Year) %>%
  mutate(
    replicate_id = row_number(),
    unique_key = paste0(Final_Cleaned_ID, "_", Year, "_rep", replicate_id)
  ) %>%
  ungroup() %>%
  mutate(original_row_id = row_number())

final_dataset <- parcelles_complete %>%
  mutate(original_row_id = row_number()) %>%
  left_join(
    final_dataset %>% dplyr::select(original_row_id, replicate_id, unique_key),
    by = "original_row_id"
  )

# Add CV fold assignment
final_dataset <- final_dataset %>%
  left_join(cv_fold_assignment, by = "Final_Cleaned_ID")

# Add granulometry predictions
if ("Granulometry_Compositional" %in% names(all_results)) {
  cat("\n[COMPOSITIONAL] Adding granulometry predictions...\n")
  
  final_dataset <- final_dataset %>%
    mutate(obs_id = row_number())
  
  granulo_preds <- all_results[["Granulometry_Compositional"]]$predictions
  
  final_dataset <- final_dataset %>%
    left_join(
      granulo_preds %>% 
        dplyr::select(obs_id, matches("^Clay_|^Fine_Silt_|^Coarse_Silt_|^Fine_Sand_|^Coarse_Sand_|^ILR_")),
      by = "obs_id"
    )
  
  final_dataset <- final_dataset %>% dplyr::select(-obs_id)
  cat("  ✅ Compositional predictions added\n")
}

# Add non-compositional predictions
for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  
  if (var_name %in% names(all_results)) {
    cat("\n[", i, "/", length(NON_COMPOSITIONAL_VARIABLES), "] Adding:", var_name, "\n")
    
    result <- all_results[[var_name]]
    
    df_inla_reference <- parcelles_complete %>% 
      st_drop_geometry() %>% 
      left_join(field_mapping, by = c("Final_Cleaned_ID" = "Original_ID"), 
                relationship = "many-to-one") %>%  
      left_join(year_mapping, by = "Year", 
                relationship = "many-to-one") %>%  
      filter(!is.na(Sequential_ID), !is.na(ID.Year)) %>% 
      group_by(Final_Cleaned_ID, Year) %>%
      mutate(
        replicate_id = row_number(),
        unique_key = paste0(Final_Cleaned_ID, "_", Year, "_rep", replicate_id)
      ) %>%
      ungroup() %>%
      mutate(original_row_id = row_number()) %>% 
      arrange(Sequential_ID, ID.Year, replicate_id) %>%
      mutate(inla_row_id = row_number())
    
    if (nrow(df_inla_reference) == nrow(result$predictions)) {
      predictions_aligned <- df_inla_reference %>% 
        dplyr::select(Final_Cleaned_ID, Year, replicate_id, unique_key, original_row_id) %>% 
        bind_cols(
          result$predictions %>% 
            dplyr::select(matches(paste0("^", var_name, "_")))
        )
      
      final_dataset <- final_dataset %>% 
        left_join(
          predictions_aligned %>% dplyr::select(-Final_Cleaned_ID, -Year, -replicate_id, -original_row_id),
          by = "unique_key"
        )
      
      first_pred_col <- paste0(var_name, "_predicted")
      if (first_pred_col %in% colnames(final_dataset)) {
        non_null_count <- sum(!is.na(final_dataset[[first_pred_col]]))
        cat("  ✅ Successfully added", non_null_count, "predictions for", var_name, "\n")
      }
    }
    
    gc()
  }
}

final_dataset <- final_dataset %>% 
  dplyr::select(-original_row_id)

cat("\n✅ Final combined dataset created\n")
cat("• Total observations:", nrow(final_dataset), "\n")
cat("• Total columns:", ncol(final_dataset), "\n")

gc()

# ============================================================================
# STEP 10: SAVE COMPREHENSIVE OUTPUTS
# ============================================================================
cat("\n💾 SAVING COMPREHENSIVE OUTPUTS:\n")
cat("=", rep("=", 35), "\n", sep="")

base_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/"

gpkg_output <- paste0(base_path, "parcels_ALL_VARIABLES_CV_predictions_PARALLEL.gpkg")
st_write(final_dataset, gpkg_output, delete_dsn = TRUE, quiet = TRUE)
cat("✅ Complete dataset saved:", basename(gpkg_output), "\n")

csv_output <- paste0(base_path, "parcels_ALL_VARIABLES_CV_attributes_PARALLEL.csv")
st_drop_geometry(final_dataset) %>% write_csv(csv_output)
cat("✅ Attributes saved:", basename(csv_output), "\n")

metrics_output <- paste0(base_path, "ALL_VARIABLES_CV_metrics_summary_PARALLEL.csv")
write_csv(all_metrics, metrics_output)
cat("✅ CV metrics summary saved:", basename(metrics_output), "\n")

# Save CV fold assignment
cv_folds_output <- paste0(base_path, "CV_fold_assignment.csv")
write_csv(cv_fold_assignment, cv_folds_output)
cat("✅ CV fold assignment saved:", basename(cv_folds_output), "\n")
```

# II. Part 2 : Granulo Fractions Analysis

## II.1. Data Import

```{r}
# ============================================================================
# STEP 1: LOAD THE COMPLETE DATASET AND IDENTIFY SOIL VARIABLES
# ============================================================================
cat("📂 LOADING COMPLETE DATASET AND IDENTIFYING SOIL VARIABLES:\n")
cat("=", rep("=", 60), "\n", sep="")

gpkg_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/parcels_with_soil_complete.gpkg"

if (file.exists(gpkg_path)) {
  parcelles_complete <- st_read(gpkg_path, quiet = TRUE)
  cat("✅ Complete dataset loaded:", nrow(parcelles_complete), "observations\n")
} else {
  cat("❌ Complete dataset not found. Please run the enhanced join script first.\n")
  stop("Dataset not available")
}

# Define compositional and non-compositional variables
GRANULO_FRACTIONS <- c(
  "Clay_percent",
  "Fine_Silt_percent",
  "Coarse_Silt_percent",
  "Fine_Sand_percent",
  "Coarse_Sand_percent"
)

NON_COMPOSITIONAL_VARIABLES <- c(
  "pH",
  "CN_Ratio", 
  "Electrical_Conductivity_mmho_cm",
  "Exchangeable_Calcium_meq_100g",
  "Exchangeable_Magnesium_meq_100g", 
  "Exchangeable_Sodium_meq_100g",
  "Exchangeable_Potassium_meq_100g",
  "Available_P2O5_mg_kg",
  "Organic_Carbon_per_mille",
  "Total_Nitrogen_per_mille",
  "Total_Silt_percent",
  "Total_Sand_percent", 
  "Organic_Matter_percent",
  "Sum_of_Exchangeable_Bases_meq_100g",
  "CaO_g_100g",
  "Mg_Ca_Ratio",
  "Mg_K_Ratio",
  "Compaction_Index"
)

cat("Compositional variables (granulometry):", length(GRANULO_FRACTIONS), "\n")
cat("Non-compositional variables:", length(NON_COMPOSITIONAL_VARIABLES), "\n")

# ============================================================================
# ANALYZE GRANULOMETRY FRACTION SUMS
# ============================================================================
cat("\n📊 ANALYZING GRANULOMETRY FRACTION SUMS:\n")
cat("=", rep("=", 60), "\n", sep="")

# Convert granulo columns to numeric and calculate sum
parcelles_complete$granulo_sum <- rowSums(
  st_drop_geometry(parcelles_complete)[, GRANULO_FRACTIONS] %>% 
    mutate(across(everything(), as.numeric)),
  na.rm = FALSE
)

# Categorize observations
parcelles_complete$sum_category <- cut(
  parcelles_complete$granulo_sum,
  breaks = c(-Inf, 96, 104, Inf),
  labels = c("< 96%", "96-104%", "> 104%"),
  right = FALSE
)

# Summary statistics
cat("\n📈 Summary of granulometry sums:\n")
print(summary(parcelles_complete$granulo_sum))

cat("\n📊 Distribution by category:\n")
sum_table <- table(parcelles_complete$sum_category, useNA = "ifany")
print(sum_table)
cat("\nPercentages:\n")
print(round(prop.table(sum_table) * 100, 2))
```

## II.2. Data Cleaning

```{r}
# ============================================================================
# CLEAN DATA AND ANALYZE LOW SUM OBSERVATIONS
# ============================================================================
cat("\n🧹 CLEANING DATA AND ANALYZING LOW GRANULOMETRY SUMS:\n")
cat("=", rep("=", 60), "\n", sep="")

# Load required libraries
library(ggplot2)
library(dplyr)
library(sf)

# Create results folder
results_folder <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/granulo_analysis_results"
dir.create(results_folder, showWarnings = FALSE, recursive = TRUE)
cat("📁 Results folder created:", results_folder, "\n")

# Set granulometry fractions to NA for observations > 104%
obs_to_clean <- which(parcelles_complete$granulo_sum > 104)
cat("🗑️ Setting granulometry fractions to NA for", length(obs_to_clean), 
    "observations with sum > 104%\n")

parcelles_filtered <- parcelles_complete
for (var in GRANULO_FRACTIONS) {
  parcelles_filtered[obs_to_clean, var] <- NA
}

# Recalculate granulo_sum after cleaning
parcelles_filtered$granulo_sum <- rowSums(
  st_drop_geometry(parcelles_filtered)[, GRANULO_FRACTIONS] %>% 
    mutate(across(everything(), as.numeric)),
  na.rm = FALSE
)

# Recalculate categories
parcelles_filtered$sum_category <- cut(
  parcelles_filtered$granulo_sum,
  breaks = c(-Inf, 96, 104, Inf),
  labels = c("< 96%", "96-104%", "> 104%"),
  right = FALSE
)

cat("📊 Total observations retained:", nrow(parcelles_filtered), "\n")

# Focus on observations < 96%
low_sum_obs <- parcelles_filtered %>%
  filter(granulo_sum < 96)

cat("\n📉 Analyzing", nrow(low_sum_obs), "observations with sum < 96%\n")

# ============================================================================
# TEMPORAL AND SPATIAL ANALYSIS
# ============================================================================

# Summary by Year
year_summary <- low_sum_obs %>%
  st_drop_geometry() %>%
  group_by(Year) %>%
  summarise(
    count = n(),
    mean_sum = mean(granulo_sum, na.rm = TRUE),
    min_sum = min(granulo_sum, na.rm = TRUE),
    max_sum = max(granulo_sum, na.rm = TRUE)
  ) %>%
  arrange(Year)

cat("\n📅 Distribution by Year:\n")
print(year_summary)

# Summary by Bloc
bloc_summary <- low_sum_obs %>%
  st_drop_geometry() %>%
  group_by(Bloc) %>%
  summarise(
    count = n(),
    mean_sum = mean(granulo_sum, na.rm = TRUE),
    min_sum = min(granulo_sum, na.rm = TRUE),
    max_sum = max(granulo_sum, na.rm = TRUE)
  ) %>%
  arrange(desc(count))

cat("\n🗺️ Distribution by Bloc:\n")
print(bloc_summary)

# Summary by Year and Bloc
year_bloc_summary <- low_sum_obs %>%
  st_drop_geometry() %>%
  group_by(Year, Bloc) %>%
  summarise(count = n(), .groups = "drop")

cat("\n📊 Distribution by Year and Bloc:\n")
print(year_bloc_summary)

# Save summaries
write.csv(year_summary, 
          file.path(results_folder, "low_sum_by_year.csv"), 
          row.names = FALSE)
write.csv(bloc_summary, 
          file.path(results_folder, "low_sum_by_bloc.csv"), 
          row.names = FALSE)
write.csv(year_bloc_summary, 
          file.path(results_folder, "low_sum_by_year_bloc.csv"), 
          row.names = FALSE)

# ============================================================================
# CREATE HEATMAP
# ============================================================================

# Prepare data for heatmap - ORDER BY irrg_ty (GRV first), THEN alphabetically by Bloc
bloc_order_data <- parcelles_filtered %>%
  st_drop_geometry() %>%
  select(Bloc, irrg_ty) %>%
  distinct() %>%
  mutate(irrg_ty_order = ifelse(irrg_ty == "GRV", 1, 2)) %>%  # GRV = 1, others = 2
  arrange(irrg_ty_order, Bloc) %>%  # Order by GRV first, then alphabetically
  select(-irrg_ty_order)

cat("\n📋 Bloc ordering (GRV first, then others, all alphabetically):\n")
print(bloc_order_data)

# Add irrg_ty to year_bloc_summary for the heatmap
year_bloc_summary <- year_bloc_summary %>%
  left_join(
    parcelles_filtered %>% 
      st_drop_geometry() %>% 
      select(Bloc, irrg_ty) %>% 
      distinct(),
    by = "Bloc"
  )

# Create ordered factor for Bloc with the correct order
year_bloc_summary$Bloc <- factor(
  year_bloc_summary$Bloc,
  levels = rev(bloc_order_data$Bloc)  # Reverse so GRV appears at top of plot
)

# Create color vectors for axes
bloc_colors <- ifelse(bloc_order_data$irrg_ty == "GRV", "#2E7D32", "#D32F2F")
names(bloc_colors) <- bloc_order_data$Bloc

# Create heatmap with white background and larger fonts
p <- ggplot(year_bloc_summary, aes(x = as.factor(Year), y = Bloc, fill = count)) +
  geom_tile(color = "white", size = 0.5) +
  scale_fill_gradient(low = "#FFF5E6", high = "#D32F2F", 
                      name = "Count",
                      na.value = "grey90") +
  labs(
    title = "Heatmap of Observations with Granulometry Sum < 96%",
    subtitle = "Distribution across Years and Blocs (Green = GRV, Red = Other)",
    x = "Year",
    y = "Bloc"
  ) +
  theme_bw() +
  theme(
    panel.background = element_rect(fill = "white", color = NA),
    plot.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_rect(color = "black", fill = NA, size = 1),
    axis.text.x = element_text(angle = 45, hjust = 1, size = 14, color = "black"),
    axis.text.y = element_text(size = 12, color = bloc_colors[rev(bloc_order_data$Bloc)], face = "bold"),
    axis.title.x = element_text(size = 16, face = "bold", color = "black"),
    axis.title.y = element_text(size = 16, face = "bold", color = "black"),
    plot.title = element_text(face = "bold", size = 18, color = "black"),
    plot.subtitle = element_text(size = 14, color = "grey40"),
    legend.position = "right",
    legend.background = element_rect(fill = "white", color = "black"),
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14, face = "bold")
  ) +
  geom_text(aes(label = count), color = "black", size = 5, fontface = "bold")

# Save heatmap
ggsave(
  filename = file.path(results_folder, "heatmap_low_sum_year_bloc.png"),
  plot = p,
  width = 14,
  height = 12,
  dpi = 300,
  bg = "white"
)

cat("\n✅ Heatmap saved to:", file.path(results_folder, "heatmap_low_sum_year_bloc.png"), "\n")
```

## II.3. SPATIAL MAPS FOR ALL GRANULOMETRY SUM CATEGORIES

```{r}
# ============================================================================
# SPATIAL MAPS FOR ALL GRANULOMETRY SUM CATEGORIES
# ============================================================================
cat("\n🗺️ CREATING SPATIAL MAPS FOR ALL GRANULOMETRY CATEGORIES:\n")
cat("=", rep("=", 60), "\n", sep="")

# Load required libraries
library(sf)
library(ggplot2)
library(dplyr)
library(viridis)

# Create results folder
results_folder <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/granulo_analysis_results"
dir.create(results_folder, showWarnings = FALSE, recursive = TRUE)

# ============================================================================
# COUNT OBSERVATIONS PER PARCEL FOR EACH CATEGORY
# ============================================================================

# Category 1: < 96%
parcel_low_count <- parcelles_filtered %>%
  st_drop_geometry() %>%
  filter(granulo_sum < 96) %>%
  group_by(field) %>%
  summarise(
    count_below_96 = n(),
    .groups = "drop"
  )

cat("\n📊 Category < 96%:\n")
cat("Parcels with observations:", nrow(parcel_low_count), "\n")
cat("Range:", min(parcel_low_count$count_below_96), 
    "to", max(parcel_low_count$count_below_96), "\n")

# Category 2: 96-104%
parcel_normal_count <- parcelles_filtered %>%
  st_drop_geometry() %>%
  filter(granulo_sum >= 96 & granulo_sum <= 104) %>%
  group_by(field) %>%
  summarise(
    count_96_104 = n(),
    .groups = "drop"
  )

cat("\n📊 Category 96-104%:\n")
cat("Parcels with observations:", nrow(parcel_normal_count), "\n")
cat("Range:", min(parcel_normal_count$count_96_104), 
    "to", max(parcel_normal_count$count_96_104), "\n")

# Category 3: > 104%
parcel_high_count <- parcelles_filtered %>%
  st_drop_geometry() %>%
  filter(granulo_sum > 104) %>%
  group_by(field) %>%
  summarise(
    count_above_104 = n(),
    .groups = "drop"
  )

cat("\n📊 Category > 104%:\n")
cat("Parcels with observations:", nrow(parcel_high_count), "\n")
cat("Range:", min(parcel_high_count$count_above_104), 
    "to", max(parcel_high_count$count_above_104), "\n")

# Get unique parcels with geometry
parcels_spatial <- parcelles_filtered %>%
  group_by(field) %>%
  slice(1) %>%
  ungroup() %>%
  select(field, Bloc, irrg_ty)

# Join all count data to spatial data
parcels_map <- parcels_spatial %>%
  left_join(parcel_low_count, by = "field") %>%
  left_join(parcel_normal_count, by = "field") %>%
  left_join(parcel_high_count, by = "field") %>%
  mutate(
    count_below_96 = ifelse(is.na(count_below_96), 0, count_below_96),
    count_96_104 = ifelse(is.na(count_96_104), 0, count_96_104),
    count_above_104 = ifelse(is.na(count_above_104), 0, count_above_104)
  )

# ============================================================================
# MAP 1: GRANULOMETRY SUM < 96%
# ============================================================================

map1 <- ggplot() +
  geom_sf(data = parcels_map %>% filter(count_below_96 == 0), 
          fill = "#F5F5F5", 
          color = "#E0E0E0", 
          size = 0.1) +
  geom_sf(data = parcels_map %>% filter(count_below_96 > 0), 
          aes(fill = count_below_96), 
          color = "#FFFFFF", 
          size = 0.2) +
  scale_fill_viridis(
    option = "magma",
    name = "Count of samples\nwith sum < 96%",
    begin = 0.1,
    end = 0.95,
    direction = 1,
    breaks = function(x) unique(floor(pretty(x))),
    guide = guide_colorbar(
      barwidth = 1.5,
      barheight = 12,
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      ticks.linewidth = 0.5
    )
  ) +
  labs(
    title = "Spatial distribution of low granulometry sum observations",
    subtitle = "Count of sampling events with granulometry sum < 96% per parcel",
    caption = "Gray parcels indicate no low granulometry observations"
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "sans", color = "black"),
    plot.title = element_text(size = 14, face = "bold", 
                              margin = margin(b = 8), hjust = 0),
    plot.subtitle = element_text(size = 11, color = "grey30", 
                                 margin = margin(b = 12), hjust = 0),
    plot.caption = element_text(size = 9, color = "grey50", 
                                margin = margin(t = 12), hjust = 0),
    legend.position = "right",
    legend.title = element_text(size = 10, face = "bold", lineheight = 1.2),
    legend.text = element_text(size = 9),
    legend.background = element_rect(fill = "white", color = NA),
    legend.key = element_rect(color = "black", linewidth = 0.3),
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "#F0F0F0", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(15, 15, 15, 15),
    axis.text = element_text(size = 8, color = "grey50"),
    axis.title = element_blank()
  )

ggsave(
  filename = file.path(results_folder, "spatial_map_below_96.png"),
  plot = map1,
  width = 10,
  height = 8,
  dpi = 600,
  bg = "white"
)

ggsave(
  filename = file.path(results_folder, "spatial_map_below_96.pdf"),
  plot = map1,
  width = 10,
  height = 8,
  device = cairo_pdf,
  bg = "white"
)

cat("\n✅ Map 1 (< 96%) saved\n")

# ============================================================================
# MAP 2: GRANULOMETRY SUM 96-104%
# ============================================================================

map2 <- ggplot() +
  geom_sf(data = parcels_map %>% filter(count_96_104 == 0), 
          fill = "#F5F5F5", 
          color = "#E0E0E0", 
          size = 0.1) +
  geom_sf(data = parcels_map %>% filter(count_96_104 > 0), 
          aes(fill = count_96_104), 
          color = "#FFFFFF", 
          size = 0.2) +
  scale_fill_viridis(
    option = "magma",
    name = "Count of samples\nwith sum 96-104%",
    begin = 0.1,
    end = 0.95,
    direction = 1,
    breaks = function(x) unique(floor(pretty(x))),
    guide = guide_colorbar(
      barwidth = 1.5,
      barheight = 12,
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      ticks.linewidth = 0.5
    )
  ) +
  labs(
    title = "Spatial distribution of normal granulometry sum observations",
    subtitle = "Count of sampling events with granulometry sum 96-104% per parcel",
    caption = "Gray parcels indicate no observations in this range"
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "sans", color = "black"),
    plot.title = element_text(size = 14, face = "bold", 
                              margin = margin(b = 8), hjust = 0),
    plot.subtitle = element_text(size = 11, color = "grey30", 
                                 margin = margin(b = 12), hjust = 0),
    plot.caption = element_text(size = 9, color = "grey50", 
                                margin = margin(t = 12), hjust = 0),
    legend.position = "right",
    legend.title = element_text(size = 10, face = "bold", lineheight = 1.2),
    legend.text = element_text(size = 9),
    legend.background = element_rect(fill = "white", color = NA),
    legend.key = element_rect(color = "black", linewidth = 0.3),
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "#F0F0F0", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(15, 15, 15, 15),
    axis.text = element_text(size = 8, color = "grey50"),
    axis.title = element_blank()
  )

ggsave(
  filename = file.path(results_folder, "spatial_map_96_104.png"),
  plot = map2,
  width = 10,
  height = 8,
  dpi = 600,
  bg = "white"
)

ggsave(
  filename = file.path(results_folder, "spatial_map_96_104.pdf"),
  plot = map2,
  width = 10,
  height = 8,
  device = cairo_pdf,
  bg = "white"
)

cat("✅ Map 2 (96-104%) saved\n")

# ============================================================================
# MAP 3: GRANULOMETRY SUM > 104%
# ============================================================================

map3 <- ggplot() +
  geom_sf(data = parcels_map %>% filter(count_above_104 == 0), 
          fill = "#F5F5F5", 
          color = "#E0E0E0", 
          size = 0.1) +
  geom_sf(data = parcels_map %>% filter(count_above_104 > 0), 
          aes(fill = count_above_104), 
          color = "#FFFFFF", 
          size = 0.2) +
  scale_fill_viridis(
    option = "magma",
    name = "Count of samples\nwith sum > 104%",
    begin = 0.1,
    end = 0.95,
    direction = 1,
    breaks = function(x) unique(floor(pretty(x))),
    guide = guide_colorbar(
      barwidth = 1.5,
      barheight = 12,
      title.position = "top",
      title.hjust = 0.5,
      frame.colour = "black",
      ticks.colour = "black",
      ticks.linewidth = 0.5
    )
  ) +
  labs(
    title = "Spatial distribution of high granulometry sum observations",
    subtitle = "Count of sampling events with granulometry sum > 104% per parcel",
    caption = "Gray parcels indicate no high granulometry observations"
  ) +
  theme_minimal() +
  theme(
    text = element_text(family = "sans", color = "black"),
    plot.title = element_text(size = 14, face = "bold", 
                              margin = margin(b = 8), hjust = 0),
    plot.subtitle = element_text(size = 11, color = "grey30", 
                                 margin = margin(b = 12), hjust = 0),
    plot.caption = element_text(size = 9, color = "grey50", 
                                margin = margin(t = 12), hjust = 0),
    legend.position = "right",
    legend.title = element_text(size = 10, face = "bold", lineheight = 1.2),
    legend.text = element_text(size = 9),
    legend.background = element_rect(fill = "white", color = NA),
    legend.key = element_rect(color = "black", linewidth = 0.3),
    panel.background = element_rect(fill = "white", color = NA),
    panel.grid.major = element_line(color = "#F0F0F0", linewidth = 0.3),
    panel.grid.minor = element_blank(),
    plot.background = element_rect(fill = "white", color = NA),
    plot.margin = margin(15, 15, 15, 15),
    axis.text = element_text(size = 8, color = "grey50"),
    axis.title = element_blank()
  )

ggsave(
  filename = file.path(results_folder, "spatial_map_above_104.png"),
  plot = map3,
  width = 10,
  height = 8,
  dpi = 600,
  bg = "white"
)

ggsave(
  filename = file.path(results_folder, "spatial_map_above_104.pdf"),
  plot = map3,
  width = 10,
  height = 8,
  device = cairo_pdf,
  bg = "white"
)

cat("✅ Map 3 (> 104%) saved\n")

# ============================================================================
# SAVE SUMMARY STATISTICS FOR ALL CATEGORIES
# ============================================================================

summary_all <- data.frame(
  Category = rep(c("< 96%", "96-104%", "> 104%"), each = 8),
  Metric = rep(c(
    "Total parcels",
    "Parcels with observations",
    "Parcels without observations",
    "Percentage with observations",
    "Min count",
    "Max count",
    "Mean count",
    "Median count"
  ), 3),
  Value = c(
    # < 96%
    nrow(parcels_map),
    sum(parcels_map$count_below_96 > 0),
    sum(parcels_map$count_below_96 == 0),
    paste0(round(sum(parcels_map$count_below_96 > 0) / nrow(parcels_map) * 100, 2), "%"),
    ifelse(nrow(parcel_low_count) > 0, min(parcel_low_count$count_below_96), 0),
    ifelse(nrow(parcel_low_count) > 0, max(parcel_low_count$count_below_96), 0),
    ifelse(nrow(parcel_low_count) > 0, round(mean(parcel_low_count$count_below_96), 2), 0),
    ifelse(nrow(parcel_low_count) > 0, median(parcel_low_count$count_below_96), 0),
    
    # 96-104%
    nrow(parcels_map),
    sum(parcels_map$count_96_104 > 0),
    sum(parcels_map$count_96_104 == 0),
    paste0(round(sum(parcels_map$count_96_104 > 0) / nrow(parcels_map) * 100, 2), "%"),
    ifelse(nrow(parcel_normal_count) > 0, min(parcel_normal_count$count_96_104), 0),
    ifelse(nrow(parcel_normal_count) > 0, max(parcel_normal_count$count_96_104), 0),
    ifelse(nrow(parcel_normal_count) > 0, round(mean(parcel_normal_count$count_96_104), 2), 0),
    ifelse(nrow(parcel_normal_count) > 0, median(parcel_normal_count$count_96_104), 0),
    
    # > 104%
    nrow(parcels_map),
    sum(parcels_map$count_above_104 > 0),
    sum(parcels_map$count_above_104 == 0),
    paste0(round(sum(parcels_map$count_above_104 > 0) / nrow(parcels_map) * 100, 2), "%"),
    ifelse(nrow(parcel_high_count) > 0, min(parcel_high_count$count_above_104), 0),
    ifelse(nrow(parcel_high_count) > 0, max(parcel_high_count$count_above_104), 0),
    ifelse(nrow(parcel_high_count) > 0, round(mean(parcel_high_count$count_above_104), 2), 0),
    ifelse(nrow(parcel_high_count) > 0, median(parcel_high_count$count_above_104), 0)
  )
)

write.csv(summary_all, 
          file.path(results_folder, "spatial_summary_all_categories.csv"), 
          row.names = FALSE)

```

```{r}
# ============================================================================
# CREATE CLEAN GEOPACKAGE - REMOVE GRANULOMETRY DATA FOR SUM < 96% OR > 104%
# ============================================================================
cat("\n🧹 CREATING CLEAN GEOPACKAGE:\n")
cat("=", rep("=", 60), "\n", sep="")

# Create a copy of the filtered data
parcelles_clean <- parcelles_filtered

# Identify observations to clean (sum < 96% OR sum > 104%)
obs_to_clean_low <- which(parcelles_clean$granulo_sum < 96)
obs_to_clean_high <- which(parcelles_clean$granulo_sum > 104)
obs_to_clean_all <- unique(c(obs_to_clean_low, obs_to_clean_high))

cat("📊 Observations to clean:\n")
cat("  - Sum < 96%:", length(obs_to_clean_low), "observations\n")
cat("  - Sum > 104%:", length(obs_to_clean_high), "observations\n")
cat("  - Total to clean:", length(obs_to_clean_all), "observations\n")

# Set granulometry fractions to NA for problematic observations
for (var in GRANULO_FRACTIONS) {
  parcelles_clean[obs_to_clean_all, var] <- NA
}

# Recalculate granulo_sum after cleaning
parcelles_clean$granulo_sum <- rowSums(
  st_drop_geometry(parcelles_clean)[, GRANULO_FRACTIONS] %>% 
    mutate(across(everything(), as.numeric)),
  na.rm = FALSE
)

# Recalculate categories
parcelles_clean$sum_category <- cut(
  parcelles_clean$granulo_sum,
  breaks = c(-Inf, 96, 104, Inf),
  labels = c("< 96%", "96-104%", "> 104%"),
  right = FALSE
)

# Check results
cat("\n✅ After cleaning - Distribution by category:\n")
sum_table_clean <- table(parcelles_clean$sum_category, useNA = "ifany")
print(sum_table_clean)
cat("\nPercentages:\n")
print(round(prop.table(sum_table_clean) * 100, 2))

# Count valid granulometry observations
valid_granulo <- sum(!is.na(parcelles_clean$granulo_sum))
cat("\n📊 Valid granulometry observations (96-104%):", valid_granulo, "\n")
cat("📊 Invalid granulometry (set to NA):", length(obs_to_clean_all), "\n")

# Save clean geopackage
clean_gpkg_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/parcels_with_soil_clean_granulo.gpkg"

st_write(
  parcelles_clean,
  clean_gpkg_path,
  driver = "GPKG",
  delete_dsn = TRUE,  # Overwrite if exists
  quiet = FALSE
)

cat("\n✅ Clean geopackage saved to:\n")
cat("  ", clean_gpkg_path, "\n")

# ============================================================================
# SUMMARY STATISTICS OF CLEAN DATASET
# ============================================================================
cat("\n📋 CLEAN DATASET SUMMARY:\n")
cat("=", rep("=", 60), "\n", sep="")
cat("Total observations:", nrow(parcelles_clean), "\n")
cat("Observations with valid granulometry (96-104%):", valid_granulo, 
    "(", round(valid_granulo/nrow(parcelles_clean)*100, 2), "%)\n")
cat("Observations with NA granulometry:", length(obs_to_clean_all),
    "(", round(length(obs_to_clean_all)/nrow(parcelles_clean)*100, 2), "%)\n")

# Show summary of valid granulometry sums
cat("\n📊 Summary of valid granulometry sums (96-104%):\n")
print(summary(parcelles_clean$granulo_sum[parcelles_clean$granulo_sum >= 96 & 
                                          parcelles_clean$granulo_sum <= 104]))

```

## II.4. NORMALIZING GRANULOMETRY FRACTIONS TO 100%

```{r}
# ============================================================================
# NORMALIZE GRANULOMETRY FRACTIONS TO 100%
# ============================================================================
cat("\n⚖️ NORMALIZING GRANULOMETRY FRACTIONS TO 100%:\n")
cat("=", rep("=", 60), "\n", sep="")

# ============================================================================
# METHOD 1: PROPORTIONAL CLOSURE 
# ============================================================================
# This is the standard method in compositional data analysis (Aitchison, 1986)
# It preserves the relative proportions between fractions
# Formula: normalized_value = (original_value / sum) * 100

cat("\n📐 METHOD 1: Proportional Closure (Aitchison Closure)\n")
cat("Reference: Aitchison, J. (1986). The Statistical Analysis of Compositional Data.\n\n")

# Work with a non-spatial version for easier manipulation
data_df <- st_drop_geometry(parcelles_clean)

# Only normalize observations with valid granulometry (96-104%)
valid_obs <- which(data_df$granulo_sum >= 96 & data_df$granulo_sum <= 104)

cat("Normalizing", length(valid_obs), "observations with sum between 96-104%\n")

for (i in valid_obs) {
  # Get current sum
  current_sum <- data_df$granulo_sum[i]
  
  # Calculate normalization factor
  norm_factor <- 100 / current_sum
  
  # Apply to each fraction
  for (var in GRANULO_FRACTIONS) {
    current_value <- as.numeric(data_df[i, var])
    if (length(current_value) == 1 && !is.na(current_value)) {
      data_df[i, var] <- current_value * norm_factor
    }
  }
}

# Recalculate sums to verify
data_df$granulo_sum_normalized <- rowSums(
  data_df[, GRANULO_FRACTIONS] %>% 
    mutate(across(everything(), as.numeric)),
  na.rm = FALSE
)

cat("\n✅ Verification of normalized sums:\n")
print(summary(data_df$granulo_sum_normalized[valid_obs]))

# Check if all are 100% (within rounding error)
all_100 <- all(abs(data_df$granulo_sum_normalized[valid_obs] - 100) < 0.001, 
               na.rm = TRUE)
cat("All normalized sums equal 100%:", all_100, "\n")

# Recreate sf object with normalized data
parcelles_normalized <- st_sf(data_df, geometry = st_geometry(parcelles_clean))

# ============================================================================
# SAVE NORMALIZED GEOPACKAGE
# ============================================================================
normalized_gpkg_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/parcels_with_soil_normalized_granulo.gpkg"

st_write(
  parcelles_normalized,
  normalized_gpkg_path,
  driver = "GPKG",
  delete_dsn = TRUE,
  quiet = FALSE
)

cat("\n✅ Normalized geopackage saved to:\n")
cat("  ", normalized_gpkg_path, "\n")

# ============================================================================
# COMPARISON: BEFORE vs AFTER NORMALIZATION
# ============================================================================
cat("\n📊 COMPARISON OF EXAMPLE OBSERVATIONS:\n")
cat("=", rep("=", 60), "\n", sep="")

# Show 5 random examples
set.seed(123)
sample_obs <- sample(valid_obs, min(5, length(valid_obs)))

data_clean <- st_drop_geometry(parcelles_clean)
data_norm <- st_drop_geometry(parcelles_normalized)

for (idx in sample_obs) {
  cat("\nObservation", idx, ":\n")
  cat("  Original sum:", round(data_clean$granulo_sum[idx], 2), "%\n")
  cat("  Normalized sum:", round(data_norm$granulo_sum_normalized[idx], 2), "%\n")
  
  for (var in GRANULO_FRACTIONS) {
    orig <- as.numeric(data_clean[idx, var])
    norm <- as.numeric(data_norm[idx, var])
    cat(sprintf("    %s: %.2f%% → %.2f%%\n", var, orig, norm))
  }
}
```

# III. INLA Modelling

```{r}
# ============================================================================
# MULTI-VARIABLE ENHANCED SPATIO-TEMPORAL IMPUTATION WITH TRAIN/VALIDATION SPLIT
# Compositional treatment for granulometry fractions
# ============================================================================

library(sf)
library(INLA)
library(dplyr)
library(ggplot2)
library(viridis)
library(spdep)
library(readr)
library(purrr)
library(tibble)

cat("========================================\n")
cat("  MULTI-VARIABLE SPATIO-TEMPORAL       \n")
cat("  IMPUTATION WITH TRAIN/VAL SPLIT      \n")
cat("========================================\n")

# ============================================================================
# STEP 1: LOAD THE COMPLETE DATASET AND IDENTIFY SOIL VARIABLES
# ============================================================================
cat("📂 LOADING COMPLETE DATASET AND IDENTIFYING SOIL VARIABLES:\n")
cat("=", rep("=", 60), "\n", sep="")

gpkg_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/parcels_with_soil_normalized_granulo.gpkg"

if (file.exists(gpkg_path)) {
  parcelles_complete <- st_read(gpkg_path, quiet = TRUE)
  cat("✅ Complete dataset loaded:", nrow(parcelles_complete), "observations\n")
} else {
  cat("❌ Complete dataset not found. Please run the enhanced join script first.\n")
  stop("Dataset not available")
}

# Define compositional and non-compositional variables
GRANULO_FRACTIONS <- c(
  "Clay_percent",
  "Fine_Silt_percent",
  "Coarse_Silt_percent",
  "Fine_Sand_percent",
  "Coarse_Sand_percent"
)

NON_COMPOSITIONAL_VARIABLES <- c(
  "pH",
  "CN_Ratio", 
  "Electrical_Conductivity_mmho_cm",
  "Exchangeable_Calcium_meq_100g",
  "Exchangeable_Magnesium_meq_100g", 
  "Exchangeable_Sodium_meq_100g",
  "Exchangeable_Potassium_meq_100g",
  "Available_P2O5_mg_kg",
  "Organic_Carbon_per_mille",
  "Total_Nitrogen_per_mille",
  "Total_Silt_percent",
  "Total_Sand_percent", 
  "Organic_Matter_percent",
  "Sum_of_Exchangeable_Bases_meq_100g",
  "CaO_g_100g",
  "Mg_Ca_Ratio",
  "Mg_K_Ratio",
  "Compaction_Index"
)

cat("Compositional variables (granulometry):", length(GRANULO_FRACTIONS), "\n")
cat("Non-compositional variables:", length(NON_COMPOSITIONAL_VARIABLES), "\n")

for (i in seq_along(GRANULO_FRACTIONS)) {
  var_name <- GRANULO_FRACTIONS[i]
  n_available <- sum(!is.na(parcelles_complete[[var_name]]))
  pct_available <- round(n_available/nrow(parcelles_complete)*100, 1)
  cat(sprintf("• %2d. %-35s: %4d available (%5.1f%%), %4d missing\n", 
              i, var_name, n_available, pct_available, sum(is.na(parcelles_complete[[var_name]]))))
}

for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  n_available <- sum(!is.na(parcelles_complete[[var_name]]))
  pct_available <- round(n_available/nrow(parcelles_complete)*100, 1)
  cat(sprintf("• %2d. %-35s: %4d available (%5.1f%%), %4d missing\n", 
              length(GRANULO_FRACTIONS) + i, var_name, n_available, pct_available, sum(is.na(parcelles_complete[[var_name]]))))
}

cat("\nDataset structure:\n")
cat("• Total observations:", nrow(parcelles_complete), "\n")
cat("• Unique parcels:", length(unique(parcelles_complete$Final_Cleaned_ID)), "\n")
cat("• Year range:", min(parcelles_complete$Year, na.rm = TRUE), "-", max(parcelles_complete$Year, na.rm = TRUE), "\n")

# Validate compositional data
parcelles_complete <- parcelles_complete %>%
  mutate(
    has_complete_granulo = !is.na(Clay_percent) & 
                          !is.na(Fine_Silt_percent) & 
                          !is.na(Coarse_Silt_percent) & 
                          !is.na(Fine_Sand_percent) & 
                          !is.na(Coarse_Sand_percent),
    granulo_sum = Clay_percent + Fine_Silt_percent + Coarse_Silt_percent + 
                  Fine_Sand_percent + Coarse_Sand_percent
  )

complete_granulo <- sum(parcelles_complete$has_complete_granulo, na.rm = TRUE)
cat("\n🔍 COMPOSITIONAL DATA VALIDATION:\n")
cat("• Observations with complete granulometry:", complete_granulo, "\n")

if(complete_granulo > 0) {
  sum_stats <- parcelles_complete %>%
    filter(has_complete_granulo) %>%
    pull(granulo_sum) %>%
    summary()
  cat("• Sum range:", round(sum_stats[1], 2), "to", round(sum_stats[6], 2), "%\n")
}

# ============================================================================
# STEP 2: PREPARE SHARED SPATIAL STRUCTURE
# ============================================================================
cat("\n🗺️ PREPARING SHARED SPATIAL STRUCTURE:\n")
cat("=", rep("=", 40), "\n", sep="")

unique_parcels_spatial <- parcelles_complete %>%
  group_by(Final_Cleaned_ID) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(Final_Cleaned_ID)

cat("Unique parcels for spatial analysis:", nrow(unique_parcels_spatial), "\n")

unique_parcels_spatial <- st_make_valid(unique_parcels_spatial)
cat("Geometries validated\n")

field_mapping <- data.frame(
  Original_ID = unique_parcels_spatial$Final_Cleaned_ID,
  Sequential_ID = 1:nrow(unique_parcels_spatial)
)

cat("Field mapping created: 1 to", nrow(field_mapping), "\n")

unique_parcels_spatial <- unique_parcels_spatial %>%
  left_join(field_mapping, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
  arrange(Sequential_ID)

# ============================================================================
# STEP 3: CREATE SPATIAL WEIGHTS MATRIX W.P1
# ============================================================================
cat("\n🔗 CREATING SHARED SPATIAL WEIGHTS MATRIX:\n")
cat("=", rep("=", 42), "\n", sep="")

cat("Computing boundary-to-boundary distances (this may take time)...\n")

boundaries <- st_boundary(unique_parcels_spatial)
dist_matrix <- st_distance(boundaries, boundaries)
dist_matrix <- units::drop_units(dist_matrix)
diag(dist_matrix) <- NA

cat("Distance matrix calculated:", nrow(dist_matrix), "x", ncol(dist_matrix), "\n")

DISTANCE_THRESHOLD <- 110
cat("Using distance threshold:", DISTANCE_THRESHOLD, "meters\n")

W.P1_complete <- matrix(0, nrow = nrow(unique_parcels_spatial), ncol = nrow(unique_parcels_spatial))
W.P1_complete[dist_matrix <= DISTANCE_THRESHOLD & !is.na(dist_matrix)] <- 1
diag(W.P1_complete) <- 0
rownames(W.P1_complete) <- colnames(W.P1_complete) <- 1:nrow(W.P1_complete)

n_connections <- sum(W.P1_complete)
avg_neighbors <- n_connections / nrow(unique_parcels_spatial)

cat("✅ W.P1 matrix created:\n")
cat("• Dimensions:", nrow(W.P1_complete), "x", ncol(W.P1_complete), "\n")
cat("• Total connections:", n_connections, "\n")
cat("• Average neighbors per parcel:", round(avg_neighbors, 2), "\n")

# ============================================================================
# STEP 4: PREPARE SHARED TEMPORAL STRUCTURE
# ============================================================================
years_unique <- sort(unique(parcelles_complete$Year))
n_years <- length(years_unique)
year_mapping <- data.frame(
  Year = years_unique,
  ID.Year = 1:n_years
)

cat("Years mapped:", min(years_unique), "to", max(years_unique), "(", n_years, "years)\n")

# ============================================================================
# ILR TRANSFORMATION FUNCTIONS
# ============================================================================

create_ilr_basis <- function(D) {
  V <- matrix(0, nrow = D - 1, ncol = D)
  for (i in 1:(D - 1)) {
    V[i, 1:i] <- 1/i
    V[i, i + 1] <- -1
    V[i, ] <- V[i, ] * sqrt(i / (i + 1))
  }
  return(V)
}

ilr_transform <- function(composition_matrix) {
  D <- ncol(composition_matrix)
  V <- create_ilr_basis(D)
  log_comp <- log(composition_matrix)
  ilr_coords <- log_comp %*% t(V)
  return(ilr_coords)
}

ilr_inverse <- function(ilr_matrix) {
  D <- ncol(ilr_matrix) + 1
  V <- create_ilr_basis(D)
  log_comp <- ilr_matrix %*% V
  composition <- exp(log_comp)
  row_sums <- rowSums(composition)
  composition <- composition / row_sums
  return(composition)
}

# ============================================================================
# STEP 5: DEFINE IMPUTATION FUNCTIONS
# ============================================================================

# Non-compositional imputation (same as before)
impute_single_variable <- function(variable_name, parcelles_data, W_matrix, field_map, year_map, verbose = TRUE) {
  
  if (verbose) {
    cat("\n", "=", rep("=", 60), "\n", sep="")
    cat("🔬 PROCESSING VARIABLE:", variable_name, "\n")
    cat("=", rep("=", 60), "\n", sep="")
  }
  
  var_values <- parcelles_data[[variable_name]]
  n_total <- length(var_values)
  n_available <- sum(!is.na(var_values))
  n_missing <- n_total - n_available
  pct_available <- round(n_available/n_total*100, 1)
  
  if (verbose) {
    cat("Variable statistics:\n")
    cat("• Total observations:", n_total, "\n")
    cat("• Available (non-NA):", n_available, "(", pct_available, "%)\n")
    cat("• Missing (NA):", n_missing, "(", round(n_missing/n_total*100, 1), "%)\n")
    
    if (n_available > 0) {
      cat("• Min value:", round(min(var_values, na.rm = TRUE), 3), "\n")
      cat("• Max value:", round(max(var_values, na.rm = TRUE), 3), "\n")
      cat("• Mean value:", round(mean(var_values, na.rm = TRUE), 3), "\n")
      cat("• Std deviation:", round(sd(var_values, na.rm = TRUE), 3), "\n")
    }
  }
  
  if (n_available < 10) {
    if (verbose) cat("❌ Insufficient data (< 10 observations). Skipping variable.\n")
    return(NULL)
  }
  
  observed_data <- parcelles_data %>% filter(!is.na(.data[[variable_name]]))
  
  set.seed(123)
  unique_parcels_with_data <- observed_data %>%
    st_drop_geometry() %>%
    dplyr::select(Final_Cleaned_ID) %>%
    distinct() %>%
    mutate(split = sample(c("train", "val"), n(), replace = TRUE, prob = c(0.9, 0.1)))
  
  parcelles_data_split <- parcelles_data %>%
    left_join(unique_parcels_with_data, by = "Final_Cleaned_ID") %>%
    mutate(
      data_split = case_when(
        is.na(.data[[variable_name]]) ~ "predict",
        split == "train" ~ "train",
        split == "val" ~ "validation",
        TRUE ~ "predict"
      )
    ) %>%
    dplyr::select(-split)
  
  train_obs <- sum(parcelles_data_split$data_split == "train")
  val_obs <- sum(parcelles_data_split$data_split == "validation")
  pred_obs <- sum(parcelles_data_split$data_split == "predict")
  
  if (verbose) {
    cat("Data split:\n")
    cat("• Training:", train_obs, "\n")
    cat("• Validation:", val_obs, "\n")
    cat("• To predict:", pred_obs, "\n")
  }
  
  df_inla <- parcelles_data_split %>%
    st_drop_geometry() %>%
    left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
    left_join(year_map, by = "Year") %>%
    mutate(
      Y = case_when(
        data_split == "train" ~ .data[[variable_name]],
        TRUE ~ NA_real_
      ),
      Y_true = .data[[variable_name]],
      ID.Field = Sequential_ID,
      ID.Year = as.integer(ID.Year),
      Month = as.integer(ifelse(is.na(Month), 6, Month)),
      Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
      Var = as.factor(ifelse(is.na(Var), "Unknown", as.character(Var))),
      irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
    ) %>%
    filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
    arrange(ID.Field, ID.Year)
  
  if (max(df_inla$ID.Field) > nrow(W_matrix)) {
    if (verbose) cat("❌ Spatial structure incompatible. Skipping variable.\n")
    return(NULL)
  }
  
  if (verbose) cat("🚀 Fitting spatio-temporal model...\n")
  
  start_time <- Sys.time()
  
  fit_result <- tryCatch({
    prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
    
    inla(
      Y ~ 1 + Bloc + irrg_ty +
        f(ID.Field, model = "besag", graph = W_matrix,
          group = ID.Year, 
          control.group = list(model = "ar1"),
          hyper = list(prec = prec_prior),
          constr = TRUE),
      data = df_inla,
      family = "gaussian",
      control.predictor = list(compute = TRUE, link = 1),
      control.compute = list(dic = TRUE, waic = TRUE, cpo = TRUE),
      control.inla = list(strategy = "adaptive", int.strategy = "eb"),
      verbose = FALSE
    )
  }, error = function(e) {
    if (verbose) cat("❌ Full model failed. Trying simpler model...\n")
    
    tryCatch({
      inla(
        Y ~ 1 + Bloc + f(ID.Field, model = "besag", graph = W_matrix, constr = TRUE),
        data = df_inla,
        family = "gaussian",
        control.predictor = list(compute = TRUE, link = 1),
        control.compute = list(dic = TRUE),
        verbose = FALSE
      )
    }, error = function(e2) {
      if (verbose) cat("❌ Simple model also failed:", e2$message, "\n")
      return(NULL)
    })
  })
  
  end_time <- Sys.time()
  fitting_time <- as.numeric(difftime(end_time, start_time, units = "mins"))
  
  if (is.null(fit_result) || is.null(fit_result$summary.fitted.values)) {
    if (verbose) cat("❌ Model fitting failed completely\n")
    return(NULL)
  }
  
  if (verbose) {
    cat("✅ Model successful! Fitting time:", round(fitting_time, 1), "minutes\n")
    if (!is.null(fit_result$dic)) cat("DIC:", round(fit_result$dic$dic, 2), "\n")
    if (!is.null(fit_result$waic)) cat("WAIC:", round(fit_result$waic$waic, 2), "\n")
  }
  
  predictions <- fit_result$summary.fitted.values
  
  df_predicted <- df_inla %>%
    mutate(
      !!paste0(variable_name, "_predicted") := predictions$mean,
      !!paste0(variable_name, "_predicted_sd") := predictions$sd,
      !!paste0(variable_name, "_predicted_lower") := predictions$`0.025quant`,
      !!paste0(variable_name, "_predicted_upper") := predictions$`0.975quant`,
      !!paste0(variable_name, "_complete") := case_when(
        data_split == "train" ~ Y,
        TRUE ~ predictions$mean
      ),
      !!paste0(variable_name, "_prediction_type") := case_when(
        data_split == "train" ~ "Training",
        data_split == "validation" ~ "Validation", 
        data_split == "predict" ~ "Predicted"
      )
    )
  
  validation_data <- df_predicted %>%
    filter(data_split == "validation", !is.na(Y_true), !is.na(predictions$mean))
  
  metrics <- list(variable = variable_name)
  
  if (nrow(validation_data) > 5) {
    pred_vals <- validation_data[[paste0(variable_name, "_predicted")]]
    true_vals <- validation_data$Y_true
    
    ss_res <- sum((true_vals - pred_vals)^2)
    ss_tot <- sum((true_vals - mean(true_vals))^2)
    r_squared <- 1 - (ss_res / ss_tot)
    
    metrics$r_squared <- r_squared
    metrics$correlation <- cor(true_vals, pred_vals)
    metrics$rmse <- sqrt(mean((true_vals - pred_vals)^2))
    metrics$mae <- mean(abs(true_vals - pred_vals))
    metrics$mean_error <- mean(true_vals - pred_vals)
    metrics$n_validation <- nrow(validation_data)
    
    in_interval <- sum(validation_data$Y_true >= validation_data[[paste0(variable_name, "_predicted_lower")]] & 
                      validation_data$Y_true <= validation_data[[paste0(variable_name, "_predicted_upper")]])
    metrics$coverage_prob <- in_interval / nrow(validation_data)
    
    if (verbose) {
      cat("🎯 Validation Results:\n")
      cat("• Validation obs:", metrics$n_validation, "\n")
      cat("• R²:", round(metrics$r_squared, 4), "\n")
      cat("• Correlation:", round(metrics$correlation, 4), "\n") 
      cat("• RMSE:", round(metrics$rmse, 4), "\n")
      cat("• MAE:", round(metrics$mae, 4), "\n")
      cat("• Coverage:", round(metrics$coverage_prob * 100, 1), "%\n")
    }
  } else {
    metrics$r_squared <- NA
    metrics$correlation <- NA
    metrics$rmse <- NA
    metrics$mae <- NA
    metrics$mean_error <- NA
    metrics$n_validation <- nrow(validation_data)
    metrics$coverage_prob <- NA
    
    if (verbose) cat("❌ Insufficient validation data for metrics\n")
  }
  
  result <- list(
    variable_name = variable_name,
    model_fit = fit_result,
    predictions = df_predicted,
    metrics = metrics,
    fitting_time = fitting_time,
    n_total = n_total,
    n_available = n_available,
    n_missing = n_missing,
    pct_available = pct_available
  )
  
  return(result)
}

# Compositional imputation for granulometry
impute_compositional_granulometry <- function(parcelles_data, W_matrix, field_map, year_map, verbose = TRUE) {
  
  if (verbose) {
    cat("\n", "=", rep("=", 60), "\n", sep="")
    cat("🔬 PROCESSING COMPOSITIONAL GRANULOMETRY (ILR)\n")
    cat("=", rep("=", 60), "\n", sep="")
  }
  
  parcelles_data <- parcelles_data %>%
    mutate(obs_id = row_number())
  
  sampled_obs <- parcelles_data %>%
    filter(has_complete_granulo)
  
  unsampled_obs <- parcelles_data %>%
    filter(!has_complete_granulo)
  
  n_sampled <- nrow(sampled_obs)
  n_unsampled <- nrow(unsampled_obs)
  
  if (verbose) {
    cat("• Sampled observations:", n_sampled, "\n")
    cat("• Unsampled observations:", n_unsampled, "\n")
  }
  
  if (n_sampled < 20) {
    if (verbose) cat("❌ Insufficient sampled observations\n")
    return(NULL)
  }
  
  comp_matrix <- sampled_obs %>%
    st_drop_geometry() %>%
    dplyr::select(all_of(GRANULO_FRACTIONS)) %>%
    as.matrix()
  
  comp_matrix[comp_matrix <= 0] <- 0.01
  row_sums <- rowSums(comp_matrix)
  comp_matrix <- comp_matrix / row_sums * 100
  
  ilr_coords <- ilr_transform(comp_matrix / 100)
  colnames(ilr_coords) <- paste0("ILR_", 1:4)
  
  if (verbose) {
    cat("✅ ILR transformation: 5 fractions -> 4 orthonormal coordinates\n")
  }
  
  sampled_obs_ilr <- sampled_obs %>%
    bind_cols(as.data.frame(ilr_coords))
  
  set.seed(123)
  unique_sampled_parcels <- sampled_obs_ilr %>%
    st_drop_geometry() %>%
    dplyr::select(Final_Cleaned_ID) %>%
    distinct() %>%
    mutate(split = sample(c("train", "val"), n(), replace = TRUE, prob = c(0.85, 0.15)))
  
  cat("• Unique parcels in training:", sum(unique_sampled_parcels$split == "train"), "\n")
  cat("• Unique parcels in validation:", sum(unique_sampled_parcels$split == "val"), "\n")
  
  sampled_obs_split <- sampled_obs_ilr %>%
    left_join(unique_sampled_parcels, by = "Final_Cleaned_ID") %>%
    mutate(data_split = ifelse(split == "train", "train", "validation"))
  
  all_obs_for_prediction <- bind_rows(
    sampled_obs_split %>% mutate(obs_type = "sampled"),
    unsampled_obs %>% mutate(obs_type = "unsampled", data_split = "predict")
  )
  
  train_obs <- sum(sampled_obs_split$data_split == "train")
  val_obs <- sum(sampled_obs_split$data_split == "validation")
  
  if (verbose) {
    cat("• Training observations:", train_obs, "\n")
    cat("• Validation observations:", val_obs, "\n")
    cat("• To predict (unsampled):", n_unsampled, "\n")
  }
  
  ilr_results <- list()
  
  for (i in 1:4) {
    ilr_var <- paste0("ILR_", i)
    
    if (verbose) cat("\n  Fitting", ilr_var, "...\n")
    
    df_inla <- all_obs_for_prediction %>%
      st_drop_geometry() %>%
      left_join(field_map, by = c("Final_Cleaned_ID" = "Original_ID")) %>%
      left_join(year_map, by = "Year") %>%
      mutate(
        Y = case_when(
          data_split == "train" ~ .data[[ilr_var]],
          TRUE ~ NA_real_
        ),
        Y_true = .data[[ilr_var]],
        ID.Field = Sequential_ID,
        ID.Year = as.integer(ID.Year),
        Month = as.integer(ifelse(is.na(Month), 6, Month)),
        Bloc = as.factor(ifelse(is.na(Bloc), "Unknown", as.character(Bloc))),
        irrg_ty = as.factor(ifelse(is.na(irrg_ty), "Unknown", as.character(irrg_ty)))
      ) %>%
      filter(!is.na(ID.Field), !is.na(ID.Year)) %>%
      arrange(ID.Field, ID.Year)
    
    fit_result <- tryCatch({
      prec_prior <- list(theta = list(prior = "loggamma", param = c(1, 0.005)))
      
      inla(
        Y ~ 1 + Bloc + irrg_ty +
          f(ID.Field, model = "besag", graph = W_matrix,
            group = ID.Year, 
            control.group = list(model = "ar1"),
            hyper = list(prec = prec_prior),
            constr = TRUE),
        data = df_inla,
        family = "gaussian",
        control.predictor = list(compute = TRUE, link = 1),
        control.compute = list(dic = TRUE, waic = TRUE),
        control.inla = list(strategy = "adaptive", int.strategy = "eb"),
        verbose = FALSE
      )
    }, error = function(e) {
      tryCatch({
        inla(
          Y ~ 1 + Bloc + f(ID.Field, model = "besag", graph = W_matrix, constr = TRUE),
          data = df_inla,
          family = "gaussian",
          control.predictor = list(compute = TRUE, link = 1),
          control.compute = list(dic = TRUE),
          verbose = FALSE
        )
      }, error = function(e2) NULL)
    })
    
    if (!is.null(fit_result)) {
      predictions <- fit_result$summary.fitted.values
      
      df_predicted <- df_inla %>%
        mutate(
          !!paste0(ilr_var, "_predicted") := predictions$mean,
          !!paste0(ilr_var, "_predicted_sd") := predictions$sd
        )
      
      val_data <- df_predicted %>%
        filter(data_split == "validation", !is.na(Y_true))
      
      metrics <- list()
      if (nrow(val_data) > 5) {
        pred_vals <- val_data[[paste0(ilr_var, "_predicted")]]
        true_vals <- val_data$Y_true
        
        ss_res <- sum((true_vals - pred_vals)^2)
        ss_tot <- sum((true_vals - mean(true_vals))^2)
        
        metrics$r_squared <- 1 - (ss_res / ss_tot)
        metrics$rmse <- sqrt(mean((true_vals - pred_vals)^2))
        metrics$n_val <- nrow(val_data)
      }
      
      ilr_results[[ilr_var]] <- list(
        model = fit_result,
        predictions = df_predicted,
        metrics = metrics
      )
      
      if (verbose && !is.null(metrics$r_squared)) {
        cat("    ✅ R² (validation):", round(metrics$r_squared, 4), 
            "| n =", metrics$n_val, "\n")
      }
    } else {
      if (verbose) cat("    ❌ Failed\n")
      return(NULL)
    }
  }
  
  if (verbose) cat("\n🔄 Transforming back to composition...\n")
  
  all_predictions <- ilr_results[[1]]$predictions %>%
    dplyr::select(obs_id, Final_Cleaned_ID, Year, data_split, obs_type,
                  ILR_1_predicted, ILR_1_predicted_sd)
  
  for (i in 2:4) {
    ilr_var <- paste0("ILR_", i)
    pred_cols <- ilr_results[[ilr_var]]$predictions %>%
      dplyr::select(obs_id, !!paste0(ilr_var, "_predicted"), !!paste0(ilr_var, "_predicted_sd"))
    
    all_predictions <- all_predictions %>%
      left_join(pred_cols, by = "obs_id")
  }
  
  ilr_pred_matrix <- all_predictions %>%
    dplyr::select(ILR_1_predicted, ILR_2_predicted, ILR_3_predicted, ILR_4_predicted) %>%
    as.matrix()
  
  comp_predicted <- ilr_inverse(ilr_pred_matrix)
  comp_predicted_pct <- comp_predicted * 100
  colnames(comp_predicted_pct) <- GRANULO_FRACTIONS
  
  if (verbose) {
    cat("✅ ILR back-transformation complete\n")
    sums_check <- rowSums(comp_predicted_pct)
    cat("• Sum range:", round(min(sums_check), 2), "to", round(max(sums_check), 2), "%\n")
  }
  
  final_results <- all_predictions %>%
    bind_cols(as.data.frame(comp_predicted_pct) %>%
                rename_with(~paste0(., "_predicted"), all_of(GRANULO_FRACTIONS)))
  
  for (frac in GRANULO_FRACTIONS) {
    original_map <- all_obs_for_prediction %>%
      st_drop_geometry() %>%
      dplyr::select(obs_id, !!frac) %>%
      rename(original_value = !!frac)
    
    final_results <- final_results %>%
      left_join(original_map, by = "obs_id")
    
    pred_col <- paste0(frac, "_predicted")
    
    final_results[[paste0(frac, "_complete")]] <- ifelse(
      final_results$data_split == "train",
      final_results$original_value,
      final_results[[pred_col]]
    )
    
    final_results[[paste0(frac, "_prediction_type")]] <- final_results$data_split
    
    final_results <- final_results %>% dplyr::select(-original_value)
  }
  
  overall_metrics <- list()
  val_subset <- final_results %>% filter(data_split == "validation")
  
  if (nrow(val_subset) > 5) {
    frac_metrics <- list()
    
    for (frac in GRANULO_FRACTIONS) {
      true_map <- sampled_obs_split %>%
        filter(data_split == "validation") %>%
        st_drop_geometry() %>%
        dplyr::select(obs_id, !!frac) %>%
        rename(true_val = !!frac)
      
      val_comparison <- val_subset %>%
        left_join(true_map, by = "obs_id")
      
      if (sum(!is.na(val_comparison$true_val)) > 5) {
        true_vals <- val_comparison$true_val
        pred_vals <- val_comparison[[paste0(frac, "_predicted")]]
        
        valid_idx <- !is.na(true_vals) & !is.na(pred_vals)
        true_vals <- true_vals[valid_idx]
        pred_vals <- pred_vals[valid_idx]
        
        if (length(true_vals) > 0) {
          ss_res <- sum((true_vals - pred_vals)^2)
          ss_tot <- sum((true_vals - mean(true_vals))^2)
          
          frac_metrics[[frac]] <- list(
            r_squared = 1 - (ss_res / ss_tot),
            rmse = sqrt(mean((true_vals - pred_vals)^2)),
            mae = mean(abs(true_vals - pred_vals)),
            n = length(true_vals)
          )
        }
      }
    }
    
    overall_metrics$fraction_metrics <- frac_metrics
    overall_metrics$n_validation <- nrow(val_subset)
    
    if (verbose) {
      cat("\n🎯 Validation Results (Back-transformed Fractions):\n")
      cat("• Validation observations:", nrow(val_subset), "\n")
      for (frac in GRANULO_FRACTIONS) {
        if (!is.null(frac_metrics[[frac]])) {
          cat("• ", frac, ": R²=", round(frac_metrics[[frac]]$r_squared, 4),
              " RMSE=", round(frac_metrics[[frac]]$rmse, 2), "% (n=", 
              frac_metrics[[frac]]$n, ")\n", sep="")
        }
      }
    }
  }
  
  return(list(
    predictions = final_results,
    ilr_results = ilr_results,
    metrics = overall_metrics,
    n_sampled = n_sampled,
    n_unsampled = n_unsampled
  ))
}

# ============================================================================
# STEP 6: PROCESS ALL VARIABLES
# ============================================================================
cat("\n🔄 PROCESSING ALL VARIABLES:\n")
cat("=", rep("=", 50), "\n", sep="")

all_results <- list()
all_metrics <- data.frame()
failed_variables <- character(0)

# Process compositional granulometry first
cat("\n[COMPOSITIONAL] Processing granulometry fractions...\n")

granulo_result <- impute_compositional_granulometry(
  parcelles_data = parcelles_complete,
  W_matrix = W.P1_complete,
  field_map = field_mapping,
  year_map = year_mapping,
  verbose = TRUE
)

if (!is.null(granulo_result)) {
  all_results[["Granulometry_Compositional"]] <- granulo_result
  
  if (!is.null(granulo_result$metrics$fraction_metrics)) {
    for (frac in GRANULO_FRACTIONS) {
      if (!is.null(granulo_result$metrics$fraction_metrics[[frac]])) {
        metrics_row <- data.frame(
          Variable = frac,
          N_Total = nrow(parcelles_complete),
          N_Available = granulo_result$n_sampled,
          N_Missing = granulo_result$n_unsampled,
          Pct_Available = round(granulo_result$n_sampled / nrow(parcelles_complete) * 100, 1),
          Fitting_Time_Min = NA,
          R_Squared = granulo_result$metrics$fraction_metrics[[frac]]$r_squared,
          Correlation = NA,
          RMSE = granulo_result$metrics$fraction_metrics[[frac]]$rmse,
          MAE = granulo_result$metrics$fraction_metrics[[frac]]$mae,
          Coverage_Prob = NA,
          N_Validation = granulo_result$metrics$fraction_metrics[[frac]]$n,
          DIC = NA,
          WAIC = NA,
          stringsAsFactors = FALSE
        )
        
        all_metrics <- rbind(all_metrics, metrics_row)
      }
    }
  }
  cat("✅ SUCCESS for compositional granulometry\n")
} else {
  failed_variables <- c(failed_variables, GRANULO_FRACTIONS)
  cat("❌ FAILED for compositional granulometry\n")
}

# Process non-compositional variables
for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  
  cat("\n[", i, "/", length(NON_COMPOSITIONAL_VARIABLES), "] Processing:", var_name, "\n")
  
  result <- impute_single_variable(
    variable_name = var_name,
    parcelles_data = parcelles_complete,
    W_matrix = W.P1_complete,
    field_map = field_mapping,
    year_map = year_mapping,
    verbose = TRUE
  )
  
  if (!is.null(result)) {
    all_results[[var_name]] <- result
    
    metrics_row <- data.frame(
      Variable = var_name,
      N_Total = result$n_total,
      N_Available = result$n_available,
      N_Missing = result$n_missing,
      Pct_Available = result$pct_available,
      Fitting_Time_Min = result$fitting_time,
      R_Squared = ifelse(is.null(result$metrics$r_squared), NA, result$metrics$r_squared),
      Correlation = ifelse(is.null(result$metrics$correlation), NA, result$metrics$correlation),
      RMSE = ifelse(is.null(result$metrics$rmse), NA, result$metrics$rmse),
      MAE = ifelse(is.null(result$metrics$mae), NA, result$metrics$mae),
      Coverage_Prob = ifelse(is.null(result$metrics$coverage_prob), NA, result$metrics$coverage_prob),
      N_Validation = ifelse(is.null(result$metrics$n_validation), 0, result$metrics$n_validation),
      DIC = ifelse(is.null(result$model_fit$dic), NA, result$model_fit$dic$dic),
      WAIC = ifelse(is.null(result$model_fit$waic), NA, result$model_fit$waic$waic),
      stringsAsFactors = FALSE
    )
    
    all_metrics <- rbind(all_metrics, metrics_row)
    cat("✅ SUCCESS for", var_name, "\n")
  } else {
    failed_variables <- c(failed_variables, var_name)
    cat("❌ FAILED for", var_name, "\n")
  }
}

# ============================================================================
# STEP 7: COMBINE ALL PREDICTIONS
# ============================================================================
cat("\n🔗 COMBINING ALL PREDICTIONS:\n")
cat("=", rep("=", 50), "\n", sep="")

successful_variables <- names(all_results)
cat("Successfully processed:", length(successful_variables), "variable groups\n")

final_dataset <- parcelles_complete %>% 
  st_drop_geometry() %>%
  group_by(Final_Cleaned_ID, Year) %>%
  mutate(
    replicate_id = row_number(),
    unique_key = paste0(Final_Cleaned_ID, "_", Year, "_rep", replicate_id)
  ) %>%
  ungroup() %>%
  mutate(original_row_id = row_number())

final_dataset <- parcelles_complete %>%
  mutate(original_row_id = row_number()) %>%
  left_join(
    final_dataset %>% dplyr::select(original_row_id, replicate_id, unique_key),
    by = "original_row_id"
  )

replicate_summary <- final_dataset %>% 
  st_drop_geometry() %>%
  count(Final_Cleaned_ID, Year) %>% 
  count(n, name = "combinations")

cat("📊 REPLICATE SUMMARY:\n")
for(i in 1:nrow(replicate_summary)) {
  n_reps <- replicate_summary$n[i]
  n_combinations <- replicate_summary$combinations[i]
  cat("•", n_combinations, "combinations have", n_reps, "replicate(s) each\n")
}

# Add granulometry predictions
if ("Granulometry_Compositional" %in% names(all_results)) {
  cat("\n[COMPOSITIONAL] Adding granulometry predictions...\n")
  
  final_dataset <- final_dataset %>%
    mutate(obs_id = row_number())
  
  granulo_preds <- all_results[["Granulometry_Compositional"]]$predictions
  
  final_dataset <- final_dataset %>%
    left_join(
      granulo_preds %>% 
        dplyr::select(obs_id, matches("^Clay_|^Fine_Silt_|^Coarse_Silt_|^Fine_Sand_|^Coarse_Sand_|^ILR_")),
      by = "obs_id"
    )
  
  check_cols <- paste0(GRANULO_FRACTIONS, "_complete")
  if (all(check_cols %in% colnames(final_dataset))) {
    sums <- final_dataset %>%
      st_drop_geometry() %>%
      filter(!is.na(Clay_percent_complete)) %>%
      rowwise() %>%
      mutate(sum_fractions = sum(c_across(all_of(check_cols)))) %>%
      pull(sum_fractions)
    
    if (length(sums) > 0) {
      cat("  ✅ Closure verified: sum range [", 
          round(min(sums), 2), ", ", round(max(sums), 2), "]%\n", sep="")
    }
  }
  
  final_dataset <- final_dataset %>% dplyr::select(-obs_id)
  cat("  ✅ Compositional predictions added\n")
}

# Add non-compositional predictions
for (i in seq_along(NON_COMPOSITIONAL_VARIABLES)) {
  var_name <- NON_COMPOSITIONAL_VARIABLES[i]
  
  if (var_name %in% names(all_results)) {
    cat("\n[", i, "/", length(NON_COMPOSITIONAL_VARIABLES), "] Processing:", var_name, "\n")
    
    result <- all_results[[var_name]]
    
    df_inla_reference <- parcelles_complete %>% 
      st_drop_geometry() %>% 
      left_join(field_mapping, by = c("Final_Cleaned_ID" = "Original_ID"), 
                relationship = "many-to-one") %>%  
      left_join(year_mapping, by = "Year", 
                relationship = "many-to-one") %>%  
      filter(!is.na(Sequential_ID), !is.na(ID.Year)) %>% 
      group_by(Final_Cleaned_ID, Year) %>%
      mutate(
        replicate_id = row_number(),
        unique_key = paste0(Final_Cleaned_ID, "_", Year, "_rep", replicate_id)
      ) %>%
      ungroup() %>%
      mutate(original_row_id = row_number()) %>% 
      arrange(Sequential_ID, ID.Year, replicate_id) %>%
      mutate(inla_row_id = row_number())
    
    cat("  Reference dataset rows:", nrow(df_inla_reference), "\n")
    cat("  Prediction rows:", nrow(result$predictions), "\n")
    
    if (nrow(df_inla_reference) == nrow(result$predictions)) {
      predictions_aligned <- df_inla_reference %>% 
        dplyr::select(Final_Cleaned_ID, Year, replicate_id, unique_key, original_row_id) %>% 
        bind_cols(
          result$predictions %>% 
            dplyr::select(matches(paste0("^", var_name, "_")))
        )
      
      if (any(duplicated(predictions_aligned$unique_key))) {
        cat("  ❌ Still have duplicates in unique_key - this shouldn't happen!\n")
        next
      }
      
      final_dataset <- final_dataset %>% 
        left_join(
          predictions_aligned %>% dplyr::select(-Final_Cleaned_ID, -Year, -replicate_id, -original_row_id),
          by = "unique_key"
        )
      
      first_pred_col <- paste0(var_name, "_predicted")
      var_cols <- colnames(final_dataset)[grepl(paste0("^", var_name), colnames(final_dataset))]
      
      if (first_pred_col %in% colnames(final_dataset)) {
        non_null_count <- sum(!is.na(final_dataset[[first_pred_col]]))
        cat("  ✅ Successfully added", non_null_count, "predictions for", var_name, "\n")
        check_col <- first_pred_col
      } else if (length(var_cols) > 0) {
        cat("  📋 Found columns:", paste(var_cols, collapse = ", "), "\n")
        non_null_count <- sum(!is.na(final_dataset[[var_cols[1]]]))
        cat("  ✅ Successfully added", non_null_count, "predictions for", var_name, "\n")
        check_col <- var_cols[1]
      } else {
        cat("  ❌ No prediction columns found for", var_name, "\n")
        check_col <- NULL
      }
      
      if (!is.null(check_col)) {
        replicate_dist <- final_dataset %>% 
          st_drop_geometry() %>%
          filter(!is.na(.data[[check_col]])) %>%
          count(replicate_id) %>%
          arrange(replicate_id)
        
        if (nrow(replicate_dist) > 0) {
          cat("    Predictions by replicate:", paste(paste0("rep", replicate_dist$replicate_id, ":", replicate_dist$n), collapse = ", "), "\n")
        }
      }
      
    } else {
      cat("  ❌ Row count mismatch for", var_name, "- skipping\n")
      cat("    Reference rows:", nrow(df_inla_reference), "\n")
      cat("    Prediction rows:", nrow(result$predictions), "\n")
    }
    
    gc()
  }
}

final_dataset <- final_dataset %>% 
  dplyr::select(-original_row_id)

cat("\n✅ Final combined dataset created\n")
cat("• Total observations:", nrow(final_dataset), "\n")
cat("• Total columns:", ncol(final_dataset), "\n")
cat("• Unique combinations (Final_Cleaned_ID + Year):", length(unique(paste0(final_dataset$Final_Cleaned_ID, "_", final_dataset$Year))), "\n")
cat("• Total replicates:", nrow(final_dataset), "\n")

prediction_cols <- colnames(final_dataset)[grepl("_predicted$|_complete$", colnames(final_dataset))]

if (length(prediction_cols) > 0) {
  cat("\n📊 PREDICTION SUMMARY:\n")
  for (col in prediction_cols) {
    non_null_count <- sum(!is.na(final_dataset[[col]]))
    cat("•", col, ":", non_null_count, "non-null values\n")
    
    if (non_null_count > 0) {
      rep_dist <- final_dataset %>% 
        st_drop_geometry() %>%
        filter(!is.na(.data[[col]])) %>%
        count(replicate_id, name = "count") %>%
        arrange(replicate_id)
      
      cat("  Distribution:", paste(paste0("rep", rep_dist$replicate_id, ":", rep_dist$count), collapse = ", "), "\n")
    }
  }
} else {
  cat("\n⚠️  No prediction columns found with standard naming pattern\n")
  cat("Available columns containing prediction variables:\n")
  
  all_var_names <- c(GRANULO_FRACTIONS, NON_COMPOSITIONAL_VARIABLES)
  for (var_name in all_var_names) {
    var_cols <- colnames(final_dataset)[grepl(var_name, colnames(final_dataset))]
    if (length(var_cols) > 0) {
      cat("• Variable", var_name, ":", paste(var_cols, collapse = ", "), "\n")
      first_col <- var_cols[1]
      non_null_vals <- final_dataset[[first_col]][!is.na(final_dataset[[first_col]])]
      if (length(non_null_vals) > 0) {
        cat("  Sample values:", paste(head(non_null_vals, 3), collapse = ", "), "...\n")
        
        rep_dist <- final_dataset %>% 
          st_drop_geometry() %>%
          filter(!is.na(.data[[first_col]])) %>%
          count(replicate_id, name = "count") %>%
          arrange(replicate_id)
        
        if (nrow(rep_dist) > 0) {
          cat("  Distribution:", paste(paste0("rep", rep_dist$replicate_id, ":", rep_dist$count), collapse = ", "), "\n")
        }
      }
    }
  }
}

gc()

# ============================================================================
# STEP 8: SAVE COMPREHENSIVE OUTPUTS
# ============================================================================
cat("\n💾 SAVING COMPREHENSIVE OUTPUTS:\n")
cat("=", rep("=", 35), "\n", sep="")

base_path <- "C:/Users/U078-N800/Desktop/Ml_Internship/Copy/clean_soil_dataset/"

gpkg_output <- paste0(base_path, "parcels_ALL_VARIABLES_enhanced_predictions2.gpkg")
st_write(final_dataset, gpkg_output, delete_dsn = TRUE, quiet = TRUE)
cat("✅ Complete dataset saved:", basename(gpkg_output), "\n")

csv_output <- paste0(base_path, "parcels_ALL_VARIABLES_predictions_attributes.csv")
st_drop_geometry(final_dataset) %>% write_csv(csv_output)
cat("✅ Attributes saved:", basename(csv_output), "\n")

metrics_output <- paste0(base_path, "ALL_VARIABLES_validation_metrics_summary.csv")
write_csv(all_metrics, metrics_output)
cat("✅ Metrics summary saved:", basename(metrics_output), "\n"
```
